### 2014-05-27 PALSE/28_05_2014_meeting.tex
PALSE meeting St Etienne. Looking for a yield stress.

\documentclass[xcolor=table]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[francais]{babel}
%\usetheme{Boadilla}
%\usecolortheme{rose}
%\usecolortheme{crane}
\usefonttheme{structuresmallcapsserif}
\setbeamertemplate{navigation symbols}{}

\definecolor{Main}{rgb}{0.74, 0.13, 0.19}
\definecolor{Accent1}{rgb}{0.76,0.36,0.13}
\definecolor{Accent2}{rgb}{0.54,0.1,0.4}

\usecolortheme{rose}
%\useinnertheme[shadow]{circles}
\usecolortheme{whale}
%\useoutertheme{infolines}

\usecolortheme[named=Accent1]{structure}

\setbeamercolor{alerted text}{fg=Accent2}
%\setbeamercolor{palette primary}{fg=white}
%\setbeamercolor{palette secondary}{bg=Accent1}
%\setbeamercolor{palette tertiary}{bg=Accent2,fg=white}

\setbeamerfont{page number in head/foot}{size=\large}
\setbeamercolor{page number in head/foot}{fg=Main}
% page/total
%\setbeamertemplate{footline}[frame number]
% pas de total
\setbeamertemplate{footline}{%
    	\hfill%
	\usebeamercolor[fg]{page number in head/foot}%
	\usebeamerfont{page number in head/foot}%
	\insertframenumber\kern1em\vskip2pt%
}

\setbeamersize{text margin left=1em}
\setbeamersize{text margin right=1em}

%font
\usepackage[T1]{fontenc}
\usepackage[oldstylenums]{kpfonts}

%proper math and math symbols
%\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{datenumber,fp}

\usepackage{siunitx}

\usepackage{tabu}
\usepackage{multirow}
\usepackage{booktabs}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

\usepackage{pgfplots}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{external}

%beamer related package

\usepackage{todonotes}
\presetkeys{todonotes}{inline}{}

%bibliography
\usepackage[style=authoryear-comp, language=british,eprint=false, url=false, doi=false, sortcites=true, sorting=none, isbn=false, firstinits=true,maxcitenames=6]{biblatex}
%minimal citations
\AtEveryCitekey{%
	\clearfield{title}
	\clearfield{pages}
	\clearfield{volume}
	\clearfield{number}
	\clearfield{month}}
\newcommand{\myfullcite}[1]{{\scriptsize\fullcite{#1}}}
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}
%\bibliography{biblio}

\newcolumntype{P}[1]{>{\raggedright}p{#1}}

\begin{document}
\tikzset{every mark/.append style={scale=0.8}}
\pgfplotsset{compat=1.3, every axis/.append style={footnotesize}}

\begin{frame}{Looking for a yield stress}

\begin{columns}
\column{0.5\textwidth}

\begin{block}{Upside-down test}
\begin{tabu}{b{3em}X[b]}
\begin{tikzpicture}
\fill[Accent1!50] (0,1em) rectangle (1em, 4em);
\draw (0,0) -- (0,4em) -- (1em, 4em) -- (1em,0);
\draw[->, Main] (0.5em,1.5em) -- +(0,-1em) node[below] {$\sigma$};
\draw[<->] (1.25em, 1em) -- +(0,3em) node[midway, right] {$h$};
\end{tikzpicture} &
Flows if $\alert{\sigma_Y} < \sigma = \rho g h$\linebreak
Actually, more complicated because surface tension
\end{tabu}
\end{block}

\structure{Cone-plate geometry}

\begin{center}
\begin{tikzpicture}
%sample
\fill[Accent1!50] (3em,0) arc (0:10:3em)  -- (170:3em) arc (170:180:3em) --cycle;
%cone
\fill[gray] (10:0.1em) -- (10:3em) -- (170:3em) -- (170:0.1em)  -- cycle;
\filldraw[gray]  (10:3em) +(0,0.1em) rectangle (170:3em);
%tige
\fill[gray] +(-0.5em,0.5em) rectangle +(0.5em, 3em);
\draw[->, Accent2] (-1em, 3em) arc (-180:0:1em and 0.5em) node[below right] {$\dot{\gamma}$};
%plate
\fill[gray] (-4em,0) rectangle (4em, -1em);
\end{tikzpicture}
\end{center}

\begin{itemize}
\item Can't spin too fast
\item Can't use ultrasound velocimetry
\end{itemize}

\column{0.5\textwidth}
Carbopol microgel\\
\textit{Divoux et al. 2012}
\includegraphics[width=\textwidth]{Divoux_Herschel–Bulkley.png}

\begin{block}{Herschel-Bulkley}
\[ \sigma = \alert{\sigma_Y} + K \dot{\gamma}^n \]
\end{block}
\end{columns}
\end{frame}

\end{document}

### 2014-05-27 PALSE/28_05_2014_meeting.tex
Herschel-Bukley, power law, Wall slip, must use Couette

@@
 
+\begin{frame}{Herschel-Bulkley fluid?}
+\[ \sigma = \alert{\sigma_Y} + K \dot{\gamma}^n \]
+\begin{columns}
+\column{0.5\textwidth}
+\includegraphics[width=\textwidth]{HershelBuckley_only_r1}
+\column{0.5\textwidth}
+\includegraphics[width=\textwidth]{HershelBuckley_loglog_r1}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Or power-law fluid?}
+\[ \sigma = K \dot{\gamma}^n\]
+\begin{columns}
+\column{0.5\textwidth}
+\includegraphics[width=\textwidth]{Power-law_logx_r1}
+\column{0.5\textwidth}
+\includegraphics[width=\textwidth]{Power-law_r1}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Or Herschel-Bulkley with slip at the wall?}
+\[ \sigma = \alert{\sigma_Y} + K \dot{\gamma}^n  \qquad\text{for}\quad\dot{\gamma}>\SI{1}{\per\second}\]
+\begin{columns}
+\column{0.5\textwidth}
+\includegraphics[width=\textwidth]{HershelBuckley_r1}
+\column{0.5\textwidth}
+\includegraphics[width=\textwidth]{HershelBuckley_slip_loglog_r1}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Yield stress vs concentration}
+\begin{center}
+\includegraphics[height=\textheight-3\baselineskip]{yield_stress_r1}
+\end{center}
+\end{frame}
+
+\begin{frame}{Time to use Couette geometry}
+\begin{columns}
+\column{0.35\textwidth}
+\begin{tikzpicture}
+	\draw circle[x radius=0.5\textwidth, y radius=0.1\textwidth] +(0,-5\baselineskip) circle;
+	\fill (0,0.015\textwidth) circle [x radius=0.4\textwidth, y radius=0.08\textwidth];
+	\draw[->, white, x radius=0.2\textwidth, y radius=0.05\textwidth]  (0,0.03\textwidth) +(90:0.05\textwidth) arc[start angle=90, end angle=270] node[right=1mm, font=\scriptsize] {$\dot{\gamma}$};
+	\draw (0.5\textwidth,0) -- +(0,-5\baselineskip) arc[x radius=0.5\textwidth, y radius=0.1\textwidth, start angle=0, end angle=-180] -- +(0,5\baselineskip);
+	\draw[dotted] (0.4\textwidth,0.015\textwidth) -- +(0,-5\baselineskip) arc[x radius=0.4\textwidth, y radius=0.09\textwidth, start angle=0, end angle=-180] -- +(0,5\baselineskip);
+\end{tikzpicture}
+
+\column{0.65\textwidth}
+\begin{itemize}
+\item Higher shear rates
+\item Less evaporation
+\item Possible to use Ultrasound velocimetry
+\begin{itemize}
+\item To quantify wall slip
+\item To observe instabilities at high shear rate
+\end{itemize}
+\item[But] much more sample needed (\SI{20}{\milli\litre})
+\end{itemize}
+\end{columns}
+\end{frame}
+
 \end{document}

### 2014-11-04 PALSE/head2body.tex
head to body ArXiV

\RequirePackage{fixltx2e}
\documentclass[prl,a4paper,twocolumn,superscriptaddress,showkeys]{revtex4}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
%\usepackage{kpfonts}

\usepackage{siunitx}
\usepackage{hyperref}

\begin{document}
\author{Hassan Srour}
\affiliation{Laboratoire de Chimie UMR CNRS 5182,
Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon, 
46 Allée d’Italie, 69007 Lyon, France}
\author{Olivier Ratel}
\affiliation{Ingénierie des Matériaux Polymères UMR CNRS 5223 
Universite Jean Monnet de Saint-Etienne/ Université de Lyon; 42023, Saint-Etienne, France}
\author{Mathieu Leocmach}
\affiliation{Laboratoire de Physique UMR CNRS 5672,
Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon,
46 Allée d’Italie, 69007 Lyon, France}
\author{Emma A. Adams}
\author{Sandrine Denis-Quanquin}
\author{Vinukrishnan Appukuttan}
\affiliation{Laboratoire de Chimie UMR CNRS 5182,
Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon, 
46 Allée d’Italie, 69007 Lyon, France}
\author{Nicolas Taberlet}
\author{Sébastien Manneville}
\affiliation{Laboratoire de Physique UMR CNRS 5672,
Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon,
46 Allée d’Italie, 69007 Lyon, France}
\author{Jean-Charles Majesté}
\author{Christian Carrot}
\affiliation{Ingénierie des Matériaux Polymères UMR CNRS 5223 
Universite Jean Monnet de Saint-Etienne/ Université de Lyon; 42023, Saint-Etienne, France}
\author{Chantal Andraud}
\author{Cyrille Monnereau}
\affiliation{Laboratoire de Chimie UMR CNRS 5182,
Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon, 
46 Allée d’Italie, 69007 Lyon, France}
\email{cyrille.monnereau@ens-lyon.fr}

\title{Mediating gel formation from structurally controlled poly(electrolytes) through multiple “head-to-body” electrostatic interactions}

\keywords{Polymers; ionic liquid; gel; self-assembly; rheology}

\begin{abstract}
Tuning the chain-end functionality of a short-chain cationic homopolymer, owing to the nature of the initiator used in the ATRP polymerisation step, can be used to mediate the formation of a gel of this poly(electrolyte) in water. While a neutral end group gives a solution of low viscosity, a highly homogeneous gel is obtained with a phosphonate anionic moiety, as characterized by rheometry and diffusion NMR. This novel type of supramolecular control over poly(electrolytic) gel formation could find potential use in a variety of applications in the field of electroactive materials.
\end{abstract}

\maketitle

Ionic liquids have been used for decades as solvents for a variety of synthetic applications, and have recently experienced a renewed interest in material chemistry~\cite{Antonietti2004,Hapiot2008,Ichikawa2011,MacFarlane2014}. Poly(ionic liquids) (PILs), in particular, show promising characteristics as electro-active materials, more specifically in the field of energy storage application (batteries, dye-sensitized solar cells or fuel cells) \cite{Ohno2001,Kim2011b,Mecerreyes2011,Zhao2011,Yuan2013}. The synthesis of structurally-controlled polymer architectures that incorporate Ionic-Liquid (IL) moieties, and the study of the electrochemical properties of such polymers have become very active topics, as witnessed by the impressive number of recent reports dealing with these two aspects~\cite{Hirao2000,Yuan2011,Matyjaszewski1999,Appukuttan2012}.
The incorporation of PILs into functional devices commonly requires their shaping into processable materials, such as thin films, membranes, or gels. The latter constitute an increasingly sought-after category of materials, as gel poly(electrolytes) are expected to substitute currently used volatile, flammable and leakage-prone organic solvent-based systems in future energy storage devices~\cite{Wang2003,Choudhury2009,Lewandowski2004}.

Different strategies have been proposed to mediate gel formation from a variety of polyelectrolytes~\cite{Katsampas2005,Xiong2012,Hu2014}. In most cases these strategies rely on the physical interaction of colloidal nanobjects (nanoparticles, micelles, highly crosslinked polymer networks) resulting in scattering colloidal gels with relatively poor mechanical and diffusion properties. By comparison, gels based on the supramolecular interactions of short polymeric chains constitute a very promising class of materials for practical applications, as they combine the advantages of polymeric materials (processability, solid-like behavior) and individual molecule solutions (efficient diffusion of solvent molecules, dynamic reorganization, and self-healing properties)\cite{Noro2011,Li2014,Furusho2014}. So far the few reported examples dealing with the formation of such ``supramolecular'' gels from polycations (anions) have all relied on a similar strategy, i.e. the addition of a molecule or a short polymer bearing multiple anionic (cationic) groups, which induces gel formation by the occurrence of a non-covalent crosslinking network~\cite{Zhao2013,Becht2011}.

In this paper, we propose a markedly different and more straightforward conceptual approach for the making of supramolecular homopolymer gels, based on electrostatic interactions between the positive charges held by the repeating cationic units (imidazolium) and a terminal anionic group (phosphonate) introduced in the initiation step of the atom transfer radical polymerisation (ATRP). We bring conclusive evidence that the single, relatively weak interaction between both entities is sufficient to mediate the formation of a gel in water, while in the absence of this interaction the aqueous polymer solution behaves like a Newtonian fluid of low viscosity.

\begin{figure}
\includegraphics[]{fig1}
\caption{Syntheses of PIm1 and PIm2 and of their intermediate.}
\label{fig:reaction}
\end{figure}

Tunability of the polymer chain end, which is a keystone element in the obtention of those gels, is a direct benefit of the methodology we specifically developed for the synthesis of our PILs: as an intrinsic feature of the ATRP used in the first step of our synthesis relies it is indeed possible to control the chemical nature of the polymer chain-end thanks to the molecular structure of the so-called polymerization initiator making it possible to readily incorporate any functionality. Thus, we chose to work with two different bromoisobutyryl ester derivatives: the phosphonate terminated one (In1) is used to provide the anion/cation interaction that is required to mediate gel formation, while the ethyl terminated one (In2) serves as a blank. In the case of In1, the phosphonate moiety was introduced in its ethyl-protected form, using a straightforward two-step methodology. Then, the ethylphosphonate- (PIm1) and ethyl-terminated (PIm2) polymers were synthesized according to the same general methodology as described in a previous report\cite{Appukuttan2012}. This methodology relies on the following three-step sequence consisting of i) ATRP of hydroxyethylacrylate followed by ii) nucleophilic substitution of the pendant hydroxyl group with bromide ions and iii) addition of methylimidazole, resulting in quaternarization of the nucleophile (Figure~\ref{fig:reaction}). Polymerization times were adjusted to get similar polymer chain lengths in both cases (n \textit{ca} 70 for POH1 and POH2), which was confirmed by \textsuperscript{1}H NMR and Steric exclusion chromatography (SEC) experiments. Deprotection of the phosphonic ester in POH1 could be readily and quantitatively achieved at the same step as the nucleophile OH-to-Br substitution, making the whole procedure leading to PBr1 extremely straightforward. POH2 was converted to PBr2 using this same procedure. After introduction of the imidazolium group, PIm1 and PIm2 could be isolated as white rubbery materials. Experimental details are provided as SI (S3-7)

\begin{figure}
\includegraphics[]{fig2}
\caption{Pictures of 10\% weight solutions of PIm1 and PIm2 in water before (left) and after (right) flipping of the vial.}
\label{fig:flip}
\end{figure}

When mixed with water at a weight concentration between typically 5\% to 30\%, it was noticed that PIm1 forms a very homogeneous and optically transparent (non-scattering) gel which does not flow under its own weight whereas PIm2 seems to behave like pure water (Figure~\ref{fig:flip}).

In order to characterize this behavior standard rheometry experiments were conducted. Figure~\ref{fig:yieldstress} shows the viscosity $\eta$ at different applied shear rates  in a cone-and-plate geometry at a fixed temperature of \SI{25}{\celsius}\cite{Macosko1994}. The difference between PIm1 and PIm2 is obvious. The viscosity of PIm2 is constant and close to that of pure water at \SI{25}{\celsius} showing that PIm2 behaves as a low-viscosity Newtonian fluid. Conversely, the viscosity of PIm1 always remains several orders of magnitude higher than that of PIm2, increases with decreasing shear rate (a behavior referred to as ``shear-thinning'') and keeps increasing without showing any sign of a Newtonian plateau at low shear rates. This divergence of the viscosity at vanishingly small shear rates indicates a solid-like behavior of the sample at rest as typically found in gels. More precisely, the inset in figure~\ref{fig:yieldstress} shows the shear stress $\sigma=\eta\dot{\gamma}$ as a function of the applied shear rate. Contrary to the case of PIm2, the shear stress for PIm1 tends towards a finite value $\sigma_c$ at low shear rates. This indicates that a critical force ($\sigma_c$) is required for the material to flow. This critical value is referred to as the ``yield stress'' in rheology\cite{Barnes1999} and ascertains the gel-like behavior of PIm1. The solid lines in the inset are the best fits to the data using a Herschel-Buckley law $\sigma = \sigma_c + A\dot{\gamma}^n$ and allow one to measure the values of the yield stress. It was found that PIm1 is a weak gel, with $\sigma_c = \SI{4.7}{\pascal}$ for a weight concentration of 22\% and $\sigma_c = \SI{0.4}{\pascal}$ for 8\%.

\begin{figure}
\includegraphics{fig3}
\caption{Shear rate dependence of the viscosity for PIm1 at 22\% (black) and 8\% (red) weight concentration and PIm2 at 10\% (blue). The black dashed line is the viscosity of pure water. Inset: Shear rate dependence of the shear stress. Continuous lines are best fits by a Herschel-Buckley model with a concentration-dependent yield stress. Note that the measurement range for PIm2 is smaller than that of PIm1 due to the resolution limit of the instrument (for stresses below \SI{1e-2}{\pascal}).}
\label{fig:yieldstress}
\end{figure}

The gel-like nature of PIm1 deduced from steady shear was also confirmed through oscillatory shear rheology using both amplitude sweeps at constant frequency and frequency sweeps within the linear regime. At a fixed angular frequency of \SI{0.1}{\radian\per\second}, PIm1 at 20 \% wt. shows an elasticity-dominated behavior ($G^\prime\gg G^{\prime\prime}$) until a stress amplitude of nearly \SI{20}{\pascal} and a liquid viscosity-dominated behavior ($G^\prime\ll G^{\prime\prime}$) above this value (Figure~\ref{fig:linearrheol}). At a stress amplitude of \SI{1}{\pascal}, a frequency sweep on PIm1 at 20 \% shows a behavior typical of a polymer network or soft glassy system\cite{Larson1999}, with nearly constant storage and loss moduli ($G^\prime\approx \SI{1000}{\pascal}$ and $G^{\prime\prime} \approx \SI{100}{\pascal}$ respectively, inset of Figure~\ref{fig:linearrheol}). Note that a shear stress of \SI{1}{\pascal} is well below $\sigma_c$, consistent with the solid-like behavior shown in Figure~\ref{fig:yieldstress}. A discrepancy still exists between the critical shear stress observed in steady shear and that observed in amplitude sweeps. This could be partly due to the definition chosen for the yield stress and to the different mode of solicitation\cite{Coussot2014}. In particular the loss modulus $G^{\prime\prime}$ is seen to increase significantly for stresses amplitudes above \SI{5+-1}{\pascal}, which is closer to the yield stress estimated above. In other systems such an increase followed by a maximum of the loss modulus has been attributed to structural rearrangements before the system yields\cite{Hyun2011}. In any case Figure~\ref{fig:linearrheol} clearly demonstrates the existence of a weak gel whose non-covalent crosslinks are greatly strain/stress-sensitive as similarly observed for filled polymers\cite{Harwood1965}. This property can be advantageously used, as it allows processing of the material through simple injection (video 1).

\begin{figure}
\includegraphics[]{fig4}
\caption{Storage ($G^\prime$) and loss ($G^{\prime\prime}$) moduli during oscillatory shear experiments for  PIm1 at 20 \%wt. $G^\prime$ (red) and $G^{\prime\prime}$ (black) vs stress amplitude during an amplitude sweep at a fixed angular frequency $\omega=\SI{0.1}{\radian\per\second}$. Inset: $G^\prime$ (red) and $G^{\prime\prime}$ (black) vs angular frequency $\omega$ during a frequency sweep at a fixed strain amplitude of 0.1\% (corresponding to a stress amplitude of \SI{1}{\pascal}).}
\label{fig:linearrheol}
\end{figure}

This striking difference in the behavior of PIm1 and PIm2 upon solubilization in water could be further demonstrated and characterized using 2D diffusion NMR (DOSY) (Figure~\ref{fig:DOSY}).

Both hydroxyl (POH1 and POH2, S9) and brominated precursors (PBr1 and PBr2) show diffusion coefficients that are, within error margin, undistinguishable from one another. From the diffusion data, it is possible to extract polydispersity indexes (PDIs)\cite{Vieville2011}. PDIs of 1.24 and 1.25 are found for POH1 and POH2 in D\textsubscript{2}O, respectively; this is slightly higher than what is given by SEC analyses in DMSO (1.14 and 1.16, resp.; S8). Similarly, values of 1.15 and 1.14 are measured for PBr1 and PBr2 in d6-DMSO.

\begin{figure}
\includegraphics[]{fig5}
\caption{DOSY spectra of \SI{10}{\gram\per\litre} solutions of PBr1 and PBr2 in d6-DMSO (top), PIm1 and PIm2 in D\textsubscript{2}O.}
\label{fig:DOSY}
\end{figure}

In contrast, DOSY spectra of PIm1 and PIm2 show marked differences. First the average diffusion coefficient of the diffusing polymer chain species of PIm1 is decreased by a 1.4 ratio (from \SIrange{35}{25}{\square\metre\per\second}) which clearly exceeds the error margin of the experiment. A possible explanation would be the significant difference in the weight of the two polymers. For a linear polymer chain of molecular weight $M$, the diffusion coefficient $D$ is assumed to obey the relation $M\propto D^{-d_F}$ with $d_F$, the fractal dimension of the polymer chain ($d_F$ values between $5/3$ for stretched-out, $3$ for collapsed polymer chains).  Considering a typical value $d_F = 2$, this result would suggest an $M$ value about twice as high for PIm1 as for PIm2.

Since PIm1 and PIm2 have similar molecular weights and compositions, attested by multiple SEC and NMR analyses of their synthesis precursors, the difference between their diffusion coefficients shows that PIm1 does not behave like a single, linear non-interacting polymer. For this sample the measured diffusion coefficient corresponds to a weighted average between the free and interacting chains, the exchange rate being fast compared to the diffusing time (\SI{200}{\milli\second}). Furthermore the presence of non-diffusing signals suggests that a significant proportion of the polymer chains are already immobilized, although DOSY experiments were recorded at polymer weight concentrations in D\textsubscript{2}O that were significantly lower than the critical gel formation concentration.

Altogether, these results suggest that the presence of this single terminal phosphonic acid at the extremity of each short polymer chain is enough to mediate efficient gel formation. This behavior can be rationalized on the basis of the model illustrated in Figure~\ref{fig:model}: owing to a single electrostatic interaction between their terminal phosphonate and lateral imidazolium group, each polymer chain is involved in a global supramolecular network; this gives rise, above a critical concentration limit, to cohesion forces that ``freeze'' the system and prevent it from flowing.

\begin{figure}
\includegraphics[]{fig6}
\caption{Proposed model accounting for the difference in behavior between PIm1 (left) and PIm2 (right); when the polyelectrolyte is terminated by a phosphonate ion, interaction with imidazolium moieties from proximal polymer chains induces the formation of a supramolecular network. In the absence of this interaction, polymer chains are freely flowing within the solution.}
\label{fig:model}
\end{figure}

The presence, along each polymer chain, of multiple cationic anchoring points for the single anionic end allows dynamic reorganization processes within the gel network. This provides i) stabilization of the gel structure in spite of the relatively weak single electrostatic interaction that connects each polymer chain to its neighbor and ii) avoids the formation of micro-domains, which makes the gel optically transparent. The nature of these ionic interactions was further assessed by complementary experiments: it was noticed that variations of pH and/or ionic strength considerably affected the polymer organization; while moderate acidification (HCl, pH=4) preserved the gel structure, basification (NaOH, pH=10), turned it into a viscous liquid. Besides, upon using physiological serum (NaCl aq, 0.9\% wt.), no gel formation was observed and an inhomogeneous solution of low viscosity was obtained (S10). Although further investigations will be required to get more insight on the precise underlying mechanisms, these preliminary experiments strongly suggest that screening of the interacting charges by the excess ions is involved in the gel collapse process. This property might open several perspectives related to the use of similar systems as stimuli responsive materials.

We believe that the concept introduced in this paper constitutes a straightforward alternative for the production of poly(electrolyte) gels, which could find potential use in a broad range of applications related to electrolytic materials and devices. Although illustrated herein in the specific case of electrostatic interaction between complementary anionic and cationic group, we hope that it could be extended to other types of supramolecular interactions, for instance with initiator/monomer combinations involving complementary H-bond donor/acceptor units.

\begin{acknowledgments}
The authors thank the Region Rhône Alpes and the Programme d'avenir Lyon- Saint Etienne (PALSE NoGELPo) for fundings and postoctoral grants for HS and ML. The Region Rhône Alpes (ARC 6-Energies program) is also acknowledged for a doctoral grant (OR). SM and NT acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013) / ERC grant agreement No. 258803. Dr. Marc-André Delsuc is warmly acknowledged for discussions regarding DOSY results.
\end{acknowledgments}

\bibliography{../../bib/PALSE.bib}

\end{document}

### 2015-12-09 ion_pairing.tex
Generalities and interpretation of moduli.

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{kpfonts}
\usepackage[version=4]{mhchem}
\usepackage{siunitx}

\author{Mathieu Leocmach}
\title{Ion pairing rheology}

\begin{document}
\maketitle

We are dealing with low-salt aqueous solutions of polyelectrolytes with cationic moieties along the chain and a single anionic head (Phosphonate). A chain has $N_0 = 70$ units from NMR measurements. The cationic moieties are either all Pyrrolidinium (\ce{Pyr+}) or all Imidazolium (\ce{Im+}) and we note the polymers \ce{PPyr+X-} and \ce{PIm+X-} respectively. The counterion \ce{X-} can be either Fluoride (\ce{F-}), Chloride (\ce{Cl-}), Bromide (\ce{Br-}) or Iodide (\ce{I-}). According to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected for harder cation (\ce{Pyr+}) and softer anion (\ce{F-}). Therefore \ce{PPyr+I-} should be the most charged polymer and \ce{PIm+F-} the least charged.

Experimentally, we are able to dissolve the five polymers that are the most charged, but neither the two fluorides nor \ce{PIm+Cl-}. This means that in absence of charge dissociation the chain is in poor solvent. Furthermore, we held the two fluoride polymers in water at boiling temperature overnight without any dissolution. The $\Theta$ temperature is thus higher than \SI{100}{\celsius}. Nonetheless, due to charge dissociation, most of our polymers are soluble at \SI{25}{\celsius}.

Due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible. If every chain has a single foreign head attached to its body, every polymer is linked to two others and we obtain a single effective chain of effective polymerisation index $n N_0$ where $n$ is the number of polymers in the effective chain. Here we have supposed that in order to minimize the inter-chain repulsion between charged cationic groups, heads are preferentially attached to the tail of their neighbour.

If as least two foreign heads attach to the same body, we obtain an effective cross-linking point. When the probability of such a configuration is non-zero, we obtain a physically cross-linked gel. Indeed, we observe that our 5 soluble polymers form soft solid when mixed with water. Submitted to sufficient shear stress these gels flow but conserve their rheological properties before and after the flow, confirming the physical nature of the gel.

Let us now discuss the rheological measurements and link them to microstructure. Amplitude sweep measurements at \SI{1}{\hertz} in a cone-plate geometry give access to two main observables: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e. the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, the end of the linear regime and the first broken bonds.

The spatial conformation of a polyelectrolyte (super)chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions. Between each cross-link point the (super)chain can be considered as an entropic spring made of $N=n N_0$ monomers. The shear modulus of the gel is then given by:
\begin{equation}
G = \frac{c}{N}k_\mathrm{B}T,
\label{eq:G}
\end{equation}
where $c$ is the monomer number density, $k_\mathrm{B}$ is the Boltzmann constant and $T$ the temperature. Knowing the molecular mass $M$ of the chains, density $d$ of the solvent and the weight fraction $w$ of polymer, we obtain
\begin{equation}
n = \frac{\mathcal{N_A}}{M} w d \frac{k_\mathrm{B}T}{G}.
\end{equation}

We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 780 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation. This is coherent with a higher probability of attaching two or more heads on a highly dissociated body, i.e. a large cross-linking ratio.

\end{document}

### 2015-12-11 PALSE/ion_pairing.tex
Interpretation of gamma_c, relation between n and A, figures

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{kpfonts}
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage[table]{xcolor}
\usepackage{pgfplots}
\usetikzlibrary{calc}

\author{Mathieu Leocmach}
\title{Ion pairing rheology}

\begin{document}
\maketitle

We are dealing with low-salt aqueous solutions of polyelectrolytes with cationic moieties along the chain and a single anionic head (Phosphonate, \ce{\bond{-}PO3^2-}). A chain has $N_0 = 70$ units from NMR measurements. The cationic moieties are either all Pyrrolidinium (\ce{Pyr+}) or all Imidazolium (\ce{Im+}) and we note the polymers \ce{PPyr+X-} and \ce{PIm+X-} respectively. The counterion \ce{X-} can be either Fluoride (\ce{F-}), Chloride (\ce{Cl-}), Bromide (\ce{Br-}) or Iodide (\ce{I-}). According to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected for harder cation (\ce{Pyr+}) and softer anion (\ce{F-}). Therefore \ce{PPyr+I-} should be the most charged polymer and \ce{PIm+F-} the least charged.

Experimentally, we are able to dissolve the five polymers that are the most charged, but neither the two fluorides nor \ce{PIm+Cl-}. This means that in absence of charge dissociation the chain is in poor solvent. Furthermore, we held the two fluoride polymers in water at boiling temperature overnight without any dissolution. The $\Theta$ temperature is thus higher than \SI{100}{\celsius}. Nonetheless, due to charge dissociation, most of our polymers are soluble at \SI{25}{\celsius}. The reduced temperature $\tau = 1 - T/\Theta$ is thus at least 0.2.

Due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible. If every chain has a single foreign head attached to its body, every polymer is linked to two others and we obtain a single effective chain of effective polymerisation index $n N_0$ where $n$ is the number of polymers in the effective chain. Here we have supposed that in order to minimize the inter-chain repulsion between charged cationic groups, heads are preferentially attached to the tail of their neighbour. We also neglect the possibility of the two charges of the phosphonate head to attach to two separate chains.

If as least two foreign heads attach to the same body, we obtain an effective cross-linking point. When the probability of such a configuration is non-zero, we obtain a physically cross-linked gel. Indeed, we observe that our 5 soluble polymers form soft solid when mixed with water. Submitted to sufficient shear stress these gels flow but conserve their rheological properties before and after the flow, confirming the physical nature of the gel.

Let us now discuss the rheological measurements and link them to microstructure. Amplitude sweep measurements at \SI{1}{\hertz} in a cone-plate geometry give access to two main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e. the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, the end of the linear regime and the first broken cross-links.

\begin{table}
\begin{tabular}{l|ll|lll}
& $G^\prime(\gamma\rightarrow 0)$ & $\gamma_c$ & $n$ & $E_c$ & $A$\\
&	\si{\pascal} &  & & \si{\kilo\joule/\mol} & \\\hline&&&&\\[-10pt]
\ce{PPyr+I-}	& 8800	&	0.004	&	1	&	\cellcolor{gray!25}0.004	&	1\\
\ce{PPyr+Br-}	& 9800 	& 	0.0016	&	1.05&	\cellcolor{gray!25}0.001	&	1.1\\
\ce{PPyr+Cl-}	& 270 	&	0.009	&	45	&	\cellcolor{gray!25}0.02	&	7\\
\ce{PIm+I-}	& 78	&	0.016	&	113	&	\cellcolor{gray!25}0.063	&	9.6\\
\ce{PIm+Br-}	& 13	&	8	&	790	&	158	&	\cellcolor{gray!25}3.4\\
\end{tabular}
\caption{Results. Incoherent values are on gray background.}
\label{tab:results}
\end{table}

The spatial conformation of a polyelectrolyte (super)chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions, see Figure~\ref{fig:persistent}. Between each cross-link point the (super)chain can be considered as an entropic spring made of $N=n N_0$ monomers. The shear modulus of the gel is then given by:
\begin{equation}
G = \frac{c}{N}k_\mathrm{B}T,
\label{eq:G}
\end{equation}
where $c$ is the monomer number density, $k_\mathrm{B}$ is the Boltzmann constant and $T$ the temperature. Knowing the molecular mass $M$ of the chains, density $d$ of the solvent and the weight fraction $w$ of polymer, we obtain
\begin{equation}
n = \frac{\mathcal{N_A}}{M} w d \frac{k_\mathrm{B}T}{G}.
\end{equation}

We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 780 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}. This is coherent with a higher probability of attaching two or more heads on a highly dissociated body, i.e. a large cross-linking ratio.

Considering that each super chain is an entropic spring in its linear domain, the mechanical energy stored by a single super chain at a strain $\gamma$ is $E = \gamma^2 k_\mathrm{B}T$. If we suppose that the ionic bonds break before the complete extension of the (super)chains, we evaluate the head-to-body bonding energy to
\begin{equation}
E_c = \gamma_c^2 k_\mathrm{B}T.
\end{equation}

For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 64 k_\mathrm{B}T$ per bond or \SI{158}{\kilo\joule\per\mol}, coherent with typical energies of ionic bonds. However for all other samples the critical strain is between 2\% and 20\%, leading to incoherent values for the bonding energies between $10^{-3}k_\mathrm{B}T$ and $10^{-1}k_\mathrm{B}T$ per bond, see Table~\ref{tab:results}. Therefore, the hypothesis of a linear spring fails for these four samples, meaning that the super chain is fully extended at $\gamma_c$.

\begin{figure}
\begin{center}
\begin{tikzpicture}
\begin{axis}[
	scale only axis,
	width=12\baselineskip, height=6\baselineskip, 
	axis lines=none,
	xmin=0.5, xmax=4.5,
	ymin=-0.7, ymax=1.3,
	]
\addplot[only marks, mark=o, mark size=2] table {/home/mathieu/tex/PALSE/persistent.xy} coordinate[pos=0.43] (a) coordinate[pos=0.57] (b) coordinate[pos=0.81](c);
\draw[<->] ($(a)+(-0.5em,0)$) -- ($(b)+(-0.5em,0)$) node[midway, left]{$r_\mathrm{scr}$};
\draw[<->] ($(c)+(-0.25em,-0.5em)$) -- ($(c)+(0.25em,-0.5em)$) node[midway, below]{$D$};
\end{axis}
\end{tikzpicture}
\end{center}
\caption{Sketch of the persistent random walk.}
\label{fig:persistent}
\end{figure}

For polyelectrolytes in bad solvent the persistent random walk in made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:persistent}:
\begin{align}
g_e &= \frac{A^2}{u}\tau\label{eq:ge},\\
D &= b \left(\frac{A^2}{u}\right)^{1/3}\label{eq:D}.
\end{align}
Here $A$ is the number of monomers between uncondensed charges, $\tau$ the reduced temperature and $u = \ell_\mathrm{B}/b \approx 2.7$ is the ratio of Bjerrum length $\ell_\mathrm{B}$ and monomer size $b\approx \SI{0.26}{\nano\metre}$.

Extending these electrostatic blobs means exposing more monomers to the bad solvent and typically requires energies at least an order of magnitude higher than for the extension of the persistent random walk. We will thus consider that the exit from the linear domain of the material corresponds to the full extension of the persistent random walk:
\begin{equation}
\gamma_c = \frac{R_0}{R} - 1
\end{equation}
where $R$ and $R_0$ are respectively the end-to-end distance and the contour length of the persistent random walk.

A segment of the persistent random walk is $r_\mathrm{scr}$ long and thus contains $g_\mathrm{scr} = g_e r_\mathrm{scr}/D$ monomers. We thus have
\begin{align}
R &= r_\mathrm{scr} \left(\frac{N}{g_\mathrm{scr}}\right)^{1/2},\\
R_0 &= r_\mathrm{scr} \frac{N}{g_\mathrm{scr}},\\
\gamma_c &= \left(\frac{N}{g_\mathrm{scr}}\right)^{1/2} -1\label{eq:gamma0}.
\end{align}

One can express the screening length function of $A$ and $\tau$ as
\begin{equation}
r_\mathrm{scr} = \left(\frac{A^2}{u}\right)^{1/3} \left(\frac{\tau}{cb}\right)^{1/2}
\label{eq:rsrc}
\end{equation}

From equations (\ref{eq:ge}), (\ref{eq:D}) and (\ref{eq:rsrc}) we deduce
\begin{equation}
g_\mathrm{scr} = \frac{A^2}{u} \left(\frac{\tau^3}{c b^3}\right)^{1/2}\label{eq:gscr}.
\end{equation}

From equations (\ref{eq:gamma0}) and (\ref{eq:gscr}) we can finally relate the measured critical strain to the microscopic parameter $A$:
\begin{equation}
A = \frac{\left(N u\right)^{1/2}}{\gamma_c+1}\left(\frac{c b^3}{\tau^3}\right)^{1/4}.
\end{equation}

The only unknown here is the reduced temperature $0.2<\tau<1$, possibly different between Poly(pyrrolidinium) based and Poly(immidazolium) based polymers. Setting $\tau\approx 0.235$, corresponding to $\Theta\approx\SI{116}{\celsius}$, yields $A=1$ for both \ce{PPyr+I-} and \ce{PPyr+Br-}, indicating full dissociation of the charges. By contrast \ce{PPyr+Cl-} has a dissociated charge every 7 monomers. Supposing the same $\tau$, \ce{PIm+I-} has a dissociated charge every 10 monomers, see Table~\ref{tab:results}.

There must be a link between the charge dissociation and and the amount of physical cross-link. We assume that the probability of having $k$ charge dissociated on a polymer follows a Poisson distribution of mean $\lambda = N_0/A$. Having 2 charges dissociated on a chain implies a cross-link. Therefore, we the cross-link probability $1/n$ is proportional to $\lambda^2$, yielding
\begin{equation}
n \sim A^2.
\end{equation}

This is indeed what we observe, as shown in Figure~\ref{fig:nA}.

\begin{figure}
\begin{center}
\begin{tikzpicture}
\begin{loglogaxis}[
	xlabel = $A$,
	ylabel = $n$,
	]
	\addplot[only marks] coordinates {(1.090233298, 72.59) (6.98656403, 3141.79) (1.000001279, 69.17) (9.589224893, 7876.54)};
	\addplot[no marks, domain=1:10] {70*x^2};
\end{loglogaxis}
\end{tikzpicture}
\end{center}
\caption{Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points. The line corresponds to power law of exponent 2 and prefactor $N_0$ XXX not understood XXX.}
\label{fig:nA}
\end{figure}

\end{document}

### 2015-12-11 PALSE/ion_pairing.tex
Switch to revtex, add bibliography, headings. S columns. More precision on gamma_c for PImBr. Actually plot n rather than N, thus the prefactor N0 disappears. Check that the gamma_c for PImBr is actually smaller than the one predicted.

@@
-\documentclass[12pt,a4paper]{article}
+\documentclass[12pt,a4paper,prl,reprint]{revtex4-1}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 \usepackage{kpfonts}
@@
 
+\begin{document}
 \author{Mathieu Leocmach}
 \title{Ion pairing rheology}
+\maketitle
 
-\begin{document}
-\maketitle
+\section{System description}
 
@@
 
-Due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible.
+We have previously demonstrated that, due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible~\cite{Srour2014}.
 If every chain has a single foreign head attached to its body, every polymer is linked to two others and we obtain a single effective chain of effective polymerisation index $n N_0$ where $n$ is the number of polymers in the effective chain.
@@
 
+\section{Reological measurements}
+
 Let us now discuss the rheological measurements and link them to microstructure.
@@
 \begin{table}
-\begin{tabular}{l|ll|lll}
-& $G^\prime(\gamma\rightarrow 0)$ & $\gamma_c$ & $n$ & $E_c$ & $A$\\
-&	\si{\pascal} &  & & \si{\kilo\joule/\mol} & \\\hline&&&&\\[-10pt]
-\ce{PPyr+I-}	& 8800	&	0.004	&	1	&	\cellcolor{gray!25}0.004	&	1\\
-\ce{PPyr+Br-}	& 9800 	& 	0.0016	&	1.05&	\cellcolor{gray!25}0.001	&	1.1\\
-\ce{PPyr+Cl-}	& 270 	&	0.009	&	45	&	\cellcolor{gray!25}0.02	&	7\\
-\ce{PIm+I-}	& 78	&	0.016	&	113	&	\cellcolor{gray!25}0.063	&	9.6\\
-\ce{PIm+Br-}	& 13	&	8	&	790	&	158	&	\cellcolor{gray!25}3.4\\
+\begin{tabular}{l|SS|SSS}
+& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_c$} & {$n$} & {$E_c$} & {$A$}\\
+&	{\si{\pascal}} &  & & {\si{\kilo\joule/\mol}} & \\\hline&&&&\\[-10pt]
+\ce{PPyr+I-}	& 8800	&	0.004	&	1.	&	0.004\cellcolor{gray!25}	&	1.\\
+\ce{PPyr+Br-}	& 9800 	& 	0.0016	&	1.05&	0.001\cellcolor{gray!25}	&	1.1\\
+\ce{PPyr+Cl-}	& 270 	&	0.009	&	45	&	0.02\cellcolor{gray!25}	&	7.\\
+\ce{PIm+I-}	& 78	&	0.016	&	113	&	0.063\cellcolor{gray!25}	&	9.6\\
+\ce{PIm+Br-}	& 13	&	8.04	&	790	&	160.	&	3.4\cellcolor{gray!25}\\
 \end{tabular}
@@
 
-The spatial conformation of a polyelectrolyte (super)chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions, see Figure~\ref{fig:persistent}.
+\subsection{From modulus to effective chain length}
+
+The spatial conformation of a polyelectrolyte (super)chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions~\cite{Rubinstein1996}, see Figure~\ref{fig:persistent}.
 Between each cross-link point the (super)chain can be considered as an entropic spring made of $N=n N_0$ monomers.
@@
 
-We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 780 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
+We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 790 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
 This is coherent with a higher probability of attaching two or more heads on a highly dissociated body, i.e.
 a large cross-linking ratio.
+
+\subsection{Bounding energy}
 
@@
 
-For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 64 k_\mathrm{B}T$ per bond or \SI{158}{\kilo\joule\per\mol}, coherent with typical energies of ionic bonds.
+For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per bond or \SI{160}{\kilo\joule\per\mol}, coherent with typical energies of ionic bonds.
 However for all other samples the critical strain is between 2\% and 20\%, leading to incoherent values for the bonding energies between $10^{-3}k_\mathrm{B}T$ and $10^{-1}k_\mathrm{B}T$ per bond, see Table~\ref{tab:results}.
@@
 
-For polyelectrolytes in bad solvent the persistent random walk in made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:persistent}:
+\subsection{From critical strain to charge dissociation}
+
+For polyelectrolytes in bad solvent the persistent random walk in made of electrostatic blobs of diameter $D$ containing $g_e$ monomers~\cite{Rubinstein1996}, see Figure~\ref{fig:persistent}:
 \begin{align}
@@
 
-One can express the screening length function of $A$ and $\tau$ as
+Following~\cite{Rubinstein1996}, we can express the screening length function of $A$ and $\tau$ as
 \begin{equation}
@@
 A = \frac{\left(N u\right)^{1/2}}{\gamma_c+1}\left(\frac{c b^3}{\tau^3}\right)^{1/4}.
+\label{eq:A}
 \end{equation}
@@
 Supposing the same $\tau$, \ce{PIm+I-} has a dissociated charge every 10 monomers, see Table~\ref{tab:results}.
+
+\subsection{Charge dissociation and cross-linking ratio}
 
@@
 This is indeed what we observe, as shown in Figure~\ref{fig:nA}.
+We can extrapolate this law for the case of \ce{PIm+Br-} and obtain $A = 28$.
+Inverting equation~(\ref{eq:A}) yields $\gamma_c = 860\%$, indeed higher than the one observed (804\%).
+This confirms that in the \ce{PIm+Br-} strain sweep we were actually probing the strength of the ionic bonds.
 
@@
 	ylabel = $n$,
-	]
-	\addplot[only marks] coordinates {(1.090233298, 72.59) (6.98656403, 3141.79) (1.000001279, 69.17) (9.589224893, 7876.54)};
-	\addplot[no marks, domain=1:10] {70*x^2};
+	]1.04, 44.88, 0.99, 112.52
+	\addplot[only marks] coordinates {(1.090233298, 1.04) (6.98656403, 44.88) (1.000001279, 0.99) (9.589224893, 112.52)};
+	\addplot[no marks, domain=1:10] {x^2};
 \end{loglogaxis}
@@
 \end{center}
-\caption{Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points.
-The line corresponds to power law of exponent 2 and prefactor $N_0$ XXX not understood XXX.}
+\caption{Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points. The line corresponds to $A^2$.}
 \label{fig:nA}
@@
 
+\bibliography{PALSE}
+
 \end{document}

### 2015-12-17 PALSE/ion_pairing.tex
Conversion from MS Word (neither figure nor bibliography) of the no-theory version

@@
 &	{\si{\pascal}} &  & & {\si{\kilo\joule/\mol}} & \\\hline&&&&\\[-10pt]
-\ce{PPyr+I-}	& 8800	&	0.004	&	1.	&	0.004\cellcolor{gray!25}	&	1.\\
-\ce{PPyr+Br-}	& 9800 	& 	0.0016	&	1.05&	0.001\cellcolor{gray!25}	&	1.1\\
-\ce{PPyr+Cl-}	& 270 	&	0.009	&	45	&	0.02\cellcolor{gray!25}	&	7.\\
-\ce{PIm+I-}	& 78	&	0.016	&	113	&	0.063\cellcolor{gray!25}	&	9.6\\
+\ce{PPyr+I-}	& 8800	&	0.04	&	1.	&	0.004\cellcolor{gray!25}	&	1.\\
+\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.05&	0.001\cellcolor{gray!25}	&	1.1\\
+\ce{PPyr+Cl-}	& 270 	&	0.09	&	45	&	0.02\cellcolor{gray!25}	&	7.\\
+\ce{PIm+I-}	& 78	&	0.16	&	113	&	0.063\cellcolor{gray!25}	&	9.6\\
 \ce{PIm+Br-}	& 13	&	8.04	&	790	&	160.	&	3.4\cellcolor{gray!25}\\
@@
 There must be a link between the charge dissociation and and the amount of physical cross-link.
-We assume that the probability of having $k$ charge dissociated on a polymer follows a Poisson distribution of mean $\lambda = N_0/A$.
+If we assume that the probability for a charge to dissociate is independent of the number of charges already present on the polymer, we can describe the probability of having $k$ charge dissociated on a polymer by a Poisson distribution of mean $\lambda = N_0/A$.
 Having 2 charges dissociated on a chain implies a cross-link.

### 2015-12-17 ion_pairing_paper.tex
Conversion from MS Word (neither figure nor bibliography) of the no-theory version

\documentclass[journal=jacsat,manuscript=article]{achemso}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}       % Use modern font encodings
\usepackage{amssymb}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage{siunitx}

\author{Hassan Srour}
\author{Martien Duvall Deffo Ayagou}
\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}

\author{Mathieu Leocmach}
\affiliation[Institut Lumière Matière]{Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}
\email{mathieu.leocmach@univ-lyon1.fr}

\author{Thi Thanh-Tam Nguyen}
\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}

\author{Nicolas Taberlet}
\author{Sebastien Manneville}
\affiliation[Laboratoire de Physique de l'ENS de Lyon]{Laboratoire de Physique, Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon, 46 Allée d'Italie, 69007 Lyon}

\author{Chantal Andraud}
\author{Cyrille Monnereau}
\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}
\email{cyrille.monnereau@ens-lyon.fr}

\title{Controlling the rheological properties of well-defined polyelectrolytes hydrogels by fine tuning their ion pairing process}

\keywords{ATRP, hydrogel, poly(cations), rheology}

\begin{document}

\begin{tocentry}

9\,cm by 3.5\,cm

\end{tocentry}

\begin{abstract}
Starting from a common polymeric platform with lateral bromine substitution and phosphonate end group, it is possible, through two subsequent synthetic steps to generate a variety of poly(ionic liquids) with various cation/anion pair combinations. Upon water addition, the behavior of the resulting hydrogels is seen to be strongly dependent on the polymer structure, illustrating the possibility to control the gel formation properties by fine tuning of the intra and intermolecular ion pairing process.
\end{abstract}

\section{Introduction}
Since the pioneering work of Wichterle et al.,1 which established their relevance in a biomedical context, polymer based hydrogels have never ceased to be a very active field of research.2, 3 They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargos or scaffolds for wound healing or tissues repair.4 Polyelectrolytes are being extensively used in this particular context.5 Their gel formation mechanism often involves reversible-by-nature electrostatic interactions, which can be used for instance to trigger ``smart'' release of bioactive substances.6 Besides, poly(cationic) gel have been reported to combine scaffold properties for cell’s adherence and growth with antimicrobial activity, and are therefore increasingly used in clinical applications.3, 7 In this framework, injectability is a key feature, as it provides an easier way to gel delivery in vivo; thus, reversible shear thinning biocompatible hydrogels are a particularly sought after class of materials.8
We have recently reported on a highly reversible poly(cation) based hydrogel,9 which formation relies on an new concept of supramolecular  electrostatic interaction. Briefly, the atom transfer radical polymerisation (ATRP) is initiated by a phosphonate terminated ATRP initiator, eventually affording an anion terminated poly(cationic) polymer. The resulting poly(cationic) polymer, when dispersed in an aqueous medium, stabilizes hydrogel formation through the occurrence of a head(anion)-to-body(cations) supramolecular network. This highly dynamic electrostatic network provides the resulting gel with spectacular mechanical and self-healing properties. Moreover the interaction can be disrupted by various chemical stimuli, such as pH or ionic strength.
In the present report, we show that the mechanical/rheological properties of such smart hydrogels can be tuned to an extremely large extent by playing with the nature (hardness/softness) of the cationic repeating unit and the ion-paring strength of the associated counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}). The viscoelastic and diffusion properties of the resulting materials are characterized by different rheological and calorimetric analyses, evidencing a strengthening of the gel network upon reinforcing the hard character of the cation and the soft character of the anionic counterion.

\section{Results and discussion}

\subsection{Synthesis and characterizations}

The phosphonate terminated polymer \ce{PBr}, which is the common basis to all studied polymer was synthesized according to our previously reported methodology; allowing us to obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; S2, S3, Figure S1) .9, 10 All subsequent transformations of the polymer structures were achieved using the same initial batch of material. Nucleophilic addition of N-Methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF afforded the corresponding poly(imidazloium) and poly(pyrrolidinium) compounds. In these cases, as an inherent consequence of the structure of the starting material, charge neutrality is provided by bromide counter ions, and the polymer will be referred to as \ce{PIm+Br-} and \ce{PPyr+Br-}, respectively (Scheme~\ref{sch:synthesis}).

\begin{scheme}
\caption{Synthesis of \ce{PIm+Br-} and \ce{Ppyr+Br-} and their intermediate. $n=70$}
\label{sch:synthesis}
\end{scheme}

Anionic metathesis could be achieved by pouring an aqueous solution of \ce{PIm+Br-} or \ce{PPyr+Br-} into a saturated aqueous solution of the different sodium halides (\ce{NaF}, \ce{NaCl}, \ce{NaI}) (Scheme~\ref{sch:metathesis}). In the cases of \ce{PIm+F-}, \ce{PIm+Cl-} , \ce{PPyr+F-} and \ce{PPyr+Cl-}, a turbid suspension was obtained immediately after addition. After extensive dialysis of the resulting mixture against deionized water and lyophilization,, the different \ce{PIm+X-} and \ce{PPyr+X-} (where \ce{X}=\ce{F}, \ce{Cl}, \ce{I}) were recovered in high yields. In order to bring evidence for the efficiency of the ionic metathesis, the resulting materials were submitted to negative mode high resolution mass spectrometry (HRMS, S3). Displacement of bromide counterions could be unambiguously assessed by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Figures S8-9). Substitution by iodide was further confirmed by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Figures S6-7), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (see supporting information).

\begin{scheme}
\caption{Anionic metathesis reaction.}
\label{sch:metathesis}
\end{scheme}

Calorimetric data were then recorded in order to characterize the effect of counterion substitution. In solids, melting points are inherently connected to the reticular enthalpy of the crystal lattice, hence primarily to ion pairing strength.11 This last parameter is directly related to the diffuse and hindered nature of the interacting charges, bulkier and softer ions providing in the most general case weaker interactions.12 For typical ``ionic liquids'' such as imidazolium or pyrollidinium halides, melting points will thus gradually increase as the size of the halide counterion decreases, ie $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$.13 Although the trend is less obvious in amorphous polymeric systems, in which phase transitions result from the interplay of multiple parameters, ionic substitution is however expected to have an impact on the melting temperature of the material.14

All polymers were thus studied by Differential Scanning Calorimetry (DSC). As expected, the lowest melting temperature was achieved for polymers with iodide counterions. Quite remarkably, a similar melting temperature (84-\SI{85}{\celsius}) was found for \ce{PIm+I-} and \ce{PPyr+I-}. When bromide counterions are present, a significant increase of the melting temperature was observed (114-\SI{115}{\celsius}) but again, a similar value was found for \ce{PIm+Br-} and \ce{PPyr+Br-}. In the case of fluoride and chloride, no phase transition was seen below decomposition temperature of the polymer (\SI{200}{\celsius}), which indicates a much stronger ion pairing in these materials (Figure~\ref{fig:dsc}, S4 Figures S2-5).

\begin{figure}
\caption{DSC analysis of a crystalline polymer samples showing the presence of melting point.}
\label{fig:dsc}
\end{figure}

\subsection{Gels preparation}
All materials were placed in vials, and water was added until a final aqueous dispersion of each polymer at a 8\%wt. concentration was obtained. Swelling properties turned out to be extremely dependent on the polymer compositions. With fluoride counterions, no observable swelling or dissolution took place and the suspension was recovered as a biphasic mixture of its individual components. Results were very similar for the chloride containg Poly(immidazolium) polymer, whereas a limited swelling was observed for its poly(pyrrolidinium) counterpart, resulting in a granular, inhomogeneous gel. When bromide was present, both polypyrrolidinium and polyimmidazolium polymers afforded a gel upon swelling with water. However, their appearance markedly differed, as the pyrrolidinium based system gave rise to a much harder and glassy-like material, indicating a consistently stronger interchain network. A similar gel behavior was found with \ce{PPyr+I-}, while \ce{PIm+I-}, although very similar in appearance to \ce{PIm+Br-} (homogeneous and optically transparent) turned out to be significantly harder when submitted to mechanical shearing (Figure~\ref{fig:pictures}).

\begin{figure}
\caption{pictures of 8\%wt water solutions of \ce{PIm+X-} and \ce{PPyr+X-} in water.}
\label{fig:pictures}
\end{figure}

Rheological measurements. In order to refine these visual observations, we performed rheological studies of the water-swelling materials. Briefly, data were recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}.15, 16 The sample was placed on the plate, the cone was then lowered to the measuring position, spreading the sample in the process. Excess material was removed, so that the sample exactly filled the gap. Due to Weissenberg effect at high shear rate, no preshear was applied before measurement. To minimize water absorption, the geometry was covered with a solvent trap, using light mineral oil as a liquid seal between the rotor and the cap. After one minute of equilibration we performed an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$). For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile was that of pure water which confirms the visual observation. However as shown in Figure~\ref{fig:strain}, all other samples are solidlike ($G^\prime \gg G"$) at small strain and flow at large strains ($G" \gg G^\prime$). Solidlike behavior at low strain was confirmed for all accessible frequencies (see supplementary material S6 Fig. 1). We define the critical strain value $\gamma_c$ between these two regimes when $G"$ reaches a maximum. We checked that the linear mechanical properties are unchanged by the flow history given a few minutes of rest. %To understand the influence of the ion pairing on the mechanical properties, we compute the cohesive energy (equation 1) that should probe the strength of the physical supramolecular bonds.15,16

\begin{figure}
\caption{Storage modulus $G^\prime$ (filled symbols and solid lines) and loss modulus $G"$ (open symbols and dashed lines) measured through oscillatory shear experiments for (a) \ce{PPyr+Cl-} ($\bullet$), \ce{PPyr+Br-} ($\blacksquare$) and \ce{PPyr+I-} ($\blacktriangle$), (b) \ce{PIm+Br-} ($\blacklozenge$) and \ce{PPyr+Br-} ($\blacksquare$) and (c) \ce{PIm+I-} ($\blacktriangledown$) and \ce{PPyr+I-} ($\blacktriangle$). $G^\prime$  and $G"$ measured during an amplitude sweep at a fixed frequency $f=\SI{1}{\hertz}$ are plotted against the strain amplitude $\gamma$. All samples are at 8\%wt }
\label{fig:strain}
\end{figure}

In Figure~\ref{fig:strain}a we compare the rheological responses of Poly(pyrrolidinium) based hydrogels for the three counterions allowing dissolution in water. Gels with bromide and iodide counterions showed roughly the same mechanical behavior in the linear regime but \ce{PPyr+I-} yields at slightly higher strains ($\gamma_c=4\%$ instead of $3\%$). By contrast, the gel with chloride counterions is weaker by more than one order of magnitude. The cohesive energy goes down with smaller counterion indicating weaker intermolecular bonds (see supplementary material S6 table 1).

We also compare the rheology of \ce{PIm+X-} and \ce{PPyr+X-} with the same counterion \ce{X} in Figure~\ref{fig:strain}b,c. For both bromide (Figure~\ref{fig:strain}b, X=Br) and iodide (Figure~\ref{fig:strain}c, X=I), imidazolium based gels are much weaker than their pyrrolidinium counterpart.

\subsection{Discussion}

The above dependence of the gel parameters on the nature of the counterions could be rationalized on the basis of ion-pairing processes. The unsolubility/unswellability of \ce{PIm+F-}, \ce{PPyr+F-} and \ce{PIm+Cl-} suggest that ion pairs are not efficiently dissociated in aqueous medium, thus hindering the solvation process. By comparison, \ce{PIm+Br-}, \ce{PIm+I-}, \ce{PPyr+Br-} and \ce{PPyr+I-} form homogeneous materials when suspended in water, which one can correlate with efficient dissociation of the halide counterions from the polymer main chain is fully consistency with decreased ionic bond strength. Such dissociation, in turns makes it possible for the ``hard'' phosphonate group on each polymer chain's end to interact with the ``free'' cationic groups of the side chains, which we showed in a previous paper to be at the origin of the gel formation mechanism.9

The similarity in gel parameters observed for \ce{PPyr+Br-} and \ce{PPyr+I-} ($G^\prime\approx\SI{8500}{\pascal}$) suggests that full ion pair dissociation is achieved in the latter, or at least that this dissociation is enough in both cases to provide maximal interaction between the terminal phosphonate and the pyrrolidinium side groups. Conversely the significantly different behavior between \ce{PIm+I-} ($G^\prime= \SI{78}{\pascal}$) and \ce{PIm+Br-} ($G^\prime= \SI{13}{\pascal}$) indicates that dissociation is only partial and occurs to a lesser extent in the latter than in the former. The same conclusion can be drawn for \ce{PPyr+Cl-}, in which the elastic modulus ($G^\prime= \SI{270}{\pascal}$) value drops by more than one order of magnitude as compared to its bromide and iodide counterparts.

As already mentioned (vide supra) another striking feature of these gels is that in similar conditions poly(pyrrolidinium) based gels are systematically much stronger than those formed with poly(imidazolium). Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge; consequently, according to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected for the latter with the ``soft'' halide counterions.17 As a consequence, for a given halide, dissociation is less favored in the case of poly(imidazolium), as clearly illustrated by the difference in solvation between \ce{PIm+Cl-} and \ce{PPyr+Cl-}. Conversely, association of the lateral cations with the ``hard'' phosphonate moieties at the polymer chain’s end is expected to be stronger with pyrrolidinium, hence a systematically larger elastic modulus. The interplay of these two effects contributes in making poly(pyrrolidinium) based gels much stronger than their poly(imidazolium) counterparts; comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ value (13 and \SI{8100}{\pascal}, resp.) differ by almost three orders of magnitude.

\section{Conclusion}
With this work we demonstrated the versatility of our ``head-to-body'' electrostatic approach in the fabrication of hydrogels with readily tunable rheological properties. Thus, depending on the nature of the nucleophile used in the quaternarization step of the polymer, but also on that of the counterion which can be modified in the course of its purification process, we showed that it is possible to manipulate almost at will the gel formability and respective magnitudes of the shear modulus and of the shear stress. We don’t observe the formation of a shear thinning gel when the halide counterion is ``hard.'' Consistent, the strongest gels are obtained when a combination of ``hard'' cations and ``soft'' halides are used, which gives rise, in line with the HSAB theory, to maximal charge dissociation in water: this in turn facilitates charge pairing of the peripheral, iterative, cations with the terminal anion. It is thus possible to obtain robust hydrogels with short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials.

\begin{acknowledgement}
The authors are grateful for the financial support of Université de Lyon through the program ``Investissements d'Avenir'' ( ANR-1 1-IDEX-0007). S.M. and N.T. acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007–2013)/ERC grant agreement No.258803.
\end{acknowledgement}

\begin{suppinfo}

Materials and methods, detailed synthesis and characterizations of gels, GPC, HRMS and DSC data, calculated cohesion energy, and frequency sweep curves.

\end{suppinfo}

\end{document}

### 2015-12-20 ion_pairing_paper.tex
include synthesis.pdf, generated from synthesis.pgf
Metathesis

@@
 The phosphonate terminated polymer \ce{PBr}, which is the common basis to all studied polymer was synthesized according to our previously reported methodology; allowing us to obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; S2, S3, Figure S1) .9, 10 All subsequent transformations of the polymer structures were achieved using the same initial batch of material.
-Nucleophilic addition of N-Methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF afforded the corresponding poly(imidazloium) and poly(pyrrolidinium) compounds.
+Nucleophilic addition of N-Methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF afforded the corresponding poly(imidazolium) and poly(pyrrolidinium) compounds.
 In these cases, as an inherent consequence of the structure of the starting material, charge neutrality is provided by bromide counter ions, and the polymer will be referred to as \ce{PIm+Br-} and \ce{PPyr+Br-}, respectively (Scheme~\ref{sch:synthesis}).
@@
 \begin{scheme}
+\includegraphics{synthesis.pdf}
 \caption{Synthesis of \ce{PIm+Br-} and \ce{Ppyr+Br-} and their intermediate. $n=70$}
@@
 \begin{scheme}
-\caption{Anionic metathesis reaction.}
+\includegraphics{metathesis.pdf}
+\caption{Anionic metathesis reaction. As above, \ce{Nu+} corresponds either to pyrrolidinium or imidazolium.}
 \label{sch:metathesis}

### 2015-12-21 PALSE/ion_pairing_paper.tex
fig:strain, as from python script, except switch b and c and no dashed lines

--- 2015-12-20 ion_pairing_paper.tex
@@
 \usepackage{siunitx}
+\usepackage{xcolor}
 
@@
 For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile was that of pure water which confirms the visual observation.
-However as shown in Figure~\ref{fig:strain}, all other samples are solidlike ($G^\prime \gg G"$) at small strain and flow at large strains ($G" \gg G^\prime$).
+However as shown in Figure~\ref{fig:strain}, all other samples are solidlike ($G^\prime \gg G^{\prime\prime}$) at small strain and flow at large strains ($G^{\prime\prime} \gg G^\prime$).
 Solidlike behavior at low strain was confirmed for all accessible frequencies (see supplementary material S6 Fig.
 1).
-We define the critical strain value $\gamma_c$ between these two regimes when $G"$ reaches a maximum.
+We define the critical strain value $\gamma_c$ between these two regimes when $G^{\prime\prime}$ reaches a maximum.
 We checked that the linear mechanical properties are unchanged by the flow history given a few minutes of rest.
@@
 \begin{figure}
-\caption{Storage modulus $G^\prime$ (filled symbols and solid lines) and loss modulus $G"$ (open symbols and dashed lines) measured through oscillatory shear experiments for (a) \ce{PPyr+Cl-} ($\bullet$), \ce{PPyr+Br-} ($\blacksquare$) and \ce{PPyr+I-} ($\blacktriangle$), (b) \ce{PIm+Br-} ($\blacklozenge$) and \ce{PPyr+Br-} ($\blacksquare$) and (c) \ce{PIm+I-} ($\blacktriangledown$) and \ce{PPyr+I-} ($\blacktriangle$).
-$G^\prime$  and $G"$ measured during an amplitude sweep at a fixed frequency $f=\SI{1}{\hertz}$ are plotted against the strain amplitude $\gamma$.
+\includegraphics{strain.pdf}
+\caption{Storage modulus $G^\prime$ (filled symbols) and loss modulus $G^{\prime\prime}$ (open symbols) measured through oscillatory shear experiments for (a) \ce{PPyr+Cl-} (\textcolor{lightgray}{$\bullet$}), \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}) and \ce{PPyr+I-} ($\blacktriangle$), (b) \ce{PIm+I-} ($\blacktriangledown$) and \ce{PPyr+I-} ($\blacktriangle$) and (c) \ce{PIm+Br-} (\textcolor{gray}{$\blacklozenge$}) and \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}).
+$G^\prime$  and $G^{\prime\prime}$ measured during an amplitude sweep at a fixed frequency $f=\SI{1}{\hertz}$ are plotted against the strain amplitude $\gamma$.
 All samples are at 8\%wt }
@@
 We also compare the rheology of \ce{PIm+X-} and \ce{PPyr+X-} with the same counterion \ce{X} in Figure~\ref{fig:strain}b,c.
-For both bromide (Figure~\ref{fig:strain}b, X=Br) and iodide (Figure~\ref{fig:strain}c, X=I), imidazolium based gels are much weaker than their pyrrolidinium counterpart.
+For both iodide (Figure~\ref{fig:strain}b, X=I) and bromide (Figure~\ref{fig:strain}c, X=Br), imidazolium based gels are much weaker than their pyrrolidinium counterpart.
 

### 2016-01-04 PALSE/ion_pairing_paper.tex
bibliography (generated by Mendeley)

@@
 \section{Introduction}
-Since the pioneering work of Wichterle et al.,1 which established their relevance in a biomedical context, polymer based hydrogels have never ceased to be a very active field of research.2, 3 They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargos or scaffolds for wound healing or tissues repair.4 Polyelectrolytes are being extensively used in this particular context.5 Their gel formation mechanism often involves reversible-by-nature electrostatic interactions, which can be used for instance to trigger ``smart'' release of bioactive substances.6 Besides, poly(cationic) gel have been reported to combine scaffold properties for cell’s adherence and growth with antimicrobial activity, and are therefore increasingly used in clinical applications.3, 7 In this framework, injectability is a key feature, as it provides an easier way to gel delivery in vivo; thus, reversible shear thinning biocompatible hydrogels are a particularly sought after class of materials.8
-We have recently reported on a highly reversible poly(cation) based hydrogel,9 which formation relies on an new concept of supramolecular  electrostatic interaction.
+Since the pioneering work of Wichterle et al.,\cite{Wichterle1960} which established their relevance in a biomedical context, polymer based hydrogels have never ceased to be a very active field of research.\cite{Calo2015,Hoffman2001} They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargos or scaffolds for wound healing or tissues repair.\cite{Vashist2014,Ratner2004} Polyelectrolytes are being extensively used in this particular context.\cite{Rosso2003,BinImran2014} Their gel formation mechanism often involves reversible-by-nature electrostatic interactions, which can be used for instance to trigger ``smart'' release of bioactive substances.\cite{Khare1993,Lockwood2007} Besides, poly(cationic) gel have been reported to combine scaffold properties for cell’s adherence and growth with antimicrobial activity, and are therefore increasingly used in clinical applications.\cite{Hoffman2001,Landers2002} In this framework, injectability is a key feature, as it provides an easier way to gel delivery in vivo; thus, reversible shear thinning biocompatible hydrogels are a particularly sought after class of materials.\cite{Yu2008}
+
+We have recently reported on a highly reversible poly(cation) based hydrogel,\cite{Srour2014} which formation relies on an new concept of supramolecular  electrostatic interaction.
 Briefly, the atom transfer radical polymerisation (ATRP) is initiated by a phosphonate terminated ATRP initiator, eventually affording an anion terminated poly(cationic) polymer.
@@
 Moreover the interaction can be disrupted by various chemical stimuli, such as pH or ionic strength.
+
 In the present report, we show that the mechanical/rheological properties of such smart hydrogels can be tuned to an extremely large extent by playing with the nature (hardness/softness) of the cationic repeating unit and the ion-paring strength of the associated counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}).
@@
 
-The phosphonate terminated polymer \ce{PBr}, which is the common basis to all studied polymer was synthesized according to our previously reported methodology; allowing us to obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; S2, S3, Figure S1) .9, 10 All subsequent transformations of the polymer structures were achieved using the same initial batch of material.
+The phosphonate terminated polymer \ce{PBr}, which is the common basis to all studied polymer was synthesized according to our previously reported methodology; allowing us to obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; S2, S3, Figure S1) .\cite{Srour2014,Appukuttan2012} All subsequent transformations of the polymer structures were achieved using the same initial batch of material.
 Nucleophilic addition of N-Methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF afforded the corresponding poly(imidazolium) and poly(pyrrolidinium) compounds.
@@
 Calorimetric data were then recorded in order to characterize the effect of counterion substitution.
-In solids, melting points are inherently connected to the reticular enthalpy of the crystal lattice, hence primarily to ion pairing strength.11 This last parameter is directly related to the diffuse and hindered nature of the interacting charges, bulkier and softer ions providing in the most general case weaker interactions.12 For typical ``ionic liquids'' such as imidazolium or pyrollidinium halides, melting points will thus gradually increase as the size of the halide counterion decreases, ie $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$.13 Although the trend is less obvious in amorphous polymeric systems, in which phase transitions result from the interplay of multiple parameters, ionic substitution is however expected to have an impact on the melting temperature of the material.14
+In solids, melting points are inherently connected to the reticular enthalpy of the crystal lattice, hence primarily to ion pairing strength.\cite{Sherman1932} This last parameter is directly related to the diffuse and hindered nature of the interacting charges, bulkier and softer ions providing in the most general case weaker interactions.\cite{Krossing2006} For typical ``ionic liquids'' such as imidazolium or pyrollidinium halides, melting points will thus gradually increase as the size of the halide counterion decreases, ie $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$.\cite{Dean2010} Although the trend is less obvious in amorphous polymeric systems, in which phase transitions result from the interplay of multiple parameters, ionic substitution is however expected to have an impact on the melting temperature of the material.\cite{Orler1994}
 
@@
 In order to refine these visual observations, we performed rheological studies of the water-swelling materials.
-Briefly, data were recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}.15, 16 The sample was placed on the plate, the cone was then lowered to the measuring position, spreading the sample in the process.
+Briefly, data were recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}.\cite{Macosko1994,Larson1999} The sample was placed on the plate, the cone was then lowered to the measuring position, spreading the sample in the process.
 Excess material was removed, so that the sample exactly filled the gap.
@@
 \includegraphics{strain.pdf}
-\caption{Storage modulus $G^\prime$ (filled symbols) and loss modulus $G^{\prime\prime}$ (open symbols) measured through oscillatory shear experiments for (a) \ce{PPyr+Cl-} (\textcolor{lightgray}{$\bullet$}), \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}) and \ce{PPyr+I-} ($\blacktriangle$), (b) \ce{PIm+I-} ($\blacktriangledown$) and \ce{PPyr+I-} ($\blacktriangle$) and (c) \ce{PIm+Br-} (\textcolor{gray}{$\blacklozenge$}) and \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}).
-$G^\prime$  and $G^{\prime\prime}$ measured during an amplitude sweep at a fixed frequency $f=\SI{1}{\hertz}$ are plotted against the strain amplitude $\gamma$.
+\caption{Storage modulus $G^\prime$ (filled symbols) and loss modulus $G^{\prime\prime}$ (open symbols) measured through oscillatory shear experiments plotted against the strain amplitude $\gamma$ for (a) \ce{PPyr+Cl-} (\textcolor{lightgray}{$\bullet$}), \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}) and \ce{PPyr+I-} ($\blacktriangle$), (b) \ce{PIm+I-} ($\blacktriangledown$) and \ce{PPyr+I-} ($\blacktriangle$) and (c) \ce{PIm+Br-} (\textcolor{gray}{$\blacklozenge$}) and \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}).
+The fixed frequency is $f=\SI{1}{\hertz}$.
 All samples are at 8\%wt }
@@
 By comparison, \ce{PIm+Br-}, \ce{PIm+I-}, \ce{PPyr+Br-} and \ce{PPyr+I-} form homogeneous materials when suspended in water, which one can correlate with efficient dissociation of the halide counterions from the polymer main chain is fully consistency with decreased ionic bond strength.
-Such dissociation, in turns makes it possible for the ``hard'' phosphonate group on each polymer chain's end to interact with the ``free'' cationic groups of the side chains, which we showed in a previous paper to be at the origin of the gel formation mechanism.9
+Such dissociation, in turns makes it possible for the ``hard'' phosphonate group on each polymer chain's end to interact with the ``free'' cationic groups of the side chains, which we showed in a previous paper to be at the origin of the gel formation mechanism.\cite{Srour2014}
 
@@
 As already mentioned (vide supra) another striking feature of these gels is that in similar conditions poly(pyrrolidinium) based gels are systematically much stronger than those formed with poly(imidazolium).
-Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge; consequently, according to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected for the latter with the ``soft'' halide counterions.17 As a consequence, for a given halide, dissociation is less favored in the case of poly(imidazolium), as clearly illustrated by the difference in solvation between \ce{PIm+Cl-} and \ce{PPyr+Cl-}.
+Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge; consequently, according to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected for the latter with the ``soft'' halide counterions.\cite{Goossens2009} As a consequence, for a given halide, dissociation is less favored in the case of poly(imidazolium), as clearly illustrated by the difference in solvation between \ce{PIm+Cl-} and \ce{PPyr+Cl-}.
 Conversely, association of the lateral cations with the ``hard'' phosphonate moieties at the polymer chain’s end is expected to be stronger with pyrrolidinium, hence a systematically larger elastic modulus.
@@
 
+\bibliography{PALSE.bib}
+
 \end{document}

### 2016-01-06 PALSE/ion_pairing_paper.tex
fig:dsc
caption of fig:dsc

@@
 \includegraphics{synthesis.pdf}
-\caption{Synthesis of \ce{PIm+Br-} and \ce{Ppyr+Br-} and their intermediate. $n=70$}
+\caption{Synthesis of \ce{PIm+Br-} and \ce{PPyr+Br-} and their intermediates \ce{POH} and \ce{PBr}. $n=70$}
 \label{sch:synthesis}
@@
 \begin{figure}
-\caption{DSC analysis of a crystalline polymer samples showing the presence of melting point.}
+\includegraphics{dsc.pdf}
+\caption{DSC measurements of polymer samples showing the presence of a melting point. Curves are shifted vertically by arbitrary amounts for clarity.}
 \label{fig:dsc}

### 2016-01-06 PALSE/ion_pairing_paper.tex
rewrite abstract and title with "crocodile line"

@@
 
-\title{Controlling the rheological properties of well-defined polyelectrolytes hydrogels by fine tuning their ion pairing process}
+\title{Supramolecular crocodile line controls the rheological properties of polyelectrolytes hydrogels}
+%\title{Controlling the rheological properties of well-defined polyelectrolytes hydrogels by fine tuning their ion pairing process}
 
@@
 \begin{abstract}
-Starting from a common polymeric platform with lateral bromine substitution and phosphonate end group, it is possible, through two subsequent synthetic steps to generate a variety of poly(ionic liquids) with various cation/anion pair combinations.
-Upon water addition, the behavior of the resulting hydrogels is seen to be strongly dependent on the polymer structure, illustrating the possibility to control the gel formation properties by fine tuning of the intra and intermolecular ion pairing process.
+We demonstrated recently that polyelectrolytes with cationic moieties along the chain and a single anionic head are able to form physical hydrogels due to the reversible nature of the head-to-body ionic bond.
+Here we generate a variety of such polyelectrolytes with various cationic moieties and counter-ions combinations starting from a common polymeric platform.
+We show that the rheological properties (shear modulus, critical strain) of the final hydrogels vary by three orders of magnitude depending on the cation/pair.
+Our data are well described by a theoretical model involving a supramolecular crocodile line between cross-links, the length of the line being set by the amount of charge dissociation.
 \end{abstract}

### 2016-01-08 ion_pairing.tex
label eq:Ec

@@
 \usepackage{siunitx}
+\usepackage[table]{xcolor}
+\usepackage{pgfplots}
+\usetikzlibrary{calc}
 
+\begin{document}
 \author{Mathieu Leocmach}
 \title{Ion pairing rheology}
-
-\begin{document}
 \maketitle
 
-We are dealing with low-salt aqueous solutions of polyelectrolytes with cationic moieties along the chain and a single anionic head (Phosphonate).
+\section{System description}
+
+We are dealing with low-salt aqueous solutions of polyelectrolytes with cationic moieties along the chain and a single anionic head (Phosphonate, \ce{\bond{-}PO3^2-}).
 A chain has $N_0 = 70$ units from NMR measurements.
@@
 Nonetheless, due to charge dissociation, most of our polymers are soluble at \SI{25}{\celsius}.
+The reduced temperature $\tau = 1 - T/\Theta$ is thus at least 0.2.
 
-Due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible.
+We have previously demonstrated that, due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible~\cite{Srour2014}.
 If every chain has a single foreign head attached to its body, every polymer is linked to two others and we obtain a single effective chain of effective polymerisation index $n N_0$ where $n$ is the number of polymers in the effective chain.
 Here we have supposed that in order to minimize the inter-chain repulsion between charged cationic groups, heads are preferentially attached to the tail of their neighbour.
+We also neglect the possibility of the two charges of the phosphonate head to attach to two separate chains.
 
@@
 
+\section{Reological measurements}
+
 Let us now discuss the rheological measurements and link them to microstructure.
-Amplitude sweep measurements at \SI{1}{\hertz} in a cone-plate geometry give access to two main observables: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
-the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, the end of the linear regime and the first broken bonds.
+Amplitude sweep measurements at \SI{1}{\hertz} in a cone-plate geometry give access to two main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
+the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, the end of the linear regime and the first broken cross-links.
 
-The spatial conformation of a polyelectrolyte (super)chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions.
+\begin{table}
+\begin{tabular}{l|SS|SSS}
+& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_c$} & {$n$} & {$E_c$} & {$A$}\\
+&	{\si{\pascal}} &  & & {\si{\kilo\joule/\mol}} & \\\hline&&&&\\[-10pt]
+\ce{PPyr+I-}	& 8800	&	0.04	&	1.	&	0.004\cellcolor{gray!25}	&	1.\\
+\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.05&	0.001\cellcolor{gray!25}	&	1.1\\
+\ce{PPyr+Cl-}	& 270 	&	0.09	&	45	&	0.02\cellcolor{gray!25}	&	7.\\
+\ce{PIm+I-}	& 78	&	0.16	&	113	&	0.063\cellcolor{gray!25}	&	9.6\\
+\ce{PIm+Br-}	& 13	&	8.04	&	790	&	160.	&	3.4\cellcolor{gray!25}\\
+\end{tabular}
+\caption{Results. Incoherent values are on gray background.}
+\label{tab:results}
+\end{table}
+
+\subsection{From modulus to effective chain length}
+
+The spatial conformation of a polyelectrolyte (super)chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions~\cite{Rubinstein1996}, see Figure~\ref{fig:persistent}.
 Between each cross-link point the (super)chain can be considered as an entropic spring made of $N=n N_0$ monomers.
@@
 
-We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 780 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation.
+We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 790 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
 This is coherent with a higher probability of attaching two or more heads on a highly dissociated body, i.e.
@@
 
+\subsection{Bounding energy}
+
+Considering that each super chain is an entropic spring in its linear domain, the mechanical energy stored by a single super chain at a strain $\gamma$ is $E = \gamma^2 k_\mathrm{B}T$.
+If we suppose that the ionic bonds break before the complete extension of the (super)chains, we evaluate the head-to-body bonding energy to
+\begin{equation}
+E_c = \gamma_c^2 k_\mathrm{B}T.
+\label{eq:Ec}
+\end{equation}
+
+For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per bond or \SI{160}{\kilo\joule\per\mol}, coherent with typical energies of ionic bonds.
+However for all other samples the critical strain is between 2\% and 20\%, leading to incoherent values for the bonding energies between $10^{-3}k_\mathrm{B}T$ and $10^{-1}k_\mathrm{B}T$ per bond, see Table~\ref{tab:results}.
+Therefore, the hypothesis of a linear spring fails for these four samples, meaning that the super chain is fully extended at $\gamma_c$.
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+\begin{axis}[
+	scale only axis,
+	width=12\baselineskip, height=6\baselineskip, 
+	axis lines=none,
+	xmin=0.5, xmax=4.5,
+	ymin=-0.7, ymax=1.3,
+	]
+\addplot[only marks, mark=o, mark size=2] table {/home/mathieu/tex/PALSE/persistent.xy} coordinate[pos=0.43] (a) coordinate[pos=0.57] (b) coordinate[pos=0.81](c);
+\draw[<->] ($(a)+(-0.5em,0)$) -- ($(b)+(-0.5em,0)$) node[midway, left]{$r_\mathrm{scr}$};
+\draw[<->] ($(c)+(-0.25em,-0.5em)$) -- ($(c)+(0.25em,-0.5em)$) node[midway, below]{$D$};
+\end{axis}
+\end{tikzpicture}
+\end{center}
+\caption{Sketch of the persistent random walk.}
+\label{fig:persistent}
+\end{figure}
+
+\subsection{From critical strain to charge dissociation}
+
+For polyelectrolytes in bad solvent the persistent random walk in made of electrostatic blobs of diameter $D$ containing $g_e$ monomers~\cite{Rubinstein1996}, see Figure~\ref{fig:persistent}:
+\begin{align}
+g_e &= \frac{A^2}{u}\tau\label{eq:ge},\\
+D &= b \left(\frac{A^2}{u}\right)^{1/3}\label{eq:D}.
+\end{align}
+Here $A$ is the number of monomers between uncondensed charges, $\tau$ the reduced temperature and $u = \ell_\mathrm{B}/b \approx 2.7$ is the ratio of Bjerrum length $\ell_\mathrm{B}$ and monomer size $b\approx \SI{0.26}{\nano\metre}$.
+
+Extending these electrostatic blobs means exposing more monomers to the bad solvent and typically requires energies at least an order of magnitude higher than for the extension of the persistent random walk.
+We will thus consider that the exit from the linear domain of the material corresponds to the full extension of the persistent random walk:
+\begin{equation}
+\gamma_c = \frac{R_0}{R} - 1
+\end{equation}
+where $R$ and $R_0$ are respectively the end-to-end distance and the contour length of the persistent random walk.
+
+A segment of the persistent random walk is $r_\mathrm{scr}$ long and thus contains $g_\mathrm{scr} = g_e r_\mathrm{scr}/D$ monomers. We thus have
+\begin{align}
+R &= r_\mathrm{scr} \left(\frac{N}{g_\mathrm{scr}}\right)^{1/2},\\
+R_0 &= r_\mathrm{scr} \frac{N}{g_\mathrm{scr}},\\
+\gamma_c &= \left(\frac{N}{g_\mathrm{scr}}\right)^{1/2} -1\label{eq:gamma0}.
+\end{align}
+
+Following~\cite{Rubinstein1996}, we can express the screening length function of $A$ and $\tau$ as
+\begin{equation}
+r_\mathrm{scr} = \left(\frac{A^2}{u}\right)^{1/3} \left(\frac{\tau}{cb}\right)^{1/2}
+\label{eq:rsrc}
+\end{equation}
+
+From equations (\ref{eq:ge}), (\ref{eq:D}) and (\ref{eq:rsrc}) we deduce
+\begin{equation}
+g_\mathrm{scr} = \frac{A^2}{u} \left(\frac{\tau^3}{c b^3}\right)^{1/2}\label{eq:gscr}.
+\end{equation}
+
+From equations (\ref{eq:gamma0}) and (\ref{eq:gscr}) we can finally relate the measured critical strain to the microscopic parameter $A$:
+\begin{equation}
+A = \frac{\left(N u\right)^{1/2}}{\gamma_c+1}\left(\frac{c b^3}{\tau^3}\right)^{1/4}.
+\label{eq:A}
+\end{equation}
+
+The only unknown here is the reduced temperature $0.2<\tau<1$, possibly different between Poly(pyrrolidinium) based and Poly(immidazolium) based polymers.
+Setting $\tau\approx 0.235$, corresponding to $\Theta\approx\SI{116}{\celsius}$, yields $A=1$ for both \ce{PPyr+I-} and \ce{PPyr+Br-}, indicating full dissociation of the charges.
+By contrast \ce{PPyr+Cl-} has a dissociated charge every 7 monomers.
+Supposing the same $\tau$, \ce{PIm+I-} has a dissociated charge every 10 monomers, see Table~\ref{tab:results}.
+
+\subsection{Charge dissociation and cross-linking ratio}
+
+There must be a link between the charge dissociation and and the amount of physical cross-link.
+If we assume that the probability for a charge to dissociate is independent of the number of charges already present on the polymer, we can describe the probability of having $k$ charge dissociated on a polymer by a Poisson distribution of mean $\lambda = N_0/A$.
+Having 2 charges dissociated on a chain implies a cross-link.
+Therefore, we the cross-link probability $1/n$ is proportional to $\lambda^2$, yielding
+\begin{equation}
+n \sim A^2.
+\end{equation}
+
+This is indeed what we observe, as shown in Figure~\ref{fig:nA}.
+We can extrapolate this law for the case of \ce{PIm+Br-} and obtain $A = 28$.
+Inverting equation~(\ref{eq:A}) yields $\gamma_c = 860\%$, indeed higher than the one observed (804\%).
+This confirms that in the \ce{PIm+Br-} strain sweep we were actually probing the strength of the ionic bonds.
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	xlabel = $A$,
+	ylabel = $n$,
+	]1.04, 44.88, 0.99, 112.52
+	\addplot[only marks] coordinates {(1.090233298, 1.04) (6.98656403, 44.88) (1.000001279, 0.99) (9.589224893, 112.52)};
+	\addplot[no marks, domain=1:10] {x^2};
+\end{loglogaxis}
+\end{tikzpicture}
+\end{center}
+\caption{Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points. The line corresponds to $A^2$.}
+\label{fig:nA}
+\end{figure}
+
+\bibliography{PALSE}
+
 \end{document}

### 2016-01-08 PALSE/ion_pairing_paper.tex
change the gel preparation § into a water swelling and theta temperature § fix citations vs punctuation Passive -> active past -> present

@@
 \usepackage{siunitx}
-\usepackage{xcolor}
+\usepackage[table]{xcolor}
 
@@
 
-\title{Supramolecular crocodile line controls the rheological properties of polyelectrolytes hydrogels}
+%\title{Supramolecular crocodile line controls the rheological properties of polyelectrolytes hydrogels}
+\title{Ion pairing controls rheological properties of polyelectrolyte crocodile line hydrogels}
+%\title{Ion pairing in crocodile line hydrogels controls rheological properties}
+%\title{In crocodile line hydrogels ion pairing controls rheological properties}
 %\title{Controlling the rheological properties of well-defined polyelectrolytes hydrogels by fine tuning their ion pairing process}
@@
 \section{Introduction}
-Since the pioneering work of Wichterle et al.,\cite{Wichterle1960} which established their relevance in a biomedical context, polymer based hydrogels have never ceased to be a very active field of research.\cite{Calo2015,Hoffman2001} They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargos or scaffolds for wound healing or tissues repair.\cite{Vashist2014,Ratner2004} Polyelectrolytes are being extensively used in this particular context.\cite{Rosso2003,BinImran2014} Their gel formation mechanism often involves reversible-by-nature electrostatic interactions, which can be used for instance to trigger ``smart'' release of bioactive substances.\cite{Khare1993,Lockwood2007} Besides, poly(cationic) gel have been reported to combine scaffold properties for cell’s adherence and growth with antimicrobial activity, and are therefore increasingly used in clinical applications.\cite{Hoffman2001,Landers2002} In this framework, injectability is a key feature, as it provides an easier way to gel delivery in vivo; thus, reversible shear thinning biocompatible hydrogels are a particularly sought after class of materials.\cite{Yu2008}
+Since the pioneering work of \citeauthor{Wichterle1960} which established their relevance in a biomedical context, polymer based hydrogels have never ceased to be a very active field of research\cite{Calo2015,Hoffman2001}.
+They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargos or scaffolds for wound healing or tissues repair\cite{Vashist2014,Ratner2004}.
+Polyelectrolytes are being extensively used in this particular context\cite{Rosso2003,BinImran2014}.
+Their gel formation mechanism often involves reversible-by-nature electrostatic interactions, which can be used for instance to trigger ``smart'' release of bioactive substances\cite{Khare1993,Lockwood2007}.
+Besides, poly(cationic) gel have been reported to combine scaffold properties for cell’s adherence and growth with antimicrobial activity, and are therefore increasingly used in clinical applications\cite{Hoffman2001,Landers2002}.
+In this framework, injectability is a key feature, as it provides an easier way to gel delivery in vivo; thus, reversible shear thinning biocompatible hydrogels are a particularly sought after class of materials\cite{Yu2008}.
 
-We have recently reported on a highly reversible poly(cation) based hydrogel,\cite{Srour2014} which formation relies on an new concept of supramolecular  electrostatic interaction.
+We have recently reported on a highly reversible poly(cation) based hydrogel\cite{Srour2014}, which formation relies on an new concept of supramolecular  electrostatic interaction.
 Briefly, the atom transfer radical polymerisation (ATRP) is initiated by a phosphonate terminated ATRP initiator, eventually affording an anion terminated poly(cationic) polymer.
@@
 
-The phosphonate terminated polymer \ce{PBr}, which is the common basis to all studied polymer was synthesized according to our previously reported methodology; allowing us to obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; S2, S3, Figure S1) .\cite{Srour2014,Appukuttan2012} All subsequent transformations of the polymer structures were achieved using the same initial batch of material.
-Nucleophilic addition of N-Methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF afforded the corresponding poly(imidazolium) and poly(pyrrolidinium) compounds.
+We synthesise the phosphonate terminated polymer \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
+We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; S2, S3, Figure S1), which is the common basis to all studied polymers.
+We achieve all subsequent transformations of the polymer structures using the same initial batch of material.
+Nucleophilic addition of N-Methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF affords the corresponding poly(imidazolium) and poly(pyrrolidinium) compounds.
 In these cases, as an inherent consequence of the structure of the starting material, charge neutrality is provided by bromide counter ions, and the polymer will be referred to as \ce{PIm+Br-} and \ce{PPyr+Br-}, respectively (Scheme~\ref{sch:synthesis}).
@@
 
-Anionic metathesis could be achieved by pouring an aqueous solution of \ce{PIm+Br-} or \ce{PPyr+Br-} into a saturated aqueous solution of the different sodium halides (\ce{NaF}, \ce{NaCl}, \ce{NaI}) (Scheme~\ref{sch:metathesis}).
-In the cases of \ce{PIm+F-}, \ce{PIm+Cl-} , \ce{PPyr+F-} and \ce{PPyr+Cl-}, a turbid suspension was obtained immediately after addition.
-After extensive dialysis of the resulting mixture against deionized water and lyophilization,, the different \ce{PIm+X-} and \ce{PPyr+X-} (where \ce{X}=\ce{F}, \ce{Cl}, \ce{I}) were recovered in high yields.
-In order to bring evidence for the efficiency of the ionic metathesis, the resulting materials were submitted to negative mode high resolution mass spectrometry (HRMS, S3).
-Displacement of bromide counterions could be unambiguously assessed by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Figures S8-9).
-Substitution by iodide was further confirmed by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Figures S6-7), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (see supporting information).
+We performed anionic metathesis by pouring an aqueous solution of \ce{PIm+Br-} or \ce{PPyr+Br-} into a saturated aqueous solution of the different sodium halides (\ce{NaF}, \ce{NaCl}, \ce{NaI}) (Scheme~\ref{sch:metathesis}).
+In the cases of \ce{PIm+F-}, \ce{PIm+Cl-} , \ce{PPyr+F-} and \ce{PPyr+Cl-}, we obtain a turbid suspension immediately after addition.
+After extensive dialysis of the resulting mixture against deionized water and lyophilization we recover in high yields the different \ce{PIm+X-} and \ce{PPyr+X-} (where \ce{X}=\ce{F}, \ce{Cl}, \ce{I}).
+In order to bring evidence for the efficiency of the ionic metathesis, we submit the resulting materials to negative mode high resolution mass spectrometry (HRMS, S3).
+We unambiguously assess displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Figures S8-9).
+We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Figures S6-7), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (see supporting information).
 
@@
 
-Calorimetric data were then recorded in order to characterize the effect of counterion substitution.
-In solids, melting points are inherently connected to the reticular enthalpy of the crystal lattice, hence primarily to ion pairing strength.\cite{Sherman1932} This last parameter is directly related to the diffuse and hindered nature of the interacting charges, bulkier and softer ions providing in the most general case weaker interactions.\cite{Krossing2006} For typical ``ionic liquids'' such as imidazolium or pyrollidinium halides, melting points will thus gradually increase as the size of the halide counterion decreases, ie $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$.\cite{Dean2010} Although the trend is less obvious in amorphous polymeric systems, in which phase transitions result from the interplay of multiple parameters, ionic substitution is however expected to have an impact on the melting temperature of the material.\cite{Orler1994}
+In order to characterize the effect of counterion substitution, we study all polymers by Differential Scanning Calorimetry (DSC).
+In solids, melting points are inherently connected to the reticular enthalpy of the crystal lattice, hence primarily to ion pairing strength\cite{Sherman1932}.
+This last parameter is directly related to the diffuse and hindered nature of the interacting charges, bulkier and softer ions providing in the most general case weaker interactions\cite{Krossing2006}.
+For typical ``ionic liquids'' such as imidazolium or pyrollidinium halides, melting points will thus gradually increase as the size of the halide counterion decreases, ie $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$\cite{Dean2010}.
+Although the trend is less obvious in amorphous polymeric systems, in which phase transitions result from the interplay of multiple parameters, ionic substitution is however expected to have an impact on the melting temperature of the material\cite{Orler1994}.
 
-All polymers were thus studied by Differential Scanning Calorimetry (DSC).
-As expected, the lowest melting temperature was achieved for polymers with iodide counterions.
-Quite remarkably, a similar melting temperature (84-\SI{85}{\celsius}) was found for \ce{PIm+I-} and \ce{PPyr+I-}.
-When bromide counterions are present, a significant increase of the melting temperature was observed (114-\SI{115}{\celsius}) but again, a similar value was found for \ce{PIm+Br-} and \ce{PPyr+Br-}.
-In the case of fluoride and chloride, no phase transition was seen below decomposition temperature of the polymer (\SI{200}{\celsius}), which indicates a much stronger ion pairing in these materials (Figure~\ref{fig:dsc}, S4 Figures S2-5).
+As expected, the lowest melting temperature was recorded for polymers with iodide counterions.
+Quite remarkably,we found a similar melting temperature (84-\SI{85}{\celsius}) for \ce{PIm+I-} and \ce{PPyr+I-}, see Figure~\ref{fig:dsc}.
+When bromide counterions are present, we observe a significant increase of the melting temperature (114-\SI{115}{\celsius}) but again, with a similar value between \ce{PIm+Br-} and \ce{PPyr+Br-} as shown in Figure~\ref{fig:dsc}.
+In the case of fluoride and chloride, we observe no phase transition below decomposition temperature of the polymer (\SI{200}{\celsius}), which indicates a much stronger ion pairing in these materials.
+% no need of fig S2-5 for me, but a wy not a supplementary figure with chloride and fluoride.
 
@@
 \includegraphics{dsc.pdf}
-\caption{DSC measurements of polymer samples showing the presence of a melting point. Curves are shifted vertically by arbitrary amounts for clarity.}
+\caption{DSC measurements of polymer samples showing the presence of a melting point.
+Curves are shifted vertically by arbitrary amounts for clarity.
+Vertical lines indicate the melting points at \SI{84}{\celsius} for the two iodides in black and \SI{114}{\celsius} for the two bromides in gray.}
 \label{fig:dsc}
@@
 
-\subsection{Gels preparation}
-All materials were placed in vials, and water was added until a final aqueous dispersion of each polymer at a 8\%wt.
-concentration was obtained.
-Swelling properties turned out to be extremely dependent on the polymer compositions.
-With fluoride counterions, no observable swelling or dissolution took place and the suspension was recovered as a biphasic mixture of its individual components.
-Results were very similar for the chloride containg Poly(immidazolium) polymer, whereas a limited swelling was observed for its poly(pyrrolidinium) counterpart, resulting in a granular, inhomogeneous gel.
-When bromide was present, both polypyrrolidinium and polyimmidazolium polymers afforded a gel upon swelling with water.
-However, their appearance markedly differed, as the pyrrolidinium based system gave rise to a much harder and glassy-like material, indicating a consistently stronger interchain network.
-A similar gel behavior was found with \ce{PPyr+I-}, while \ce{PIm+I-}, although very similar in appearance to \ce{PIm+Br-} (homogeneous and optically transparent) turned out to be significantly harder when submitted to mechanical shearing (Figure~\ref{fig:pictures}).
+\subsection{Water swelling and solvent quality}
+Water-swelling properties turn out to be extremely dependent on the polymer compositions.
+With bromide or iodide conterions, both poly(pyrrolidinium) and poly(immidazolium) polymers affords homogeneous and optically transparent gels upon swelling with deionised water.
+We observe a limited swelling for \ce{PPyr+Cl-}, resulting in a granular, inhomogeneous gel.
+With fluoride counterions, or in the case of \ce{PIm+Cl-}, we observe neither swelling nor dissolution.
+We recover a 8\%wt suspension as a biphasic mixture of its individual components.
 
-\begin{figure}
-\caption{pictures of 8\%wt water solutions of \ce{PIm+X-} and \ce{PPyr+X-} in water.}
-\label{fig:pictures}
-\end{figure}
+We thus are able to dissolve the five polymers that have the weakest ion pairing, but neither the two fluorides nor \ce{PIm+Cl-} where counterions are expected to be strongly bounded to the chain.
+This means that in absence of charge dissociation the chain is in poor solvent.
+Furthermore, we held the two fluoride polymers in water at boiling temperature overnight without any dissolution.
+The $\Theta$ temperature is thus higher than \SI{100}{\celsius}.
+Nonetheless, due to charge dissociation, most of our polymers are soluble at \SI{25}{\celsius}.
 
-Rheological measurements.
-In order to refine these visual observations, we performed rheological studies of the water-swelling materials.
-Briefly, data were recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}.\cite{Macosko1994,Larson1999} The sample was placed on the plate, the cone was then lowered to the measuring position, spreading the sample in the process.
+%\begin{figure}
+%\caption{pictures of 8\%wt water solutions of \ce{PIm+X-} and \ce{PPyr+X-} in water.}
+%\label{fig:pictures}
+%\end{figure}
+
+\subsection{Rheological measurements}
+We perform rheological studies of the water-swelling materials.
+Briefly, data were recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}\cite{Macosko1994,Larson1999}.
+The sample was placed on the plate, the cone was then lowered to the measuring position, spreading the sample in the process.
 Excess material was removed, so that the sample exactly filled the gap.
@@
 By contrast, the gel with chloride counterions is weaker by more than one order of magnitude.
-The cohesive energy goes down with smaller counterion indicating weaker intermolecular bonds (see supplementary material S6 table 1).
 
@@
 By comparison, \ce{PIm+Br-}, \ce{PIm+I-}, \ce{PPyr+Br-} and \ce{PPyr+I-} form homogeneous materials when suspended in water, which one can correlate with efficient dissociation of the halide counterions from the polymer main chain is fully consistency with decreased ionic bond strength.
-Such dissociation, in turns makes it possible for the ``hard'' phosphonate group on each polymer chain's end to interact with the ``free'' cationic groups of the side chains, which we showed in a previous paper to be at the origin of the gel formation mechanism.\cite{Srour2014}
+Such dissociation, in turns makes it possible for the ``hard'' phosphonate group on each polymer chain's end to interact with the ``free'' cationic groups of the side chains, which we showed in a previous paper to be at the origin of the gel formation mechanism\cite{Srour2014}.
 
@@
 As already mentioned (vide supra) another striking feature of these gels is that in similar conditions poly(pyrrolidinium) based gels are systematically much stronger than those formed with poly(imidazolium).
-Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge; consequently, according to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected for the latter with the ``soft'' halide counterions.\cite{Goossens2009} As a consequence, for a given halide, dissociation is less favored in the case of poly(imidazolium), as clearly illustrated by the difference in solvation between \ce{PIm+Cl-} and \ce{PPyr+Cl-}.
+Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge; consequently, according to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected for the latter with the ``soft'' halide counterions\cite{Goossens2009}.
+As a consequence, for a given halide, dissociation is less favored in the case of poly(imidazolium), as clearly illustrated by the difference in solvation between \ce{PIm+Cl-} and \ce{PPyr+Cl-}.
 Conversely, association of the lateral cations with the ``hard'' phosphonate moieties at the polymer chain’s end is expected to be stronger with pyrrolidinium, hence a systematically larger elastic modulus.

### 2016-01-08 ion_pairing_paper.tex
Insert Crocodile chain model from ion_pairing.tex Compact a bit Remove the independent schematic figure (to make anew).

@@
 \usepackage{siunitx}
+\usepackage[table]{xcolor}
 
@@
 
-\title{Controlling the rheological properties of well-defined polyelectrolytes hydrogels by fine tuning their ion pairing process}
+%\title{Supramolecular crocodile line controls the rheological properties of polyelectrolytes hydrogels}
+\title{Ion pairing controls rheological properties of polyelectrolyte crocodile line hydrogels}
+%\title{Ion pairing in crocodile line hydrogels controls rheological properties}
+%\title{In crocodile line hydrogels ion pairing controls rheological properties}
+%\title{Controlling the rheological properties of well-defined polyelectrolytes hydrogels by fine tuning their ion pairing process}
 
@@
 \section{Introduction}
-Since the pioneering work of Wichterle et al.,1 which established their relevance in a biomedical context, polymer based hydrogels have never ceased to be a very active field of research.2, 3 They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargos or scaffolds for wound healing or tissues repair.4 Polyelectrolytes are being extensively used in this particular context.5 Their gel formation mechanism often involves reversible-by-nature electrostatic interactions, which can be used for instance to trigger ``smart'' release of bioactive substances.6 Besides, poly(cationic) gel have been reported to combine scaffold properties for cell’s adherence and growth with antimicrobial activity, and are therefore increasingly used in clinical applications.3, 7 In this framework, injectability is a key feature, as it provides an easier way to gel delivery in vivo; thus, reversible shear thinning biocompatible hydrogels are a particularly sought after class of materials.8
-We have recently reported on a highly reversible poly(cation) based hydrogel,9 which formation relies on an new concept of supramolecular  electrostatic interaction.
+Since the pioneering work of \citeauthor{Wichterle1960} which established their relevance in a biomedical context, polymer based hydrogels have never ceased to be a very active field of research\cite{Calo2015,Hoffman2001}.
+They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargos or scaffolds for wound healing or tissues repair\cite{Vashist2014,Ratner2004}.
+Polyelectrolytes are being extensively used in this particular context\cite{Rosso2003,BinImran2014}.
+Their gel formation mechanism often involves reversible-by-nature electrostatic interactions, which can be used for instance to trigger ``smart'' release of bioactive substances\cite{Khare1993,Lockwood2007}.
+Besides, poly(cationic) gel have been reported to combine scaffold properties for cell’s adherence and growth with antimicrobial activity, and are therefore increasingly used in clinical applications\cite{Hoffman2001,Landers2002}.
+In this framework, injectability is a key feature, as it provides an easier way to gel delivery in vivo; thus, reversible shear thinning biocompatible hydrogels are a particularly sought after class of materials\cite{Yu2008}.
+
+We have recently reported on a highly reversible poly(cation) based hydrogel\cite{Srour2014}, which formation relies on an new concept of supramolecular  electrostatic interaction.
 Briefly, the atom transfer radical polymerisation (ATRP) is initiated by a phosphonate terminated ATRP initiator, eventually affording an anion terminated poly(cationic) polymer.
@@
 
-The phosphonate terminated polymer \ce{PBr}, which is the common basis to all studied polymer was synthesized according to our previously reported methodology; allowing us to obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; S2, S3, Figure S1) .9, 10 All subsequent transformations of the polymer structures were achieved using the same initial batch of material.
-Nucleophilic addition of N-Methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF afforded the corresponding poly(imidazolium) and poly(pyrrolidinium) compounds.
+We synthesise the phosphonate terminated polymer \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
+We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; S2, S3, Figure S1), which is the common basis to all studied polymers.
+We achieve all subsequent transformations of the polymer structures using the same initial batch of material.
+Nucleophilic addition of N-Methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF affords the corresponding poly(imidazolium) and poly(pyrrolidinium) compounds.
 In these cases, as an inherent consequence of the structure of the starting material, charge neutrality is provided by bromide counter ions, and the polymer will be referred to as \ce{PIm+Br-} and \ce{PPyr+Br-}, respectively (Scheme~\ref{sch:synthesis}).
@@
 
-Calorimetric data were then recorded in order to characterize the effect of counterion substitution.
-In solids, melting points are inherently connected to the reticular enthalpy of the crystal lattice, hence primarily to ion pairing strength.11 This last parameter is directly related to the diffuse and hindered nature of the interacting charges, bulkier and softer ions providing in the most general case weaker interactions.12 For typical ``ionic liquids'' such as imidazolium or pyrollidinium halides, melting points will thus gradually increase as the size of the halide counterion decreases, ie $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$.13 Although the trend is less obvious in amorphous polymeric systems, in which phase transitions result from the interplay of multiple parameters, ionic substitution is however expected to have an impact on the melting temperature of the material.14
+In order to characterize the effect of counterion substitution, we study all polymers by Differential Scanning Calorimetry (DSC).
+In solids, melting points are inherently connected to the reticular enthalpy of the crystal lattice, hence primarily to ion pairing strength\cite{Sherman1932}.
+This last parameter is directly related to the diffuse and hindered nature of the interacting charges, bulkier and softer ions providing in the most general case weaker interactions\cite{Krossing2006}.
+For typical ``ionic liquids'' such as imidazolium or pyrollidinium halides, melting points will thus gradually increase as the size of the halide counterion decreases, ie $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$\cite{Dean2010}.
+Although the trend is less obvious in amorphous polymeric systems, in which phase transitions result from the interplay of multiple parameters, ionic substitution is however expected to have an impact on the melting temperature of the material\cite{Orler1994}.
 
-All polymers were thus studied by Differential Scanning Calorimetry (DSC).
-As expected, the lowest melting temperature was achieved for polymers with iodide counterions.
-Quite remarkably, a similar melting temperature (84-\SI{85}{\celsius}) was found for \ce{PIm+I-} and \ce{PPyr+I-}.
-When bromide counterions are present, a significant increase of the melting temperature was observed (114-\SI{115}{\celsius}) but again, a similar value was found for \ce{PIm+Br-} and \ce{PPyr+Br-}.
-In the case of fluoride and chloride, no phase transition was seen below decomposition temperature of the polymer (\SI{200}{\celsius}), which indicates a much stronger ion pairing in these materials (Figure~\ref{fig:dsc}, S4 Figures S2-5).
+As expected, the lowest melting temperature was recorded for polymers with iodide counterions.
+Quite remarkably,we found a similar melting temperature (84-\SI{85}{\celsius}) for \ce{PIm+I-} and \ce{PPyr+I-}, see Figure~\ref{fig:dsc}.
+When bromide counterions are present, we observe a significant increase of the melting temperature (114-\SI{115}{\celsius}) but again, with a similar value between \ce{PIm+Br-} and \ce{PPyr+Br-} as shown in Figure~\ref{fig:dsc}.
+In the case of fluoride and chloride, we observe no phase transition below decomposition temperature of the polymer (\SI{200}{\celsius}), which indicates a much stronger ion pairing in these materials.
+% no need of fig S2-5 for me, but a wy not a supplementary figure with chloride and fluoride.
 
 \begin{figure}
-\caption{DSC analysis of a crystalline polymer samples showing the presence of melting point.}
+\includegraphics{dsc.pdf}
+\caption{DSC measurements of polymer samples showing the presence of a melting point.
+Curves are shifted vertically by arbitrary amounts for clarity.
+Vertical lines indicate the melting points at \SI{84}{\celsius} for the two iodides in black and \SI{114}{\celsius} for the two bromides in gray.}
 \label{fig:dsc}
@@
 
-\subsection{Gels preparation}
-All materials were placed in vials, and water was added until a final aqueous dispersion of each polymer at a 8\%wt.
-concentration was obtained.
-Swelling properties turned out to be extremely dependent on the polymer compositions.
-With fluoride counterions, no observable swelling or dissolution took place and the suspension was recovered as a biphasic mixture of its individual components.
-Results were very similar for the chloride containg Poly(immidazolium) polymer, whereas a limited swelling was observed for its poly(pyrrolidinium) counterpart, resulting in a granular, inhomogeneous gel.
-When bromide was present, both polypyrrolidinium and polyimmidazolium polymers afforded a gel upon swelling with water.
-However, their appearance markedly differed, as the pyrrolidinium based system gave rise to a much harder and glassy-like material, indicating a consistently stronger interchain network.
-A similar gel behavior was found with \ce{PPyr+I-}, while \ce{PIm+I-}, although very similar in appearance to \ce{PIm+Br-} (homogeneous and optically transparent) turned out to be significantly harder when submitted to mechanical shearing (Figure~\ref{fig:pictures}).
+\subsection{Water swelling and solvent quality}
+Water-swelling properties turn out to be extremely dependent on the polymer compositions.
+With bromide or iodide conterions, both poly(pyrrolidinium) and poly(immidazolium) polymers afford homogeneous and optically transparent gels upon swelling with deionised water.
+We observe a limited swelling for \ce{PPyr+Cl-}, resulting in a granular, inhomogeneous gel.
+With fluoride counterions, or in the case of \ce{PIm+Cl-}, we observe neither swelling nor dissolution.
+We recover a 8\%wt suspension as a biphasic mixture of its individual components.
 
-\begin{figure}
-\caption{pictures of 8\%wt water solutions of \ce{PIm+X-} and \ce{PPyr+X-} in water.}
-\label{fig:pictures}
-\end{figure}
+Therefore, we are able to dissolve the five polymers that have the weakest ion pairing, but neither the two fluorides nor \ce{PIm+Cl-} where counterions are expected to be strongly bounded to the chain.
+This means that in absence of charge dissociation the chain is in poor solvent.
+Furthermore, we held the two fluoride polymers in water at boiling temperature overnight without any dissolution, implying that the $\Theta$ temperature is higher than \SI{100}{\celsius}.
+Nonetheless, due to charge dissociation, most of our polymers are soluble at \SI{25}{\celsius}.
 
-Rheological measurements.
-In order to refine these visual observations, we performed rheological studies of the water-swelling materials.
-Briefly, data were recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}.15, 16 The sample was placed on the plate, the cone was then lowered to the measuring position, spreading the sample in the process.
-Excess material was removed, so that the sample exactly filled the gap.
-Due to Weissenberg effect at high shear rate, no preshear was applied before measurement.
-To minimize water absorption, the geometry was covered with a solvent trap, using light mineral oil as a liquid seal between the rotor and the cap.
-After one minute of equilibration we performed an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$).
-For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile was that of pure water which confirms the visual observation.
-However as shown in Figure~\ref{fig:strain}, all other samples are solidlike ($G^\prime \gg G"$) at small strain and flow at large strains ($G" \gg G^\prime$).
-Solidlike behavior at low strain was confirmed for all accessible frequencies (see supplementary material S6 Fig.
+%\begin{figure}
+%\caption{pictures of 8\%wt water solutions of \ce{PIm+X-} and \ce{PPyr+X-} in water.}
+%\label{fig:pictures}
+%\end{figure}
+
+\subsection{Rheological measurements}
+We perform rheological studies of the water-swelling materials.
+Briefly, data are recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}\cite{Macosko1994,Larson1999}.
+We place the sample on the plate, and then lower the cone to the measuring position, spreading the sample in the process.
+We remove excess material, so that the sample exactly fills the gap.
+Due to Weissenberg effect at high shear rate, we apply no preshear before measurement.
+To minimize water absorption, we cover the geometry with a solvent trap, using light mineral oil as a liquid seal between the rotor and the cap.
+After one minute of equilibration we perform an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$).
+For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile is that of pure water which confirms the visual observation.
+However as shown in Figure~\ref{fig:strain}, all other samples are solidlike ($G^\prime \gg G^{\prime\prime}$) at small strain and flow at large strains ($G^{\prime\prime} \gg G^\prime$).
+We confirm solidlike behavior at low strain for all accessible frequencies (see supplementary material S6 Fig.
 1).
-We define the critical strain value $\gamma_c$ between these two regimes when $G"$ reaches a maximum.
+We define the critical strain value $\gamma_c$ between these two regimes when $G^{\prime\prime}$ reaches a maximum.
 We checked that the linear mechanical properties are unchanged by the flow history given a few minutes of rest.
@@
 \begin{figure}
-\caption{Storage modulus $G^\prime$ (filled symbols and solid lines) and loss modulus $G"$ (open symbols and dashed lines) measured through oscillatory shear experiments for (a) \ce{PPyr+Cl-} ($\bullet$), \ce{PPyr+Br-} ($\blacksquare$) and \ce{PPyr+I-} ($\blacktriangle$), (b) \ce{PIm+Br-} ($\blacklozenge$) and \ce{PPyr+Br-} ($\blacksquare$) and (c) \ce{PIm+I-} ($\blacktriangledown$) and \ce{PPyr+I-} ($\blacktriangle$).
-$G^\prime$  and $G"$ measured during an amplitude sweep at a fixed frequency $f=\SI{1}{\hertz}$ are plotted against the strain amplitude $\gamma$.
+\includegraphics{strain.pdf}
+\caption{Storage modulus $G^\prime$ (filled symbols) and loss modulus $G^{\prime\prime}$ (open symbols) measured through oscillatory shear experiments plotted against the strain amplitude $\gamma$ for (a) \ce{PPyr+Cl-} (\textcolor{lightgray}{$\bullet$}), \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}) and \ce{PPyr+I-} ($\blacktriangle$), (b) \ce{PIm+I-} ($\blacktriangledown$) and \ce{PPyr+I-} ($\blacktriangle$) and (c) \ce{PIm+Br-} (\textcolor{gray}{$\blacklozenge$}) and \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}).
+The fixed frequency is $f=\SI{1}{\hertz}$.
 All samples are at 8\%wt }
@@
 In Figure~\ref{fig:strain}a we compare the rheological responses of Poly(pyrrolidinium) based hydrogels for the three counterions allowing dissolution in water.
-Gels with bromide and iodide counterions showed roughly the same mechanical behavior in the linear regime but \ce{PPyr+I-} yields at slightly higher strains ($\gamma_c=4\%$ instead of $3\%$).
+Gels with bromide and iodide counterions show roughly the same mechanical behavior in the linear regime but \ce{PPyr+I-} yields at slightly higher strains ($\gamma_c=4\%$ instead of $3\%$).
 By contrast, the gel with chloride counterions is weaker by more than one order of magnitude.
-The cohesive energy goes down with smaller counterion indicating weaker intermolecular bonds (see supplementary material S6 table 1).
 
 We also compare the rheology of \ce{PIm+X-} and \ce{PPyr+X-} with the same counterion \ce{X} in Figure~\ref{fig:strain}b,c.
-For both bromide (Figure~\ref{fig:strain}b, X=Br) and iodide (Figure~\ref{fig:strain}c, X=I), imidazolium based gels are much weaker than their pyrrolidinium counterpart.
+For both iodide (Figure~\ref{fig:strain}b, X=I) and bromide (Figure~\ref{fig:strain}c, X=Br), imidazolium based gels are much weaker than their pyrrolidinium counterpart.
+
+\subsection{Crocodile chain model}
+We have previously demonstrated that, due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible~\cite{Srour2014}.
+If at least two foreign heads attach to the same body, we obtain an effective cross-linking point.
+When the probability of such a configuration is non-zero, we obtain a physically cross-linked gel.
+
+Here we argue that if every chain has a single foreign head attached to its body, every polymer is linked to two others in a crocodile line.
+We neglect the possibility of the two charges of the phosphonate head to attach to two separate chains.
+In addition, if we suppose that in order to minimize the inter-chain repulsion between charged cationic groups, heads are preferentially attached to the tail of their neighbour, we obtain a linear chain of effective polymerisation index $n N_0$ where $n$ is the number of polymers in the crocodile line.
+
+We thus have two limiting cases: (i) every chain has at least two heads attached and we have roughly $N_0$ monomers between cross-linking points; (ii) every chain has a single head attached and we have a single supramolecular chain in the system.
+The former case should be observed at high charge dissociation when many moieties are ionised along the chain providing many sites for a phosphonate head to bound to.
+The latter case should be observed at negligible charge dissociation.
+In between we should observe cross-linking points separated by crocodile lines of $n N_0$ monomers.
+
+Let us now discuss the rheological measurements and link them to microstructure.
+Amplitude sweep measurements at \SI{1}{\hertz} in a cone-plate geometry give access to two main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
+the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, the end of the linear regime and the first broken cross-links.
+
+\begin{table}
+\begin{tabular}{l|SS|SSS}
+& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_c$} & {$n$} & {$E_c$} & {$A$}\\
+&	{\si{\pascal}} &  & & {\si{\kilo\joule/\mol}} & \\\hline&&&&\\[-10pt]
+\ce{PPyr+I-}	& 8800	&	0.04	&	1.	&	0.004\cellcolor{gray!25}	&	1.\\
+\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.05&	0.001\cellcolor{gray!25}	&	1.1\\
+\ce{PPyr+Cl-}	& 270 	&	0.09	&	45	&	0.02\cellcolor{gray!25}	&	7.\\
+\ce{PIm+I-}	& 78	&	0.16	&	113	&	0.063\cellcolor{gray!25}	&	9.6\\
+\ce{PIm+Br-}	& 13	&	8.04	&	790	&	160.	&	3.4\cellcolor{gray!25}\\
+\end{tabular}
+\caption{Summary of rheological measurements and microscopic values deduced from the model.
+Gray background indicates incoherent values obtained either by (Eq.~\ref{eq:Ec}) or by (Eq.~\ref{eq:A}) outside of their respective validity domain.}
+\label{tab:results}
+\end{table}
+
+\subsubsection{From modulus to effective chain length}
+
+Between two cross-links the crocodile chain can be considered as an entropic spring made of $N=n N_0$ monomers. The shear modulus of the gel is then given by:
+\begin{equation}
+G = \frac{c}{N}k_\mathrm{B}T,
+\label{eq:G}
+\end{equation}
+where $c$ is the monomer number density, $k_\mathrm{B}$ is the Boltzmann constant and $T$ the temperature.
+Knowing the molecular mass $M$ of the chains, density $d$ of the solvent and the weight fraction $w$ of polymer, we obtain
+\begin{equation}
+n = \frac{\mathcal{N_A}}{M} w d \frac{k_\mathrm{B}T}{G}.
+\end{equation}
+
+We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 790 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
+This is coherent with a higher probability of attaching two or more heads on a highly dissociated body, i.e.
+a large cross-linking ratio.
+
+\subsubsection{Bounding energy}
+
+Considering that each crocodile chain is an entropic spring in its linear domain, the mechanical energy stored by a single crocodile chain at a strain $\gamma$ is $E = \gamma^2 k_\mathrm{B}T$.
+If we suppose that the ionic bonds breaks before the complete extension of the crocodile chains, we evaluate the head-to-body bonding energy to
+\begin{equation}
+E_c = \gamma_c^2 k_\mathrm{B}T.
+\label{eq:Ec}
+\end{equation}
+
+For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per bond or \SI{160}{\kilo\joule\per\mol}, coherent with typical energies of ionic bonds.
+However for all other samples the critical strain is between 2\% and 20\%, leading to incoherent values for the bonding energies between $10^{-3}k_\mathrm{B}T$ and $10^{-1}k_\mathrm{B}T$ per bond, see Table~\ref{tab:results}.
+Therefore, the hypothesis of a linear spring fails for these four samples, meaning that the crocodile chain is fully extended at $\gamma_c$.
+
+%\begin{figure}
+%\begin{center}
+%\begin{tikzpicture}
+%\begin{axis}[
+%	scale only axis,
+%	width=12\baselineskip, height=6\baselineskip, 
+%	axis lines=none,
+%	xmin=0.5, xmax=4.5,
+%	ymin=-0.7, ymax=1.3,
+%	]
+%\addplot[only marks, mark=o, mark size=2] table {/home/mathieu/tex/PALSE/persistent.xy} coordinate[pos=0.43] (a) coordinate[pos=0.57] (b) coordinate[pos=0.81](c);
+%\draw[<->] ($(a)+(-0.5em,0)$) -- ($(b)+(-0.5em,0)$) node[midway, left]{$r_\mathrm{scr}$};
+%\draw[<->] ($(c)+(-0.25em,-0.5em)$) -- ($(c)+(0.25em,-0.5em)$) node[midway, below]{$D$};
+%\end{axis}
+%\end{tikzpicture}
+%\end{center}
+%\caption{Sketch of the persistent random walk.}
+%\label{fig:persistent}
+%\end{figure}
+
+\subsubsection{From critical strain to charge dissociation}
+
+The spatial conformation of a polyelectrolyte chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions~\cite{Rubinstein1996}.
+In bad solvent, the persistent random walk is made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:persistent}:
+\begin{align}
+g_e &= \frac{A^2}{u}\tau\label{eq:ge},\\
+D &= b \left(\frac{A^2}{u}\right)^{1/3}\label{eq:D}.
+\end{align}
+Here $A$ is the number of monomers between uncondensed charges, $\tau = 1 - T/\Theta$ the reduced temperature and $u = \ell_\mathrm{B}/b \approx 2.7$ is the ratio of Bjerrum length $\ell_\mathrm{B}$ and monomer size $b\approx \SI{0.26}{\nano\metre}$.
+
+Extending these electrostatic blobs means exposing more monomers to the bad solvent and typically requires energies at least an order of magnitude higher than for the extension of the persistent random walk.
+We will thus consider that the exit from the linear domain of the material corresponds to the full extension of the persistent random walk:
+\begin{equation}
+\gamma_c = \frac{R_0}{R} - 1
+\end{equation}
+where $R$ and $R_0$ are respectively the end-to-end distance and the contour length of the persistent random walk.
+
+A segment of the persistent random walk is $r_\mathrm{scr}$ long and thus contains $g_\mathrm{scr} = g_e r_\mathrm{scr}/D$ monomers. We thus have
+\begin{equation}
+\left. \begin{array}{ll}
+R &= r_\mathrm{scr} \left(\frac{N}{g_\mathrm{scr}}\right)^{1/2}\\
+R_0 &= r_\mathrm{scr} \frac{N}{g_\mathrm{scr}}
+\end{array}\right\rbrace\quad
+\gamma_c = \left(\frac{N}{g_\mathrm{scr}}\right)^{1/2} -1\label{eq:gamma0}.
+\end{equation}
+
+\citeauthor{Rubinstein1996} express the screening length as
+$r_\mathrm{scr} = \left(A^2 / u\right)^{1/3} \left(\tau /cb \right)^{1/2}$. Combining with (Eq.~\ref{eq:ge}) and (Eq.~\ref{eq:D}) we deduce
+\begin{equation}
+g_\mathrm{scr} = \frac{A^2}{u} \left(\frac{\tau^3}{c b^3}\right)^{1/2}\label{eq:gscr}.
+\end{equation}
+
+From (Eq.~\ref{eq:gamma0}) and (Eq.~\ref{eq:gscr}) we can finally relate the measured critical strain to the microscopic parameter $A$:
+\begin{equation}
+A = \frac{\left(N u\right)^{1/2}}{\gamma_c+1}\left(\frac{c b^3}{\tau^3}\right)^{1/4}.
+\label{eq:A}
+\end{equation}
+
+The only unknown here is the reduced temperature $\tau$, possibly different between poly(pyrrolidinium) based and poly(immidazolium) based polymers.
+We know that the undissociated chain has a $\Theta$ temperature higher than \SI{100}{\celsius}, thus $0.2<\tau<1$.
+Setting $\tau\approx 0.235$, corresponding to $\Theta\approx\SI{116}{\celsius}$, yields $A=1$ for both \ce{PPyr+I-} and \ce{PPyr+Br-}, indicating full dissociation of the charges.
+By contrast \ce{PPyr+Cl-} has a dissociated charge every 7 monomers.
+Supposing the same $\tau$, \ce{PIm+I-} has a dissociated charge every 10 monomers, see Table~\ref{tab:results}.
+
+\subsubsection{Charge dissociation and cross-linking ratio}
+
+Finally we cross check our model by linking the charge dissociation and and the amount of physical cross-link by an independent argument.
+If we assume that the probability for a charge to dissociate is independent of the number of charges already present on the polymer, we can describe the probability of having $k$ charge dissociated on a polymer by a Poisson distribution of mean $\lambda = N_0/A$.
+Having 2 charges dissociated on a chain implies a cross-link.
+Therefore, we the cross-link probability $1/n$ is proportional to $\lambda^2$, yielding $n \sim A^2$.
+
+This is indeed what we observe, as shown in Figure~\ref{fig:dissociation}.
+We can extrapolate this law for the case of \ce{PIm+Br-} and obtain $A = 28$.
+Inverting equation~(\ref{eq:A}) yields $\gamma_c = 860\%$, indeed higher than the one observed (804\%).
+This confirms that in the \ce{PIm+Br-} strain sweep we were actually probing the strength of the ionic bonds.
+
+\begin{figure}
+\begin{center}
+\includegraphics{dissociation.pdf}
+\end{center}
+\caption{Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points. The line corresponds to $n = A^2$.}
+\label{fig:dissociation}
+\end{figure}
 
@@
 By comparison, \ce{PIm+Br-}, \ce{PIm+I-}, \ce{PPyr+Br-} and \ce{PPyr+I-} form homogeneous materials when suspended in water, which one can correlate with efficient dissociation of the halide counterions from the polymer main chain is fully consistency with decreased ionic bond strength.
-Such dissociation, in turns makes it possible for the ``hard'' phosphonate group on each polymer chain's end to interact with the ``free'' cationic groups of the side chains, which we showed in a previous paper to be at the origin of the gel formation mechanism.9
+Such dissociation, in turns makes it possible for the ``hard'' phosphonate group on each polymer chain's end to interact with the ``free'' cationic groups of the side chains, which we showed in a previous paper to be at the origin of the gel formation mechanism\cite{Srour2014}.
 
@@
 As already mentioned (vide supra) another striking feature of these gels is that in similar conditions poly(pyrrolidinium) based gels are systematically much stronger than those formed with poly(imidazolium).
-Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge; consequently, according to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected for the latter with the ``soft'' halide counterions.17 As a consequence, for a given halide, dissociation is less favored in the case of poly(imidazolium), as clearly illustrated by the difference in solvation between \ce{PIm+Cl-} and \ce{PPyr+Cl-}.
+Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge; consequently, according to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected for the latter with the ``soft'' halide counterions\cite{Goossens2009}.
+As a consequence, for a given halide, dissociation is less favored in the case of poly(imidazolium), as clearly illustrated by the difference in solvation between \ce{PIm+Cl-} and \ce{PPyr+Cl-}.
 Conversely, association of the lateral cations with the ``hard'' phosphonate moieties at the polymer chain’s end is expected to be stronger with pyrrolidinium, hence a systematically larger elastic modulus.

### 2016-01-08 PALSE/ion_pairing_paper.tex
Summary of the polymers derived from \ce{PBr}

@@
 \includegraphics{metathesis.pdf}
-\caption{Anionic metathesis reaction. As above, \ce{Nu+} corresponds either to pyrrolidinium or imidazolium.}
+\caption{(a) Anionic metathesis reaction. As above, \ce{Nu+} corresponds either to pyrrolidinium or imidazolium. (b) Summary of the polymers derived from \ce{PBr}.}
 \label{sch:metathesis}
@@
 Water-swelling properties turn out to be extremely dependent on the polymer compositions.
-With bromide or iodide conterions, both poly(pyrrolidinium) and poly(immidazolium) polymers affords homogeneous and optically transparent gels upon swelling with deionised water.
+With bromide or iodide conterions, both poly(pyrrolidinium) and poly(immidazolium) polymers afford homogeneous and optically transparent gels upon swelling with deionised water.
 We observe a limited swelling for \ce{PPyr+Cl-}, resulting in a granular, inhomogeneous gel.
@@
 
-We thus are able to dissolve the five polymers that have the weakest ion pairing, but neither the two fluorides nor \ce{PIm+Cl-} where counterions are expected to be strongly bounded to the chain.
+Therefore, we are able to dissolve the five polymers that have the weakest ion pairing, but neither the two fluorides nor \ce{PIm+Cl-} where counterions are expected to be strongly bounded to the chain.
 This means that in absence of charge dissociation the chain is in poor solvent.
-Furthermore, we held the two fluoride polymers in water at boiling temperature overnight without any dissolution.
-The $\Theta$ temperature is thus higher than \SI{100}{\celsius}.
+Furthermore, we held the two fluoride polymers in water at boiling temperature overnight without any dissolution, implying that the $\Theta$ temperature is higher than \SI{100}{\celsius}.
 Nonetheless, due to charge dissociation, most of our polymers are soluble at \SI{25}{\celsius}.
@@
 We perform rheological studies of the water-swelling materials.
-Briefly, data were recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}\cite{Macosko1994,Larson1999}.
-The sample was placed on the plate, the cone was then lowered to the measuring position, spreading the sample in the process.
-Excess material was removed, so that the sample exactly filled the gap.
-Due to Weissenberg effect at high shear rate, no preshear was applied before measurement.
-To minimize water absorption, the geometry was covered with a solvent trap, using light mineral oil as a liquid seal between the rotor and the cap.
-After one minute of equilibration we performed an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$).
-For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile was that of pure water which confirms the visual observation.
+Briefly, data are recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}\cite{Macosko1994,Larson1999}.
+We place the sample on the plate, and then lower the cone to the measuring position, spreading the sample in the process.
+We remove excess material, so that the sample exactly fills the gap.
+Due to Weissenberg effect at high shear rate, we apply no preshear before measurement.
+To minimize water absorption, we cover the geometry with a solvent trap, using light mineral oil as a liquid seal between the rotor and the cap.
+After one minute of equilibration we perform an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$).
+For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile is that of pure water which confirms the visual observation.
 However as shown in Figure~\ref{fig:strain}, all other samples are solidlike ($G^\prime \gg G^{\prime\prime}$) at small strain and flow at large strains ($G^{\prime\prime} \gg G^\prime$).
-Solidlike behavior at low strain was confirmed for all accessible frequencies (see supplementary material S6 Fig.
+We confirm solidlike behavior at low strain for all accessible frequencies (see supplementary material S6 Fig.
 1).
@@
 In Figure~\ref{fig:strain}a we compare the rheological responses of Poly(pyrrolidinium) based hydrogels for the three counterions allowing dissolution in water.
-Gels with bromide and iodide counterions showed roughly the same mechanical behavior in the linear regime but \ce{PPyr+I-} yields at slightly higher strains ($\gamma_c=4\%$ instead of $3\%$).
+Gels with bromide and iodide counterions show roughly the same mechanical behavior in the linear regime but \ce{PPyr+I-} yields at slightly higher strains ($\gamma_c=4\%$ instead of $3\%$).
 By contrast, the gel with chloride counterions is weaker by more than one order of magnitude.
@@
 For both iodide (Figure~\ref{fig:strain}b, X=I) and bromide (Figure~\ref{fig:strain}c, X=Br), imidazolium based gels are much weaker than their pyrrolidinium counterpart.
+
+\subsection{Crocodile chain model}
+We have previously demonstrated that, due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible~\cite{Srour2014}.
+If at least two foreign heads attach to the same body, we obtain an effective cross-linking point.
+When the probability of such a configuration is non-zero, we obtain a physically cross-linked gel.
+
+Here we argue that if every chain has a single foreign head attached to its body, every polymer is linked to two others in a crocodile line.
+We neglect the possibility of the two charges of the phosphonate head to attach to two separate chains.
+In addition, if we suppose that in order to minimize the inter-chain repulsion between charged cationic groups, heads are preferentially attached to the tail of their neighbour, we obtain a linear chain of effective polymerisation index $n N_0$ where $n$ is the number of polymers in the crocodile line.
+
+We thus have two limiting cases: (i) every chain has at least two heads attached and we have roughly $N_0$ monomers between cross-linking points; (ii) every chain has a single head attached and we have a single supramolecular chain in the system.
+The former case should be observed at high charge dissociation when many moieties are ionised along the chain providing many sites for a phosphonate head to bound to.
+The latter case should be observed at negligible charge dissociation.
+In between we should observe cross-linking points separated by crocodile lines of $n N_0$ monomers.
+
+Let us now discuss the rheological measurements and link them to microstructure.
+Amplitude sweep measurements at \SI{1}{\hertz} in a cone-plate geometry give access to two main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
+the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, the end of the linear regime and the first broken cross-links.
+
+\begin{table}
+\begin{tabular}{l|SS|SSS}
+& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_c$} & {$n$} & {$E_c$} & {$A$}\\
+&	{\si{\pascal}} &  & & {\si{\kilo\joule/\mol}} & \\\hline&&&&\\[-10pt]
+\ce{PPyr+I-}	& 8800	&	0.04	&	1.	&	0.004\cellcolor{gray!25}	&	1.\\
+\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.05&	0.001\cellcolor{gray!25}	&	1.1\\
+\ce{PPyr+Cl-}	& 270 	&	0.09	&	45	&	0.02\cellcolor{gray!25}	&	7.\\
+\ce{PIm+I-}	& 78	&	0.16	&	113	&	0.063\cellcolor{gray!25}	&	9.6\\
+\ce{PIm+Br-}	& 13	&	8.04	&	790	&	160.	&	3.4\cellcolor{gray!25}\\
+\end{tabular}
+\caption{Summary of rheological measurements and microscopic values deduced from the model.
+Gray background indicates incoherent values obtained either by (Eq.~\ref{eq:Ec}) or by (Eq.~\ref{eq:A}) outside of their respective validity domain.}
+\label{tab:results}
+\end{table}
+
+\subsubsection{From modulus to effective chain length}
+
+Between two cross-links the crocodile chain can be considered as an entropic spring made of $N=n N_0$ monomers. The shear modulus of the gel is then given by:
+\begin{equation}
+G = \frac{c}{N}k_\mathrm{B}T,
+\label{eq:G}
+\end{equation}
+where $c$ is the monomer number density, $k_\mathrm{B}$ is the Boltzmann constant and $T$ the temperature.
+Knowing the molecular mass $M$ of the chains, density $d$ of the solvent and the weight fraction $w$ of polymer, we obtain
+\begin{equation}
+n = \frac{\mathcal{N_A}}{M} w d \frac{k_\mathrm{B}T}{G}.
+\end{equation}
+
+We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 790 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
+This is coherent with a higher probability of attaching two or more heads on a highly dissociated body, i.e.
+a large cross-linking ratio.
+
+\subsubsection{Bounding energy}
+
+Considering that each crocodile chain is an entropic spring in its linear domain, the mechanical energy stored by a single crocodile chain at a strain $\gamma$ is $E = \gamma^2 k_\mathrm{B}T$.
+If we suppose that the ionic bonds breaks before the complete extension of the crocodile chains, we evaluate the head-to-body bonding energy to
+\begin{equation}
+E_c = \gamma_c^2 k_\mathrm{B}T.
+\label{eq:Ec}
+\end{equation}
+
+For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per bond or \SI{160}{\kilo\joule\per\mol}, coherent with typical energies of ionic bonds.
+However for all other samples the critical strain is between 2\% and 20\%, leading to incoherent values for the bonding energies between $10^{-3}k_\mathrm{B}T$ and $10^{-1}k_\mathrm{B}T$ per bond, see Table~\ref{tab:results}.
+Therefore, the hypothesis of a linear spring fails for these four samples, meaning that the crocodile chain is fully extended at $\gamma_c$.
+
+%\begin{figure}
+%\begin{center}
+%\begin{tikzpicture}
+%\begin{axis}[
+%	scale only axis,
+%	width=12\baselineskip, height=6\baselineskip, 
+%	axis lines=none,
+%	xmin=0.5, xmax=4.5,
+%	ymin=-0.7, ymax=1.3,
+%	]
+%\addplot[only marks, mark=o, mark size=2] table {/home/mathieu/tex/PALSE/persistent.xy} coordinate[pos=0.43] (a) coordinate[pos=0.57] (b) coordinate[pos=0.81](c);
+%\draw[<->] ($(a)+(-0.5em,0)$) -- ($(b)+(-0.5em,0)$) node[midway, left]{$r_\mathrm{scr}$};
+%\draw[<->] ($(c)+(-0.25em,-0.5em)$) -- ($(c)+(0.25em,-0.5em)$) node[midway, below]{$D$};
+%\end{axis}
+%\end{tikzpicture}
+%\end{center}
+%\caption{Sketch of the persistent random walk.}
+%\label{fig:persistent}
+%\end{figure}
+
+\subsubsection{From critical strain to charge dissociation}
+
+The spatial conformation of a polyelectrolyte chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions~\cite{Rubinstein1996}.
+In bad solvent, the persistent random walk is made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:persistent}:
+\begin{align}
+g_e &= \frac{A^2}{u}\tau\label{eq:ge},\\
+D &= b \left(\frac{A^2}{u}\right)^{1/3}\label{eq:D}.
+\end{align}
+Here $A$ is the number of monomers between uncondensed charges, $\tau = 1 - T/\Theta$ the reduced temperature and $u = \ell_\mathrm{B}/b \approx 2.7$ is the ratio of Bjerrum length $\ell_\mathrm{B}$ and monomer size $b\approx \SI{0.26}{\nano\metre}$.
+
+Extending these electrostatic blobs means exposing more monomers to the bad solvent and typically requires energies at least an order of magnitude higher than for the extension of the persistent random walk.
+We will thus consider that the exit from the linear domain of the material corresponds to the full extension of the persistent random walk:
+\begin{equation}
+\gamma_c = \frac{R_0}{R} - 1
+\end{equation}
+where $R$ and $R_0$ are respectively the end-to-end distance and the contour length of the persistent random walk.
+
+A segment of the persistent random walk is $r_\mathrm{scr}$ long and thus contains $g_\mathrm{scr} = g_e r_\mathrm{scr}/D$ monomers. We thus have
+\begin{equation}
+\left. \begin{array}{ll}
+R &= r_\mathrm{scr} \left(\frac{N}{g_\mathrm{scr}}\right)^{1/2}\\
+R_0 &= r_\mathrm{scr} \frac{N}{g_\mathrm{scr}}
+\end{array}\right\rbrace\quad
+\gamma_c = \left(\frac{N}{g_\mathrm{scr}}\right)^{1/2} -1\label{eq:gamma0}.
+\end{equation}
+
+\citeauthor{Rubinstein1996} express the screening length as
+$r_\mathrm{scr} = \left(A^2 / u\right)^{1/3} \left(\tau /cb \right)^{1/2}$. Combining with (Eq.~\ref{eq:ge}) and (Eq.~\ref{eq:D}) we deduce
+\begin{equation}
+g_\mathrm{scr} = \frac{A^2}{u} \left(\frac{\tau^3}{c b^3}\right)^{1/2}\label{eq:gscr}.
+\end{equation}
+
+From (Eq.~\ref{eq:gamma0}) and (Eq.~\ref{eq:gscr}) we can finally relate the measured critical strain to the microscopic parameter $A$:
+\begin{equation}
+A = \frac{\left(N u\right)^{1/2}}{\gamma_c+1}\left(\frac{c b^3}{\tau^3}\right)^{1/4}.
+\label{eq:A}
+\end{equation}
+
+The only unknown here is the reduced temperature $\tau$, possibly different between poly(pyrrolidinium) based and poly(immidazolium) based polymers.
+We know that the undissociated chain has a $\Theta$ temperature higher than \SI{100}{\celsius}, thus $0.2<\tau<1$.
+Setting $\tau\approx 0.235$, corresponding to $\Theta\approx\SI{116}{\celsius}$, yields $A=1$ for both \ce{PPyr+I-} and \ce{PPyr+Br-}, indicating full dissociation of the charges.
+By contrast \ce{PPyr+Cl-} has a dissociated charge every 7 monomers.
+Supposing the same $\tau$, \ce{PIm+I-} has a dissociated charge every 10 monomers, see Table~\ref{tab:results}.
+
+\subsubsection{Charge dissociation and cross-linking ratio}
+
+Finally we cross check our model by linking the charge dissociation and and the amount of physical cross-link by an independent argument.
+If we assume that the probability for a charge to dissociate is independent of the number of charges already present on the polymer, we can describe the probability of having $k$ charge dissociated on a polymer by a Poisson distribution of mean $\lambda = N_0/A$.
+Having 2 charges dissociated on a chain implies a cross-link.
+Therefore, we the cross-link probability $1/n$ is proportional to $\lambda^2$, yielding $n \sim A^2$.
+
+This is indeed what we observe, as shown in Figure~\ref{fig:dissociation}.
+We can extrapolate this law for the case of \ce{PIm+Br-} and obtain $A = 28$.
+Inverting equation~(\ref{eq:A}) yields $\gamma_c = 860\%$, indeed higher than the one observed (804\%).
+This confirms that in the \ce{PIm+Br-} strain sweep we were actually probing the strength of the ionic bonds.
+
+\begin{figure}
+\begin{center}
+\includegraphics{dissociation.pdf}
+\end{center}
+\caption{Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points. The line corresponds to $n = A^2$.}
+\label{fig:dissociation}
+\end{figure}
 

### 2016-01-08 ion_pairing_paper.tex
polymerisation index n -> N_0

@@
 We synthesise the phosphonate terminated polymer \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
-We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; S2, S3, Figure S1), which is the common basis to all studied polymers.
+We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; S2, S3, Figure S1), of polymerisation index $N_0=70$, which is the common basis to all studied polymers.
 We achieve all subsequent transformations of the polymer structures using the same initial batch of material.
@@
 \includegraphics{synthesis.pdf}
-\caption{Synthesis of \ce{PIm+Br-} and \ce{PPyr+Br-} and their intermediates \ce{POH} and \ce{PBr}. $n=70$}
+\caption{Synthesis of \ce{PIm+Br-} and \ce{PPyr+Br-} and their intermediates \ce{POH} and \ce{PBr}. $N_0=70$}
 \label{sch:synthesis}
@@
 \includegraphics{metathesis.pdf}
-\caption{Anionic metathesis reaction. As above, \ce{Nu+} corresponds either to pyrrolidinium or imidazolium.}
+\caption{(a) Anionic metathesis reaction. As above, \ce{Nu+} corresponds either to pyrrolidinium or imidazolium. (b) Summary of the polymers derived from a single batch of \ce{PBr}.}
 \label{sch:metathesis}

### 2016-01-11 PALSE/ion_pairing_paper.tex
Add the sketch of the network in fig:dissociation and update the text accordingly

@@
 \includegraphics{metathesis.pdf}
-\caption{(a) Anionic metathesis reaction. As above, \ce{Nu+} corresponds either to pyrrolidinium or imidazolium. (b) Summary of the polymers derived from \ce{PBr}.}
+\caption{(a) Anionic metathesis reaction. As above, \ce{Nu+} corresponds either to pyrrolidinium or imidazolium. (b) Summary of the polymers derived from a single batch of \ce{PBr}.}
 \label{sch:metathesis}
@@
 \subsection{Crocodile chain model}
+
+\begin{figure}
+\includegraphics{dissociation.pdf}
+\caption{Crocodile-line model.
+(a) Sketch of the network in the case of a crocodile line of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{src}\approx 3D$.
+Empty circles are not individual monomers but electrostatic blobs of size $D$.
+Electrostatic blobs containing an anionic head are shown as black filled circles.
+(b) Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points.
+The line corresponds to $n = A^2$.}
+\label{fig:dissociation}
+\end{figure}
+
 We have previously demonstrated that, due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible~\cite{Srour2014}.
@@
 The latter case should be observed at negligible charge dissociation.
-In between we should observe cross-linking points separated by crocodile lines of $n N_0$ monomers.
+In between we should observe cross-linking points separated by crocodile lines of $n N_0$ monomers as sketched in Figure~\ref{fig:dissociation}a.
 
@@
 the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, the end of the linear regime and the first broken cross-links.
-
-\begin{table}
-\begin{tabular}{l|SS|SSS}
-& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_c$} & {$n$} & {$E_c$} & {$A$}\\
-&	{\si{\pascal}} &  & & {\si{\kilo\joule/\mol}} & \\\hline&&&&\\[-10pt]
-\ce{PPyr+I-}	& 8800	&	0.04	&	1.	&	0.004\cellcolor{gray!25}	&	1.\\
-\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.05&	0.001\cellcolor{gray!25}	&	1.1\\
-\ce{PPyr+Cl-}	& 270 	&	0.09	&	45	&	0.02\cellcolor{gray!25}	&	7.\\
-\ce{PIm+I-}	& 78	&	0.16	&	113	&	0.063\cellcolor{gray!25}	&	9.6\\
-\ce{PIm+Br-}	& 13	&	8.04	&	790	&	160.	&	3.4\cellcolor{gray!25}\\
-\end{tabular}
-\caption{Summary of rheological measurements and microscopic values deduced from the model.
-Gray background indicates incoherent values obtained either by (Eq.~\ref{eq:Ec}) or by (Eq.~\ref{eq:A}) outside of their respective validity domain.}
-\label{tab:results}
-\end{table}
 
@@
 
+\begin{table}
+\begin{tabular}{l|SS|SSSS}
+& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_c$} & {$n$} & \multicolumn{2}{c}{$E_c$} & {$A$}\\
+&	{\si{\pascal}} &  & & {$k_\mathrm{B}T$} & {\si{\kilo\joule/\mol}} & \\\hline&&&&\\[-10pt]
+\ce{PPyr+I-}	& 8800	&	0.04	&	1.	&	2e-3\cellcolor{gray!25}	&	0.004\cellcolor{gray!25}	&	1.\\
+\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.05&	3e-4\cellcolor{gray!25}	&	0.001\cellcolor{gray!25}	&	1.1\\
+\ce{PPyr+Cl-}	& 270 	&	0.09	&	45	&	8e-3\cellcolor{gray!25}	&	0.02\cellcolor{gray!25}	&	7.\\
+\ce{PIm+I-}	& 78	&	0.16	&	113	&	3e-2\cellcolor{gray!25}	&	0.063\cellcolor{gray!25}	&	9.6\\
+\ce{PIm+Br-}	& 13	&	8.04	&	790	&	64	&	160.	&	3.4\cellcolor{gray!25}\\
+\end{tabular}
+\caption{Summary of rheological measurements and microscopic values deduced from the model.
+Gray background indicates incoherent values obtained either by (Eq.~\ref{eq:Ec}) or by (Eq.~\ref{eq:A}) outside of their respective validity domain.}
+\label{tab:results}
+\end{table}
+
 \subsubsection{Bounding energy}
@@
 
-%\begin{figure}
-%\begin{center}
-%\begin{tikzpicture}
-%\begin{axis}[
-%	scale only axis,
-%	width=12\baselineskip, height=6\baselineskip, 
-%	axis lines=none,
-%	xmin=0.5, xmax=4.5,
-%	ymin=-0.7, ymax=1.3,
-%	]
-%\addplot[only marks, mark=o, mark size=2] table {/home/mathieu/tex/PALSE/persistent.xy} coordinate[pos=0.43] (a) coordinate[pos=0.57] (b) coordinate[pos=0.81](c);
-%\draw[<->] ($(a)+(-0.5em,0)$) -- ($(b)+(-0.5em,0)$) node[midway, left]{$r_\mathrm{scr}$};
-%\draw[<->] ($(c)+(-0.25em,-0.5em)$) -- ($(c)+(0.25em,-0.5em)$) node[midway, below]{$D$};
-%\end{axis}
-%\end{tikzpicture}
-%\end{center}
-%\caption{Sketch of the persistent random walk.}
-%\label{fig:persistent}
-%\end{figure}
-
 \subsubsection{From critical strain to charge dissociation}
@@
 The spatial conformation of a polyelectrolyte chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions~\cite{Rubinstein1996}.
-In bad solvent, the persistent random walk is made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:persistent}:
+In bad solvent, the persistent random walk is made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:dissociation}a:
 \begin{align}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics{dissociation.pdf}
-\end{center}
-\caption{Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points. The line corresponds to $n = A^2$.}
-\label{fig:dissociation}
-\end{figure}
-
 \subsection{Discussion}

### 2016-01-11 PALSE/ion_pairing_paper.tex
Last paragraph of the intro. Plan and introduce crocodile line.

@@
 
-In the present report, we show that the mechanical/rheological properties of such smart hydrogels can be tuned to an extremely large extent by playing with the nature (hardness/softness) of the cationic repeating unit and the ion-paring strength of the associated counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}).
-The viscoelastic and diffusion properties of the resulting materials are characterized by different rheological and calorimetric analyses, evidencing a strengthening of the gel network upon reinforcing the hard character of the cation and the soft character of the anionic counterion.
+In the present report, we investigate the role of the dissociation rate of the poly(cationic) body on the mechanical/rheological properties of the hydrogel.
+In a first part we describe how we take advantage of our well-controlled synthesis to play systematically with the nature (hardness/softness) of the cationic repeating unit and the ion-paring strength of the associated counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}).
+In a second part, we show the qualitative and quantitative variations of mechanical properties of the aqueous dispersion obtained from these well-characterized polymers.
+In a third part, we develop a microscopic model based on the idea that at low dissociation rates, cross-links are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular crocodile line.
+Finally, we discuss our results and possible applications of such smart hydrogels.
 
@@
 
-\subsection{Crocodile chain model}
+\subsection{Crocodile line model}
 
@@
 \includegraphics{dissociation.pdf}
-\caption{Crocodile-line model.
+\caption{Crocodile line model.
 (a) Sketch of the network in the case of a crocodile line of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{src}\approx 3D$.

### 2016-01-12 ion_pairing_paper.tex
Bonding energies in units of kT in tab:results
Conclusion

@@
 
-\subsection{Crocodile chain model}
+\subsection{Crocodile line model}
+
+\begin{figure}
+\includegraphics{dissociation.pdf}
+\caption{Crocodile line model.
+(a) Sketch of the network in the case of a crocodile line of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{src}\approx 3D$.
+Empty circles are not individual monomers but electrostatic blobs of size $D$.
+Electrostatic blobs containing an anionic head are shown as black filled circles.
+(b) Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points.
+The line corresponds to $n = A^2$.}
+\label{fig:dissociation}
+\end{figure}
+
 We have previously demonstrated that, due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible~\cite{Srour2014}.
@@
 With this work we demonstrated the versatility of our ``head-to-body'' electrostatic approach in the fabrication of hydrogels with readily tunable rheological properties.
-Thus, depending on the nature of the nucleophile used in the quaternarization step of the polymer, but also on that of the counterion which can be modified in the course of its purification process, we showed that it is possible to manipulate almost at will the gel formability and respective magnitudes of the shear modulus and of the shear stress.
-We don’t observe the formation of a shear thinning gel when the halide counterion is ``hard.'' Consistent, the strongest gels are obtained when a combination of ``hard'' cations and ``soft'' halides are used, which gives rise, in line with the HSAB theory, to maximal charge dissociation in water: this in turn facilitates charge pairing of the peripheral, iterative, cations with the terminal anion.
-It is thus possible to obtain robust hydrogels with short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials.
+Thus, depending on the nature of the nucleophile used in the quaternarization step of the polymer, but also on that of the counterion which can be modified in the course of its purification process, we showed that it is possible to manipulate almost at will the gel formability, the density of physical cross-links and the respective magnitudes of the shear modulus and of the shear stress.
+
+When the halide counterion is ``hard'', there is not enough charge dissociation to allow dissolution and gel formation.
+On the other hand of the spectrum outlined by HSAB theory, the hardest gels with the narrowest linear domain are obtained when a combination of ``hard'' cations and ``soft'' halides are used.
+This gives rise to maximal charge dissociation in water, facilitates charge pairing of the peripheral, iterative, cations with the terminal anion and thus afford a very high density of cross-links.
+In between, in the case of ``hard'' cations and not so hard halide (\ce{Cl-}) or in the ``soft-soft'' case, charge dissociation is limited, polymers associate in crocodile lines and cross-links are few.
+We thus obtain soft, highly deformable gels.
+
+It is thus possible to obtain robust, highly tunable hydrogels with short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials.
 

### 2016-01-12 ion_pairing_paper.tex
Dispatch former discussion into results to introduce properly HSAB theory upfront.

@@
 
-\subsection{Water swelling and solvent quality}
+\subsection{Water swelling, ion-pairing and solvent quality}
 Water-swelling properties turn out to be extremely dependent on the polymer compositions.
@@
 
-Therefore, we are able to dissolve the five polymers that have the weakest ion pairing, but neither the two fluorides nor \ce{PIm+Cl-} where counterions are expected to be strongly bounded to the chain.
-This means that in absence of charge dissociation the chain is in poor solvent.
-Furthermore, we held the two fluoride polymers in water at boiling temperature overnight without any dissolution, implying that the $\Theta$ temperature is higher than \SI{100}{\celsius}.
-Nonetheless, due to charge dissociation, most of our polymers are soluble at \SI{25}{\celsius}.
+The above dependence of gel formability on the nature of the counterions could be rationalized on the basis of ion-pairing processes.
+According to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected between a ``softer'' cationic moiety and a ``harder'' halide counterions\cite{Goossens2009}.
+Therefore charge dissociation in water and thus solubility of the polymer should be easier with larger halide and pyrrolidinium.
+Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge.
+That is why we are able to dissolve the five polymers that have the weakest ion pairing, but neither the two fluorides nor \ce{PIm+Cl-} where counterions are expected to be strongly bounded to the chain.
+
+A fortiori, the insolubility in absence of charge dissociation indicates that the chain is in poor solvent.
+To estimate $\Theta$ temperature, we held the two fluoride polymers in water at boiling temperature overnight without any dissolution, implying $\Theta>\SI{100}{\celsius}$.
+Nonetheless, due to charge dissociation, five of our polyelectrolytes are soluble at \SI{25}{\celsius}.
 
@@
 By contrast, the gel with chloride counterions is weaker by more than one order of magnitude.
+In other words, a ``soft'' halide counterion, efficiently dissociated, is correlated to a harder gel.
 
 We also compare the rheology of \ce{PIm+X-} and \ce{PPyr+X-} with the same counterion \ce{X} in Figure~\ref{fig:strain}b,c.
-For both iodide (Figure~\ref{fig:strain}b, X=I) and bromide (Figure~\ref{fig:strain}c, X=Br), imidazolium based gels are much weaker than their pyrrolidinium counterpart.
+For both iodide (Figure~\ref{fig:strain}b, X=I) and bromide (Figure~\ref{fig:strain}c, X=Br), imidazolium based gels are much weaker than their pyrrolidinium counterpart with a broader linear regime.
+Comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ value (13 and \SI{8100}{\pascal}, resp.) and $\gamma_c$ value (804\% and 1.6\%, resp.) differ by almost three orders of magnitude.
+Here also, gel hardness is correlated to efficient charge dissociation, this time via ``harder'' cationic moieties.
+In the following we explain quantitatively this correlation from a microscopic model.
 
@@
 
-\subsection{Discussion}
-
-The above dependence of the gel parameters on the nature of the counterions could be rationalized on the basis of ion-pairing processes.
+%\subsection{Discussion}
+%
+%The above dependence of the gel parameters on the nature of the counterions could be rationalized on the basis of ion-pairing processes.
 The unsolubility/unswellability of \ce{PIm+F-}, \ce{PPyr+F-} and \ce{PIm+Cl-} suggest that ion pairs are not efficiently dissociated in aqueous medium, thus hindering the solvation process.
@@
 Such dissociation, in turns makes it possible for the ``hard'' phosphonate group on each polymer chain's end to interact with the ``free'' cationic groups of the side chains, which we showed in a previous paper to be at the origin of the gel formation mechanism\cite{Srour2014}.
-
-The similarity in gel parameters observed for \ce{PPyr+Br-} and \ce{PPyr+I-} ($G^\prime\approx\SI{8500}{\pascal}$) suggests that full ion pair dissociation is achieved in the latter, or at least that this dissociation is enough in both cases to provide maximal interaction between the terminal phosphonate and the pyrrolidinium side groups.
+%
+%The similarity in gel parameters observed for \ce{PPyr+Br-} and \ce{PPyr+I-} ($G^\prime\approx\SI{8500}{\pascal}$) suggests that full ion pair dissociation is achieved in the latter, or at least that this dissociation is enough in both cases to provide maximal interaction between the terminal phosphonate and the pyrrolidinium side groups.
 Conversely the significantly different behavior between \ce{PIm+I-} ($G^\prime= \SI{78}{\pascal}$) and \ce{PIm+Br-} ($G^\prime= \SI{13}{\pascal}$) indicates that dissociation is only partial and occurs to a lesser extent in the latter than in the former.
 The same conclusion can be drawn for \ce{PPyr+Cl-}, in which the elastic modulus ($G^\prime= \SI{270}{\pascal}$) value drops by more than one order of magnitude as compared to its bromide and iodide counterparts.
-
-As already mentioned (vide supra) another striking feature of these gels is that in similar conditions poly(pyrrolidinium) based gels are systematically much stronger than those formed with poly(imidazolium).
+%
+%As already mentioned (vide supra) another striking feature of these gels is that in similar conditions poly(pyrrolidinium) based gels are systematically much stronger than those formed with poly(imidazolium).
 Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge; consequently, according to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected for the latter with the ``soft'' halide counterions\cite{Goossens2009}.

### 2016-01-12 PALSE/ion_pairing_paper.tex
remoe repetitions in the model

@@
 Let us now discuss the rheological measurements and link them to microstructure.
-Amplitude sweep measurements at \SI{1}{\hertz} in a cone-plate geometry give access to two main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
+We reduce our measurements to two main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
 the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, the end of the linear regime and the first broken cross-links.
@@
 
-This is indeed what we observe, as shown in Figure~\ref{fig:dissociation}.
+As shown in Figure~\ref{fig:dissociation}b this relation describes very well our data.
 We can extrapolate this law for the case of \ce{PIm+Br-} and obtain $A = 28$.

### 2016-01-14 PALSE/ion_pairing_paper.tex
discuss 160kJ/mol > 2* 5kJ/mol (ionic bonding in water)

@@
 
-For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per bond or \SI{160}{\kilo\joule\per\mol}, coherent with typical energies of ionic bonds.
+For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per head-to-body bond or \SI{160}{\kilo\joule\per\mol}.
+This value is much higher than the $E_i\approx\SI{5}{\kilo\joule\per\mol}$ typical of ionic bonds in water\cite{Schneider1992}, even when adding up the two bonds of the bivalent phosphonate head.
+This implies that the electrostatic bond is reinforced by other supramolecular interactions, probably hydrophobic interactions between the surrounding uncharged monomers of both chains and $\pi$-stacking of the imidazolium\cite{Rehm2010}.
+
 However for all other samples the critical strain is between 2\% and 20\%, leading to incoherent values for the bonding energies between $10^{-3}k_\mathrm{B}T$ and $10^{-1}k_\mathrm{B}T$ per bond, see Table~\ref{tab:results}.
@@
 
+%We can now estimate the hydrophobic interaction energy $E_h$ by considering that to separate two chains two interfaces with water of area $\pi D^2/4$ have to be created.
+Surface tension is $\tau^2 k_\mathrm{B}T / b^2$\cite{Rubinstein1996}.
+Combining with (Eq.~\ref{eq:D}) we find
+%\begin{equation}
+%\frac{E_i}{k_\mathrm{B}T} = \frac{\pi}{2}\tau^2\left(\frac{A^2}{u}\right)^{2/3} \approx 4
+%\end{equation}
+
 %\subsection{Discussion}

### 2016-01-14 PALSE/ion_pairing_paper.tex
apolar micro environement

@@
 
-Between two cross-links the crocodile chain can be considered as an entropic spring made of $N=n N_0$ monomers. The shear modulus of the gel is then given by:
+Between two cross-links the crocodile chain can be considered as an entropic spring made of $N=n N_0$ monomers. The shear modulus of the gel is then given by\cite{Rubinstein1996}:
 \begin{equation}
@@
 This value is much higher than the $E_i\approx\SI{5}{\kilo\joule\per\mol}$ typical of ionic bonds in water\cite{Schneider1992}, even when adding up the two bonds of the bivalent phosphonate head.
-This implies that the electrostatic bond is reinforced by other supramolecular interactions, probably hydrophobic interactions between the surrounding uncharged monomers of both chains and $\pi$-stacking of the imidazolium\cite{Rehm2010}.
+However the surrounding medium is not water but the hydrophobic chains.
+Such an apolar microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}.
+The local dielectric constant can easily be 16 times lower than water, justifying the large bonding energy we measure.
 
@@
 
-The spatial conformation of a polyelectrolyte chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions~\cite{Rubinstein1996}.
+The spatial conformation of a polyelectrolyte chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions\cite{Rubinstein1996}.
 In bad solvent, the persistent random walk is made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:dissociation}a:

### 2016-01-22 ion_pairing_paper.tex
Insert Cyrille changes in DSC part, but in the present and active voice.

@@
 
-In order to characterize the effect of counterion substitution, we study all polymers by Differential Scanning Calorimetry (DSC).
-In solids, melting points are inherently connected to the reticular enthalpy of the crystal lattice, hence primarily to ion pairing strength\cite{Sherman1932}.
+In ionic solids, melting points are inherently connected to the reticular enthalpy of the crystal lattice, hence primarily to ion pairing strength\cite{Sherman1932}.
 This last parameter is directly related to the diffuse and hindered nature of the interacting charges, bulkier and softer ions providing in the most general case weaker interactions\cite{Krossing2006}.
 For typical ``ionic liquids'' such as imidazolium or pyrollidinium halides, melting points will thus gradually increase as the size of the halide counterion decreases, ie $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$\cite{Dean2010}.
-Although the trend is less obvious in amorphous polymeric systems, in which phase transitions result from the interplay of multiple parameters, ionic substitution is however expected to have an impact on the melting temperature of the material\cite{Orler1994}.
+Although analysis is less straightforward in the case of amorphous polymer electrolytes, ionic substitution is however expected to have an impact on the phase transition temperatures of the material\cite{Orler1994}.
 
-As expected, the lowest melting temperature was recorded for polymers with iodide counterions.
-Quite remarkably,we found a similar melting temperature (84-\SI{85}{\celsius}) for \ce{PIm+I-} and \ce{PPyr+I-}, see Figure~\ref{fig:dsc}.
-When bromide counterions are present, we observe a significant increase of the melting temperature (114-\SI{115}{\celsius}) but again, with a similar value between \ce{PIm+Br-} and \ce{PPyr+Br-} as shown in Figure~\ref{fig:dsc}.
-In the case of fluoride and chloride, we observe no phase transition below decomposition temperature of the polymer (\SI{200}{\celsius}), which indicates a much stronger ion pairing in these materials.
-% no need of fig S2-5 for me, but a wy not a supplementary figure with chloride and fluoride.
+In order to characterize the effect of counterion substitution, we study all polymers by Differential Scanning Calorimetry (DSC).
+Although we observe no glass transition in the studied temperature range for any of the sample, comparison of the DSC data reveal significant differences.
+
+In the case polyelectrolytes with fluoride and chloride counterion, we observe no exo or endothermic transition below \SI{200}{\celsius}.
+By contrast, we observe a broad endothermic peak when iodide and bromide are used as counterions.
+Quite remarkably,we found a similar peak temperature (84-\SI{85}{\celsius}) for \ce{PIm+I-} and \ce{PPyr+I-}, see Figure~\ref{fig:dsc}.
+When bromide counterions are present, we observe a significant increase of the peak temperature (114-\SI{115}{\celsius}) but again, with a similar value between \ce{PIm+Br-} and \ce{PPyr+Br-} as shown in Figure~\ref{fig:dsc}.
+The overall shape and position of the DSC peaks is very reminiscent of previously reported data for various naturally occuring or synthetic poly(electrolytes).[ Li, L.; Hsieh, Y.-L., Ultra-fine polyelectrolyte fibers from electrospinning of poly(acrylic acid).
+Polymer 2005, 46 (14), 5133-5139.
+; Ostrowska-Czubenko, J.; Gierszewska-Drużyńska, M., Effect of ionic crosslinking on the water state in hydrogel chitosan membranes.
+Carbohydrate Polymers 2009, 77 (3), 590-598., Deb, T.
+K.; Ramireddy, B.; Moin, A.; Shivakumar, H.
+G., In vitro-in vivo evaluation of xanthan gum and eudragit inter polyelectrolyte complex based sustained release tablets.
+International Journal of Pharmaceutical Investigation 2015, 5 (1), 65-72.), Sarmento, B.; Ferreira, D.; Veiga, F.; Ribeiro, A., Characterization of insulin-loaded alginate nanoparticles produced by ionotropic pre-gelation through DSC and FTIR studies.
+Carbohydrate Polymers 2006, 66 (1), 1-7.) It is generally attributed to desorption of weakly bound water from the polymer network.
+Its presence in the polymer samples where bromide and iodide are used as couterions suggests that adsorbed water could play a role in solvating the dissociated ion pairs in the latter; conversely, its absence for chloride and fluoride counterions suggests that dissociation is not effective because of stronger ion pair interaction.
 
@@
 
-For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per bond or \SI{160}{\kilo\joule\per\mol}, coherent with typical energies of ionic bonds.
+For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per head-to-body bond or \SI{160}{\kilo\joule\per\mol}.
+This value is much higher than the $E_i\approx\SI{5}{\kilo\joule\per\mol}$ typical of ionic bonds in water\cite{Schneider1992}, even when adding up the two bonds of the bivalent phosphonate head.
+However the surrounding medium is not water but the hydrophobic chains.
+Such an apolar microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}.
+The local dielectric constant can easily be 16 times lower than water, justifying the large bonding energy we measure.
+
 However for all other samples the critical strain is between 2\% and 20\%, leading to incoherent values for the bonding energies between $10^{-3}k_\mathrm{B}T$ and $10^{-1}k_\mathrm{B}T$ per bond, see Table~\ref{tab:results}.
@@
 This confirms that in the \ce{PIm+Br-} strain sweep we were actually probing the strength of the ionic bonds.
+
+%We can now estimate the hydrophobic interaction energy $E_h$ by considering that to separate two chains two interfaces with water of area $\pi D^2/4$ have to be created.
+Surface tension is $\tau^2 k_\mathrm{B}T / b^2$\cite{Rubinstein1996}.
+Combining with (Eq.~\ref{eq:D}) we find
+%\begin{equation}
+%\frac{E_i}{k_\mathrm{B}T} = \frac{\pi}{2}\tau^2\left(\frac{A^2}{u}\right)^{2/3} \approx 4
+%\end{equation}
 

### 2016-01-22 ion_pairing_paper.tex
Cyrille citations of DSC and water sorbtion

@@
 When bromide counterions are present, we observe a significant increase of the peak temperature (114-\SI{115}{\celsius}) but again, with a similar value between \ce{PIm+Br-} and \ce{PPyr+Br-} as shown in Figure~\ref{fig:dsc}.
-The overall shape and position of the DSC peaks is very reminiscent of previously reported data for various naturally occuring or synthetic poly(electrolytes).[ Li, L.; Hsieh, Y.-L., Ultra-fine polyelectrolyte fibers from electrospinning of poly(acrylic acid).
-Polymer 2005, 46 (14), 5133-5139.
-; Ostrowska-Czubenko, J.; Gierszewska-Drużyńska, M., Effect of ionic crosslinking on the water state in hydrogel chitosan membranes.
-Carbohydrate Polymers 2009, 77 (3), 590-598., Deb, T.
-K.; Ramireddy, B.; Moin, A.; Shivakumar, H.
-G., In vitro-in vivo evaluation of xanthan gum and eudragit inter polyelectrolyte complex based sustained release tablets.
-International Journal of Pharmaceutical Investigation 2015, 5 (1), 65-72.), Sarmento, B.; Ferreira, D.; Veiga, F.; Ribeiro, A., Characterization of insulin-loaded alginate nanoparticles produced by ionotropic pre-gelation through DSC and FTIR studies.
-Carbohydrate Polymers 2006, 66 (1), 1-7.) It is generally attributed to desorption of weakly bound water from the polymer network.
+The overall shape and position of the DSC peaks is very reminiscent of previously reported data for various naturally occuring or synthetic poly(electrolytes)\cite{Li2005,Sarmento2006,Ostrowska-Czubenko2009a,Moin2015}.
+It is generally attributed to desorption of weakly bound water from the polymer network.
 Its presence in the polymer samples where bromide and iodide are used as couterions suggests that adsorbed water could play a role in solvating the dissociated ion pairs in the latter; conversely, its absence for chloride and fluoride counterions suggests that dissociation is not effective because of stronger ion pair interaction.

### 2016-01-22 PALSE/ion_pairing_paper.tex
Remove "melting" from any conclusion or description of the data.

@@
 
-In order to characterize the effect of counterion substitution, we study all polymers by Differential Scanning Calorimetry (DSC).
-In solids, melting points are inherently connected to the reticular enthalpy of the crystal lattice, hence primarily to ion pairing strength\cite{Sherman1932}.
+In ionic solids, melting points are inherently connected to the reticular enthalpy of the crystal lattice, hence primarily to ion pairing strength\cite{Sherman1932}.
 This last parameter is directly related to the diffuse and hindered nature of the interacting charges, bulkier and softer ions providing in the most general case weaker interactions\cite{Krossing2006}.
 For typical ``ionic liquids'' such as imidazolium or pyrollidinium halides, melting points will thus gradually increase as the size of the halide counterion decreases, ie $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$\cite{Dean2010}.
-Although the trend is less obvious in amorphous polymeric systems, in which phase transitions result from the interplay of multiple parameters, ionic substitution is however expected to have an impact on the melting temperature of the material\cite{Orler1994}.
+Although analysis is less straightforward in the case of amorphous polymer electrolytes, ionic substitution is however expected to have an impact on the phase transition temperatures of the material\cite{Orler1994}.
 
-As expected, the lowest melting temperature was recorded for polymers with iodide counterions.
-Quite remarkably,we found a similar melting temperature (84-\SI{85}{\celsius}) for \ce{PIm+I-} and \ce{PPyr+I-}, see Figure~\ref{fig:dsc}.
-When bromide counterions are present, we observe a significant increase of the melting temperature (114-\SI{115}{\celsius}) but again, with a similar value between \ce{PIm+Br-} and \ce{PPyr+Br-} as shown in Figure~\ref{fig:dsc}.
-In the case of fluoride and chloride, we observe no phase transition below decomposition temperature of the polymer (\SI{200}{\celsius}), which indicates a much stronger ion pairing in these materials.
-% no need of fig S2-5 for me, but a wy not a supplementary figure with chloride and fluoride.
+In order to characterize the effect of counterion substitution, we study all polymers by Differential Scanning Calorimetry (DSC).
+Although we observe no glass transition in the studied temperature range for any of the sample, comparison of the DSC data reveal significant differences.
+
+In the case polyelectrolytes with fluoride and chloride counterion, we observe no exo or endothermic transition below \SI{200}{\celsius}.
+By contrast, we observe a broad endothermic peak when iodide and bromide are used as counterions.
+Quite remarkably,we found a similar peak temperature (84-\SI{85}{\celsius}) for \ce{PIm+I-} and \ce{PPyr+I-}, see Figure~\ref{fig:dsc}.
+When bromide counterions are present, we observe a significant increase of the peak temperature (114-\SI{115}{\celsius}) but again, with a similar value between \ce{PIm+Br-} and \ce{PPyr+Br-} as shown in Figure~\ref{fig:dsc}.
+The overall shape and position of the DSC peaks is very reminiscent of previously reported data for various naturally occuring or synthetic poly(electrolytes)\cite{Li2005,Sarmento2006,Ostrowska-Czubenko2009a,Moin2015}.
+It is generally attributed to desorption of weakly bound water from the polymer network.
+Its presence in the polymer samples where bromide and iodide are used as couterions suggests that adsorbed water could play a role in solvating the dissociated ion pairs in the latter; conversely, its absence for chloride and fluoride counterions suggests that dissociation is not effective because of stronger ion pair interaction.
 
@@
 \includegraphics{dsc.pdf}
-\caption{DSC measurements of polymer samples showing the presence of a melting point.
+\caption{DSC measurements of polymer samples showing the presence of a endothermic peak.
 Curves are shifted vertically by arbitrary amounts for clarity.
-Vertical lines indicate the melting points at \SI{84}{\celsius} for the two iodides in black and \SI{114}{\celsius} for the two bromides in gray.}
+Vertical lines indicate the peak temperatures at \SI{84}{\celsius} for the two iodides in black and \SI{114}{\celsius} for the two bromides in gray.}
 \label{fig:dsc}

### 2016-01-28 ion_pairing_paper.tex
Tibbitt2016

@@
 Besides, poly(cationic) gel have been reported to combine scaffold properties for cell’s adherence and growth with antimicrobial activity, and are therefore increasingly used in clinical applications\cite{Hoffman2001,Landers2002}.
-In this framework, injectability is a key feature, as it provides an easier way to gel delivery in vivo; thus, reversible shear thinning biocompatible hydrogels are a particularly sought after class of materials\cite{Yu2008}.
+In this framework, injectability is a key feature, as it provides an easier way to gel delivery in vivo\cite{Tibbitt2016}; thus, reversible shear thinning biocompatible hydrogels are a particularly sought after class of materials\cite{Yu2008}.
 

### 2016-01-29 PALSE/ion_pairing_paper.tex
Title order and authors order

@@
 \author{Hassan Srour}
-\author{Martien Duvall Deffo Ayagou}
 \affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}
@@
 
+\author{Martien Duvall Deffo Ayagou}
 \author{Thi Thanh-Tam Nguyen}
@@
 \author{Nicolas Taberlet}
-\author{Sebastien Manneville}
+\author{Sébastien Manneville}
 \affiliation[Laboratoire de Physique de l'ENS de Lyon]{Laboratoire de Physique, Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon, 46 Allée d'Italie, 69007 Lyon}
@@
 %\title{Supramolecular crocodile line controls the rheological properties of polyelectrolytes hydrogels}
-\title{Ion pairing controls rheological properties of polyelectrolyte crocodile line hydrogels}
+\title{Ion pairing controls rheological properties of ``crocodile-line'' polyelectrolyte hydrogels}
 %\title{Ion pairing in crocodile line hydrogels controls rheological properties}
@@
 
-It is thus possible to obtain robust, highly tunable hydrogels with short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials.
+It is thus possible to obtain robust, highly tunable hydrogels from short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials.
 
 \begin{acknowledgement}
-The authors are grateful for the financial support of Université de Lyon through the program ``Investissements d'Avenir'' ( ANR-1 1-IDEX-0007).
+The authors are grateful for the financial support of Université de Lyon through the program ``Investissements d'Avenir'' (ANR-1 1-IDEX-0007).
 S.M.

### 2016-01-29 PALSE/ion_pairing_paper.tex
Corrections Seb.

@@
 Here we generate a variety of such polyelectrolytes with various cationic moieties and counter-ions combinations starting from a common polymeric platform.
-We show that the rheological properties (shear modulus, critical strain) of the final hydrogels vary by three orders of magnitude depending on the cation/pair.
+We show that the rheological properties (shear modulus, critical strain) of the final hydrogels vary by three orders of magnitude depending on the cation/anion pair.
 Our data are well described by a theoretical model involving a supramolecular crocodile line between cross-links, the length of the line being set by the amount of charge dissociation.
@@
 \section{Introduction}
-Since the pioneering work of \citeauthor{Wichterle1960} which established their relevance in a biomedical context, polymer based hydrogels have never ceased to be a very active field of research\cite{Calo2015,Hoffman2001}.
-They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargos or scaffolds for wound healing or tissues repair\cite{Vashist2014,Ratner2004}.
-Polyelectrolytes are being extensively used in this particular context\cite{Rosso2003,BinImran2014}.
+Since the pioneering work of \citet{Wichterle1960} which established their relevance in a biomedical context, polymer-based hydrogels have never ceased to be a very active field of research\cite{Calo2015,Hoffman2001}.
+They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargos or scaffolds for wound healing or tissue repair\cite{Vashist2014,Ratner2004}.
+Polyelectrolytes are being extensively put to use in this particular context\cite{Rosso2003,BinImran2014}.
 Their gel formation mechanism often involves reversible-by-nature electrostatic interactions, which can be used for instance to trigger ``smart'' release of bioactive substances\cite{Khare1993,Lockwood2007}.
-Besides, poly(cationic) gel have been reported to combine scaffold properties for cell’s adherence and growth with antimicrobial activity, and are therefore increasingly used in clinical applications\cite{Hoffman2001,Landers2002}.
+Besides, poly(cationic) gels have been reported to combine scaffold properties for cell’s adherence and growth with antimicrobial activity, and are therefore increasingly used in clinical applications\cite{Hoffman2001,Landers2002}.
 In this framework, injectability is a key feature, as it provides an easier way to gel delivery in vivo\cite{Tibbitt2016}; thus, reversible shear thinning biocompatible hydrogels are a particularly sought after class of materials\cite{Yu2008}.
@@
 We have recently reported on a highly reversible poly(cation) based hydrogel\cite{Srour2014}, which formation relies on an new concept of supramolecular  electrostatic interaction.
-Briefly, the atom transfer radical polymerisation (ATRP) is initiated by a phosphonate terminated ATRP initiator, eventually affording an anion terminated poly(cationic) polymer.
+Briefly, the atom transfer radical polymerisation (ATRP) is initiated by a phosphonate-terminated ATRP initiator, eventually affording an anion-terminated poly(cationic) polymer.
 The resulting poly(cationic) polymer, when dispersed in an aqueous medium, stabilizes hydrogel formation through the occurrence of a head(anion)-to-body(cations) supramolecular network.
@@
 
-In the present report, we investigate the role of the dissociation rate of the poly(cationic) body on the mechanical/rheological properties of the hydrogel.
-In a first part we describe how we take advantage of our well-controlled synthesis to play systematically with the nature (hardness/softness) of the cationic repeating unit and the ion-paring strength of the associated counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}).
-In a second part, we show the qualitative and quantitative variations of mechanical properties of the aqueous dispersion obtained from these well-characterized polymers.
+In the present report, we investigate the role of the dissociation rate of the poly(cationic) body on the mechanical properties of the hydrogel.
+In a first part we describe how we take advantage of our well-controlled synthesis to play systematically with the nature (hardness/softness) of the cationic repeating unit and the ion-paring strength of the associated counter-ion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}).
+In a second part, we show the qualitative and quantitative variations of mechanical properties of the aqueous dispersions obtained from these well-characterized polymers.
 In a third part, we develop a microscopic model based on the idea that at low dissociation rates, cross-links are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular crocodile line.
-Finally, we discuss our results and possible applications of such smart hydrogels.
+%Finally, we discuss our results and possible applications of such smart hydrogels.
 
@@
 
-We synthesise the phosphonate terminated polymer \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
-We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; S2, S3, Figure S1), of polymerisation index $N_0=70$, which is the common basis to all studied polymers.
+We synthesise the phosphonate terminated polymer hereafter reffered to as \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
+We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; see Supplementary Materials S2, S3, Figure S1), of polymerisation index $N_0=70$, which is the common basis to all studied polymers.
 We achieve all subsequent transformations of the polymer structures using the same initial batch of material.
-Nucleophilic addition of N-Methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF affords the corresponding poly(imidazolium) and poly(pyrrolidinium) compounds.
+Nucleophilic addition of N-methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF affords the corresponding poly(imidazolium) and poly(pyrrolidinium) compounds.
 In these cases, as an inherent consequence of the structure of the starting material, charge neutrality is provided by bromide counter ions, and the polymer will be referred to as \ce{PIm+Br-} and \ce{PPyr+Br-}, respectively (Scheme~\ref{sch:synthesis}).
@@
 \includegraphics{synthesis.pdf}
-\caption{Synthesis of \ce{PIm+Br-} and \ce{PPyr+Br-} and their intermediates \ce{POH} and \ce{PBr}. $N_0=70$}
+\caption{Synthesis of \ce{PIm+Br-} and \ce{PPyr+Br-} and their intermediates \ce{POH} and \ce{PBr} with $N_0=70$.}
 \label{sch:synthesis}
@@
 After extensive dialysis of the resulting mixture against deionized water and lyophilization we recover in high yields the different \ce{PIm+X-} and \ce{PPyr+X-} (where \ce{X}=\ce{F}, \ce{Cl}, \ce{I}).
-In order to bring evidence for the efficiency of the ionic metathesis, we submit the resulting materials to negative mode high resolution mass spectrometry (HRMS, S3).
+In order to bring evidence for the efficiency of the ionic metathesis, we submit the resulting materials to negative-mode high-resolution mass spectrometry (HRMS, see Supplementary Materials S3).
 We unambiguously assess displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Figures S8-9).
-We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Figures S6-7), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (see supporting information).
+We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures S6-7), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (see supporting information XXX S?
+XXX).
 
@@
 This last parameter is directly related to the diffuse and hindered nature of the interacting charges, bulkier and softer ions providing in the most general case weaker interactions\cite{Krossing2006}.
-For typical ``ionic liquids'' such as imidazolium or pyrollidinium halides, melting points will thus gradually increase as the size of the halide counterion decreases, ie $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$\cite{Dean2010}.
+For typical ``ionic liquids'' such as imidazolium or pyrrolidinium halides, melting points will thus gradually increase as the size of the halide counterion decreases, ie $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$\cite{Dean2010}.
 Although analysis is less straightforward in the case of amorphous polymer electrolytes, ionic substitution is however expected to have an impact on the phase transition temperatures of the material\cite{Orler1994}.
@@
 
-In the case polyelectrolytes with fluoride and chloride counterion, we observe no exo or endothermic transition below \SI{200}{\celsius}.
+In the case polyelectrolytes with fluoride and chloride counterion, we observe no exo- or endothermic transition below \SI{200}{\celsius}.
 By contrast, we observe a broad endothermic peak when iodide and bromide are used as counterions.
-Quite remarkably,we found a similar peak temperature (84-\SI{85}{\celsius}) for \ce{PIm+I-} and \ce{PPyr+I-}, see Figure~\ref{fig:dsc}.
+Quite remarkably, we found a similar peak temperature (84-\SI{85}{\celsius}) for \ce{PIm+I-} and \ce{PPyr+I-}, see Figure~\ref{fig:dsc}.
 When bromide counterions are present, we observe a significant increase of the peak temperature (114-\SI{115}{\celsius}) but again, with a similar value between \ce{PIm+Br-} and \ce{PPyr+Br-} as shown in Figure~\ref{fig:dsc}.
@@
 It is generally attributed to desorption of weakly bound water from the polymer network.
-Its presence in the polymer samples where bromide and iodide are used as couterions suggests that adsorbed water could play a role in solvating the dissociated ion pairs in the latter; conversely, its absence for chloride and fluoride counterions suggests that dissociation is not effective because of stronger ion pair interaction.
+Its presence in the polymer samples where bromide and iodide are used as couterions suggests that adsorbed water could play a role in solvating the dissociated ion pairs in the latter.
+Conversely, its absence for chloride and fluoride counterions suggests that dissociation is not effective because of stronger ion pair interaction.
 
@@
 \includegraphics{dsc.pdf}
-\caption{DSC measurements of polymer samples showing the presence of a endothermic peak.
+\caption{DSC measurements of polymer samples showing the presence of an endothermic peak.
 Curves are shifted vertically by arbitrary amounts for clarity.
@@
 \subsection{Water swelling, ion-pairing and solvent quality}
-Water-swelling properties turn out to be extremely dependent on the polymer compositions.
+Water-swelling properties turn out to be extremely dependent on the polymer composition.
 With bromide or iodide conterions, both poly(pyrrolidinium) and poly(immidazolium) polymers afford homogeneous and optically transparent gels upon swelling with deionised water.
@@
 Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge.
-That is why we are able to dissolve the five polymers that have the weakest ion pairing, but neither the two fluorides nor \ce{PIm+Cl-} where counterions are expected to be strongly bounded to the chain.
+That is why we are able to dissolve the five polymers that have the weakest ion pairing, but neither the two fluorides nor \ce{PIm+Cl-} where counterions are expected to be strongly bound to the chain.
 
 A fortiori, the insolubility in absence of charge dissociation indicates that the chain is in poor solvent.
-To estimate $\Theta$ temperature, we held the two fluoride polymers in water at boiling temperature overnight without any dissolution, implying $\Theta>\SI{100}{\celsius}$.
+To estimate the $\Theta$-temperature, we held the two fluoride polymers in water at boiling temperature overnight without any dissolution, implying that $\Theta>\SI{100}{\celsius}$.
 Nonetheless, due to charge dissociation, five of our polyelectrolytes are soluble at \SI{25}{\celsius}.
@@
 
-In Figure~\ref{fig:strain}a we compare the rheological responses of Poly(pyrrolidinium) based hydrogels for the three counterions allowing dissolution in water.
+In Figure~\ref{fig:strain}a we compare the rheological responses of Poly(pyrrolidinium)-based hydrogels for the three counterions allowing dissolution in water.
 Gels with bromide and iodide counterions show roughly the same mechanical behavior in the linear regime but \ce{PPyr+I-} yields at slightly higher strains ($\gamma_c=4\%$ instead of $3\%$).
@@
 
-We have previously demonstrated that, due to the opposite charge of the phosphonate head, head-to-body ionic bonds are possible~\cite{Srour2014}.
-If at least two foreign heads attach to the same body, we obtain an effective cross-linking point.
+We have previously demonstrated that, due to the opposite charges of the phosphonate head, head-to-body ionic bonds are possible~\cite{Srour2014}.
+On the one hand, if at least two foreign heads attach to the same body, we obtain an effective cross-link point.
 When the probability of such a configuration is non-zero, we obtain a physically cross-linked gel.
 
-Here we argue that if every chain has a single foreign head attached to its body, every polymer is linked to two others in a crocodile line.
-We neglect the possibility of the two charges of the phosphonate head to attach to two separate chains.
+On the other hand, if every chain has a single foreign head attached to its body, every polymer is linked to two others in a crocodile line\bibnote{We neglect the possibility of the two charges of the phosphonate head to attach to two separate chains}.
 In addition, if we suppose that in order to minimize the inter-chain repulsion between charged cationic groups, heads are preferentially attached to the tail of their neighbour, we obtain a linear chain of effective polymerisation index $n N_0$ where $n$ is the number of polymers in the crocodile line.
 
-We thus have two limiting cases: (i) every chain has at least two heads attached and we have roughly $N_0$ monomers between cross-linking points; (ii) every chain has a single head attached and we have a single supramolecular chain in the system.
+We thus have two limiting cases: (i) every chain has at least two heads attached and we have roughly $N_0$ monomers between cross-link points; (ii) every chain has a single head attached and we have a single supramolecular chain in the system.
 The former case should be observed at high charge dissociation when many moieties are ionised along the chain providing many sites for a phosphonate head to bound to.
 The latter case should be observed at negligible charge dissociation.
-In between we should observe cross-linking points separated by crocodile lines of $n N_0$ monomers as sketched in Figure~\ref{fig:dissociation}a.
+In between these two limiting cases, we should observe cross-link points separated by crocodile lines of $n N_0$ monomers as sketched in Figure~\ref{fig:dissociation}a.
 
-Let us now discuss the rheological measurements and link them to microstructure.
+Let us now discuss the rheological measurements and link them to the hydrogel microstructure.
 We reduce our measurements to two main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
-the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, the end of the linear regime and the first broken cross-links.
+the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, that we consider as representative of the end of the linear regime and of the first broken cross-links.
+XXX Ref XXX
 
@@
 We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 790 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
-This is coherent with a higher probability of attaching two or more heads on a highly dissociated body, i.e.
-a large cross-linking ratio.
+This is consistent with a higher probability of attaching two or more heads on a highly dissociated body, i.e.
+a large cross-link ratio.
 
 \begin{table}
-\begin{tabular}{l|SS|SSSS}
-& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_c$} & {$n$} & \multicolumn{2}{c}{$E_c$} & {$A$}\\
-&	{\si{\pascal}} &  & & {$k_\mathrm{B}T$} & {\si{\kilo\joule/\mol}} & \\\hline&&&&\\[-10pt]
-\ce{PPyr+I-}	& 8800	&	0.04	&	1.	&	2e-3\cellcolor{gray!25}	&	0.004\cellcolor{gray!25}	&	1.\\
-\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.05&	3e-4\cellcolor{gray!25}	&	0.001\cellcolor{gray!25}	&	1.1\\
-\ce{PPyr+Cl-}	& 270 	&	0.09	&	45	&	8e-3\cellcolor{gray!25}	&	0.02\cellcolor{gray!25}	&	7.\\
-\ce{PIm+I-}	& 78	&	0.16	&	113	&	3e-2\cellcolor{gray!25}	&	0.063\cellcolor{gray!25}	&	9.6\\
-\ce{PIm+Br-}	& 13	&	8.04	&	790	&	64	&	160.	&	3.4\cellcolor{gray!25}\\
+\begin{tabular}{l|SS|SSSSS}
+& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_c$} & {$n$} & \multicolumn{2}{c}{$E_c$} & {$A$} & {$\sqrt{n}$}\\
+&	{\si{\pascal}} &  & & {$k_\mathrm{B}T$} & {\si{\kilo\joule/\mol}} & \\\hline&&&&&\\[-10pt]
+\ce{PPyr+I-}	& 8800	&	0.04	&	1.	&	2e-3\cellcolor{gray!25}	&	0.004\cellcolor{gray!25}	&	1.	&	1.\\
+\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.05&	3e-4\cellcolor{gray!25}	&	0.001\cellcolor{gray!25}	&	1.1	&	1.\\
+\ce{PPyr+Cl-}	& 270 	&	0.09	&	45	&	8e-3\cellcolor{gray!25}	&	0.02\cellcolor{gray!25}	&	7.	&	6.7\\
+\ce{PIm+I-}	& 78	&	0.16	&	113	&	3e-2\cellcolor{gray!25}	&	0.063\cellcolor{gray!25}	&	9.6	&	10.6\\
+\ce{PIm+Br-}	& 13	&	8.04	&	790	&	64	&	160.	&	3.4\cellcolor{gray!25}	&	28\\
 \end{tabular}
 \caption{Summary of rheological measurements and microscopic values deduced from the model.
-Gray background indicates incoherent values obtained either by (Eq.~\ref{eq:Ec}) or by (Eq.~\ref{eq:A}) outside of their respective validity domain.}
+Gray background indicates inconsistent values obtained either by (Eq.~\ref{eq:Ec}) or by (Eq.~\ref{eq:A}) outside of their respective validity domain.}
 \label{tab:results}
@@
 
-Considering that each crocodile chain is an entropic spring in its linear domain, the mechanical energy stored by a single crocodile chain at a strain $\gamma$ is $E = \gamma^2 k_\mathrm{B}T$.
-If we suppose that the ionic bonds breaks before the complete extension of the crocodile chains, we evaluate the head-to-body bonding energy to
+Considering that each crocodile chain is an entropic spring in its linear domain, the mechanical energy stored by a single crocodile chain under a strain $\gamma$ is $E = \gamma^2 k_\mathrm{B}T$.
+If we suppose that ionic bonds break before the complete extension of the crocodile chains, we evaluate the head-to-body bonding energy to
 \begin{equation}
@@
 For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per head-to-body bond or \SI{160}{\kilo\joule\per\mol}.
-This value is much higher than the $E_i\approx\SI{5}{\kilo\joule\per\mol}$ typical of ionic bonds in water\cite{Schneider1992}, even when adding up the two bonds of the bivalent phosphonate head.
-However the surrounding medium is not water but the hydrophobic chains.
+This value is much higher than the $E_i\approx\SI{5}{\kilo\joule\per\mol}$ typical of ionic bonds in water\cite{Schneider1992}, even when adding up the two bonds of the bivalent phosphonate head\bibnote{Since the two minus charges are born by the same atom, chelation effect may enhance bounding energy, but not by an order of magnitude}.
+However, we have to remember that the chain is in bad solvent, thus forming locally compact blobs almost devoid of water.
+Therefore, the medium surrounding the ionic bond is not water but the hydrophobic chains.
 Such an apolar microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}.
-The local dielectric constant can easily be 16 times lower than water, justifying the large bonding energy we measure.
+Quantitatively, the bonding energy of an ionic bond is inversely proportional to the relative dielectric constant $\epsilon_r$, therefore we need to suppose a local $\epsilon_r$ around 5 (instead of 80 in water) to recover the large bonding energy we measure.
 
-However for all other samples the critical strain is between 2\% and 20\%, leading to incoherent values for the bonding energies between $10^{-3}k_\mathrm{B}T$ and $10^{-1}k_\mathrm{B}T$ per bond, see Table~\ref{tab:results}.
-Therefore, the hypothesis of a linear spring fails for these four samples, meaning that the crocodile chain is fully extended at $\gamma_c$.
+However for all other samples the critical strain is between 2\% and 20\%, leading to inconsistent values for the bonding energies ranging between $10^{-3}k_\mathrm{B}T$ and $10^{-1}k_\mathrm{B}T$ per bond, see Table~\ref{tab:results}.
+Therefore, the end of the linear regime cannot correspond to bond breaking but rather to the point where extended chains become nonlinear.
 
@@
 \end{align}
-Here $A$ is the number of monomers between uncondensed charges, $\tau = 1 - T/\Theta$ the reduced temperature and $u = \ell_\mathrm{B}/b \approx 2.7$ is the ratio of Bjerrum length $\ell_\mathrm{B}$ and monomer size $b\approx \SI{0.26}{\nano\metre}$.
+Here $A$ is the number of monomers between dissociated ion pairs, $\tau = 1 - T/\Theta$ the reduced temperature and $u = \ell_\mathrm{B}/b \approx 2.7$ is the ratio of the Bjerrum length in water $\ell_\mathrm{B}\approx \SI{0.7}{\nano\metre}$ and monomer size $b\approx \SI{0.26}{\nano\metre}$.
 
@@
 We know that the undissociated chain has a $\Theta$ temperature higher than \SI{100}{\celsius}, thus $0.2<\tau<1$.
-Setting $\tau\approx 0.235$, corresponding to $\Theta\approx\SI{116}{\celsius}$, yields $A=1$ for both \ce{PPyr+I-} and \ce{PPyr+Br-}, indicating full dissociation of the charges.
-By contrast \ce{PPyr+Cl-} has a dissociated charge every 7 monomers.
-Supposing the same $\tau$, \ce{PIm+I-} has a dissociated charge every 10 monomers, see Table~\ref{tab:results}.
+In addition, we must have $A\geq 1$, which implies through (Eq.~\ref{eq:A}) $\tau\leq 0.235$, corresponding to $\Theta\approx\SI{116}{\celsius}$.
+This limit value yields $A=1$ for both \ce{PPyr+I-} and \ce{PPyr+Br-}, indicating full dissociation of the cationic moieties.
+By contrast \ce{PPyr+Cl-} has a charge every 7 monomers.
+Supposing the same $\tau$, \ce{PIm+I-} has a dissociated ion pair every 10 monomers, see Table~\ref{tab:results}.
+Of course analysing \ce{PIm+Br-} based on the value of $\gamma_c$ gives an inconsistent value $A=3.4$ indicating that the hydrogel doesn't belong to the same regime.
 
@@
 
-Finally we cross check our model by linking the charge dissociation and and the amount of physical cross-link by an independent argument.
-If we assume that the probability for a charge to dissociate is independent of the number of charges already present on the polymer, we can describe the probability of having $k$ charge dissociated on a polymer by a Poisson distribution of mean $\lambda = N_0/A$.
+Finally we cross-check our model by linking the ion pair dissociation and the amount of physical cross-links by an independent argument.
+If we assume that the probability for an ion pair to dissociate is independent of the number of dissociated pairs already present on the polymer, we can describe the probability of having $k$ dissociated pairs on a polymer by a Poisson distribution of mean $\lambda = N_0/A$.
 Having 2 charges dissociated on a chain implies a cross-link.
-Therefore, we the cross-link probability $1/n$ is proportional to $\lambda^2$, yielding $n \sim A^2$.
+Therefore, the cross-link probability $1/n$ is proportional to $\lambda^2$, yielding $n \sim A^2$.
 
-As shown in Figure~\ref{fig:dissociation}b this relation describes very well our data.
+As shown in Figure~\ref{fig:dissociation}b and the last two columns of Table~\ref{tab:results} this relation describes very well our data.
 We can extrapolate this law for the case of \ce{PIm+Br-} and obtain $A = 28$.
-Inverting equation~(\ref{eq:A}) yields $\gamma_c = 860\%$, indeed higher than the one observed (804\%).
-This confirms that in the \ce{PIm+Br-} strain sweep we were actually probing the strength of the ionic bonds.
+Inverting (Eq.~\ref{eq:A}) yields $R_0/R - 1 = 860\%$, indeed higher than the observed $\gamma_c\approx 804\%$.
+This confirms that in the \ce{PIm+Br-} strain sweep $\gamma_c$ actually results from the strength of the ionic bonds.
 
@@
 On the other hand of the spectrum outlined by HSAB theory, the hardest gels with the narrowest linear domain are obtained when a combination of ``hard'' cations and ``soft'' halides are used.
-This gives rise to maximal charge dissociation in water, facilitates charge pairing of the peripheral, iterative, cations with the terminal anion and thus afford a very high density of cross-links.
-In between, in the case of ``hard'' cations and not so hard halide (\ce{Cl-}) or in the ``soft-soft'' case, charge dissociation is limited, polymers associate in crocodile lines and cross-links are few.
+This gives rise to maximal charge dissociation in water, facilitates charge pairing of the peripheral, iterative cations with the terminal anion and thus affords a very high density of cross-links.
+In between, in the case of ``hard'' cations and not-so-hard halide (\ce{Cl-}) or in the ``soft-soft'' case, charge dissociation is limited, polymers associate in crocodile lines and cross-links are few.
 We thus obtain soft, highly deformable gels.
 
-It is thus possible to obtain robust, highly tunable hydrogels from short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials.
+To conclude, our procedure yields robust, highly tunable hydrogels from short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials.
 

### 2016-01-29 ion_pairing_paper.tex
weight fraction vs molar concentration paves the way to supra-molecular assembly

@@
 where $c$ is the monomer number density, $k_\mathrm{B}$ is the Boltzmann constant and $T$ the temperature.
-Knowing the molecular mass $M$ of the chains, density $d$ of the solvent and the weight fraction $w$ of polymer, we obtain
+Because of the different weights of the pyrrolidinium and imidazolium moieties and of the different counter-ions, $c$ is not constant across our experiments done at constant weight fraction $w$ of polymer.
+This explains why the respective modulus or critical strains of \ce{PPyr+I-} (heavier, lower number density) and \ce{PPyr+Br-} are in reverse order with respect to the \textit{a priori} ranking of charge dissociation.
+Taking into account the molecular mass $M$ of each polymer and  knowing the density $d$ of the solvent, we obtain
 \begin{equation}
 n = \frac{\mathcal{N_A}}{M} w d \frac{k_\mathrm{B}T}{G}.
+\label{eq:n}
 \end{equation}
@@
 On the other hand of the spectrum outlined by HSAB theory, the hardest gels with the narrowest linear domain are obtained when a combination of ``hard'' cations and ``soft'' halides are used.
-This gives rise to maximal charge dissociation in water, facilitates charge pairing of the peripheral, iterative, cations with the terminal anion and thus afford a very high density of cross-links.
-In between, in the case of ``hard'' cations and not so hard halide (\ce{Cl-}) or in the ``soft-soft'' case, charge dissociation is limited, polymers associate in crocodile lines and cross-links are few.
+This gives rise to maximal charge dissociation in water, facilitates charge pairing of the peripheral, iterative cations with the terminal anion and thus affords a very high density of cross-links.
+In between, in the case of ``hard'' cations and not-so-hard halide (\ce{Cl-}) or in the ``soft-soft'' case, charge dissociation is limited, polymers associate in crocodile lines and cross-links are few.
 We thus obtain soft, highly deformable gels.
 
-It is thus possible to obtain robust, highly tunable hydrogels with short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials.
+To conclude, our procedure yields robust, highly tunable hydrogels from short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials and paves the way for future works on supra-molecular assembly.
 
 \begin{acknowledgement}
-The authors are grateful for the financial support of Université de Lyon through the program ``Investissements d'Avenir'' ( ANR-1 1-IDEX-0007).
+The authors are grateful for the financial support of Université de Lyon through the program ``Investissements d'Avenir'' (ANR-1 1-IDEX-0007).
 S.M.

### 2016-01-29 PALSE/ion_pairing_paper.tex
Show extrapolation for PImBr in fig:dissociation

@@
 (b) Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points.
-The line corresponds to $n = A^2$.}
+Full circles are obtained from (Eq.~\ref{eq:A}) in its validity domain.
+Empty circles are obtained for \ce{PIm+Br-} either via (Eq.~\ref{eq:A}) outside its validity or by extrapolation of the solid line that corresponds to $n = A^2$.}
 \label{fig:dissociation}
@@
 
-To conclude, our procedure yields robust, highly tunable hydrogels from short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials.
+To conclude, our procedure yields robust, highly tunable hydrogels from short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials and paves the way for future works on supra-molecular assembly.
 

### 2016-01-29 PALSE/ion_pairing_paper.tex
no empty circle in fig:dissociation better caption

@@
 (b) Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points.
-Full circles are obtained from (Eq.~\ref{eq:A}) in its validity domain.
-Empty circles are obtained for \ce{PIm+Br-} either via (Eq.~\ref{eq:A}) outside its validity or by extrapolation of the solid line that corresponds to $n = A^2$.}
+The solid line corresponds to $n = A^2$.
+Full circles are obtained from (Eq.~\ref{eq:A}).
+In the case of \ce{PIm+Br-}, we are not in the validity regime of (Eq.~\ref{eq:A}), and we have to extrapolate the solid line to estimate $A$ (dotted arrow).}
 \label{fig:dissociation}
@@
 Supposing the same $\tau$, \ce{PIm+I-} has a dissociated ion pair every 10 monomers, see Table~\ref{tab:results}.
-Of course analysing \ce{PIm+Br-} based on the value of $\gamma_c$ gives an inconsistent value $A=3.4$ indicating that the hydrogel doesn't belong to the same regime.
+Of course analysing \ce{PIm+Br-} based on the value of $\gamma_c$ gives an inconsistent value $A=3.4$ indicating that this hydrogel doesn't belong to the same regime.
 

### 2016-02-01 ion_pairing_paper.tex
A word about collapsed blob and relevant citations when discussing F-

@@
 A fortiori, the insolubility in absence of charge dissociation indicates that the chain is in poor solvent.
+Locally, the monomers are condensed in collapsed globules to minimize the contact with solvent, and solubility on larger scales can only be achieved if the polymer bears enough charges\cite{Khokhlov1980,Raphael1990}.
 To estimate the $\Theta$-temperature, we held the two fluoride polymers in water at boiling temperature overnight without any dissolution, implying that $\Theta>\SI{100}{\celsius}$.
@@
 (b) Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points.
-The line corresponds to $n = A^2$.}
+The solid line corresponds to $n = A^2$.
+Full circles are obtained from (Eq.~\ref{eq:A}).
+In the case of \ce{PIm+Br-}, we are not in the validity regime of (Eq.~\ref{eq:A}), and we have to extrapolate the solid line to estimate $A$ (dotted arrow).}
 \label{fig:dissociation}
@@
 Therefore, the medium surrounding the ionic bond is not water but the hydrophobic chains.
-Such an apolar microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}.
+Such a low polarity microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}.
 Quantitatively, the bonding energy of an ionic bond is inversely proportional to the relative dielectric constant $\epsilon_r$, therefore we need to suppose a local $\epsilon_r$ around 5 (instead of 80 in water) to recover the large bonding energy we measure.

### 2016-02-01 PALSE/ion_pairing_paper.tex
cite Hyun2011 for maximum of loss modulus

@@
 We reduce our measurements to two main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
-the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, that we consider as representative of the end of the linear regime and of the first broken cross-links.
-XXX Ref XXX
+the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, that we consider as representative of the end of the linear regime and of the first broken cross-links\cite{Hyun2011}.
 

### 2016-02-01 PALSE/ion_pairing_SI.tex
Supplementary method with no repetition with the main text. Mass spectra as scans

\documentclass[journal=jacsat,manuscript=article]{achemso}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}       % Use modern font encodings
\usepackage{amssymb}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage{siunitx}
\usepackage[table]{xcolor}

\usepackage{xr,xc}
\externaldocument[M-]{ion_pairing_paper}
\externalcitedocument{ion_pairing_paper}

% Hack for making figures Say \figurename S\thefigure, e.g. Figure S1:
\makeatletter
\makeatletter \renewcommand{\fnum@figure}
{\textbf{Figure~S\thefigure}}
\makeatother

\renewcommand{\thetable}{\arabic{table}}
\makeatletter
\makeatletter \renewcommand{\fnum@table}
{\textbf{Tab.~S\thetable}}
\makeatother

\author{Hassan Srour}
\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}

\author{Mathieu Leocmach}
\affiliation[Institut Lumière Matière]{Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}
\email{mathieu.leocmach@univ-lyon1.fr}

\author{Martien Duvall Deffo Ayagou}
\author{Thi Thanh-Tam Nguyen}
\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}

\author{Nicolas Taberlet}
\author{Sébastien Manneville}
\affiliation[Laboratoire de Physique de l'ENS de Lyon]{Laboratoire de Physique, Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon, 46 Allée d'Italie, 69007 Lyon}

\author{Chantal Andraud}
\author{Cyrille Monnereau}
\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}
\email{cyrille.monnereau@ens-lyon.fr}

%\title{Supramolecular crocodile line controls the rheological properties of polyelectrolytes hydrogels}
\title{Supplementary Informations to Ion pairing controls rheological properties of ``crocodile-line'' polyelectrolyte hydrogels}
%\title{Ion pairing in crocodile line hydrogels controls rheological properties}
%\title{In crocodile line hydrogels ion pairing controls rheological properties}
%\title{Controlling the rheological properties of well-defined polyelectrolytes hydrogels by fine tuning their ion pairing process}

\keywords{ATRP, hydrogel, poly(cations), rheology}

\begin{document}

%\section{Materials and methods}
All reactions were performed under an argon atmosphere using schlenk techniques. CH2Cl2 and THF were dried and degassed on a solvent station by passage through an activated alumina column followed by argon flush. All other solvents were used without additional purification. HEA purchased from Alfa Aesar was purified as reported in Ref~\cite{Srour2014} of the main text. Additional chemicals were obtained from Sigma Aldrich, Acros or Alfa Aesar and underwent no further purification.

SEC analyses were carried out on ca \SI{5}{\milli\gram\per\milli\litre} solutions of each polymer in DMSO LiBr (\SI{0.01}{M}), using a MALVERN VISCOTEK 430  system equipped with three PSS Gram columns (Polyester Copolymer Network) set in series at \SI{80}{\celsius}, coupled to a Viscotek HT-GPC Module 350 A (refractive and viscosimeter) detector (at \SI{80}{\celsius}) . The polydispersity indexes (PDI = Mw/Mn) of the samples were derived from the RI signal by a universal calibration curve by pullulan. The number-average molar masses (Mn) were calculated from RI signals with the OmniSEC 4.6 software.

The high resolution mass spectra (MS QTOF) were recorded in positive and negative ion modes on a hybrid quadrupole time-of-flight mass spectrometer (MicroTOFQ-II, Bruker Daltonics, Bremen) with an electrospray ionization (ESI) ion source. The gas flow of the spray gas is \SI{0.6}{\bar} and the capillary voltage is $\pm\SI{4.5}{\kilo\volt}$. The solutions are infused at \SI{180}{\micro\litre\per\hour}. The mass range of the analysis is 50-1000~m/z and the calibration was done with sodium formate.

\begin{figure}
\includegraphics[width=\textwidth]{sec_spectra.png}
\caption{SEC Spectra of \ce{POH}.}
\end{figure}

\begin{figure}
\includegraphics[height=\textheight-2\baselineskip]{mass_PImBr.png}
\caption{High resolution mass spectra of \ce{PIm+Br-}. Bromide peak exist at m/z=81.0.}
\end{figure}

\begin{figure}
\includegraphics[height=\textheight-2\baselineskip]{mass_PPyBr.png}
\caption{High resolution mass spectra of \ce{PPy+Br-}. Bromide peak exist at m/z=79-81.}
\end{figure}

\begin{figure}
\includegraphics[height=\textheight-2\baselineskip]{mass_PImI.png}
\caption{High resolution mass spectra of \ce{PIm+I-}. Iodide peak exist at m/z=126.9.}
\end{figure}

\begin{figure}
\includegraphics[height=\textheight-2\baselineskip]{mass_PPyI.png}
\caption{High resolution mass spectra of \ce{PPy+I-}. Iodide peak exist at m/z=127.1.}
\end{figure}

\end{document}

### 2016-02-01 ion_pairing_paper.tex
Supplementary method with no repetition with the main text. Mass spectra as scans

@@
 
-We synthesise the phosphonate terminated polymer hereafter reffered to as \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
-We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; see Supplementary Materials S2, S3, Figure S1), of polymerisation index $N_0=70$, which is the common basis to all studied polymers.
+We synthesise the phosphonate terminated polymer hereafter referred to as \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
+We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; see Supplementary Materials~S1 and S3, Supplementary Scheme~1 and Supplementary Figure~1), of polymerisation index $N_0=70$, which is the common basis to all studied polymers.
 We achieve all subsequent transformations of the polymer structures using the same initial batch of material.
@@
 In order to bring evidence for the efficiency of the ionic metathesis, we submit the resulting materials to negative-mode high-resolution mass spectrometry (HRMS, see Supplementary Materials S3).
-We unambiguously assess displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Figures S8-9).
-We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures S6-7), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (see supporting information XXX S?
+We unambiguously assess displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Supplementary Figures~7-8).
+We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures~5-6), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (see supporting information XXX S?
 XXX).

### 2016-02-01 PALSE/ion_pairing_SI.tex
frequency sweep figure

@@
 
+\begin{figure}
+\includegraphics{frequency.pdf}
+\caption{Storage modulus $G^\prime$ (filled symbols) and loss modulus $G^{\prime\prime}$ (open symbols) measured through oscillatory shear experiments plotted against the frequency $f$ for (a) \ce{PPyr+Cl-} (\textcolor{lightgray}{$\bullet$}), \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}) and \ce{PPyr+I-} ($\blacktriangle$), (b) \ce{PIm+I-} ($\blacktriangledown$) and \ce{PPyr+I-} ($\blacktriangle$) and (c) \ce{PIm+Br-} (\textcolor{gray}{$\blacklozenge$}) and \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}).
+The fixed strain amplitude is frequency $\gamma=0.1\%$.
+All samples are at 8\%wt }
+\label{fig:frequency}
+\end{figure}
+
 \end{document}

### 2016-02-01 ion_pairing_paper.tex
proper references to the SI in the main text

@@
 \usepackage[table]{xcolor}
+
+\usepackage{xr}
+\externaldocument[S-]{ion_pairing_SI}
 
@@
 We synthesise the phosphonate terminated polymer hereafter referred to as \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
-We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; see Supplementary Materials~S1 and S3, Supplementary Scheme~1 and Supplementary Figure~1), of polymerisation index $N_0=70$, which is the common basis to all studied polymers.
+We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; see Supplementary Materials and Supplementary Figure~S\ref{S-fig:sec}), of polymerisation index $N_0=70$, which is the common basis to all studied polymers.
 We achieve all subsequent transformations of the polymer structures using the same initial batch of material.
@@
 After extensive dialysis of the resulting mixture against deionized water and lyophilization we recover in high yields the different \ce{PIm+X-} and \ce{PPyr+X-} (where \ce{X}=\ce{F}, \ce{Cl}, \ce{I}).
-In order to bring evidence for the efficiency of the ionic metathesis, we submit the resulting materials to negative-mode high-resolution mass spectrometry (HRMS, see Supplementary Materials S3).
-We unambiguously assess displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Supplementary Figures~7-8).
-We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures~5-6), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (see supporting information XXX S?
-XXX).
+In order to bring evidence for the efficiency of the ionic metathesis, we submit the resulting materials to negative-mode high-resolution mass spectrometry (HRMS, see Supplementary Materials).
+We unambiguously assess displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Supplementary Figures~S\ref{S-fig:massPImBr} and S\ref{S-fig:massPPyBr}).
+We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures~S\ref{S-fig:massPImI} and S\ref{S-fig:massPPyI}), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (not shown).
 
@@
 To minimize water absorption, we cover the geometry with a solvent trap, using light mineral oil as a liquid seal between the rotor and the cap.
-After one minute of equilibration we perform an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$).
+After one minute of equilibration we perform an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$, see Supplementary Figure~S\ref{S-fig:frequency}) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$).
 For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile is that of pure water which confirms the visual observation.

### 2016-02-02 ion_pairing_paper.tex
Correction Seb. Water insoluble polymer dielectric 2-3

@@
 Full circles are obtained from (Eq.~\ref{eq:A}).
-In the case of \ce{PIm+Br-}, we are not in the validity regime of (Eq.~\ref{eq:A}), and we have to extrapolate the solid line to estimate $A$ (dotted arrow).}
+In the case of \ce{PIm+Br-}, we are not in the validity regime of (Eq.~\ref{eq:A}), and we have to extrapolate the solid line to estimate $A$ (dotted arrow and open symbol).}
 \label{fig:dissociation}
@@
 For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per head-to-body bond or \SI{160}{\kilo\joule\per\mol}.
-This value is much higher than the $E_i\approx\SI{5}{\kilo\joule\per\mol}$ typical of ionic bonds in water\cite{Schneider1992}, even when adding up the two bonds of the bivalent phosphonate head\bibnote{Since the two minus charges are born by the same atom, chelation effect may enhance bounding energy, but not by an order of magnitude}.
-However, we have to remember that the chain is in bad solvent, thus forming locally compact blobs almost devoid of water.
+This value is much higher than the $E_i\approx\SI{5}{\kilo\joule\per\mol}$ typical of ionic bonds in water\cite{Schneider1992}, even when adding up the two bonds of the bivalent phosphonate head\bibnote{Since the two minus charges are borne by the same atom, chelation effect may enhance bounding energy, but not by an order of magnitude}.
+However, one has to keep in mind that the chain is in bad solvent, thus forming locally compact blobs of low water content.
 Therefore, the medium surrounding the ionic bond is not water but the hydrophobic chains.
 Such a low polarity microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}.
-Quantitatively, the bonding energy of an ionic bond is inversely proportional to the relative dielectric constant $\epsilon_r$, therefore we need to suppose a local $\epsilon_r$ around 5 (instead of 80 in water) to recover the large bonding energy we measure.
+Quantitatively, the bonding energy of an ionic bond is inversely proportional to the relative dielectric constant $\epsilon_r$, therefore we need to suppose a local $\epsilon_r$ around 5 (between 2-3 typical of water-insoluble polymers and 80 for water) to recover the large bonding energy that we measure.
 
@@
 We know that the undissociated chain has a $\Theta$ temperature higher than \SI{100}{\celsius}, thus $0.2<\tau<1$.
-In addition, we must have $A\geq 1$, which implies through (Eq.~\ref{eq:A}) $\tau\leq 0.235$, corresponding to $\Theta\approx\SI{116}{\celsius}$.
+In addition, we must have $A\geq 1$, which implies that $\tau\leq 0.235$ through (Eq.~\ref{eq:A}), corresponding to $\Theta\approx\SI{116}{\celsius}$.
 This limit value yields $A=1$ for both \ce{PPyr+I-} and \ce{PPyr+Br-}, indicating full dissociation of the cationic moieties.
@@
 Supposing the same $\tau$, \ce{PIm+I-} has a dissociated ion pair every 10 monomers, see Table~\ref{tab:results}.
-Of course analysing \ce{PIm+Br-} based on the value of $\gamma_c$ gives an inconsistent value $A=3.4$ indicating that this hydrogel doesn't belong to the same regime.
+Of course analysing \ce{PIm+Br-} based on the value of $\gamma_c$ gives an inconsistent value $A=3.4$ indicating that this hydrogel does not belong to the same regime.
 

### 2016-02-08 ion_pairing_paper.tex
Correction Cyrille crocodile-line -> processionary

@@
 %\title{Supramolecular crocodile line controls the rheological properties of polyelectrolytes hydrogels}
-\title{Ion pairing controls rheological properties of ``crocodile-line'' polyelectrolyte hydrogels}
+\title{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}
 %\title{Ion pairing in crocodile line hydrogels controls rheological properties}
@@
 Here we generate a variety of such polyelectrolytes with various cationic moieties and counter-ions combinations starting from a common polymeric platform.
-We show that the rheological properties (shear modulus, critical strain) of the final hydrogels vary by three orders of magnitude depending on the cation/anion pair.
-Our data are well described by a theoretical model involving a supramolecular crocodile line between cross-links, the length of the line being set by the amount of charge dissociation.
+We show that the rheological properties (shear modulus, critical strain) of the final hydrogels can be modulated over three orders of magnitude depending on the cation/anion pair.
+Our data fit remarkably well within a theoretical model involving a supramolecular head-to-tail single file between cross-links, akin to the behaviour of pine-processionary caterpillar, the length of the procession being set by the amount of charge dissociation.
 \end{abstract}
@@
 
-In the present report, we investigate the role of the dissociation rate of the poly(cationic) body on the mechanical properties of the hydrogel.
-In a first part we describe how we take advantage of our well-controlled synthesis to play systematically with the nature (hardness/softness) of the cationic repeating unit and the ion-paring strength of the associated counter-ion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}).
+In the present report, we investigate the role of the ion pair dissociation equilibria rate of the poly(cationic) body on the mechanical properties of the hydrogel.
+In a first part we describe how we take advantage of our postfunctionalization approach to play systematically with the nature (hardness/softness) of the cationic repeating unit and the ion-paring strength of the associated counter-ion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}).
 In a second part, we show the qualitative and quantitative variations of mechanical properties of the aqueous dispersions obtained from these well-characterized polymers.
-In a third part, we develop a microscopic model based on the idea that at low dissociation rates, cross-links are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular crocodile line.
-%Finally, we discuss our results and possible applications of such smart hydrogels.
+In a third part, we develop a microscopic model based on the idea that at low dissociation rates, cross-links (defined as a polymer that is connected to three or more others) are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular chain.
+We call this behaviour ``processionary'' in analogy to pine processionary caterpillar (\textit{Thaumetopoea pityocampa}) behaviour\cite{Fabre1916}.%Finally, we discuss our results and possible applications of such smart hydrogels.
 
@@
 We synthesise the phosphonate terminated polymer hereafter referred to as \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
-We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; see Supplementary Materials and Supplementary Figure~S\ref{S-fig:sec}), of polymerisation index $N_0=70$, which is the common basis to all studied polymers.
-We achieve all subsequent transformations of the polymer structures using the same initial batch of material.
+We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; see Supplementary Materials and Supplementary Figure~S\ref{S-fig:sec}), with a degree of polymerisation $N_0=70$.
+This polymer serves as a common scaffold from which we derive  all studied systems.
 Nucleophilic addition of N-methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF affords the corresponding poly(imidazolium) and poly(pyrrolidinium) compounds.
@@
 
-The above dependence of gel formability on the nature of the counterions could be rationalized on the basis of ion-pairing processes.
+The above dependence of gel formability on the nature of the counterions could be rationalized on the basis of ion-pairing efficiency.
 According to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected between a ``softer'' cationic moiety and a ``harder'' halide counterions\cite{Goossens2009}.
+Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge.
 Therefore charge dissociation in water and thus solubility of the polymer should be easier with larger halide and pyrrolidinium.
-Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge.
-That is why we are able to dissolve the five polymers that have the weakest ion pairing, but neither the two fluorides nor \ce{PIm+Cl-} where counterions are expected to be strongly bound to the chain.
+As a consequence, while dissolution is observed for the four polymers with bromide and iodide counterions and for \ce{PPyr+Cl-}, in which ion dissociation takes place to a sufficient extent, neither \ce{PIm+Cl-} nor the two polymers with fluoride counterions, for which ion pairs are expected to be strongly bound, can be dissolved.
 
-A fortiori, the insolubility in absence of charge dissociation indicates that the chain is in poor solvent.
-Locally, the monomers are condensed in collapsed globules to minimize the contact with solvent, and solubility on larger scales can only be achieved if the polymer bears enough charges\cite{Khokhlov1980,Raphael1990}.
-To estimate the $\Theta$-temperature, we held the two fluoride polymers in water at boiling temperature overnight without any dissolution, implying that $\Theta>\SI{100}{\celsius}$.
-Nonetheless, due to charge dissociation, five of our polyelectrolytes are soluble at \SI{25}{\celsius}.
+Such an insolubility in absence of charge dissociation indicates that the chain is in poor solvent.
+Locally, the monomers are condensed in collapsed globules to minimize the contacts with solvent, and solubility on larger scales can only be achieved if the polymer bears enough charges\cite{Khokhlov1980,Raphael1990}.
+To estimate the $\Theta$-temperature, that delimit bad and good solvent behaviour, we held the two fluoride polymers in water at boiling temperature overnight.
+No dissolution or swelling was observed, implying that $\Theta>\SI{100}{\celsius}$.
+%Nonetheless, due to charge dissociation, five of our polyelectrolytes are soluble at \SI{25}{\celsius}.
 
@@
 Comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ value (13 and \SI{8100}{\pascal}, resp.) and $\gamma_c$ value (804\% and 1.6\%, resp.) differ by almost three orders of magnitude.
-Here also, gel hardness is correlated to efficient charge dissociation, this time via ``harder'' cationic moieties.
-In the following we explain quantitatively this correlation from a microscopic model.
+Here also, gel toughness is related to charge dissociation efficiency: the hard pyrrolidinium cation, with its positive charge localized on the nitrogen atom, interacts more weakly with soft halide counterions than its soft imidazolium counterpart (where the positive charge is distributed over the whole aromatic imidazolium cycle).
+In the following we show that this correlation can be explained on the basis of a microscopic model of interchain interactions.
 
-\subsection{Crocodile line model}
+\subsection{Processionary model}
 
@@
 \includegraphics{dissociation.pdf}
-\caption{Crocodile line model.
-(a) Sketch of the network in the case of a crocodile line of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{src}\approx 3D$.
+\caption{Processionary model.
+(a) Sketch of the network in the case of a procession of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{src}\approx 3D$.
 Empty circles are not individual monomers but electrostatic blobs of size $D$.
@@
 When the probability of such a configuration is non-zero, we obtain a physically cross-linked gel.
+This is the ideal situation that we described in our previous paper.
 
-On the other hand, if every chain has a single foreign head attached to its body, every polymer is linked to two others in a crocodile line\bibnote{We neglect the possibility of the two charges of the phosphonate head to attach to two separate chains}.
-In addition, if we suppose that in order to minimize the inter-chain repulsion between charged cationic groups, heads are preferentially attached to the tail of their neighbour, we obtain a linear chain of effective polymerisation index $n N_0$ where $n$ is the number of polymers in the crocodile line.
+On the other hand, if every chain has a single foreign head attached to its body, every polymer is linked to two others in a single file\bibnote{We neglect the possibility of the two charges of the phosphonate head to attach to two separate chains}.
+In addition, if we suppose that in order to minimize the inter-chain repulsion between charged cationic groups, heads are preferentially attached to the tail of their neighbour, we obtain a linear chain of effective polymerisation index $n N_0$ where $n$ is the number of polymers in the supramolecular queue.
+This behaviour bears a striking resemblance to pine processionary caterpillar behaviour first described by \citet{Fabre1916}:
+``They proceed in single file, in a continuous row, each touching with its head the rear of the one in front of it.
+[...] No Greek theoria winding its way to the Eleusinian festivals was ever more orderly.
+Hence the name of Processionary given to the gnawer of the pine.''
 
-We thus have two limiting cases: (i) every chain has at least two heads attached and we have roughly $N_0$ monomers between cross-link points; (ii) every chain has a single head attached and we have a single supramolecular chain in the system.
-The former case should be observed at high charge dissociation when many moieties are ionised along the chain providing many sites for a phosphonate head to bound to.
+We thus have two limiting cases: (i) every chain has at least two heads attached and we have roughly $N_0$ monomers between cross-link points; (ii) every chain has at the most a single head attached and we have isolated supramolecular chains in the system.
+The former case should be observed when a significant number of ion-pairs along the polymer chain are dissociated providing many sites for a phosphonate head to bound to.
 The latter case should be observed at negligible charge dissociation.
-In between these two limiting cases, we should observe cross-link points separated by crocodile lines of $n N_0$ monomers as sketched in Figure~\ref{fig:dissociation}a.
+In between these two limiting cases, we should observe cross-link points separated by processions of $n N_0$ monomers as sketched in Figure~\ref{fig:dissociation}a.
 
-Let us now discuss the rheological measurements and link them to the hydrogel microstructure.
-We reduce our measurements to two main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
+As we will discuss in the following, all rheological features of the hydrogels studied here can  be rationalised on the basis of this ``processionary'' model.
+We base our analysis on two main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
 the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, that we consider as representative of the end of the linear regime and of the first broken cross-links\cite{Hyun2011}.
+By using this two parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters as the average number of chains between cross links $n$, the head-to-body bonding energy $E_c$ or the number of monomers between dissociated ion pairs $A$.
 
@@
 
-Between two cross-links the crocodile chain can be considered as an entropic spring made of $N=n N_0$ monomers. The shear modulus of the gel is then given by\cite{Rubinstein1996}:
+Between two cross-links the ``procession'' of polymer chains can be considered as an entropic spring made of $N=n N_0$ monomers. The shear modulus of the gel is thus given by\cite{Rubinstein1996}:
 \begin{equation}
@@
 where $c$ is the monomer number density, $k_\mathrm{B}$ is the Boltzmann constant and $T$ the temperature.
-Because of the different weights of the pyrrolidinium and imidazolium moieties and of the different counter-ions, $c$ is not constant across our experiments done at constant weight fraction $w$ of polymer.
+Because of the different weights of the pyrrolidinium and imidazolium moieties and of the different counter-ions, $c$ is not constant as we chose to conduct our experiments at constant weight fraction $w$ of polymer (generally used to quantify gelation ability of a given gelator).
 This explains why the respective modulus or critical strains of \ce{PPyr+I-} (heavier, lower number density) and \ce{PPyr+Br-} are in reverse order with respect to the \textit{a priori} ranking of charge dissociation.
@@
 We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 790 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
-This is consistent with a higher probability of attaching two or more heads on a highly dissociated body, i.e.
-a large cross-link ratio.
+This larger cross-link ratio is consistent with a higher probability of attaching two or more heads on a highly dissociated body.
 
@@
 
-Considering that each crocodile chain is an entropic spring in its linear domain, the mechanical energy stored by a single crocodile chain under a strain $\gamma$ is $E = \gamma^2 k_\mathrm{B}T$.
-If we suppose that ionic bonds break before the complete extension of the crocodile chains, we evaluate the head-to-body bonding energy to
+Considering that between two cross-links a procession is an entropic spring in its linear domain, the mechanical energy stored under a strain $\gamma$ is $E = \gamma^2 k_\mathrm{B}T$.
+If we suppose that ionic bonds break before the complete extension of the procession, we evaluate the head-to-body bonding energy to
 \begin{equation}
@@
 This gives rise to maximal charge dissociation in water, facilitates charge pairing of the peripheral, iterative cations with the terminal anion and thus affords a very high density of cross-links.
-In between, in the case of ``hard'' cations and not-so-hard halide (\ce{Cl-}) or in the ``soft-soft'' case, charge dissociation is limited, polymers associate in crocodile lines and cross-links are few.
+In between, in the case of ``hard'' cations and not-so-hard halide (\ce{Cl-}) or in the ``soft-soft'' case, charge dissociation is limited, polymers associate in processions and cross-links are few.
 We thus obtain soft, highly deformable gels.
 
-To conclude, our procedure yields robust, highly tunable hydrogels from short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials and paves the way for future works on supra-molecular assembly.
+To conclude, our procedure yields robust, highly tunable hydrogels from short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials.
+Most importantly, systematic comparisons between the different poly(electrolytes) investigated in this study were used to establish a model linking the microsctructure of the gel to ion pair dissociation efficiency on the individual polymer chains.
+We hope that this model will join the procession toward future works in supramolecular assemblies.
 

### 2016-02-08 PALSE/ion_pairing_paper.tex
Graphical TOC processionaires, attributions.

@@
 
-9\,cm by 3.5\,cm
+%9\,cm by 3.5\,cm
+\includegraphics{processonaire.jpg}
 
@@
 
-In the present report, we investigate the role of the dissociation rate of the poly(cationic) body on the mechanical properties of the hydrogel.
-In a first part we describe how we take advantage of our well-controlled synthesis to play systematically with the nature (hardness/softness) of the cationic repeating unit and the ion-paring strength of the associated counter-ion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}).
-In a second part, we show the qualitative and quantitative variations of mechanical properties of the aqueous dispersions obtained from these well-characterized polymers.
-In a third part, we develop a microscopic model based on the idea that at low dissociation rates, cross-links are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular crocodile line.
-%Finally, we discuss our results and possible applications of such smart hydrogels.
+In the present report, we investigate the role of the ion pair dissociation equilibria rate of the poly(cationic) body on the mechanical properties of the hydrogel.
+In a first part we describe how we take advantage of our postfunctionalization approach to play systematically with the nature (hardness/softness) of the cationic repeating unit and the ion-paring strength of the associated counter-ion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}).
+In the second part and third part, we show respectively the qualitative change in gel formability and the quantitative variations of mechanical properties of the aqueous dispersions obtained from these well-characterized polymers.
+In a fourth part, we develop a microscopic model based on the idea that at low dissociation rates, cross-links (defined as a polymer that is connected to three or more others) are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular chain.
+We call this behaviour ``processionary'' in analogy to pine processionary caterpillar (\textit{Thaumetopoea pityocampa}) behaviour\cite{Fabre1916}.%Finally, we discuss our results and possible applications of such smart hydrogels.
 
@@
 
-We synthesise the phosphonate terminated polymer hereafter reffered to as \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
-We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; see Supplementary Materials S2, S3, Figure S1), of polymerisation index $N_0=70$, which is the common basis to all studied polymers.
-We achieve all subsequent transformations of the polymer structures using the same initial batch of material.
+We synthesise the phosphonate terminated polymer hereafter referred to as \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
+We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; see Supplementary Materials and Supplementary Figure~S\ref{S-fig:sec}), with a degree of polymerisation $N_0=70$.
+This polymer serves as a common scaffold from which we derive  all studied systems.
 Nucleophilic addition of N-methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF affords the corresponding poly(imidazolium) and poly(pyrrolidinium) compounds.
@@
 After extensive dialysis of the resulting mixture against deionized water and lyophilization we recover in high yields the different \ce{PIm+X-} and \ce{PPyr+X-} (where \ce{X}=\ce{F}, \ce{Cl}, \ce{I}).
-In order to bring evidence for the efficiency of the ionic metathesis, we submit the resulting materials to negative-mode high-resolution mass spectrometry (HRMS, see Supplementary Materials S3).
-We unambiguously assess displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Figures S8-9).
-We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures S6-7), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (see supporting information XXX S?
-XXX).
+In order to bring evidence for the efficiency of the ionic metathesis, we submit the resulting materials to negative-mode high-resolution mass spectrometry (HRMS, see Supplementary Materials).
+We unambiguously assess displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Supplementary Figures~S\ref{S-fig:massPImBr} and S\ref{S-fig:massPPyBr}).
+We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures~S\ref{S-fig:massPImI} and S\ref{S-fig:massPPyI}), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (not shown).
 
@@
 acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007–2013)/ERC grant agreement No.258803.
+The Graphical TOC entry is derived from a picture by Jean-Pol Grandmont - Own work, CC BY 3.0, \url{https://commons.wikimedia.org/w/index.php?curid=4515418}.
 \end{acknowledgement}

### 2016-02-15 PALSE/ion_pairing_paper.tex
HSAB -> Hofmeister & pi+-pi+

@@
 %9\,cm by 3.5\,cm
-\includegraphics{processonaire.jpg}
+\includegraphics{TOC2.png}
 
@@
 We demonstrated recently that polyelectrolytes with cationic moieties along the chain and a single anionic head are able to form physical hydrogels due to the reversible nature of the head-to-body ionic bond.
-Here we generate a variety of such polyelectrolytes with various cationic moieties and counter-ions combinations starting from a common polymeric platform.
+Here we generate a variety of such polyelectrolytes with various cationic moieties and counterions combinations starting from a common polymeric platform.
 We show that the rheological properties (shear modulus, critical strain) of the final hydrogels can be modulated over three orders of magnitude depending on the cation/anion pair.
@@
 
+Contrary to a small electrolyte that exhibit only dissociated ions, polyelectrolytes in solution are in general not fully dissociated, with a substantial fraction of the counterions bound to the polymer\cite{Manning1979}.
+The resulting net charge of the chain governs the physical properties of the polyelectrolytes solution\cite{Rubinstein1996}, first of all the ability of the polymer to dissolve.
+counterion condensation is the physical bounding or adsorption of counterions near the polymer chain.
+Factors influencing this process have been known since the 1880's but are still being rationalized.
+\citet{Hofmeister1888} was the first to propose a systematic ranking of ions, based on their propensity to promote the coalescence of egg white.
+This so-called Hofmeister series has since proved quite universally valid\cite{Zhang2010a}, including multi-charged polymers such as naturally occurring proteins or synthetic polyelectrolytes as well as charged colloidal particles\cite{Schwierz2010,Oncsik2015} or soft matter interfaces in general\cite{Leontidis2014}.
+However, depending on the nature of the polymer (hydrophilic or hydrophobic, anionic or cationic), the Hofmeister series can be direct (in short, harder ions promote condensation) or reverse\cite{Schwierz2010,Oncsik2015}.
+
 We have recently reported on a highly reversible poly(cation) based hydrogel\cite{Srour2014}, which formation relies on an new concept of supramolecular  electrostatic interaction.
@@
 In the present report, we investigate the role of the ion pair dissociation equilibria rate of the poly(cationic) body on the mechanical properties of the hydrogel.
-In a first part we describe how we take advantage of our postfunctionalization approach to play systematically with the nature (hardness/softness) of the cationic repeating unit and the ion-paring strength of the associated counter-ion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}).
+In a first part we describe how we take advantage of our postfunctionalization approach to play systematically with the nature (aromatic or not) of the cationic repeating unit and the associated counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}) varied along the Hofmeister series.
 In the second part and third part, we show respectively the qualitative change in gel formability and the quantitative variations of mechanical properties of the aqueous dispersions obtained from these well-characterized polymers.
@@
 
-In ionic solids, melting points are inherently connected to the reticular enthalpy of the crystal lattice, hence primarily to ion pairing strength\cite{Sherman1932}.
-This last parameter is directly related to the diffuse and hindered nature of the interacting charges, bulkier and softer ions providing in the most general case weaker interactions\cite{Krossing2006}.
-For typical ``ionic liquids'' such as imidazolium or pyrrolidinium halides, melting points will thus gradually increase as the size of the halide counterion decreases, ie $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$\cite{Dean2010}.
-Although analysis is less straightforward in the case of amorphous polymer electrolytes, ionic substitution is however expected to have an impact on the phase transition temperatures of the material\cite{Orler1994}.
-
-In order to characterize the effect of counterion substitution, we study all polymers by Differential Scanning Calorimetry (DSC).
-Although we observe no glass transition in the studied temperature range for any of the sample, comparison of the DSC data reveal significant differences.
-
-In the case polyelectrolytes with fluoride and chloride counterion, we observe no exo- or endothermic transition below \SI{200}{\celsius}.
+Differential Scanning Calorimetry (DSC) thermograms  of all synthesized polymers also reveals marked differences in their properties, that can only be explained by differences in their couterion features.
+Thus, in the case polyelectrolytes with fluoride and chloride counterion, we observe no exo or endothermic transition below \SI{200}{\celsius}.
 By contrast, we observe a broad endothermic peak when iodide and bromide are used as counterions.
@@
 It is generally attributed to desorption of weakly bound water from the polymer network.
-Its presence in the polymer samples where bromide and iodide are used as couterions suggests that adsorbed water could play a role in solvating the dissociated ion pairs in the latter.
+%Its presence in the polymer samples where bromide and iodide are used as couterions suggests that adsorbed water could play a role in solvating the dissociated ion pairs in the latter.
 Conversely, its absence for chloride and fluoride counterions suggests that dissociation is not effective because of stronger ion pair interaction.
+
+This difference in DSC profile can be understood by considering the difference in solvation of the different polymers in this study, as a function of the nature of the associated anion.
+A small anion like \ce{F-} can polarize a water molecule that is directly involved in hydrogen bonding with the ester functions displayed by our polymers\cite{Zhang2010a}.
+This water binding ability will decrease along the direct Hofmeister series $\ce{F-}>\ce{Cl-}>\ce{Br-}>\ce{I-}$.
+We are able to observe the end of this trend in the shift of peak position between bromide and iodide counterions, whereas in the case of chloride and fluoride, binding is too strong to observe water desorption.
+This hypothesis is further confirmed below, when looking at the properties of different polymers in aqueous dispersion.
 
@@
 
-\subsection{Water swelling, ion-pairing and solvent quality}
+\subsection{Water swelling, counterion condensation and solvent quality}
 Water-swelling properties turn out to be extremely dependent on the polymer composition.
@@
 
-The above dependence of gel formability on the nature of the counterions could be rationalized on the basis of ion-pairing efficiency.
-According to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected between a ``softer'' cationic moiety and a ``harder'' halide counterions\cite{Goossens2009}.
-Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge.
-Therefore charge dissociation in water and thus solubility of the polymer should be easier with larger halide and pyrrolidinium.
-As a consequence, while dissolution is observed for the four polymers with bromide and iodide counterions and for \ce{PPyr+Cl-}, in which ion dissociation takes place to a sufficient extent, neither \ce{PIm+Cl-} nor the two polymers with fluoride counterions, for which ion pairs are expected to be strongly bound, can be dissolved.
+The above dependence of solubility on the nature of the counterions follows the direct Hofmeister series.
+The strong condensation of \ce{F-} on the chain decreases the effective positive charge of the polymer and thus reduces solubility.
+Conversely, larger, poorly hydrated ions like \ce{I-} are less adsorbed, resulting in more dissociated charges and higher solubility.
+As a consequence, while dissolution is observed for the four polymers with bromide and iodide counterions, in which ion dissociation takes place to a sufficient extent, the two polymers with fluoride counterions, for which couterion condensation is expected to be strong, cannot be dissolved.
+
+However Hofmeister series alone cannot explain the difference between soluble \ce{PPyr+Cl-} and insoluble \ce{PIm+Cl-}.
+We have to suppose that counterions are more strongly condensed near imidazolium than near pyrrolidinium.
+Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties.
+Because of their aromatic ring, imidazolium are particularly prone to promote a variety of supramolecular interactions which strongly contribute to their physicochemical properties, such as \ce{\pi+-\pi} or anion$-\pi^+$.
+More recently, \ce{\pi+-\pi+} interactions have also been identified as a distinctive driving force for imidazolium dimerization\cite{Geronimo2011}.
+Although primarily weaker than their \ce{\pi-\pi} counterparts because of strong intercoulombic repulsion, it has been established that, in the presence of couterion, the stabilizing effect of \ce{\pi+-\pi+} interaction could reach magnitudes largely exceeding that of the latter\cite{Geronimo2011}.
+In particular, dimerization of imidazolium-chloride ion pairs have been the object of recent studies, and it has been shown that the involvement of the negatively charged counterion was essential in maximizing the stabilization of the \ce{\pi+-\pi+} complex by minimizing coulombic repulsion between the imidazolium moieties\cite{Matthews2014,Gao2015a}.
+It is likewise well documented that covalent incorporation of interacting group within a polymer chain strongly favours intramolecular interactions between these groups, by increasing their so-called effective molarity\cite{Li2003,Mulder2004,Huerta2013}.It is therefore not surprising that both effects participate in making the imidazolium-anion interaction stronger than pyrrolidinium at the detriment of solubility.
 
@@
 By contrast, the gel with chloride counterions is weaker by more than one order of magnitude.
-In other words, a ``soft'' halide counterion, efficiently dissociated, is correlated to a harder gel.
+In other words, counterion condensation is correlated to softer gel.
 
@@
 Comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ value (13 and \SI{8100}{\pascal}, resp.) and $\gamma_c$ value (804\% and 1.6\%, resp.) differ by almost three orders of magnitude.
-Here also, gel toughness is related to charge dissociation efficiency: the hard pyrrolidinium cation, with its positive charge localized on the nitrogen atom, interacts more weakly with soft halide counterions than its soft imidazolium counterpart (where the positive charge is distributed over the whole aromatic imidazolium cycle).
+Here also, gel toughness is related to charge dissociation efficiency.
 In the following we show that this correlation can be explained on the basis of a microscopic model of interchain interactions.
@@
 where $c$ is the monomer number density, $k_\mathrm{B}$ is the Boltzmann constant and $T$ the temperature.
-Because of the different weights of the pyrrolidinium and imidazolium moieties and of the different counter-ions, $c$ is not constant as we chose to conduct our experiments at constant weight fraction $w$ of polymer (generally used to quantify gelation ability of a given gelator).
-This explains why the respective modulus or critical strains of \ce{PPyr+I-} (heavier, lower number density) and \ce{PPyr+Br-} are in reverse order with respect to the \textit{a priori} ranking of charge dissociation.
+Because of the different weights of the pyrrolidinium and imidazolium moieties and of the different counterions, $c$ is not constant as we chose to conduct our experiments at constant weight fraction $w$ of polymer (generally used to quantify gelation ability of a given gelator).
+This explains why the respective modulus or critical strains of \ce{PPyr+I-} (heavier, lower number density) and \ce{PPyr+Br-} are in reverse order with respect to the Hofmeister series.
 Taking into account the molecular mass $M$ of each polymer and  knowing the density $d$ of the solvent, we obtain
@@
 
-When the halide counterion is ``hard'', there is not enough charge dissociation to allow dissolution and gel formation.
-On the other hand of the spectrum outlined by HSAB theory, the hardest gels with the narrowest linear domain are obtained when a combination of ``hard'' cations and ``soft'' halides are used.
+Well hydrated anions and aromatic cation favour counterion condensation, resulting in too few charges to allow dissolution and gel formation.
+The hardest gels with the narrowest linear domain are obtained with aliphatic cationic moiety and poorly hydrated counterion at the other end of the Hofmeister series.
 This gives rise to maximal charge dissociation in water, facilitates charge pairing of the peripheral, iterative cations with the terminal anion and thus affords a very high density of cross-links.
-In between, in the case of ``hard'' cations and not-so-hard halide (\ce{Cl-}) or in the ``soft-soft'' case, charge dissociation is limited, polymers associate in processions and cross-links are few.
+In between, in the case of aliphatic cations with a counterion in the middle of the Hofmeister series (\ce{Cl-}) or in the case of aromatic cation and poorly hydrated anion, charge dissociation is limited, polymers associate in processions and cross-links are few.
 We thus obtain soft, highly deformable gels.
@@
 acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007–2013)/ERC grant agreement No.258803.
-The Graphical TOC entry is derived from a picture by Jean-Pol Grandmont - Own work, CC BY 3.0, \url{https://commons.wikimedia.org/w/index.php?curid=4515418}.
+The Graphical TOC entry is derived from a picture by Arturo Reina, CC BY-SA 3.0, \url{https://commons.wikimedia.org/w/index.php?curid=282898}.
 \end{acknowledgement}

### 2016-02-15 ion_pairing_paper.tex
harder -> well hydrated. Reorganize the pi-pi vs pi+-pi+ sentence.

@@
 
-9\,cm by 3.5\,cm
+%9\,cm by 3.5\,cm
+\includegraphics{TOC2.png}
 
@@
 
+Contrary to a small electrolyte that exhibit only dissociated ions, polyelectrolytes in solution are in general not fully dissociated, with a substantial fraction of the counterions bound to the polymer\cite{Manning1979}.
+The resulting net charge of the chain governs the physical properties of the polyelectrolytes solution\cite{Rubinstein1996}, first of all the ability of the polymer to dissolve.
+counterion condensation is the physical bounding or adsorption of counterions near the polymer chain.
+Factors influencing this process have been known since the 1880's but are still being rationalized.
+\citet{Hofmeister1888} was the first to propose a systematic ranking of ions, based on their propensity to promote the coalescence of egg white.
+This so-called Hofmeister series has since proved quite universally valid\cite{Zhang2010a}, including multi-charged polymers such as naturally occurring proteins or synthetic polyelectrolytes as well as charged colloidal particles\cite{Schwierz2010,Oncsik2015} or soft matter interfaces in general\cite{Leontidis2014}.
+However, depending on the nature of the polymer (hydrophilic or hydrophobic, anionic or cationic), the Hofmeister series can be direct (in short, well hydrated ions promote condensation) or reverse\cite{Schwierz2010,Oncsik2015}.
+
 We have recently reported on a highly reversible poly(cation) based hydrogel\cite{Srour2014}, which formation relies on an new concept of supramolecular  electrostatic interaction.
@@
 In the present report, we investigate the role of the ion pair dissociation equilibria rate of the poly(cationic) body on the mechanical properties of the hydrogel.
-In a first part we describe how we take advantage of our postfunctionalization approach to play systematically with the nature (hardness/softness) of the cationic repeating unit and the ion-paring strength of the associated counter-ion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}).
-In a second part, we show the qualitative and quantitative variations of mechanical properties of the aqueous dispersions obtained from these well-characterized polymers.
-In a third part, we develop a microscopic model based on the idea that at low dissociation rates, cross-links (defined as a polymer that is connected to three or more others) are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular chain.
+In a first part we describe how we take advantage of our postfunctionalization approach to play systematically with the nature (aromatic or not) of the cationic repeating unit and the associated counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}) varied along the Hofmeister series.
+In the second part and third part, we show respectively the qualitative change in gel formability and the quantitative variations of mechanical properties of the aqueous dispersions obtained from these well-characterized polymers.
+In a fourth part, we develop a microscopic model based on the idea that at low dissociation rates, cross-links (defined as a polymer that is connected to three or more others) are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular chain.
 We call this behaviour ``processionary'' in analogy to pine processionary caterpillar (\textit{Thaumetopoea pityocampa}) behaviour\cite{Fabre1916}.%Finally, we discuss our results and possible applications of such smart hydrogels.
@@
 
-The above dependence of gel formability on the nature of the counterions could be rationalized on the basis of ion-pairing efficiency.
-According to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected between a ``softer'' cationic moiety and a ``harder'' halide counterions\cite{Goossens2009}.
-Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge.
-Therefore charge dissociation in water and thus solubility of the polymer should be easier with larger halide and pyrrolidinium.
-As a consequence, while dissolution is observed for the four polymers with bromide and iodide counterions and for \ce{PPyr+Cl-}, in which ion dissociation takes place to a sufficient extent, neither \ce{PIm+Cl-} nor the two polymers with fluoride counterions, for which ion pairs are expected to be strongly bound, can be dissolved.
+The above dependence of solubility on the nature of the counterions follows the direct Hofmeister series.
+The strong condensation of \ce{F-} on the chain decreases the effective positive charge of the polymer and thus reduces solubility.
+Conversely, larger, poorly hydrated ions like \ce{I-} are less adsorbed, resulting in more dissociated charges and higher solubility.
+As a consequence, while dissolution is observed for the four polymers with bromide and iodide counterions, in which ion dissociation takes place to a sufficient extent, the two polymers with fluoride counterions, for which couterion condensation is expected to be strong, cannot be dissolved.
+
+However Hofmeister series alone cannot explain the difference between soluble \ce{PPyr+Cl-} and insoluble \ce{PIm+Cl-}.
+We have to suppose that counterions are more strongly condensed near imidazolium than near pyrrolidinium.
+Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties.
+Because of their aromatic ring, imidazolium are particularly prone to promote a variety of supramolecular interactions which strongly contribute to their physicochemical properties, such as \ce{\pi+-\pi} or anion$-\pi^+$.
+More recently, \ce{\pi+-\pi+} interactions have also been identified as a distinctive driving force for imidazolium dimerization\cite{Geronimo2011}.
+Because of strong intercoulombic repulsion, \ce{\pi+-\pi+} are primarily weaker than their \ce{\pi-\pi} counterparts.
+However it has been established that, in the presence of couterion, the stabilizing effect of \ce{\pi+-\pi+} interaction could reach magnitudes largely exceeding that of the latter\cite{Geronimo2011}.
+In particular, dimerization of imidazolium-chloride ion pairs have been the object of recent studies, and it has been shown that the involvement of the negatively charged counterion was essential in maximizing the stabilization of the \ce{\pi+-\pi+} complex by minimizing coulombic repulsion between the imidazolium moieties\cite{Matthews2014,Gao2015a}.
+It is likewise well documented that covalent incorporation of interacting group within a polymer chain strongly favours intramolecular interactions between these groups, by increasing their so-called effective molarity\cite{Li2003,Mulder2004,Huerta2013}.
+It is therefore not surprising that both effects participate in making the imidazolium-anion interaction stronger than pyrrolidinium at the detriment of solubility.
 
 Such an insolubility in absence of charge dissociation indicates that the chain is in poor solvent.
-Locally, the monomers are condensed in collapsed globules to minimize the contacts with solvent, and solubility on larger scales can only be achieved if the polymer bears enough charges\cite{Khokhlov1980,Raphael1990}.
+Locally, the monomers are condensed in collapsed globules to minimize the contacts with water molecules, and solubility on larger scales can only be achieved if the polymer bears enough charges\cite{Khokhlov1980,Raphael1990}.
 To estimate the $\Theta$-temperature, that delimit bad and good solvent behaviour, we held the two fluoride polymers in water at boiling temperature overnight.
@@
 acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007–2013)/ERC grant agreement No.258803.
+The Graphical TOC entry is derived from a picture by Arturo Reina, CC BY-SA 3.0, \url{https://commons.wikimedia.org/w/index.php?curid=282898}.
 \end{acknowledgement}

### 2016-02-16 PALSE/ion_pairing_paper.tex
Corrections Cyrille. Why focus on ester. No "hydratation" but association with the polymer. Remove "Having 2 charges dissociated on a chain implies a cross-link." and replace by 1:1 association model.

@@
 
-Contrary to a small electrolyte that exhibit only dissociated ions, polyelectrolytes in solution are in general not fully dissociated, with a substantial fraction of the counterions bound to the polymer\cite{Manning1979}.
+Contrary to a small electrolyte for which full ion pairs dissociation is generally achieved in water, polyelectrolytes in solution are in general not fully dissociated, with a substantial fraction of the counterions bound to the polymer\cite{Manning1979}.
 The resulting net charge of the chain governs the physical properties of the polyelectrolytes solution\cite{Rubinstein1996}, first of all the ability of the polymer to dissolve.
 counterion condensation is the physical bounding or adsorption of counterions near the polymer chain.
-Factors influencing this process have been known since the 1880's but are still being rationalized.
+Factors influencing this process have been known since the 1880's but are still not fully understood.
 \citet{Hofmeister1888} was the first to propose a systematic ranking of ions, based on their propensity to promote the coalescence of egg white.
 This so-called Hofmeister series has since proved quite universally valid\cite{Zhang2010a}, including multi-charged polymers such as naturally occurring proteins or synthetic polyelectrolytes as well as charged colloidal particles\cite{Schwierz2010,Oncsik2015} or soft matter interfaces in general\cite{Leontidis2014}.
-However, depending on the nature of the polymer (hydrophilic or hydrophobic, anionic or cationic), the Hofmeister series can be direct (in short, harder ions promote condensation) or reverse\cite{Schwierz2010,Oncsik2015}.
+However, depending on the nature of the polymer (hydrophilic or hydrophobic, anionic or cationic), the Hofmeister series can be direct (in short, well hydrated ions promote condensation) or reverse\cite{Schwierz2010,Oncsik2015}.
 
@@
 In the present report, we investigate the role of the ion pair dissociation equilibria rate of the poly(cationic) body on the mechanical properties of the hydrogel.
-In a first part we describe how we take advantage of our postfunctionalization approach to play systematically with the nature (aromatic or not) of the cationic repeating unit and the associated counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}) varied along the Hofmeister series.
+In a first part we describe how we take advantage of our postfunctionalization approach to play systematically with the nature (aromatic or not) of the cationic repeating unit and the associated halide counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}) varied along the Hofmeister series.
 In the second part and third part, we show respectively the qualitative change in gel formability and the quantitative variations of mechanical properties of the aqueous dispersions obtained from these well-characterized polymers.
@@
 
-This difference in DSC profile can be understood by considering the difference in solvation of the different polymers in this study, as a function of the nature of the associated anion.
-A small anion like \ce{F-} can polarize a water molecule that is directly involved in hydrogen bonding with the ester functions displayed by our polymers\cite{Zhang2010a}.
-This water binding ability will decrease along the direct Hofmeister series $\ce{F-}>\ce{Cl-}>\ce{Br-}>\ce{I-}$.
+This difference in DSC profile can be understood by considering the water-binding ability of the different polymers in this study.
+Since the peak temperature does not depend on the presence of imidazolium or pyrrolidinium, water-binding is probably due to the only part of the size groups able to accept hydrogen bonds from water: the ester function.
+Following \citet{Zhang2010a}, we propose that these bonds can be enhanced if the involved water molecule is polarized by solvating an anion.
+The polarization, and thus the water-binding ability will decrease along the direct Hofmeister series $\ce{F-}>\ce{Cl-}>\ce{Br-}>\ce{I-}$.
 We are able to observe the end of this trend in the shift of peak position between bromide and iodide counterions, whereas in the case of chloride and fluoride, binding is too strong to observe water desorption.
@@
 The strong condensation of \ce{F-} on the chain decreases the effective positive charge of the polymer and thus reduces solubility.
-Conversely, larger, poorly hydrated ions like \ce{I-} are less adsorbed, resulting in more dissociated charges and higher solubility.
+Conversely, bulkier and softer ions like \ce{I-} are less adsorbed, resulting in more dissociated charges and higher solubility.
 As a consequence, while dissolution is observed for the four polymers with bromide and iodide counterions, in which ion dissociation takes place to a sufficient extent, the two polymers with fluoride counterions, for which couterion condensation is expected to be strong, cannot be dissolved.
@@
 More recently, \ce{\pi+-\pi+} interactions have also been identified as a distinctive driving force for imidazolium dimerization\cite{Geronimo2011}.
-Although primarily weaker than their \ce{\pi-\pi} counterparts because of strong intercoulombic repulsion, it has been established that, in the presence of couterion, the stabilizing effect of \ce{\pi+-\pi+} interaction could reach magnitudes largely exceeding that of the latter\cite{Geronimo2011}.
+Because of strong intercoulombic repulsion, \ce{\pi+-\pi+} are primarily weaker than their \ce{\pi-\pi} counterparts.
+However it has been established that, in the presence of couterion, the stabilizing effect of \ce{\pi+-\pi+} interaction could reach magnitudes largely exceeding that of the latter\cite{Geronimo2011}.
 In particular, dimerization of imidazolium-chloride ion pairs have been the object of recent studies, and it has been shown that the involvement of the negatively charged counterion was essential in maximizing the stabilization of the \ce{\pi+-\pi+} complex by minimizing coulombic repulsion between the imidazolium moieties\cite{Matthews2014,Gao2015a}.
-It is likewise well documented that covalent incorporation of interacting group within a polymer chain strongly favours intramolecular interactions between these groups, by increasing their so-called effective molarity\cite{Li2003,Mulder2004,Huerta2013}.It is therefore not surprising that both effects participate in making the imidazolium-anion interaction stronger than pyrrolidinium at the detriment of solubility.
+It is likewise well documented that covalent incorporation of interacting group within a polymer chain strongly favours intramolecular interactions between these groups, by increasing their so-called effective molarity\cite{Li2003,Mulder2004,Huerta2013}.
+It is therefore not surprising that both effects participate in making the imidazolium-anion interaction stronger than pyrrolidinium at the detriment of solubility.
 
 Such an insolubility in absence of charge dissociation indicates that the chain is in poor solvent.
-Locally, the monomers are condensed in collapsed globules to minimize the contacts with solvent, and solubility on larger scales can only be achieved if the polymer bears enough charges\cite{Khokhlov1980,Raphael1990}.
+Locally, the monomers are condensed in collapsed globules to minimize the contacts with water molecules, and solubility on larger scales can only be achieved if the polymer bears enough charges\cite{Khokhlov1980,Raphael1990}.
 To estimate the $\Theta$-temperature, that delimit bad and good solvent behaviour, we held the two fluoride polymers in water at boiling temperature overnight.
@@
 We thus have two limiting cases: (i) every chain has at least two heads attached and we have roughly $N_0$ monomers between cross-link points; (ii) every chain has at the most a single head attached and we have isolated supramolecular chains in the system.
-The former case should be observed when a significant number of ion-pairs along the polymer chain are dissociated providing many sites for a phosphonate head to bound to.
-The latter case should be observed at negligible charge dissociation.
+The former case should be observed when a significant number of ion-pairs along the polymer chain are dissociated providing a significant probability for multiple phosphonate heads binding on a single polymer chain.
+Conversely, the latter case should be observed at low charge dissociation.
 In between these two limiting cases, we should observe cross-link points separated by processions of $n N_0$ monomers as sketched in Figure~\ref{fig:dissociation}a.
@@
 Finally we cross-check our model by linking the ion pair dissociation and the amount of physical cross-links by an independent argument.
+A second charge dissociated on a chain creates a possible binding site for a second head-to-body association and thus a cross-link.
+In a 1:1 association model at low association rate, the number of association is linearly proportional to the number of available sites.
+Therefore the probability to have a cross link is proportional to the probability of having a second charge dissociation.
 If we assume that the probability for an ion pair to dissociate is independent of the number of dissociated pairs already present on the polymer, we can describe the probability of having $k$ dissociated pairs on a polymer by a Poisson distribution of mean $\lambda = N_0/A$.
-Having 2 charges dissociated on a chain implies a cross-link.
 Therefore, the cross-link probability $1/n$ is proportional to $\lambda^2$, yielding $n \sim A^2$.
 
-As shown in Figure~\ref{fig:dissociation}b and the last two columns of Table~\ref{tab:results} this relation describes very well our data.
+As shown in Figure~\ref{fig:dissociation}b and the last two columns of Table~\ref{tab:results} this relation describes very well our data with a prefactor close to 1.
 We can extrapolate this law for the case of \ce{PIm+Br-} and obtain $A = 28$.
@@
 
-Well hydrated anions and aromatic cation favour counterion condensation, resulting in too few charges to allow dissolution and gel formation.
-The hardest gels with the narrowest linear domain are obtained with aliphatic cationic moiety and poorly hydrated counterion at the other end of the Hofmeister series.
+Strongly interacting anions (i.e.
+small and hard halides, like fluoride) and aromatic cation favour counterion condensation, resulting in too few charges to allow dissolution and gel formation.
+The hardest gels with the narrowest linear domain are obtained with aliphatic cationic moiety and counterion at the other end of the Hofmeister series that are less bound to the polymer.
 This gives rise to maximal charge dissociation in water, facilitates charge pairing of the peripheral, iterative cations with the terminal anion and thus affords a very high density of cross-links.
-In between, in the case of aliphatic cations with a counterion in the middle of the Hofmeister series (\ce{Cl-}) or in the case of aromatic cation and poorly hydrated anion, charge dissociation is limited, polymers associate in processions and cross-links are few.
+In between, in the case of aliphatic cations with a counterion in the middle of the Hofmeister series (\ce{Cl-}) or in the case of aromatic cation and poorly interacting anion, charge dissociation is limited, polymers associate in processions and cross-links are few.
 We thus obtain soft, highly deformable gels.

### 2016-02-17 PALSE/ion_pairing_paper.tex
Corrections Seb, mainly typos

@@
 We demonstrated recently that polyelectrolytes with cationic moieties along the chain and a single anionic head are able to form physical hydrogels due to the reversible nature of the head-to-body ionic bond.
-Here we generate a variety of such polyelectrolytes with various cationic moieties and counterions combinations starting from a common polymeric platform.
+Here we generate a variety of such polyelectrolytes with various cationic moieties and counterion combinations starting from a common polymeric platform.
 We show that the rheological properties (shear modulus, critical strain) of the final hydrogels can be modulated over three orders of magnitude depending on the cation/anion pair.
@@
 Contrary to a small electrolyte for which full ion pairs dissociation is generally achieved in water, polyelectrolytes in solution are in general not fully dissociated, with a substantial fraction of the counterions bound to the polymer\cite{Manning1979}.
-The resulting net charge of the chain governs the physical properties of the polyelectrolytes solution\cite{Rubinstein1996}, first of all the ability of the polymer to dissolve.
-counterion condensation is the physical bounding or adsorption of counterions near the polymer chain.
+The resulting net charge of the chain governs the physical properties of the polyelectrolyte solution\cite{Rubinstein1996}, first of all the ability of the polymer to dissolve.
+Counterion condensation is the physical bounding or adsorption of counterions near the polymer chain.
 Factors influencing this process have been known since the 1880's but are still not fully understood.
@@
 
-We have recently reported on a highly reversible poly(cation) based hydrogel\cite{Srour2014}, which formation relies on an new concept of supramolecular  electrostatic interaction.
+We have recently reported on a highly reversible poly(cation) based hydrogel\cite{Srour2014}, which formation relies on a new concept of supramolecular  electrostatic interaction.
 Briefly, the atom transfer radical polymerisation (ATRP) is initiated by a phosphonate-terminated ATRP initiator, eventually affording an anion-terminated poly(cationic) polymer.
@@
 
-In the present report, we investigate the role of the ion pair dissociation equilibria rate of the poly(cationic) body on the mechanical properties of the hydrogel.
+In the present report, we investigate the role of the couterion condensation on the mechanical properties of such hydrogels.
 In a first part we describe how we take advantage of our postfunctionalization approach to play systematically with the nature (aromatic or not) of the cationic repeating unit and the associated halide counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}) varied along the Hofmeister series.
-In the second part and third part, we show respectively the qualitative change in gel formability and the quantitative variations of mechanical properties of the aqueous dispersions obtained from these well-characterized polymers.
-In a fourth part, we develop a microscopic model based on the idea that at low dissociation rates, cross-links (defined as a polymer that is connected to three or more others) are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular chain.
+In the second and third parts, we show respectively the qualitative change in gel formability and the quantitative variations of mechanical properties of the aqueous dispersions obtained from these well-characterized polymers.
+In a fourth part, we rationalize this behaviour by proposing a microscopic model based on the idea that at low dissociation rates, cross-links (defined as a point where three or more polymers meet) are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular chain.
 We call this behaviour ``processionary'' in analogy to pine processionary caterpillar (\textit{Thaumetopoea pityocampa}) behaviour\cite{Fabre1916}.%Finally, we discuss our results and possible applications of such smart hydrogels.
@@
 In order to bring evidence for the efficiency of the ionic metathesis, we submit the resulting materials to negative-mode high-resolution mass spectrometry (HRMS, see Supplementary Materials).
-We unambiguously assess displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Supplementary Figures~S\ref{S-fig:massPImBr} and S\ref{S-fig:massPPyBr}).
+We unambiguously assess complete displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Supplementary Figures~S\ref{S-fig:massPImBr} and S\ref{S-fig:massPPyBr}).
 We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures~S\ref{S-fig:massPImI} and S\ref{S-fig:massPPyI}), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (not shown).
@@
 Differential Scanning Calorimetry (DSC) thermograms  of all synthesized polymers also reveals marked differences in their properties, that can only be explained by differences in their couterion features.
-Thus, in the case polyelectrolytes with fluoride and chloride counterion, we observe no exo or endothermic transition below \SI{200}{\celsius}.
-By contrast, we observe a broad endothermic peak when iodide and bromide are used as counterions.
-Quite remarkably, we found a similar peak temperature (84-\SI{85}{\celsius}) for \ce{PIm+I-} and \ce{PPyr+I-}, see Figure~\ref{fig:dsc}.
+Thus, in the case polyelectrolytes with fluoride and chloride counterion, we observe no exo or endothermic transition below \SI{200}{\celsius} (featureless thermograms, not shown).
+By contrast, we observe a broad endothermic peak when iodide and bromide are used as counterions, see Figure~\ref{fig:dsc}.
+Quite remarkably, we found a similar peak temperature (84-\SI{85}{\celsius}) for \ce{PIm+I-} and \ce{PPyr+I-}.
 When bromide counterions are present, we observe a significant increase of the peak temperature (114-\SI{115}{\celsius}) but again, with a similar value between \ce{PIm+Br-} and \ce{PPyr+Br-} as shown in Figure~\ref{fig:dsc}.
@@
 However Hofmeister series alone cannot explain the difference between soluble \ce{PPyr+Cl-} and insoluble \ce{PIm+Cl-}.
-We have to suppose that counterions are more strongly condensed near imidazolium than near pyrrolidinium.
+This constitutes a strong indication that counterions are more strongly condensed near imidazolium than near pyrrolidinium.
 Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties.
-Because of their aromatic ring, imidazolium are particularly prone to promote a variety of supramolecular interactions which strongly contribute to their physicochemical properties, such as \ce{\pi+-\pi} or anion$-\pi^+$.
+Because of their aromatic ring, imidazolium ions are particularly prone to promote a variety of supramolecular interactions which strongly contribute to their physicochemical properties, such as \ce{\pi+-\pi} or anion$-\pi^+$.
 More recently, \ce{\pi+-\pi+} interactions have also been identified as a distinctive driving force for imidazolium dimerization\cite{Geronimo2011}.
-Because of strong intercoulombic repulsion, \ce{\pi+-\pi+} are primarily weaker than their \ce{\pi-\pi} counterparts.
-However it has been established that, in the presence of couterion, the stabilizing effect of \ce{\pi+-\pi+} interaction could reach magnitudes largely exceeding that of the latter\cite{Geronimo2011}.
+Because of strong intercoulombic repulsion, \ce{\pi+-\pi+} interactions are primarily weaker than their \ce{\pi-\pi} counterparts.
+However it has been established that, in the presence of couterions, the stabilizing effect of \ce{\pi+-\pi+} interaction could reach magnitudes largely exceeding that of the latter\cite{Geronimo2011}.
 In particular, dimerization of imidazolium-chloride ion pairs have been the object of recent studies, and it has been shown that the involvement of the negatively charged counterion was essential in maximizing the stabilization of the \ce{\pi+-\pi+} complex by minimizing coulombic repulsion between the imidazolium moieties\cite{Matthews2014,Gao2015a}.
-It is likewise well documented that covalent incorporation of interacting group within a polymer chain strongly favours intramolecular interactions between these groups, by increasing their so-called effective molarity\cite{Li2003,Mulder2004,Huerta2013}.
+It is likewise well documented that covalent incorporation of interacting groups within a polymer chain strongly favours intramolecular interactions between these groups, by increasing their so-called effective molarity\cite{Li2003,Mulder2004,Huerta2013}.
 It is therefore not surprising that both effects participate in making the imidazolium-anion interaction stronger than pyrrolidinium at the detriment of solubility.
@@
 In addition, if we suppose that in order to minimize the inter-chain repulsion between charged cationic groups, heads are preferentially attached to the tail of their neighbour, we obtain a linear chain of effective polymerisation index $n N_0$ where $n$ is the number of polymers in the supramolecular queue.
-This behaviour bears a striking resemblance to pine processionary caterpillar behaviour first described by \citet{Fabre1916}:
+This behaviour is somewhat evocative of that of the pine processionary caterpillar memorably described by \citet{Fabre1916}:
 ``They proceed in single file, in a continuous row, each touching with its head the rear of the one in front of it.
@@
 the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, that we consider as representative of the end of the linear regime and of the first broken cross-links\cite{Hyun2011}.
-By using this two parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters as the average number of chains between cross links $n$, the head-to-body bonding energy $E_c$ or the number of monomers between dissociated ion pairs $A$.
+By using these two parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters as the average number of chains between cross-links $n$, the head-to-body bonding energy $E_c$ or the number of monomers between dissociated ion pairs $A$.
 
@@
 
-\citeauthor{Rubinstein1996} express the screening length as
+\citet{Rubinstein1996} express the screening length as
 $r_\mathrm{scr} = \left(A^2 / u\right)^{1/3} \left(\tau /cb \right)^{1/2}$. Combining with (Eq.~\ref{eq:ge}) and (Eq.~\ref{eq:D}) we deduce
@@
 
-The only unknown here is the reduced temperature $\tau$, possibly different between poly(pyrrolidinium) based and poly(immidazolium) based polymers.
+The only unknown here is the reduced temperature $\tau$, possibly different between poly\-(pyrrolidinium) based and poly\-(immidazolium) based polymers.
 We know that the undissociated chain has a $\Theta$ temperature higher than \SI{100}{\celsius}, thus $0.2<\tau<1$.
@@
 A second charge dissociated on a chain creates a possible binding site for a second head-to-body association and thus a cross-link.
-In a 1:1 association model at low association rate, the number of association is linearly proportional to the number of available sites.
 Therefore the probability to have a cross link is proportional to the probability of having a second charge dissociation.
@@
 Strongly interacting anions (i.e.
-small and hard halides, like fluoride) and aromatic cation favour counterion condensation, resulting in too few charges to allow dissolution and gel formation.
+small and hard halides, like fluoride) and aromatic cations favour counterion condensation, resulting in too few charges to allow dissolution and gel formation.
 The hardest gels with the narrowest linear domain are obtained with aliphatic cationic moiety and counterion at the other end of the Hofmeister series that are less bound to the polymer.
 This gives rise to maximal charge dissociation in water, facilitates charge pairing of the peripheral, iterative cations with the terminal anion and thus affords a very high density of cross-links.
-In between, in the case of aliphatic cations with a counterion in the middle of the Hofmeister series (\ce{Cl-}) or in the case of aromatic cation and poorly interacting anion, charge dissociation is limited, polymers associate in processions and cross-links are few.
+In between, in the case of aliphatic cations with a counterion in the middle of the Hofmeister series (\ce{Cl-}) or in the case of aromatic cations and poorly interacting anion, charge dissociation is limited, polymers associate in processions and cross-links are few.
 We thus obtain soft, highly deformable gels.

### 2016-02-24 ion_pairing_paper.tex
update table

@@
 
-Contrary to a small electrolyte that exhibit only dissociated ions, polyelectrolytes in solution are in general not fully dissociated, with a substantial fraction of the counterions bound to the polymer\cite{Manning1979}.
-The resulting net charge of the chain governs the physical properties of the polyelectrolytes solution\cite{Rubinstein1996}, first of all the ability of the polymer to dissolve.
-counterion condensation is the physical bounding or adsorption of counterions near the polymer chain.
-Factors influencing this process have been known since the 1880's but are still being rationalized.
+Contrary to a small electrolyte for which full ion pairs dissociation is generally achieved in water, polyelectrolytes in solution are in general not fully dissociated, with a substantial fraction of the counterions bound to the polymer\cite{Manning1979}.
+The resulting net charge of the chain governs the physical properties of the polyelectrolyte solution\cite{Rubinstein1996}, first of all the ability of the polymer to dissolve.
+Counterion condensation is the physical bounding or adsorption of counterions near the polymer chain.
+Factors influencing this process have been known since the 1880's but are still not fully understood.
 \citet{Hofmeister1888} was the first to propose a systematic ranking of ions, based on their propensity to promote the coalescence of egg white.
@@
 
-In the present report, we investigate the role of the ion pair dissociation equilibria rate of the poly(cationic) body on the mechanical properties of the hydrogel.
-In a first part we describe how we take advantage of our postfunctionalization approach to play systematically with the nature (aromatic or not) of the cationic repeating unit and the associated counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}) varied along the Hofmeister series.
-In the second part and third part, we show respectively the qualitative change in gel formability and the quantitative variations of mechanical properties of the aqueous dispersions obtained from these well-characterized polymers.
-In a fourth part, we develop a microscopic model based on the idea that at low dissociation rates, cross-links (defined as a polymer that is connected to three or more others) are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular chain.
+In the present report, we investigate the role of the couterion condensation on the mechanical properties of such hydrogels.
+In a first part we describe how we take advantage of our postfunctionalization approach to play systematically with the nature (aromatic or not) of the cationic repeating unit and the associated halide counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}) varied along the Hofmeister series.
+In the second and third parts, we show respectively the qualitative change in gel formability and the quantitative variations of mechanical properties of the aqueous dispersions obtained from these well-characterized polymers.
+In a fourth part, we rationalize this behaviour by proposing a microscopic model based on the idea that at low dissociation rates, cross-links (defined as a point where three or more polymers meet) are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular chain.
 We call this behaviour ``processionary'' in analogy to pine processionary caterpillar (\textit{Thaumetopoea pityocampa}) behaviour\cite{Fabre1916}.%Finally, we discuss our results and possible applications of such smart hydrogels.
@@
 
-We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 790 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
+We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 880 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
 This larger cross-link ratio is consistent with a higher probability of attaching two or more heads on a highly dissociated body.
@@
 &	{\si{\pascal}} &  & & {$k_\mathrm{B}T$} & {\si{\kilo\joule/\mol}} & \\\hline&&&&&\\[-10pt]
-\ce{PPyr+I-}	& 8800	&	0.04	&	1.	&	2e-3\cellcolor{gray!25}	&	0.004\cellcolor{gray!25}	&	1.	&	1.\\
-\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.05&	3e-4\cellcolor{gray!25}	&	0.001\cellcolor{gray!25}	&	1.1	&	1.\\
-\ce{PPyr+Cl-}	& 270 	&	0.09	&	45	&	8e-3\cellcolor{gray!25}	&	0.02\cellcolor{gray!25}	&	7.	&	6.7\\
-\ce{PIm+I-}	& 78	&	0.16	&	113	&	3e-2\cellcolor{gray!25}	&	0.063\cellcolor{gray!25}	&	9.6	&	10.6\\
-\ce{PIm+Br-}	& 13	&	8.04	&	790	&	64	&	160.	&	3.4\cellcolor{gray!25}	&	28\\
+\ce{PPyr+I-}	& 8800	&	0.04	&	1.08	&	2e-3\cellcolor{gray!25}	&	0.004\cellcolor{gray!25}	&	10.6	&	1.\\
+\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.15&	3e-4\cellcolor{gray!25}	&	0.001\cellcolor{gray!25}	&	11.7	&	1.\\
+\ce{PPyr+Cl-}	& 270 	&	0.09	&	51	&	8e-3\cellcolor{gray!25}	&	0.02\cellcolor{gray!25}	&	76.2	&	6.7\\
+\ce{PIm+I-}	& 78	&	0.16	&	123	&	3e-2\cellcolor{gray!25}	&	0.063\cellcolor{gray!25}	&	101.8	&	10.6\\
+\ce{PIm+Br-}	& 13	&	8.04	&	882	&	64	&	160.	&	36.5\cellcolor{gray!25}	&	28\\
 \end{tabular}
@@
 Finally we cross-check our model by linking the ion pair dissociation and the amount of physical cross-links by an independent argument.
+A second charge dissociated on a chain creates a possible binding site for a second head-to-body association and thus a cross-link.
+Therefore the probability to have a cross link is proportional to the probability of having a second charge dissociation.
 If we assume that the probability for an ion pair to dissociate is independent of the number of dissociated pairs already present on the polymer, we can describe the probability of having $k$ dissociated pairs on a polymer by a Poisson distribution of mean $\lambda = N_0/A$.
-Having 2 charges dissociated on a chain implies a cross-link.
 Therefore, the cross-link probability $1/n$ is proportional to $\lambda^2$, yielding $n \sim A^2$.
 
-As shown in Figure~\ref{fig:dissociation}b and the last two columns of Table~\ref{tab:results} this relation describes very well our data.
+As shown in Figure~\ref{fig:dissociation}b and the last two columns of Table~\ref{tab:results} this relation describes very well our data with a prefactor close to 1.
 We can extrapolate this law for the case of \ce{PIm+Br-} and obtain $A = 28$.
@@
 
-Well hydrated anions and aromatic cation favour counterion condensation, resulting in too few charges to allow dissolution and gel formation.
-The hardest gels with the narrowest linear domain are obtained with aliphatic cationic moiety and poorly hydrated counterion at the other end of the Hofmeister series.
+Strongly interacting anions (i.e.
+small and hard halides, like fluoride) and aromatic cations favour counterion condensation, resulting in too few charges to allow dissolution and gel formation.
+The hardest gels with the narrowest linear domain are obtained with aliphatic cationic moiety and counterion at the other end of the Hofmeister series that are less bound to the polymer.
 This gives rise to maximal charge dissociation in water, facilitates charge pairing of the peripheral, iterative cations with the terminal anion and thus affords a very high density of cross-links.
-In between, in the case of aliphatic cations with a counterion in the middle of the Hofmeister series (\ce{Cl-}) or in the case of aromatic cation and poorly hydrated anion, charge dissociation is limited, polymers associate in processions and cross-links are few.
+In between, in the case of aliphatic cations with a counterion in the middle of the Hofmeister series (\ce{Cl-}) or in the case of aromatic cations and poorly interacting anion, charge dissociation is limited, polymers associate in processions and cross-links are few.
 We thus obtain soft, highly deformable gels.

### 2016-04-20 ion_pairing_paper.tex
take the comb into account

@@
 \subsection{Rheological measurements}
+\begin{figure}[t]
+\includegraphics{strain.pdf}%
+\caption{Storage modulus $G^\prime$ ($\bullet$) and loss modulus $G^{\prime\prime}$ ($\square$) measured through oscillatory shear experiments plotted against the strain amplitude $\gamma$.
+The moieties change with rows and the counterions with columns.
+The fixed frequency is $f=\SI{1}{\hertz}$.
+All samples are at 8\%wt.
+The gray area limits of the linear, plastic and shear thinning regimes.}%
+\label{fig:strain}%
+\end{figure}
+
 We perform rheological studies of the water-swelling materials.
@@
 For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile is that of pure water which confirms the visual observation.
-However as shown in Figure~\ref{fig:strain}, all other samples are solidlike ($G^\prime \gg G^{\prime\prime}$) at small strain and flow at large strains ($G^{\prime\prime} \gg G^\prime$).
+However as shown in Figure~\ref{fig:strain}, all other samples are solidlike ($G^\prime \gg G^{\prime\prime}$) at small strain and flow at large strains ($G^{\prime\prime} \gg G^\prime$) with a steep decrease of the moduli.
 We confirm solidlike behavior at low strain for all accessible frequencies (see supplementary material S6 Fig.
 1).
-We define the critical strain value $\gamma_c$ between these two regimes when $G^{\prime\prime}$ reaches a maximum.
+We checked that the linear mechanical properties are unchanged by the flow history given a few minutes of rest.
+
+The mechanical properties at low strain evolve with the factors influencing counter-ion condensation.
+Imidazolium based gels are much weaker than their pyrrolidinium counterpart.
+Comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ value (13 and \SI{8100}{\pascal}, resp.) differ by almost three orders of magnitude.
+On the bottom row of Figure~\ref{fig:strain} we can compare the rheological responses of Poly(pyrrolidinium)-based hydrogels for the three counterions allowing dissolution in water.
+Gels with bromide and iodide counterions show roughly the same modulus value at vanishing strain for both $G^\prime$ and $G^{\prime\prime}$.
+By contrast, the gel with chloride counterions is weaker by more than one order of magnitude.
+In other words, counterion condensation is correlated to softer gel.
+This trend is confirmed between \ce{PIm+Br-} and \ce{PIm+I-}, the former being an order of magnitude softer than the later.
+
+Moderately charged \ce{PIm+I-} and \ce{PPyr+Cl-} show a plateau in both moduli at low strain corresponding to the linear regime of the material; an overshoot of $G^{\prime\prime}$ and a downward slope of $G^\prime$ at intermediate strains corresponding to the onset of plasticity~\cite{Hyun2011}; and a decrease of both moduli at large strains, steeper for $G^\prime$ than for $G^{\prime\prime}$, indicating shear thinning.
+The linear regime is either extremely narrow or non existent in heavily charged \ce{PPyr+Br-} and \ce{PPyr+I-}.
+By contrast the lightly charged \ce{PIm+Br-} displays a broad intermediate regime (tree decades of strain) between the linear regime and the crossing of the moduli.
+In any case, the end of the linear regime corresponds to a strain much smaller than 1.
+
+In the following we show that the correlation between softness and counterion condensation can be explained on the basis of a microscopic model of interchain interactions.
+
+%We define the critical strain value $\gamma_c$ between these two regimes when $G^{\prime\prime}$ reaches a maximum.
 We checked that the linear mechanical properties are unchanged by the flow history given a few minutes of rest.
@@
 
-\begin{figure}
-\includegraphics{strain.pdf}
-\caption{Storage modulus $G^\prime$ (filled symbols) and loss modulus $G^{\prime\prime}$ (open symbols) measured through oscillatory shear experiments plotted against the strain amplitude $\gamma$ for (a) \ce{PPyr+Cl-} (\textcolor{lightgray}{$\bullet$}), \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}) and \ce{PPyr+I-} ($\blacktriangle$), (b) \ce{PIm+I-} ($\blacktriangledown$) and \ce{PPyr+I-} ($\blacktriangle$) and (c) \ce{PIm+Br-} (\textcolor{gray}{$\blacklozenge$}) and \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}).
-The fixed frequency is $f=\SI{1}{\hertz}$.
-All samples are at 8\%wt }
-\label{fig:strain}
-\end{figure}
-
-In Figure~\ref{fig:strain}a we compare the rheological responses of Poly(pyrrolidinium)-based hydrogels for the three counterions allowing dissolution in water.
+%In Figure~\ref{fig:strain}a we compare the rheological responses of Poly(pyrrolidinium)-based hydrogels for the three counterions allowing dissolution in water.
 Gels with bromide and iodide counterions show roughly the same mechanical behavior in the linear regime but \ce{PPyr+I-} yields at slightly higher strains ($\gamma_c=4\%$ instead of $3\%$).
@@
 
-We also compare the rheology of \ce{PIm+X-} and \ce{PPyr+X-} with the same counterion \ce{X} in Figure~\ref{fig:strain}b,c.
+%We also compare the rheology of \ce{PIm+X-} and \ce{PPyr+X-} with the same counterion \ce{X} in Figure~\ref{fig:strain}b,c.
 For both iodide (Figure~\ref{fig:strain}b, X=I) and bromide (Figure~\ref{fig:strain}c, X=Br), imidazolium based gels are much weaker than their pyrrolidinium counterpart with a broader linear regime.
@@
 
-\subsubsection{From modulus to effective chain length}
+\subsubsection{Conformation at rest}
 
-Between two cross-links the ``procession'' of polymer chains can be considered as an entropic spring made of $N=n N_0$ monomers. The shear modulus of the gel is thus given by\cite{Rubinstein1996}:
+Let us first recall the structure of a polyelectrolyte chain as described by the scaling theory of \citet{Rubinstein1996}.
+In absence of strain, a linear polyelectrolyte chain is organised at various scales.
+
+Below the thermal correlation length $\xi_T$ monomer interactions dominate over the thermal energy, thus chains behave ideally and we need to take into account the geometry of the polymer.
+Due to the relatively large lateral moieties ($s=11$ non-hydrogen atoms) with respect to the distance between them (2 carbons), the polymer can be considered as a comb polymer or a bottlebrush.
+According to \citet{Feuz2005} the persistence length is thus longer than the distance $b$ between two monomers and corresponds to the thickness $d=s^{2/3} a^{4/3}b^{-1/3}$ of the brush in ideal conditions, where $a\approx\SI{150}{\pico\metre}$ is the distance between two non-hydrogen atoms.
+Taking $b=a\sqrt{3}$, we obtain $d = \left(s/3\right)^{2/3}b \approx 2.4 b$.
+We note $g_c=d/b$ the number of monomers in the persistence length.
+
+The density in thermal blobs is set by the reduced temperature $\tau= 1-T/\Theta$ such that $(g_T/g_c)/\xi_T^3 = \tau/d^3$ and their structure is a persistent random walk $\xi_T = d (g_T/g_c)^{1/2}$.
+These conditions yields $\xi_T = d/\tau$ and $g_T = g_c/\tau$.
+
+At a scale $D$, the electrostatic energy balances the surface energy of the collapsed polymer.
+We name $g_e$ the number of monomers in the electrostatic blob.
+If there is a charge every $A$ monomers and there are $g_e/A$ charges in the electrostatic blob.
+The surface tension is proportional to the number of thermal blobs exposed on the surface $k_\mathrm{B}T/\xi_T^2$, with $k_\mathrm{B}$ is the Boltzmann constant and $T$ the temperature.
+Thus the balance of energies writes:
+\begin{equation}
+\left(\frac{D}{\xi_T}\right)^2 = \left(\frac{g_e}{A}\right)^2 \frac{\ell_\mathrm{B}}{D},
+\label{eq:electrosurface}
+\end{equation}
+where $\ell_\mathrm{B} = e^2/(4\pi\epsilon k_\mathrm{B}T)$ is the Bjerrum length with $e$ the elementary charge, $\epsilon$ the dielectric constant of the solvent.
+In water $\ell_\mathrm{B} \approx \SI{0.7}{\nano\metre}$.
+Since thermal blobs fill the volume of the electrostatic blob $D = \xi_T \left(g_e/g_T\right)^{1/3}$, we obtain
+\begin{equation}
+g_e = \frac{A^2}{u} \tau, \qquad D = b\left(\frac{A^2}{u}g_c^2\right)^{1/3},\qquad \text{with }u = \ell_\mathrm{B}/b.
+\end{equation}
+
+We define the extension parameter $B$ as the ratio between the length of a chain of $g_e$ monomers fully extended and the actual size of the electrostatic blob: 
+\begin{equation}
+B = \frac{g_e b}{D} = \left(\frac{A^2}{u g_c}\right)^{2/3} \tau
+\end{equation}
+
+At constant solvent quality and polymer architecture, $B$ is thus monotonically related to the amount of counterion condensation. In the following we will estimate $B$ from rheological measurements.
+
+Electrostatic repulsion decays with a characteristic length $r_\mathrm{src}$, the screening length.
+Following\cite{Rubinstein1996} this screening length is set by the concentration of dissociated counterions and the concentration of $c_s$ added salts:
+\begin{equation}
+r_\mathrm{src} = \left(\frac{B}{cb}\right)^{1/2} \left(1 + Bu \frac{c_s}{c}\right)^{-1/2}.
+\end{equation}
+Between $D$ and $r_\mathrm{src}$ the electrostatic blobs are organised in a linear rod containing $g_\mathrm{src}$ monomers.
+The length of this rod is $B$ times shorter than if the electrostatic blobs were extended:
+\begin{equation}
+g_\mathrm{src} = B \frac{r_\mathrm{src}}{b} = \left(\frac{B^3}{cb^3}\right)^{1/2} \left(1 + Bu \frac{c_s}{c}\right)^{-1/2}.
+\label{eq:gscr}
+\end{equation}
+
+The last length scale is the correlation length $\xi$.
+Between $r_\mathrm{src}$ and $\xi$ the chain forms a self avoiding walk of persistence length $r_\mathrm{src}$.
+Above the correlation length the polyelectrolyte chain forms a random walk of correlation blobs containing $g=c\xi^3$ monomers.
+
+\subsubsection{From modulus to procession length}
+
+If between two crosslinks or entanglements the ``procession'' of polymer chains contains more monomers than $g$ or $g_\mathrm{src}$ it performs a random or self avoiding walk respectively.
+Each chain is thus an entropic spring of constant stiffness $k_\mathrm{B}T$.
+The number density of chains is $c/N$ with $c$ the monomer concentration and $N$ the number of monomers between cross-links or entanglements.
+Therefore, the shear modulus at low strain is given by:
 \begin{equation}
@@
 \end{equation}
-where $c$ is the monomer number density, $k_\mathrm{B}$ is the Boltzmann constant and $T$ the temperature.
+where $c$ is the monomer number density.
 Because of the different weights of the pyrrolidinium and imidazolium moieties and of the different counterions, $c$ is not constant as we chose to conduct our experiments at constant weight fraction $w$ of polymer (generally used to quantify gelation ability of a given gelator).
@@
 \begin{tabular}{l|SS|SSSSS}
-& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_c$} & {$n$} & \multicolumn{2}{c}{$E_c$} & {$A$} & {$\sqrt{n}$}\\
-&	{\si{\pascal}} &  & & {$k_\mathrm{B}T$} & {\si{\kilo\joule/\mol}} & \\\hline&&&&&\\[-10pt]
-\ce{PPyr+I-}	& 8800	&	0.04	&	1.08	&	2e-3\cellcolor{gray!25}	&	0.004\cellcolor{gray!25}	&	10.6	&	1.\\
-\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.15&	3e-4\cellcolor{gray!25}	&	0.001\cellcolor{gray!25}	&	11.7	&	1.\\
-\ce{PPyr+Cl-}	& 270 	&	0.09	&	51	&	8e-3\cellcolor{gray!25}	&	0.02\cellcolor{gray!25}	&	76.2	&	6.7\\
-\ce{PIm+I-}	& 78	&	0.16	&	123	&	3e-2\cellcolor{gray!25}	&	0.063\cellcolor{gray!25}	&	101.8	&	10.6\\
-\ce{PIm+Br-}	& 13	&	8.04	&	882	&	64	&	160.	&	36.5\cellcolor{gray!25}	&	28\\
+& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_0$} & {$n$} & {$B$} & \multicolumn{2}{c}{$A$}\\
+&	{\si{\pascal}} & {\%} & & &{$\tau=0.2$}& {$\tau=1$}\\\hline&&&&&\\[-10pt]
+\ce{PIm+Br-}	& 13	&	1.1	&	860	&	740	&	1370	&	410\\
+\ce{PIm+I-}		& 78	&	1.8	&	122	&	104.8	&	317	&	94\\
+\ce{PPyr+Cl-}	& 270 	&	2.8	&	50.6	&	55.7	&	197	&	59\\
+\ce{PPyr+Br-}	& 9800 	& 	0.3	&	1.15	&	2.91	&	21	&	6.5\\
+\ce{PPyr+I-}	& 8800	&	0.0	&	1.08	&	2.63	&	19	&	6.0\\
 \end{tabular}
@@
 \end{table}
+
+\subsubsection{Limit of the linear regime}
+
+As the material is strained, the random walk is stretched, then the self-avoiding walk.
+When the correlation blobs are fully stretched, the chain is a linear assembly of electrostatic blobs.
+Stretching the chain further means exposing more thermal blobs to the solvent, a process that cannot be reduced to spring of constant stiffness.
+This condition is thus the limit between the linear and non linear regimes.
+
+For all our samples, the linear regime is narrow $\gamma_0\approx 10^{-2}\ll 1$ if existing.
+It means that procession is just long enough to start performing the self-avoiding walk.
+Incidentally, this implies that the procession is too short to perform the random walk.
+In other words the number $N$ of monomer between crosslinks is comparable to $g_\mathrm{scr}$.
+Using equation (\ref{eq:gscr}), this condition translates into
+\begin{equation}
+\left(\frac{B}{B_0}\right)^3 = 1 + \frac{B}{B_s}
+\label{eq:solveB}
+\end{equation}
+with $B_s = c/u c_s$ and $B_0 = b c^{1/3} N^{2/3} = bc \left(k_\mathrm{B}T/G\right)^{2/3}$.
+Here, the ``salt'' are the phosphonate heads, that is to say two charges per chain.
+Thus $c_s/c = 2/N_0$ and $B_s = N_0/2u \approx 23$.
+
+Analysing Equation (\ref{eq:solveB}), we identify two physically relevant limit cases: (i) if $B \ll B_s$ (low salt condition) then $B = B_0$; (ii) if $B \gg B_s$ (high salt condition) then $B = \left(B_0^3/B_s\right)^{1/2} > B_0$.
+
+Numerical solutions of Equation (\ref{eq:solveB}) are given in Table~\ref{tab:results}.
+For the same quality of solvent $B$ is monotonically related to the amount of counterion condensation.
+We thus confirm that our samples are sorted by decreasing counterion condensation (decreasing $B$).
+More quantitatively, the number $A$ of monomers between dissociated charges is
+\begin{equation}
+A = g_c u^{1/2} \left(\frac{B}{\tau}\right)^{3/4}
+\end{equation}
+We know that $\Theta>\SI{100}{\celsius}$ so $0.2<\tau<1$.
 

### 2016-05-02 ion_pairing_paper.tex
Kuhn length Stretching of the eletrostatic blobs to obtain bounding energy from strain at crossing

@@
 As we will discuss in the following, all rheological features of the hydrogels studied here can  be rationalised on the basis of this ``processionary'' model.
-We base our analysis on two main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
-the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, that we consider as representative of the end of the linear regime and of the first broken cross-links\cite{Hyun2011}.
-By using these two parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters as the average number of chains between cross-links $n$, the head-to-body bonding energy $E_c$ or the number of monomers between dissociated ion pairs $A$.
+We base our analysis on three main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
+the elasticity of the undamaged gel network, (ii) the strain amplitude $\gamma_0$ corresponding to the end of the linear regime~\cite{Hyun2011}, and (iii) the strain amplitude $\gamma_c$ corresponding to the crossing of the moduli.
+By using these three parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters as the average number of chains between cross-links $n$, the number of monomers between dissociated ion pairs $A$ and the head-to-body bonding energy $E_c$.
 
@@
 
-Let us first recall the structure of a polyelectrolyte chain as described by the scaling theory of \citet{Rubinstein1996}.
-In absence of strain, a linear polyelectrolyte chain is organised at various scales.
+The scaling theory of \citet{Rubinstein1996} describes the structure of a polyelectrolyte chain.
+In the following, we make the link between this theory and the chemical structure of our polymers.
+In absence of strain, the structure of a polyelectrolyte chain is organised at various scales.
 
-Below the thermal correlation length $\xi_T$ monomer interactions dominate over the thermal energy, thus chains behave ideally and we need to take into account the geometry of the polymer.
-Due to the relatively large lateral moieties ($s=11$ non-hydrogen atoms) with respect to the distance between them (2 carbons), the polymer can be considered as a comb polymer or a bottlebrush.
-According to \citet{Feuz2005} the persistence length is thus longer than the distance $b$ between two monomers and corresponds to the thickness $d=s^{2/3} a^{4/3}b^{-1/3}$ of the brush in ideal conditions, where $a\approx\SI{150}{\pico\metre}$ is the distance between two non-hydrogen atoms.
-Taking $b=a\sqrt{3}$, we obtain $d = \left(s/3\right)^{2/3}b \approx 2.4 b$.
-We note $g_c=d/b$ the number of monomers in the persistence length.
+The smallest scale is the Kuhn length $b$ containing $g_\mathrm{K}$ monomers.
+For a freely joint chain of tetrahedral carbons $b\approx 367$ pm which corresponds to 3 tetrahedral carbons.
+Since a monomer counts 2 carbons along the chain $g_\mathrm{K}\approx 3/2$.
 
-The density in thermal blobs is set by the reduced temperature $\tau= 1-T/\Theta$ such that $(g_T/g_c)/\xi_T^3 = \tau/d^3$ and their structure is a persistent random walk $\xi_T = d (g_T/g_c)^{1/2}$.
-These conditions yields $\xi_T = d/\tau$ and $g_T = g_c/\tau$.
+In bad solvent, the thermal length $\xi_T$ is larger than $b$.
+The thermal blob contains $g_T$ monomers and its density is set by the reduced temperature $(g_T/g_\mathrm{K})/\xi_T^3 = \tau/b^3$.
+Between $b$ and $\xi_T$ we have a persistent random walk $\xi_T = b (g_T/g_\mathrm{K})^{1/2}$.
+These conditions yields $\xi_T = b/\tau$ and $g_T = g_\mathrm{K}/\tau^2$.
 
@@
 We name $g_e$ the number of monomers in the electrostatic blob.
-If there is a charge every $A$ monomers and there are $g_e/A$ charges in the electrostatic blob.
-The surface tension is proportional to the number of thermal blobs exposed on the surface $k_\mathrm{B}T/\xi_T^2$, with $k_\mathrm{B}$ is the Boltzmann constant and $T$ the temperature.
+If there is a charge every $A$ monomers then there are $g_e/A$ charges in the electrostatic blob.
+The surface tension is proportional to the number of thermal blobs exposed on the surface $k_\mathrm{B}T/\xi_T^2$, with $k_\mathrm{B}$ the Boltzmann constant and $T$ the temperature.
 Thus the balance of energies writes:
-\begin{equation}
-\left(\frac{D}{\xi_T}\right)^2 = \left(\frac{g_e}{A}\right)^2 \frac{\ell_\mathrm{B}}{D},
+\begin{equation}\left(\frac{D}{\xi_T}\right)^2 = \left(\frac{g_e}{A}\right)^2 \frac{\ell_\mathrm{B}}{D},
 \label{eq:electrosurface}
@@
 \begin{equation}
-g_e = \frac{A^2}{u} \tau, \qquad D = b\left(\frac{A^2}{u}g_c^2\right)^{1/3},\qquad \text{with }u = \ell_\mathrm{B}/b.
+g_e = \frac{A^2}{u g_\mathrm{K}} \tau, \qquad D = b\left(\frac{A^2}{u g_\mathrm{K}^2}\right)^{1/3},\qquad \text{with }u = \ell_\mathrm{B}/b.
+\label{eq:geD}
 \end{equation}
@@
 \begin{equation}
-B = \frac{g_e b}{D} = \left(\frac{A^2}{u g_c}\right)^{2/3} \tau
+B = \frac{g_e}{g_\mathrm{K}}\frac{b}{D} = \left(\frac{A^2}{u}\right)^{2/3} g_\mathrm{K}^{-1/3}\tau
+\label{eq:B}
 \end{equation}
@@
 
-Electrostatic repulsion decays with a characteristic length $r_\mathrm{src}$, the screening length.
-Following\cite{Rubinstein1996} this screening length is set by the concentration of dissociated counterions and the concentration of $c_s$ added salts:
+In dilute solution, the chain is a linear rod of length $L= N b / B g_\mathrm{K}$.
+The overlap concentration is thus $c^* = N/L^3 = g_\mathrm{K}B / b L^2$.
+Following \cite{Dobrynin1995}, for a monomer concentration $c > c^*$ and in absence of salt the screening length is
 \begin{equation}
-r_\mathrm{src} = \left(\frac{B}{cb}\right)^{1/2} \left(1 + Bu \frac{c_s}{c}\right)^{-1/2}.
+r_\mathrm{src} = L \left(\frac{c^*}{c}\right)^{1/2} =  \left(\frac{B g_\mathrm{K}}{cb}\right)^{1/2}.
+\label{eq:rscrNoSalt}
 \end{equation}
+
+In the presence of added salt, electrostatic interactions are screened at the Debye length $r_\mathrm{src} = \left(\ell_\mathrm{B} c_s\right)^{-1/2}$. For arbitrary salt concentration,
+\begin{equation}
+r_\mathrm{src} = \left(\frac{B g_\mathrm{K}}{cb}\right)^{1/2} \left(1+ B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
+\label{eq:rscr}
+\end{equation}
+
 Between $D$ and $r_\mathrm{src}$ the electrostatic blobs are organised in a linear rod containing $g_\mathrm{src}$ monomers.
@@
 \begin{equation}
-g_\mathrm{src} = B \frac{r_\mathrm{src}}{b} = \left(\frac{B^3}{cb^3}\right)^{1/2} \left(1 + Bu \frac{c_s}{c}\right)^{-1/2}.
+g_\mathrm{src} = B \frac{r_\mathrm{src}}{b} g_\mathrm{K} = \left(\frac{B^3 g_\mathrm{K}^3}{cb^3}\right)^{1/2} \left(1 + B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
 \label{eq:gscr}
@@
 \begin{table}
-\begin{tabular}{l|SS|SSSSS}
-& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_0$} & {$n$} & {$B$} & \multicolumn{2}{c}{$A$}\\
-&	{\si{\pascal}} & {\%} & & &{$\tau=0.2$}& {$\tau=1$}\\\hline&&&&&\\[-10pt]
-\ce{PIm+Br-}	& 13	&	1.1	&	860	&	740	&	1370	&	410\\
-\ce{PIm+I-}		& 78	&	1.8	&	122	&	104.8	&	317	&	94\\
-\ce{PPyr+Cl-}	& 270 	&	2.8	&	50.6	&	55.7	&	197	&	59\\
-\ce{PPyr+Br-}	& 9800 	& 	0.3	&	1.15	&	2.91	&	21	&	6.5\\
-\ce{PPyr+I-}	& 8800	&	0.0	&	1.08	&	2.63	&	19	&	6.0\\
+\sisetup{
+separate-uncertainty,
+%table-format = 1.3(1),
+}
+\begin{tabular}{l|SSS|SSSSS}
+& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_0$} & {$\gamma_c$} & {$n$} & {$B$} & {$A$} & {$E_c$} & {$E_c$}\\
+&	{\si{\pascal}} & {\%} & {\%} & & & & {\si{\kilo\joule\per\mole}}& {$k_\mathrm{B}T$}\\\hline&&&&&\\[-10pt]
+\ce{PIm+Br-}	& 13	&	1.1	&	810	&	860	&	930	&	383	&	300	&	120\\
+\ce{PIm+I-}		& 78	&	1.8	&	18	&	122	&	126	&	85	&	116	&	47\\
+\ce{PPyr+Cl-}	& 270 	&	2.8	&	12	&	50.6	&	64	&	52	&	67	&	27\\
+\ce{PPyr+Br-}	& 9800 	& 	0.3	&	2	&	1.15	&	2.8	&	5	&	7.3	&	2.9\\
+\ce{PPyr+I-}	& 8800	&	0.0	&	89	&	1.08	&	2.5	&	4.6	&	7.2	&	2.9\\
 \end{tabular}
-\caption{Summary of rheological measurements and microscopic values deduced from the model.
-Gray background indicates inconsistent values obtained either by (Eq.~\ref{eq:Ec}) or by (Eq.~\ref{eq:A}) outside of their respective validity domain.}
+\caption{Summary of rheological measurements and microscopic values deduced from the model. $A$ and $E_c$ are obtained supposing $\tau=0.588$.}
 \label{tab:results}
@@
 Incidentally, this implies that the procession is too short to perform the random walk.
-In other words the number $N$ of monomer between crosslinks is comparable to $g_\mathrm{scr}$.
+In other words the number $N$ of monomers between crosslinks is comparable to $g_\mathrm{scr}$.
 Using equation (\ref{eq:gscr}), this condition translates into
@@
 \end{equation}
-with $B_s = c/u c_s$ and $B_0 = b c^{1/3} N^{2/3} = bc \left(k_\mathrm{B}T/G\right)^{2/3}$.
-Here, the ``salt'' are the phosphonate heads, that is to say two charges per chain.
-Thus $c_s/c = 2/N_0$ and $B_s = N_0/2u \approx 23$.
+with $B_s = c/u c_s g_\mathrm{K}$ and $B_0 = b c^{1/3} N^{2/3} g_\mathrm{K}^{-1} = (bc/g_\mathrm{K}) \left(k_\mathrm{B}T/G\right)^{2/3}$.
+Here, the ``salt'' consists in the phosphonate heads, that is to say two charges per chain.
+Thus $c_s/c = 2/N_0$ and $B_s = N_0/2u g_\mathrm{K} \approx 12$.
 
@@
 \begin{equation}
-A = g_c u^{1/2} \left(\frac{B}{\tau}\right)^{3/4}
+A = u^{1/2} \left(\frac{B}{\tau}\right)^{3/4} g_\mathrm{K}^{1/4}
+\label{eq:A}
+\end{equation}.
+
+For the theory to be correct, we need the electrostatic length to be larger than the thermal length, in other words $g_e \geq g_T$.
+Using Equations (\ref{eq:geD}) and (\ref{eq:B}), we express $g_e$ function of $B$ and $\tau$:
+\begin{equation}
+g_e = \left(\frac{B^3}{g_\mathrm{K}\tau}\right)^{1/2},
 \end{equation}
-We know that $\Theta>\SI{100}{\celsius}$ so $0.2<\tau<1$.
+which yields the condition $\tau \geq \frac{g_\mathrm{K}}{B}$.
+This condition it the most stringent for PPyrI (smallest $B$) and yields $\tau \geq 0.58$.
+In the following we set $\tau$ to this minimal value that corresponds to $\Theta=\SI{450}{\celsius}$.
 
-\subsubsection{Bounding energy}
+The values we compute for $A$ using Equation (\ref{eq:A}) are shown in Table~\ref{tab:results}.
+$A$ varies between $4.6$ for \ce{PPyr+I-}, corresponding to $15$ uncondensed counterions per polymer, to $383$ for \ce{PIm+Br-} indicating less than a single dissociated counterion every 5 polymers.
 
-Considering that between two cross-links a procession is an entropic spring in its linear domain, the mechanical energy stored under a strain $\gamma$ is $E = \gamma^2 k_\mathrm{B}T$.
-If we suppose that ionic bonds break before the complete extension of the procession, we evaluate the head-to-body bonding energy to
+\subsubsection{Extention of the electrostatic blobs}
+
+The fully extended self-avoiding walk is a cylinder of length $R_0 = Nb/Bg_\mathrm{K}$, of diameter $D$, of volume $V = R_0 D^2$ and of area $\Sigma_0 = R_0 D = (R_0 V)^{1/2}$.
+Extending the cylinder to $R = (\gamma+1) R_0$ while keeping the volume $V$ constant thus creates an extra area
+$
+\Delta\Sigma = \left(\left(\gamma +1\right)^{1/2} -1\right) \Sigma_0
+$
+that translates into a stretching energy due to the surface tension $k_\mathrm{B}T/\xi_T^2$:
 \begin{equation}
-E_c = \gamma_c^2 k_\mathrm{B}T.
-\label{eq:Ec}
+\frac{E_\mathrm{stretch}}{k_\mathrm{B}T} = \left(\left(\gamma +1\right)^{1/2} -1\right) N \left(\frac{\tau^3}{B g_\mathrm{K}^5}\right)^{1/2}.
+\label{eq:stretch}
 \end{equation}
 
-For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per head-to-body bond or \SI{160}{\kilo\joule\per\mol}.
-This value is much higher than the $E_i\approx\SI{5}{\kilo\joule\per\mol}$ typical of ionic bonds in water\cite{Schneider1992}, even when adding up the two bonds of the bivalent phosphonate head\bibnote{Since the two minus charges are borne by the same atom, chelation effect may enhance bounding energy, but not by an order of magnitude}.
-However, one has to keep in mind that the chain is in bad solvent, thus forming locally compact blobs of low water content.
-Therefore, the medium surrounding the ionic bond is not water but the hydrophobic chains.
-Such a low polarity microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}.
-Quantitatively, the bonding energy of an ionic bond is inversely proportional to the relative dielectric constant $\epsilon_r$, therefore we need to suppose a local $\epsilon_r$ around 5 (between 2-3 typical of water-insoluble polymers and 80 for water) to recover the large bonding energy that we measure.
-
-However for all other samples the critical strain is between 2\% and 20\%, leading to inconsistent values for the bonding energies ranging between $10^{-3}k_\mathrm{B}T$ and $10^{-1}k_\mathrm{B}T$ per bond, see Table~\ref{tab:results}.
-Therefore, the end of the linear regime cannot correspond to bond breaking but rather to the point where extended chains become nonlinear.
-
-\subsubsection{From critical strain to charge dissociation}
-
-The spatial conformation of a polyelectrolyte chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions\cite{Rubinstein1996}.
-In bad solvent, the persistent random walk is made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:dissociation}a:
-\begin{align}
-g_e &= \frac{A^2}{u}\tau\label{eq:ge},\\
-D &= b \left(\frac{A^2}{u}\right)^{1/3}\label{eq:D}.
-\end{align}
-Here $A$ is the number of monomers between dissociated ion pairs, $\tau = 1 - T/\Theta$ the reduced temperature and $u = \ell_\mathrm{B}/b \approx 2.7$ is the ratio of the Bjerrum length in water $\ell_\mathrm{B}\approx \SI{0.7}{\nano\metre}$ and monomer size $b\approx \SI{0.26}{\nano\metre}$.
-
-Extending these electrostatic blobs means exposing more monomers to the bad solvent and typically requires energies at least an order of magnitude higher than for the extension of the persistent random walk.
-We will thus consider that the exit from the linear domain of the material corresponds to the full extension of the persistent random walk:
-\begin{equation}
-\gamma_c = \frac{R_0}{R} - 1
-\end{equation}
-where $R$ and $R_0$ are respectively the end-to-end distance and the contour length of the persistent random walk.
-
-A segment of the persistent random walk is $r_\mathrm{scr}$ long and thus contains $g_\mathrm{scr} = g_e r_\mathrm{scr}/D$ monomers. We thus have
-\begin{equation}
-\left. \begin{array}{ll}
-R &= r_\mathrm{scr} \left(\frac{N}{g_\mathrm{scr}}\right)^{1/2}\\
-R_0 &= r_\mathrm{scr} \frac{N}{g_\mathrm{scr}}
-\end{array}\right\rbrace\quad
-\gamma_c = \left(\frac{N}{g_\mathrm{scr}}\right)^{1/2} -1\label{eq:gamma0}.
-\end{equation}
-
-\citet{Rubinstein1996} express the screening length as
-$r_\mathrm{scr} = \left(A^2 / u\right)^{1/3} \left(\tau /cb \right)^{1/2}$. Combining with (Eq.~\ref{eq:ge}) and (Eq.~\ref{eq:D}) we deduce
-\begin{equation}
-g_\mathrm{scr} = \frac{A^2}{u} \left(\frac{\tau^3}{c b^3}\right)^{1/2}\label{eq:gscr}.
-\end{equation}
-
-From (Eq.~\ref{eq:gamma0}) and (Eq.~\ref{eq:gscr}) we can finally relate the measured critical strain to the microscopic parameter $A$:
-\begin{equation}
-A = \frac{\left(N u\right)^{1/2}}{\gamma_c+1}\left(\frac{c b^3}{\tau^3}\right)^{1/4}.
-\label{eq:A}
-\end{equation}
-
-The only unknown here is the reduced temperature $\tau$, possibly different between poly\-(pyrrolidinium) based and poly\-(immidazolium) based polymers.
-We know that the undissociated chain has a $\Theta$ temperature higher than \SI{100}{\celsius}, thus $0.2<\tau<1$.
-In addition, we must have $A\geq 1$, which implies that $\tau\leq 0.235$ through (Eq.~\ref{eq:A}), corresponding to $\Theta\approx\SI{116}{\celsius}$.
-This limit value yields $A=1$ for both \ce{PPyr+I-} and \ce{PPyr+Br-}, indicating full dissociation of the cationic moieties.
-By contrast \ce{PPyr+Cl-} has a charge every 7 monomers.
-Supposing the same $\tau$, \ce{PIm+I-} has a dissociated ion pair every 10 monomers, see Table~\ref{tab:results}.
-Of course analysing \ce{PIm+Br-} based on the value of $\gamma_c$ gives an inconsistent value $A=3.4$ indicating that this hydrogel does not belong to the same regime.
-
-\subsubsection{Charge dissociation and cross-linking ratio}
-
-Finally we cross-check our model by linking the ion pair dissociation and the amount of physical cross-links by an independent argument.
-A second charge dissociated on a chain creates a possible binding site for a second head-to-body association and thus a cross-link.
-Therefore the probability to have a cross link is proportional to the probability of having a second charge dissociation.
-If we assume that the probability for an ion pair to dissociate is independent of the number of dissociated pairs already present on the polymer, we can describe the probability of having $k$ dissociated pairs on a polymer by a Poisson distribution of mean $\lambda = N_0/A$.
-Therefore, the cross-link probability $1/n$ is proportional to $\lambda^2$, yielding $n \sim A^2$.
-
-As shown in Figure~\ref{fig:dissociation}b and the last two columns of Table~\ref{tab:results} this relation describes very well our data with a prefactor close to 1.
-We can extrapolate this law for the case of \ce{PIm+Br-} and obtain $A = 28$.
-Inverting (Eq.~\ref{eq:A}) yields $R_0/R - 1 = 860\%$, indeed higher than the observed $\gamma_c\approx 804\%$.
-This confirms that in the \ce{PIm+Br-} strain sweep $\gamma_c$ actually results from the strength of the ionic bonds.
-
-%We can now estimate the hydrophobic interaction energy $E_h$ by considering that to separate two chains two interfaces with water of area $\pi D^2/4$ have to be created.
-Surface tension is $\tau^2 k_\mathrm{B}T / b^2$\cite{Rubinstein1996}.
-Combining with (Eq.~\ref{eq:D}) we find
-%\begin{equation}
-%\frac{E_i}{k_\mathrm{B}T} = \frac{\pi}{2}\tau^2\left(\frac{A^2}{u}\right)^{2/3} \approx 4
-%\end{equation}
-
-%\subsection{Discussion}
-%
-%The above dependence of the gel parameters on the nature of the counterions could be rationalized on the basis of ion-pairing processes.
-The unsolubility/unswellability of \ce{PIm+F-}, \ce{PPyr+F-} and \ce{PIm+Cl-} suggest that ion pairs are not efficiently dissociated in aqueous medium, thus hindering the solvation process.
-By comparison, \ce{PIm+Br-}, \ce{PIm+I-}, \ce{PPyr+Br-} and \ce{PPyr+I-} form homogeneous materials when suspended in water, which one can correlate with efficient dissociation of the halide counterions from the polymer main chain is fully consistency with decreased ionic bond strength.
-Such dissociation, in turns makes it possible for the ``hard'' phosphonate group on each polymer chain's end to interact with the ``free'' cationic groups of the side chains, which we showed in a previous paper to be at the origin of the gel formation mechanism\cite{Srour2014}.
-%
-%The similarity in gel parameters observed for \ce{PPyr+Br-} and \ce{PPyr+I-} ($G^\prime\approx\SI{8500}{\pascal}$) suggests that full ion pair dissociation is achieved in the latter, or at least that this dissociation is enough in both cases to provide maximal interaction between the terminal phosphonate and the pyrrolidinium side groups.
-Conversely the significantly different behavior between \ce{PIm+I-} ($G^\prime= \SI{78}{\pascal}$) and \ce{PIm+Br-} ($G^\prime= \SI{13}{\pascal}$) indicates that dissociation is only partial and occurs to a lesser extent in the latter than in the former.
-The same conclusion can be drawn for \ce{PPyr+Cl-}, in which the elastic modulus ($G^\prime= \SI{270}{\pascal}$) value drops by more than one order of magnitude as compared to its bromide and iodide counterparts.
-%
-%As already mentioned (vide supra) another striking feature of these gels is that in similar conditions poly(pyrrolidinium) based gels are systematically much stronger than those formed with poly(imidazolium).
-Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge; consequently, according to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected for the latter with the ``soft'' halide counterions\cite{Goossens2009}.
-As a consequence, for a given halide, dissociation is less favored in the case of poly(imidazolium), as clearly illustrated by the difference in solvation between \ce{PIm+Cl-} and \ce{PPyr+Cl-}.
-Conversely, association of the lateral cations with the ``hard'' phosphonate moieties at the polymer chain’s end is expected to be stronger with pyrrolidinium, hence a systematically larger elastic modulus.
-The interplay of these two effects contributes in making poly(pyrrolidinium) based gels much stronger than their poly(imidazolium) counterparts; comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ value (13 and \SI{8100}{\pascal}, resp.) differ by almost three orders of magnitude.
+If we consider that the head-to-body bonds break at $\gamma_c$, we can extract their bonding energy $E_c$ from Equation (\ref{eq:stretch}).
+The results are reported in Table~\ref{tab:results}.
+In the most dissociated \ce{PPyr+I-} and \ce{PPyr+Br-}, the head-to-body bonding energy $E_c$ is compatible with typical ionic bonds, \SI{5}{\kilo\joule\per\mole} in water~\cite{Schneider1992}.
+However the bonding energy increases with increasing counterion condensation to reach 85\% of the carbon-carbon single bond (\SI{350}{\kilo\joule\per\mole} or $140 k_\mathrm{B}T$) in the case of \ce{PIm+Br-}.
+Indeed, at high counterion condensation, the electrostatic blobs are large, up to 530 polymers for \ce{PIm+Br-}, and embed head-to-body bonds into compact blobs of low water content.
+Such a low polarity microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly~\cite{Rehm2010}.
 

### 2016-05-02 PALSE/ion_pairing_paper.tex
Get the reduced temperature tau from maximal extension of PPyrI Compare D and Bjerrum

@@
 As we will discuss in the following, all rheological features of the hydrogels studied here can  be rationalised on the basis of this ``processionary'' model.
-We base our analysis on two main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
-the elasticity of the undamaged gel network, and (ii) the critical strain $\gamma_c$ corresponding to the maximum of $G"$, that we consider as representative of the end of the linear regime and of the first broken cross-links\cite{Hyun2011}.
-By using these two parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters as the average number of chains between cross-links $n$, the head-to-body bonding energy $E_c$ or the number of monomers between dissociated ion pairs $A$.
+We base our analysis on three main observables, see Table~\ref{tab:results}: (i) the value of the shear modulus at small amplitude $G^\prime(\gamma\rightarrow 0)$, i.e.
+the elasticity of the undamaged gel network, (ii) the strain amplitude $\gamma_0$ corresponding to the end of the linear regime~\cite{Hyun2011}, and (iii) the strain amplitude $\gamma_c$ corresponding to the crossing of the moduli.
+By using these three parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters as the average number of chains between cross-links $n$, the number of monomers between dissociated ion pairs $A$ and the head-to-body bonding energy $E_c$.
 
-\subsubsection{From modulus to effective chain length}
+\subsubsection{Conformation at rest}
 
-Between two cross-links the ``procession'' of polymer chains can be considered as an entropic spring made of $N=n N_0$ monomers. The shear modulus of the gel is thus given by\cite{Rubinstein1996}:
+The scaling theory of \citet{Rubinstein1996} describes the structure of a polyelectrolyte chain.
+In the following, we make the link between this theory and the chemical structure of our polymers.
+In absence of strain, the structure of a polyelectrolyte chain is organised at various scales.
+
+The smallest scale is the Kuhn length $b$ containing $g_\mathrm{K}$ monomers.
+For a freely joint chain of tetrahedral carbons $b\approx 367$ pm which corresponds to 3 tetrahedral carbons.
+Since a monomer counts 2 carbons along the chain $g_\mathrm{K}\approx 3/2$.
+
+In bad solvent, the thermal length $\xi_T$ is larger than $b$.
+The thermal blob contains $g_T$ monomers and its density is set by the reduced temperature $(g_T/g_\mathrm{K})/\xi_T^3 = \tau/b^3$.
+Between $b$ and $\xi_T$ we have a persistent random walk $\xi_T = b (g_T/g_\mathrm{K})^{1/2}$.
+These conditions yields $\xi_T = b/\tau$ and $g_T = g_\mathrm{K}/\tau^2$.
+
+At a scale $D$, the electrostatic energy balances the surface energy of the collapsed polymer.
+We name $g_e$ the number of monomers in the electrostatic blob.
+If there is a charge every $A$ monomers then there are $g_e/A$ charges in the electrostatic blob.
+The surface tension is proportional to the number of thermal blobs exposed on the surface $k_\mathrm{B}T/\xi_T^2$, with $k_\mathrm{B}$ the Boltzmann constant and $T$ the temperature.
+Thus the balance of energies writes:
+\begin{equation}\left(\frac{D}{\xi_T}\right)^2 = \left(\frac{g_e}{A}\right)^2 \frac{\ell_\mathrm{B}}{D},
+\label{eq:electrosurface}
+\end{equation}
+where $\ell_\mathrm{B} = e^2/(4\pi\epsilon k_\mathrm{B}T)$ is the Bjerrum length with $e$ the elementary charge, $\epsilon$ the dielectric constant of the solvent.
+In water $\ell_\mathrm{B} \approx \SI{0.7}{\nano\metre}$.
+Since thermal blobs fill the volume of the electrostatic blob $D = \xi_T \left(g_e/g_T\right)^{1/3}$, we obtain
+\begin{equation}
+g_e = \frac{A^2}{u g_\mathrm{K}} \tau, \qquad D = b\left(\frac{A^2}{u g_\mathrm{K}^2}\right)^{1/3},\qquad \text{with }u = \ell_\mathrm{B}/b.
+\label{eq:geD}
+\end{equation}
+
+We define the extension parameter $B$ as the ratio between the length of a chain of $g_e$ monomers fully extended and the actual size of the electrostatic blob: 
+\begin{equation}
+B = \frac{g_e}{g_\mathrm{K}}\frac{b}{D} = \left(\frac{A^2}{u}\right)^{2/3} g_\mathrm{K}^{-1/3}\tau
+\label{eq:B}
+\end{equation}
+
+At constant solvent quality and polymer architecture, $B$ is thus monotonically related to the amount of counterion condensation. In the following we will estimate $B$ from rheological measurements.
+
+In dilute solution, the chain is a linear rod of length $L= N b / B g_\mathrm{K}$.
+The overlap concentration is thus $c^* = N/L^3 = g_\mathrm{K}B / b L^2$.
+Following \cite{Dobrynin1995}, for a monomer concentration $c > c^*$ and in absence of salt the screening length is
+\begin{equation}
+r_\mathrm{src} = L \left(\frac{c^*}{c}\right)^{1/2} =  \left(\frac{B g_\mathrm{K}}{cb}\right)^{1/2}.
+\label{eq:rscrNoSalt}
+\end{equation}
+
+In the presence of added salt, electrostatic interactions are screened at the Debye length $r_\mathrm{src} = \left(\ell_\mathrm{B} c_s\right)^{-1/2}$. For arbitrary salt concentration,
+\begin{equation}
+r_\mathrm{src} = \left(\frac{B g_\mathrm{K}}{cb}\right)^{1/2} \left(1+ B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
+\label{eq:rscr}
+\end{equation}
+
+Between $D$ and $r_\mathrm{src}$ the electrostatic blobs are organised in a linear rod containing $g_\mathrm{src}$ monomers.
+The length of this rod is $B$ times shorter than if the electrostatic blobs were extended:
+\begin{equation}
+g_\mathrm{src} = B \frac{r_\mathrm{src}}{b} g_\mathrm{K} = \left(\frac{B^3 g_\mathrm{K}^3}{cb^3}\right)^{1/2} \left(1 + B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
+\label{eq:gscr}
+\end{equation}
+
+The last length scale is the correlation length $\xi$.
+Between $r_\mathrm{src}$ and $\xi$ the chain forms a self avoiding walk of persistence length $r_\mathrm{src}$.
+Above the correlation length the polyelectrolyte chain forms a random walk of correlation blobs containing $g=c\xi^3$ monomers.
+
+\subsubsection{From modulus to procession length}
+
+If between two crosslinks or entanglements the ``procession'' of polymer chains contains more monomers than $g$ or $g_\mathrm{src}$ it performs a random or self avoiding walk respectively.
+Each chain is thus an entropic spring of constant stiffness $k_\mathrm{B}T$.
+The number density of chains is $c/N$ with $c$ the monomer concentration and $N$ the number of monomers between cross-links or entanglements.
+Therefore, the shear modulus at low strain is given by:
 \begin{equation}
@@
 \begin{table}
-\begin{tabular}{l|SS|SSSSS}
-& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_c$} & {$n$} & \multicolumn{2}{c}{$E_c$} & {$A$} & {$\sqrt{n}$}\\
-&	{\si{\pascal}} &  & & {$k_\mathrm{B}T$} & {\si{\kilo\joule/\mol}} & \\\hline&&&&&\\[-10pt]
-\ce{PPyr+I-}	& 8800	&	0.04	&	1.	&	2e-3\cellcolor{gray!25}	&	0.004\cellcolor{gray!25}	&	1.	&	1.\\
-\ce{PPyr+Br-}	& 9800 	& 	0.016	&	1.05&	3e-4\cellcolor{gray!25}	&	0.001\cellcolor{gray!25}	&	1.1	&	1.\\
-\ce{PPyr+Cl-}	& 270 	&	0.09	&	45	&	8e-3\cellcolor{gray!25}	&	0.02\cellcolor{gray!25}	&	7.	&	6.7\\
-\ce{PIm+I-}	& 78	&	0.16	&	113	&	3e-2\cellcolor{gray!25}	&	0.063\cellcolor{gray!25}	&	9.6	&	10.6\\
-\ce{PIm+Br-}	& 13	&	8.04	&	790	&	64	&	160.	&	3.4\cellcolor{gray!25}	&	28\\
+\sisetup{
+separate-uncertainty,
+%table-format = 1.3(1),
+}
+\begin{tabular}{l|SSS|SSSSSS}
+& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_0$} & {$\gamma_c$} & {$n$} & {$B$} & {$E_c$} & {$E_c$} & {$D$} & {$A$}\\
+&	{\si{\pascal}} & {\%} & {\%} & & & {\si{\kilo\joule\per\mole}}& {$k_\mathrm{B}T$} & {$\ell_B$}&\\\hline&&&&&\\[-10pt]
+\ce{PIm+Br-}	& 13	&	1.1	&	810	&	860	&	930	&	210	&	84	&	13	&	460\\
+\ce{PIm+I-}		& 78	&	1.8	&	18	&	122	&	126	&	80	&	32	&	5	&	103\\
+\ce{PPyr+Cl-}	& 270 	&	2.8	&	12	&	50.6	&	64	&	46	&	19	&	3.4	&	62\\
+\ce{PPyr+Br-}	& 9800 	& 	0.3	&	2	&	1.15	&	2.8	&	5.05	&	2.02	&	0.7	&	6\\
+\ce{PPyr+I-}	& 8800	&	0.0	&	89	&	1.08	&	2.5	&	5	&	2	&	0.7	&	5.5\\
 \end{tabular}
-\caption{Summary of rheological measurements and microscopic values deduced from the model.
-Gray background indicates inconsistent values obtained either by (Eq.~\ref{eq:Ec}) or by (Eq.~\ref{eq:A}) outside of their respective validity domain.}
+\caption{Summary of rheological measurements and microscopic values deduced from the model. $E_c$, $D$ and $A$ are obtained supposing $\tau=0.46$.}
 \label{tab:results}
@@
 
-\subsubsection{Bounding energy}
+\subsubsection{Limit of the linear regime}
 
-Considering that between two cross-links a procession is an entropic spring in its linear domain, the mechanical energy stored under a strain $\gamma$ is $E = \gamma^2 k_\mathrm{B}T$.
-If we suppose that ionic bonds break before the complete extension of the procession, we evaluate the head-to-body bonding energy to
+As the material is strained, the random walk is stretched, then the self-avoiding walk.
+When the correlation blobs are fully stretched, the chain is a linear assembly of electrostatic blobs.
+Stretching the chain further means exposing more thermal blobs to the solvent, a process that cannot be reduced to spring of constant stiffness.
+This condition is thus the limit between the linear and non linear regimes.
+
+For all our samples, the linear regime is narrow $\gamma_0\approx 10^{-2}\ll 1$ if existing.
+It means that procession is just long enough to start performing the self-avoiding walk.
+Incidentally, this implies that the procession is too short to perform the random walk.
+In other words the number $N$ of monomers between crosslinks is comparable to $g_\mathrm{scr}$.
+Using equation (\ref{eq:gscr}), this condition translates into
 \begin{equation}
-E_c = \gamma_c^2 k_\mathrm{B}T.
-\label{eq:Ec}
+\left(\frac{B}{B_0}\right)^3 = 1 + \frac{B}{B_s}
+\label{eq:solveB}
+\end{equation}
+with $B_s = c/u c_s g_\mathrm{K}$ and $B_0 = b c^{1/3} N^{2/3} g_\mathrm{K}^{-1} = (bc/g_\mathrm{K}) \left(k_\mathrm{B}T/G\right)^{2/3}$.
+Here, the ``salt'' consists in the phosphonate heads, that is to say two charges per chain.
+Thus $c_s/c = 2/N_0$ and $B_s = N_0/2u g_\mathrm{K} \approx 12$.
+
+Analysing Equation (\ref{eq:solveB}), we identify two physically relevant limit cases: (i) if $B \ll B_s$ (low salt condition) then $B = B_0$; (ii) if $B \gg B_s$ (high salt condition) then $B = \left(B_0^3/B_s\right)^{1/2} > B_0$.
+
+Numerical solutions of Equation (\ref{eq:solveB}) are given in Table~\ref{tab:results}.
+For the same quality of solvent $B$ is monotonically related to the amount of counterion condensation.
+We thus confirm that our samples are sorted by decreasing counterion condensation (decreasing $B$).
+More quantitatively, the number $A$ of monomers between dissociated charges is
+\begin{equation}
+A = u^{1/2} \left(\frac{B}{\tau}\right)^{3/4} g_\mathrm{K}^{1/4}
+\label{eq:A}
+\end{equation}.
+
+We know that $\Theta>\SI{100}{\celsius}$ so $0.2<\tau<1$. In the following we will deduce a more precise measure of $\tau$ and thus values of $A$.
+
+%For the theory to be correct, we need the electrostatic length to be larger than the thermal length, in other words $g_e \geq g_T$.
+Using Equations (\ref{eq:geD}) and (\ref{eq:B}), we express $g_e$ function of $B$ and $\tau$:
+%\begin{equation}
+%g_e = \left(\frac{B^3}{g_\mathrm{K}\tau}\right)^{1/2},
+%\end{equation}
+%which yields the condition $\tau \geq \frac{g_\mathrm{K}}{B}$.
+This condition it the most stringent for PPyrI (smallest $B$) and yields $\tau \geq 0.58$.
+In the following we set $\tau$ to this minimal value that corresponds to $\Theta=\SI{450}{\celsius}$.
+%
+%The values we compute for $A$ using Equation (\ref{eq:A}) are shown in Table~\ref{tab:results}.
+$A$ varies between $4.6$ for \ce{PPyr+I-}, corresponding to $15$ uncondensed counterions per polymer, to $383$ for \ce{PIm+Br-} indicating less than a single dissociated counterion every 5 polymers.
+
+\subsubsection{Extension of the electrostatic blobs}
+
+The fully extended self-avoiding walk is a cylinder of length $R_0 = Nb/Bg_\mathrm{K}$, of diameter $D$, of volume $V = R_0 D^2$ and of area $\Sigma_0 = R_0 D = (R_0 V)^{1/2}$.
+Extending the cylinder to $R = (\gamma+1) R_0$ while keeping the volume $V$ constant thus creates an extra area
+$
+\Delta\Sigma = \left(\left(\gamma +1\right)^{1/2} -1\right) \Sigma_0
+$
+that translates into a stretching energy due to the surface tension $k_\mathrm{B}T/\xi_T^2$:
+\begin{equation}
+\frac{E_\mathrm{stretch}}{k_\mathrm{B}T} = \left(\left(\gamma +1\right)^{1/2} -1\right) N \left(\frac{\tau^3}{B g_\mathrm{K}^5}\right)^{1/2}.
+\label{eq:stretch}
 \end{equation}
 
-For \ce{PIm+Br-} we measure $\gamma_c = 800\%$ and thus $E_c \approx 65 k_\mathrm{B}T$ per head-to-body bond or \SI{160}{\kilo\joule\per\mol}.
-This value is much higher than the $E_i\approx\SI{5}{\kilo\joule\per\mol}$ typical of ionic bonds in water\cite{Schneider1992}, even when adding up the two bonds of the bivalent phosphonate head\bibnote{Since the two minus charges are borne by the same atom, chelation effect may enhance bounding energy, but not by an order of magnitude}.
-However, one has to keep in mind that the chain is in bad solvent, thus forming locally compact blobs of low water content.
-Therefore, the medium surrounding the ionic bond is not water but the hydrophobic chains.
+If we consider that the head-to-body bonds break at $\gamma_c$, we can express the reduced temperature as
+\begin{equation}
+\tau = B^{1/3} g_\mathrm{K}^{5/3} \left(\frac{1}{(\gamma_c +1)^{1/2} -1}\frac{E_c}{N k_\mathrm{B}T}\right)^{2/3}
+\label{eq:tau}
+\end{equation}
+where $E_c$ is the energy of the head-to-body bond that consists in two ionic bonds.
+In water the bonding energy between two ions is typically \SI{5}{\kilo\joule\per\mole}\cite{Schneider1992}, thus $E_c\approx 2 k_\mathrm{B}T$.
+However the medium surrounding the ionic bond is not water.
+One has to keep in mind that the chain is in bad solvent, thus electrostatic blobs have low water content.
 Such a low polarity microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}.
-Quantitatively, the bonding energy of an ionic bond is inversely proportional to the relative dielectric constant $\epsilon_r$, therefore we need to suppose a local $\epsilon_r$ around 5 (between 2-3 typical of water-insoluble polymers and 80 for water) to recover the large bonding energy that we measure.
 
-However for all other samples the critical strain is between 2\% and 20\%, leading to inconsistent values for the bonding energies ranging between $10^{-3}k_\mathrm{B}T$ and $10^{-1}k_\mathrm{B}T$ per bond, see Table~\ref{tab:results}.
-Therefore, the end of the linear regime cannot correspond to bond breaking but rather to the point where extended chains become nonlinear.
-
-\subsubsection{From critical strain to charge dissociation}
-
-The spatial conformation of a polyelectrolyte chain in semi-dilute, low salt solution is a persistent random walk where the persistence length is the screening length $r_\mathrm{src}$ due to the dissociated counter ions\cite{Rubinstein1996}.
-In bad solvent, the persistent random walk is made of electrostatic blobs of diameter $D$ containing $g_e$ monomers, see Figure~\ref{fig:dissociation}a:
-\begin{align}
-g_e &= \frac{A^2}{u}\tau\label{eq:ge},\\
-D &= b \left(\frac{A^2}{u}\right)^{1/3}\label{eq:D}.
-\end{align}
-Here $A$ is the number of monomers between dissociated ion pairs, $\tau = 1 - T/\Theta$ the reduced temperature and $u = \ell_\mathrm{B}/b \approx 2.7$ is the ratio of the Bjerrum length in water $\ell_\mathrm{B}\approx \SI{0.7}{\nano\metre}$ and monomer size $b\approx \SI{0.26}{\nano\metre}$.
-
-Extending these electrostatic blobs means exposing more monomers to the bad solvent and typically requires energies at least an order of magnitude higher than for the extension of the persistent random walk.
-We will thus consider that the exit from the linear domain of the material corresponds to the full extension of the persistent random walk:
+Using Equations~(\ref{eq:geD}) and (\ref{eq:B}) we express the size $D$ of electrostatic blobs function of $B$ and $\tau$:
 \begin{equation}
-\gamma_c = \frac{R_0}{R} - 1
-\end{equation}
-where $R$ and $R_0$ are respectively the end-to-end distance and the contour length of the persistent random walk.
-
-A segment of the persistent random walk is $r_\mathrm{scr}$ long and thus contains $g_\mathrm{scr} = g_e r_\mathrm{scr}/D$ monomers. We thus have
-\begin{equation}
-\left. \begin{array}{ll}
-R &= r_\mathrm{scr} \left(\frac{N}{g_\mathrm{scr}}\right)^{1/2}\\
-R_0 &= r_\mathrm{scr} \frac{N}{g_\mathrm{scr}}
-\end{array}\right\rbrace\quad
-\gamma_c = \left(\frac{N}{g_\mathrm{scr}}\right)^{1/2} -1\label{eq:gamma0}.
+g_e = \left(\frac{B^3}{g_\mathrm{K}\tau}\right)^{1/2},\qquad 
+D = b\left(\frac{B}{g_\mathrm{K}^3\tau}\right)^{1/2}.
+\label{eq:DB}
 \end{equation}
 
-\citet{Rubinstein1996} express the screening length as
-$r_\mathrm{scr} = \left(A^2 / u\right)^{1/3} \left(\tau /cb \right)^{1/2}$. Combining with (Eq.~\ref{eq:ge}) and (Eq.~\ref{eq:D}) we deduce
-\begin{equation}
-g_\mathrm{scr} = \frac{A^2}{u} \left(\frac{\tau^3}{c b^3}\right)^{1/2}\label{eq:gscr}.
-\end{equation}
+We thus expect stronger head-to-body bonds when $B$ is large and the low-polarity environment is extended ($D\gg\ell_\mathrm{B}$).
+A contrario, the weakest ionic links, closest to their strength in water, should be found in \ce{PPyr+I-} where $B$ is minimum (hopefully $D<\ell_\mathrm{B}$).
 
-From (Eq.~\ref{eq:gamma0}) and (Eq.~\ref{eq:gscr}) we can finally relate the measured critical strain to the microscopic parameter $A$:
-\begin{equation}
-A = \frac{\left(N u\right)^{1/2}}{\gamma_c+1}\left(\frac{c b^3}{\tau^3}\right)^{1/4}.
-\label{eq:A}
-\end{equation}
+Applying Equation~(\ref{eq:tau}) to \ce{PPyr+I-} with $E_c = 2 k_\mathrm{B}T$ yields $\tau \approx 0.46\pm 0.1$.
+Using this value and Equation (\ref{eq:stretch}) we deduce the bounding energy for each composition, as reported in Table~\ref{tab:results}.
+Indeed, the bonding energy increases with increasing counterion condensation to reach 60\% of the carbon-carbon single bond (\SI{350}{\kilo\joule\per\mole} or $140 k_\mathrm{B}T$) in the case of \ce{PIm+Br-}.
+According to Equation (\ref{eq:DB}), in this extreme case an electrostatic blob contains $\approx 500$ polymers which embed the head-to-body bonds in a low-polarity environment 13 times larger than the Bjerrum length.
+Conversely, for \ce{PPyr+I-} and \ce{PPyr+Br-} electrostatic blobs are smaller than $\ell_\mathrm{B}$, confirming an aqueous microenvironment.
+Quantitatively, the bonding energy of an ionic bond is inversely proportional to the relative dielectric constant $\epsilon_r$, therefore we need to suppose a local $\epsilon_r$ around 2 (typical of water-insoluble polymers) instead of 80 for water to recover the large bonding energy of \ce{PIm+Br-}.
 
-The only unknown here is the reduced temperature $\tau$, possibly different between poly\-(pyrrolidinium) based and poly\-(immidazolium) based polymers.
-We know that the undissociated chain has a $\Theta$ temperature higher than \SI{100}{\celsius}, thus $0.2<\tau<1$.
-In addition, we must have $A\geq 1$, which implies that $\tau\leq 0.235$ through (Eq.~\ref{eq:A}), corresponding to $\Theta\approx\SI{116}{\celsius}$.
-This limit value yields $A=1$ for both \ce{PPyr+I-} and \ce{PPyr+Br-}, indicating full dissociation of the cationic moieties.
-By contrast \ce{PPyr+Cl-} has a charge every 7 monomers.
-Supposing the same $\tau$, \ce{PIm+I-} has a dissociated ion pair every 10 monomers, see Table~\ref{tab:results}.
-Of course analysing \ce{PIm+Br-} based on the value of $\gamma_c$ gives an inconsistent value $A=3.4$ indicating that this hydrogel does not belong to the same regime.
+Table~\ref{tab:results} also shows the values of $A$ deduced from Equation~(\ref{eq:A}) and the measured value of $\tau$.
+$A$ varies between $5.5$ for \ce{PPyr+I-}, corresponding to $13$ uncondensed counterions per polymer, to $460$ for \ce{PIm+Br-} indicating a single dissociated counterion every 7 polymers.
 
-\subsubsection{Charge dissociation and cross-linking ratio}
-
-Finally we cross-check our model by linking the ion pair dissociation and the amount of physical cross-links by an independent argument.
-A second charge dissociated on a chain creates a possible binding site for a second head-to-body association and thus a cross-link.
-Therefore the probability to have a cross link is proportional to the probability of having a second charge dissociation.
-If we assume that the probability for an ion pair to dissociate is independent of the number of dissociated pairs already present on the polymer, we can describe the probability of having $k$ dissociated pairs on a polymer by a Poisson distribution of mean $\lambda = N_0/A$.
-Therefore, the cross-link probability $1/n$ is proportional to $\lambda^2$, yielding $n \sim A^2$.
-
-As shown in Figure~\ref{fig:dissociation}b and the last two columns of Table~\ref{tab:results} this relation describes very well our data with a prefactor close to 1.
-We can extrapolate this law for the case of \ce{PIm+Br-} and obtain $A = 28$.
-Inverting (Eq.~\ref{eq:A}) yields $R_0/R - 1 = 860\%$, indeed higher than the observed $\gamma_c\approx 804\%$.
-This confirms that in the \ce{PIm+Br-} strain sweep $\gamma_c$ actually results from the strength of the ionic bonds.
-
-%We can now estimate the hydrophobic interaction energy $E_h$ by considering that to separate two chains two interfaces with water of area $\pi D^2/4$ have to be created.
-Surface tension is $\tau^2 k_\mathrm{B}T / b^2$\cite{Rubinstein1996}.
-Combining with (Eq.~\ref{eq:D}) we find
-%\begin{equation}
-%\frac{E_i}{k_\mathrm{B}T} = \frac{\pi}{2}\tau^2\left(\frac{A^2}{u}\right)^{2/3} \approx 4
-%\end{equation}
-
-%\subsection{Discussion}
-%
-%The above dependence of the gel parameters on the nature of the counterions could be rationalized on the basis of ion-pairing processes.
-The unsolubility/unswellability of \ce{PIm+F-}, \ce{PPyr+F-} and \ce{PIm+Cl-} suggest that ion pairs are not efficiently dissociated in aqueous medium, thus hindering the solvation process.
-By comparison, \ce{PIm+Br-}, \ce{PIm+I-}, \ce{PPyr+Br-} and \ce{PPyr+I-} form homogeneous materials when suspended in water, which one can correlate with efficient dissociation of the halide counterions from the polymer main chain is fully consistency with decreased ionic bond strength.
-Such dissociation, in turns makes it possible for the ``hard'' phosphonate group on each polymer chain's end to interact with the ``free'' cationic groups of the side chains, which we showed in a previous paper to be at the origin of the gel formation mechanism\cite{Srour2014}.
-%
-%The similarity in gel parameters observed for \ce{PPyr+Br-} and \ce{PPyr+I-} ($G^\prime\approx\SI{8500}{\pascal}$) suggests that full ion pair dissociation is achieved in the latter, or at least that this dissociation is enough in both cases to provide maximal interaction between the terminal phosphonate and the pyrrolidinium side groups.
-Conversely the significantly different behavior between \ce{PIm+I-} ($G^\prime= \SI{78}{\pascal}$) and \ce{PIm+Br-} ($G^\prime= \SI{13}{\pascal}$) indicates that dissociation is only partial and occurs to a lesser extent in the latter than in the former.
-The same conclusion can be drawn for \ce{PPyr+Cl-}, in which the elastic modulus ($G^\prime= \SI{270}{\pascal}$) value drops by more than one order of magnitude as compared to its bromide and iodide counterparts.
-%
-%As already mentioned (vide supra) another striking feature of these gels is that in similar conditions poly(pyrrolidinium) based gels are systematically much stronger than those formed with poly(imidazolium).
-Indeed, although both cationic in nature, pyrrolidinium and imidazolium ions have markedly different properties: whereas pyrrolidinium features a ``hard'' positive charge very localized on the quaternary nitrogen, imidazolium has a much more delocalized ``softer'' charge; consequently, according to the Hard and Soft Acid and Bases (HSAB) theory, stronger association is expected for the latter with the ``soft'' halide counterions\cite{Goossens2009}.
-As a consequence, for a given halide, dissociation is less favored in the case of poly(imidazolium), as clearly illustrated by the difference in solvation between \ce{PIm+Cl-} and \ce{PPyr+Cl-}.
-Conversely, association of the lateral cations with the ``hard'' phosphonate moieties at the polymer chain’s end is expected to be stronger with pyrrolidinium, hence a systematically larger elastic modulus.
-The interplay of these two effects contributes in making poly(pyrrolidinium) based gels much stronger than their poly(imidazolium) counterparts; comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ value (13 and \SI{8100}{\pascal}, resp.) differ by almost three orders of magnitude.
+From purely rheological measurements, we have thus obtained microscopic information on the state of the gel: amount of charge condensation, head-to-body bounding energy and size of a procession.
 
@@
 This gives rise to maximal charge dissociation in water, facilitates charge pairing of the peripheral, iterative cations with the terminal anion and thus affords a very high density of cross-links.
-In between, in the case of aliphatic cations with a counterion in the middle of the Hofmeister series (\ce{Cl-}) or in the case of aromatic cations and poorly interacting anion, charge dissociation is limited, polymers associate in processions and cross-links are few.
+In between, in the case of aliphatic cations with a counterion in the middle of the Hofmeister series (\ce{Cl-}) or in the case of aromatic cations and poorly interacting anion, charge dissociation is limited, polymers associate in long processions with strong head-to-body bonds.
 We thus obtain soft, highly deformable gels.
+Coincidentally, the mesh size of our gels is always close to the procession persistence length, a regime often encountered in networks of biological semiflexible filaments as collagen or actin\cite{Meng2016}.
 

### 2016-05-02 ion_pairing_paper.tex
Simpler figure to describe procession

@@
 \begin{figure}
-\includegraphics{dissociation.pdf}
+\includegraphics{procession.pdf}
 \caption{Processionary model.
-(a) Sketch of the network in the case of a procession of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{src}\approx 3D$.
+Sketch of the network in the case of a procession of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{src}\approx 3D$.
 Empty circles are not individual monomers but electrostatic blobs of size $D$.
-Electrostatic blobs containing an anionic head are shown as black filled circles.
-(b) Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points.
-The solid line corresponds to $n = A^2$.
-Full circles are obtained from (Eq.~\ref{eq:A}).
-In the case of \ce{PIm+Br-}, we are not in the validity regime of (Eq.~\ref{eq:A}), and we have to extrapolate the solid line to estimate $A$ (dotted arrow and open symbol).}
+Electrostatic blobs containing an anionic head are shown as black filled circles.}
 \label{fig:dissociation}
@@
 }
-\begin{tabular}{l|SSS|SSSSS}
-& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_0$} & {$\gamma_c$} & {$n$} & {$B$} & {$A$} & {$E_c$} & {$E_c$}\\
-&	{\si{\pascal}} & {\%} & {\%} & & & & {\si{\kilo\joule\per\mole}}& {$k_\mathrm{B}T$}\\\hline&&&&&\\[-10pt]
-\ce{PIm+Br-}	& 13	&	1.1	&	810	&	860	&	930	&	383	&	300	&	120\\
-\ce{PIm+I-}		& 78	&	1.8	&	18	&	122	&	126	&	85	&	116	&	47\\
-\ce{PPyr+Cl-}	& 270 	&	2.8	&	12	&	50.6	&	64	&	52	&	67	&	27\\
-\ce{PPyr+Br-}	& 9800 	& 	0.3	&	2	&	1.15	&	2.8	&	5	&	7.3	&	2.9\\
-\ce{PPyr+I-}	& 8800	&	0.0	&	89	&	1.08	&	2.5	&	4.6	&	7.2	&	2.9\\
+\begin{tabular}{@{}l|SSS|SSSSSS@{}}
+& {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_0$} & {$\gamma_c$} & {$n$} & {$B$} & {$E_c$} & {$E_c$} & {$D$} & {$A$}\\
+&	{\si{\pascal}} & {\%} & {\%} & & & {\si{\kilo\joule\per\mole}}& {$k_\mathrm{B}T$} & {$\ell_B$}&\\\hline&&&&&\\[-10pt]
+\ce{PIm+Br-}	& 13	&	1.1	&	810	&	860	&	930	&	210	&	84	&	13	&	460\\
+\ce{PIm+I-}		& 78	&	1.8	&	18	&	122	&	126	&	80	&	32	&	5	&	103\\
+\ce{PPyr+Cl-}	& 270 	&	2.8	&	12	&	50.6	&	64	&	46	&	19	&	3.4	&	62\\
+\ce{PPyr+Br-}	& 9800 	& 	0.3	&	2	&	1.15	&	2.8	&	5.05	&	2.02	&	0.7	&	6\\
+\ce{PPyr+I-}	& 8800	&	0.0	&	89	&	1.08	&	2.5	&	5	&	2	&	0.7	&	5.5\\
 \end{tabular}
-\caption{Summary of rheological measurements and microscopic values deduced from the model. $A$ and $E_c$ are obtained supposing $\tau=0.588$.}
+\caption{Summary of rheological measurements and microscopic values deduced from the model. $E_c$, $D$ and $A$ are obtained supposing $\tau=0.46$.}
 \label{tab:results}
@@
 
-For the theory to be correct, we need the electrostatic length to be larger than the thermal length, in other words $g_e \geq g_T$.
+We know that $\Theta>\SI{100}{\celsius}$ so $0.2<\tau<1$. In the following we will deduce a more precise measure of $\tau$ and thus values of $A$.
+
+%For the theory to be correct, we need the electrostatic length to be larger than the thermal length, in other words $g_e \geq g_T$.
 Using Equations (\ref{eq:geD}) and (\ref{eq:B}), we express $g_e$ function of $B$ and $\tau$:
-\begin{equation}
-g_e = \left(\frac{B^3}{g_\mathrm{K}\tau}\right)^{1/2},
-\end{equation}
-which yields the condition $\tau \geq \frac{g_\mathrm{K}}{B}$.
+%\begin{equation}
+%g_e = \left(\frac{B^3}{g_\mathrm{K}\tau}\right)^{1/2},
+%\end{equation}
+%which yields the condition $\tau \geq \frac{g_\mathrm{K}}{B}$.
 This condition it the most stringent for PPyrI (smallest $B$) and yields $\tau \geq 0.58$.
 In the following we set $\tau$ to this minimal value that corresponds to $\Theta=\SI{450}{\celsius}$.
-
-The values we compute for $A$ using Equation (\ref{eq:A}) are shown in Table~\ref{tab:results}.
+%
+%The values we compute for $A$ using Equation (\ref{eq:A}) are shown in Table~\ref{tab:results}.
 $A$ varies between $4.6$ for \ce{PPyr+I-}, corresponding to $15$ uncondensed counterions per polymer, to $383$ for \ce{PIm+Br-} indicating less than a single dissociated counterion every 5 polymers.
 
-\subsubsection{Extention of the electrostatic blobs}
+\subsubsection{Extension of the electrostatic blobs}
 
@@
 
-If we consider that the head-to-body bonds break at $\gamma_c$, we can extract their bonding energy $E_c$ from Equation (\ref{eq:stretch}).
-The results are reported in Table~\ref{tab:results}.
-In the most dissociated \ce{PPyr+I-} and \ce{PPyr+Br-}, the head-to-body bonding energy $E_c$ is compatible with typical ionic bonds, \SI{5}{\kilo\joule\per\mole} in water~\cite{Schneider1992}.
-However the bonding energy increases with increasing counterion condensation to reach 85\% of the carbon-carbon single bond (\SI{350}{\kilo\joule\per\mole} or $140 k_\mathrm{B}T$) in the case of \ce{PIm+Br-}.
-Indeed, at high counterion condensation, the electrostatic blobs are large, up to 530 polymers for \ce{PIm+Br-}, and embed head-to-body bonds into compact blobs of low water content.
-Such a low polarity microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly~\cite{Rehm2010}.
+If we consider that the head-to-body bonds break at $\gamma_c$, we can express the reduced temperature as
+\begin{equation}
+\tau = B^{1/3} g_\mathrm{K}^{5/3} \left(\frac{1}{(\gamma_c +1)^{1/2} -1}\frac{E_c}{N k_\mathrm{B}T}\right)^{2/3}
+\label{eq:tau}
+\end{equation}
+where $E_c$ is the energy of the head-to-body bond that consists in two ionic bonds.
+In water the bonding energy between two ions is typically \SI{5}{\kilo\joule\per\mole}\cite{Schneider1992}, thus $E_c\approx 2 k_\mathrm{B}T$.
+However the medium surrounding the ionic bond is not water.
+One has to keep in mind that the chain is in bad solvent, thus electrostatic blobs have low water content.
+Such a low polarity microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}.
+
+Using Equations~(\ref{eq:geD}) and (\ref{eq:B}) we express the size $D$ of electrostatic blobs function of $B$ and $\tau$:
+\begin{equation}
+g_e = \left(\frac{B^3}{g_\mathrm{K}\tau}\right)^{1/2},\qquad 
+D = b\left(\frac{B}{g_\mathrm{K}^3\tau}\right)^{1/2}.
+\label{eq:DB}
+\end{equation}
+
+We thus expect stronger head-to-body bonds when $B$ is large and the low-polarity environment is extended ($D\gg\ell_\mathrm{B}$).
+A contrario, the weakest ionic links, closest to their strength in water, should be found in \ce{PPyr+I-} where $B$ is minimum (hopefully $D<\ell_\mathrm{B}$).
+
+Applying Equation~(\ref{eq:tau}) to \ce{PPyr+I-} with $E_c = 2 k_\mathrm{B}T$ yields $\tau \approx 0.46\pm 0.1$.
+Using this value and Equation (\ref{eq:stretch}) we deduce the bounding energy for each composition, as reported in Table~\ref{tab:results}.
+Indeed, the bonding energy increases with increasing counterion condensation to reach 60\% of the carbon-carbon single bond (\SI{350}{\kilo\joule\per\mole} or $140 k_\mathrm{B}T$) in the case of \ce{PIm+Br-}.
+According to Equation (\ref{eq:DB}), in this extreme case an electrostatic blob contains $\approx 500$ polymers which embed the head-to-body bonds in a low-polarity environment 13 times larger than the Bjerrum length.
+Conversely, for \ce{PPyr+I-} and \ce{PPyr+Br-} electrostatic blobs are smaller than $\ell_\mathrm{B}$, confirming an aqueous microenvironment.
+Quantitatively, the bonding energy of an ionic bond is inversely proportional to the relative dielectric constant $\epsilon_r$, therefore we need to suppose a local $\epsilon_r$ around 2 (typical of water-insoluble polymers) instead of 80 for water to recover the large bonding energy of \ce{PIm+Br-}.
+
+Table~\ref{tab:results} also shows the values of $A$ deduced from Equation~(\ref{eq:A}) and the measured value of $\tau$.
+$A$ varies between $5.5$ for \ce{PPyr+I-}, corresponding to $13$ uncondensed counterions per polymer, to $460$ for \ce{PIm+Br-} indicating a single dissociated counterion every 7 polymers.
+
+From purely rheological measurements, we have thus obtained microscopic information on the state of the gel: amount of charge condensation, head-to-body bounding energy and size of a procession.
 

### 2016-05-10 PALSE/ion_pairing_paper.tex
insert fig:phasediag bad -> poor solvent

@@
 Locally, the monomers are condensed in collapsed globules to minimize the contacts with water molecules, and solubility on larger scales can only be achieved if the polymer bears enough charges\cite{Khokhlov1980,Raphael1990}.
-To estimate the $\Theta$-temperature, that delimit bad and good solvent behaviour, we held the two fluoride polymers in water at boiling temperature overnight.
+To estimate the $\Theta$-temperature, that delimit poor and good solvent behaviour, we held the two fluoride polymers in water at boiling temperature overnight.
 No dissolution or swelling was observed, implying that $\Theta>\SI{100}{\celsius}$.
@@
 \begin{figure}
-\includegraphics{dissociation.pdf}
+\includegraphics{procession.pdf}
 \caption{Processionary model.
-(a) Sketch of the network in the case of a procession of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{src}\approx 3D$.
+Sketch of the network in the case of a procession of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{src}\approx 6D$.
 Empty circles are not individual monomers but electrostatic blobs of size $D$.
-Electrostatic blobs containing an anionic head are shown as black filled circles.
-(b) Relation between the number of monomers between uncondensed charges and the number of chains between cross-link points.
-The solid line corresponds to $n = A^2$.
-Full circles are obtained from (Eq.~\ref{eq:A}).
-In the case of \ce{PIm+Br-}, we are not in the validity regime of (Eq.~\ref{eq:A}), and we have to extrapolate the solid line to estimate $A$ (dotted arrow and open symbol).}
+Electrostatic blobs containing an anionic head are shown as black filled circles.}
 \label{fig:dissociation}
@@
 \subsubsection{Conformation at rest}
+\begin{figure}
+\includegraphics{phase_diag.pdf}
+\caption{Conformation of semi-dilute polyelectrolytes at rest at various scales function of dissociated charge frequency.
+The phase diagram is given for $\tau=0.46$ and for the monomer number density of \ce{PIm+Br-}, although $c$ is slightly different for each sample.
+Each vertical line represents a sample.}
+\end{figure}
 
@@
 
-In bad solvent, the thermal length $\xi_T$ is larger than $b$.
+In poor solvent, the thermal length $\xi_T$ is larger than $b$.
 The thermal blob contains $g_T$ monomers and its density is set by the reduced temperature $(g_T/g_\mathrm{K})/\xi_T^3 = \tau/b^3$.
@@
 In dilute solution, the chain is a linear rod of length $L= N b / B g_\mathrm{K}$.
-The overlap concentration is thus $c^* = N/L^3 = g_\mathrm{K}B / b L^2$.
+The overlap concentration is thus $c^* = N/L^3 = B g_\mathrm{K} / b L^2$.
 Following \cite{Dobrynin1995}, for a monomer concentration $c > c^*$ and in absence of salt the screening length is
@@
 }
-\begin{tabular}{l|SSS|SSSSSS}
+\begin{tabular}{@{}l|SSS|SSSSSS@{}}
 & {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_0$} & {$\gamma_c$} & {$n$} & {$B$} & {$E_c$} & {$E_c$} & {$D$} & {$A$}\\
@@
 However the medium surrounding the ionic bond is not water.
-One has to keep in mind that the chain is in bad solvent, thus electrostatic blobs have low water content.
+One has to keep in mind that the chain is in poor solvent, thus electrostatic blobs have low water content.
 Such a low polarity microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}.

### 2016-05-11 PALSE/ion_pairing_paper.tex
Fix the conversion from B to A and propagate

@@
 \caption{Processionary model.
-Sketch of the network in the case of a procession of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{src}\approx 6D$.
+Sketch of the network in the case of a procession of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{scr}\approx 6D$.
 Empty circles are not individual monomers but electrostatic blobs of size $D$.
@@
 Between $b$ and $\xi_T$ we have a persistent random walk $\xi_T = b (g_T/g_\mathrm{K})^{1/2}$.
-These conditions yields $\xi_T = b/\tau$ and $g_T = g_\mathrm{K}/\tau^2$.
+These conditions yields
+\begin{equation}
+\xi_T = b/\tau,\qquad g_T = g_\mathrm{K}/\tau^2
+\label{eq:thermal}
+\end{equation}
 
@@
 \begin{equation}
-B = \frac{g_e}{g_\mathrm{K}}\frac{b}{D} = \left(\frac{A^2}{u}\right)^{2/3} g_\mathrm{K}^{-1/3}\tau
+B = \frac{g_e}{g_\mathrm{K}}\frac{b}{D} = \left(\frac{A^2}{u g_\mathrm{K}^2}\right)^{2/3} \tau
 \label{eq:B}
@@
 
-At constant solvent quality and polymer architecture, $B$ is thus monotonically related to the amount of counterion condensation. In the following we will estimate $B$ from rheological measurements.
+At constant solvent quality and polymer architecture, $B$ is thus monotonically related to the amount of counterion condensation.
+In the following we will use $B$ as our main variable and estimate it from rheological measurements.
+We conveniently combine Equations~(\ref{eq:geD}) and (\ref{eq:B}) to express $g_e$ and $D$ function of $B$:
+\begin{equation}
+g_e = \left(\frac{B^3}{\tau}\right)^{1/2} g_\mathrm{K},\qquad D = b \left(\frac{B}{\tau}\right)^{1/2}.
+\label{eq:geDfromB}
+\end{equation}
 
@@
 \begin{equation}
-r_\mathrm{src} = L \left(\frac{c^*}{c}\right)^{1/2} =  \left(\frac{B g_\mathrm{K}}{cb}\right)^{1/2}.
+r_\mathrm{scr} = L \left(\frac{c^*}{c}\right)^{1/2} =  \left(\frac{B g_\mathrm{K}}{cb}\right)^{1/2}.
 \label{eq:rscrNoSalt}
@@
 
-In the presence of added salt, electrostatic interactions are screened at the Debye length $r_\mathrm{src} = \left(\ell_\mathrm{B} c_s\right)^{-1/2}$. For arbitrary salt concentration,
+In the presence of added salt, electrostatic interactions are screened at the Debye length $r_\mathrm{scr} = \left(\ell_\mathrm{B} c_s\right)^{-1/2}$. For arbitrary salt concentration,
 \begin{equation}
-r_\mathrm{src} = \left(\frac{B g_\mathrm{K}}{cb}\right)^{1/2} \left(1+ B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
+r_\mathrm{scr} = \left(\frac{B g_\mathrm{K}}{cb}\right)^{1/2} \left(1+ B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
 \label{eq:rscr}
@@
 
-Between $D$ and $r_\mathrm{src}$ the electrostatic blobs are organised in a linear rod containing $g_\mathrm{src}$ monomers.
+Between $D$ and $r_\mathrm{scr}$ the electrostatic blobs are organised in a linear rod containing $g_\mathrm{scr}$ monomers.
 The length of this rod is $B$ times shorter than if the electrostatic blobs were extended:
 \begin{equation}
-g_\mathrm{src} = B \frac{r_\mathrm{src}}{b} g_\mathrm{K} = \left(\frac{B^3 g_\mathrm{K}^3}{cb^3}\right)^{1/2} \left(1 + B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
+g_\mathrm{scr} = B \frac{r_\mathrm{scr}}{b} g_\mathrm{K} = \left(\frac{B^3 g_\mathrm{K}^3}{cb^3}\right)^{1/2} \left(1 + B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
 \label{eq:gscr}
@@
 The last length scale is the correlation length $\xi$.
-Between $r_\mathrm{src}$ and $\xi$ the chain forms a self avoiding walk of persistence length $r_\mathrm{src}$.
+Between $r_\mathrm{scr}$ and $\xi$ the chain forms a self avoiding walk of persistence length $r_\mathrm{scr}$.
 Above the correlation length the polyelectrolyte chain forms a random walk of correlation blobs containing $g=c\xi^3$ monomers.
@@
 
-If between two crosslinks or entanglements the ``procession'' of polymer chains contains more monomers than $g$ or $g_\mathrm{src}$ it performs a random or self avoiding walk respectively.
+If between two crosslinks or entanglements the ``procession'' of polymer chains contains more monomers than $g$ or $g_\mathrm{scr}$ it performs a random or self avoiding walk respectively.
 Each chain is thus an entropic spring of constant stiffness $k_\mathrm{B}T$.
@@
 &	{\si{\pascal}} & {\%} & {\%} & & & {\si{\kilo\joule\per\mole}}& {$k_\mathrm{B}T$} & {$\ell_B$}&\\\hline&&&&&\\[-10pt]
-\ce{PIm+Br-}	& 13	&	1.1	&	810	&	860	&	930	&	210	&	84	&	13	&	460\\
-\ce{PIm+I-}		& 78	&	1.8	&	18	&	122	&	126	&	80	&	32	&	5	&	103\\
-\ce{PPyr+Cl-}	& 270 	&	2.8	&	12	&	50.6	&	64	&	46	&	19	&	3.4	&	62\\
-\ce{PPyr+Br-}	& 9800 	& 	0.3	&	2	&	1.15	&	2.8	&	5.05	&	2.02	&	0.7	&	6\\
-\ce{PPyr+I-}	& 8800	&	0.0	&	89	&	1.08	&	2.5	&	5	&	2	&	0.7	&	5.5\\
+\ce{PIm+Br-}	& 13	&	1.1	&	810	&	860	&	930	&	210	&	84	&	29	&	850\\
+\ce{PIm+I-}		& 78	&	1.8	&	18	&	122	&	126	&	80	&	32	&	10.6	&	189\\
+\ce{PPyr+Cl-}	& 270 	&	2.8	&	12	&	50.6	&	64	&	46	&	19	&	7.6	&	115\\
+\ce{PPyr+Br-}	& 9800 	& 	0.3	&	2	&	1.15	&	2.8	&	5.05	&	2.02	&	1.6	&	11\\
+\ce{PPyr+I-}	& 8800	&	0.0	&	89	&	1.08	&	2.5	&	5	&	2	&	1.5	&	10\\
 \end{tabular}
-\caption{Summary of rheological measurements and microscopic values deduced from the model. $E_c$, $D$ and $A$ are obtained supposing $\tau=0.46$.}
+\caption{Summary of rheological measurements and microscopic values deduced from the model. $E_c$, $D$ and $A$ are obtained supposing $\tau=0.31$.}
 \label{tab:results}
@@
 \begin{equation}
-A = u^{1/2} \left(\frac{B}{\tau}\right)^{3/4} g_\mathrm{K}^{1/4}
+A = u^{1/2} \left(\frac{B}{\tau}\right)^{3/4} g_\mathrm{K}
 \label{eq:A}
@@
 
-The fully extended self-avoiding walk is a cylinder of length $R_0 = Nb/Bg_\mathrm{K}$, of diameter $D$, of volume $V = R_0 D^2$ and of area $\Sigma_0 = R_0 D = (R_0 V)^{1/2}$.
+The fully extended self-avoiding walk is a cylinder of length $R_0 = Nb/(Bg_\mathrm{K})$, of diameter $D$, of volume $V = R_0 D^2$ and of area $\Sigma_0 = R_0 D = (R_0 V)^{1/2}$.
 Extending the cylinder to $R = (\gamma+1) R_0$ while keeping the volume $V$ constant thus creates an extra area
@@
 $
-that translates into a stretching energy due to the surface tension $k_\mathrm{B}T/\xi_T^2$:
+that translates into a stretching energy due to the surface tension $k_\mathrm{B}T/\xi_T^2$, via Equations~(\ref{eq:thermal}) and (\ref{eq:geDfromB}):
 \begin{equation}
-\frac{E_\mathrm{stretch}}{k_\mathrm{B}T} = \left(\left(\gamma +1\right)^{1/2} -1\right) N \left(\frac{\tau^3}{B g_\mathrm{K}^5}\right)^{1/2}.
+\frac{E_\mathrm{stretch}}{k_\mathrm{B}T} = \left(\left(\gamma +1\right)^{1/2} -1\right) \frac{N}{g_\mathrm{K}} \left(\frac{\tau^3}{B}\right)^{1/2}.
 \label{eq:stretch}
@@
 \begin{equation}
-\tau = B^{1/3} g_\mathrm{K}^{5/3} \left(\frac{1}{(\gamma_c +1)^{1/2} -1}\frac{E_c}{N k_\mathrm{B}T}\right)^{2/3}
+\tau = B^{1/3} \left(\frac{1}{(\gamma_c +1)^{1/2} -1}\frac{g_\mathrm{K}}{N}\frac{E_c}{k_\mathrm{B}T}\right)^{2/3}
 \label{eq:tau}
@@
 
-Using Equations~(\ref{eq:geD}) and (\ref{eq:B}) we express the size $D$ of electrostatic blobs function of $B$ and $\tau$:
-\begin{equation}
-g_e = \left(\frac{B^3}{g_\mathrm{K}\tau}\right)^{1/2},\qquad 
-D = b\left(\frac{B}{g_\mathrm{K}^3\tau}\right)^{1/2}.
-\label{eq:DB}
-\end{equation}
+From Equation~(\ref{eq:geDfromB}) we know that $D$ increased with $B$.
+We thus expect stronger head-to-body bonds when $B$ is large and the low-polarity environment is extended ($D\gg\ell_\mathrm{B}$).
+A contrario, the weakest ionic links, closest to their strength in water, should be found in \ce{PPyr+I-} where $B$ is minimum ($D\approx\ell_\mathrm{B}$).
 
-We thus expect stronger head-to-body bonds when $B$ is large and the low-polarity environment is extended ($D\gg\ell_\mathrm{B}$).
-A contrario, the weakest ionic links, closest to their strength in water, should be found in \ce{PPyr+I-} where $B$ is minimum (hopefully $D<\ell_\mathrm{B}$).
-
-Applying Equation~(\ref{eq:tau}) to \ce{PPyr+I-} with $E_c = 2 k_\mathrm{B}T$ yields $\tau \approx 0.46\pm 0.1$.
+Applying Equation~(\ref{eq:tau}) to \ce{PPyr+I-} with $E_c = 2 k_\mathrm{B}T$ yields $\tau \approx 0.31\pm 0.07$, or a $\Theta$-temperature around \SI{160}{\celsius}.
 Using this value and Equation (\ref{eq:stretch}) we deduce the bounding energy for each composition, as reported in Table~\ref{tab:results}.
 Indeed, the bonding energy increases with increasing counterion condensation to reach 60\% of the carbon-carbon single bond (\SI{350}{\kilo\joule\per\mole} or $140 k_\mathrm{B}T$) in the case of \ce{PIm+Br-}.
-According to Equation (\ref{eq:DB}), in this extreme case an electrostatic blob contains $\approx 500$ polymers which embed the head-to-body bonds in a low-polarity environment 13 times larger than the Bjerrum length.
-Conversely, for \ce{PPyr+I-} and \ce{PPyr+Br-} electrostatic blobs are smaller than $\ell_\mathrm{B}$, confirming an aqueous microenvironment.
+According to Equation (\ref{eq:geDfromB}), in this extreme case an electrostatic blob contains $\approx 10^3$ polymers which embed the head-to-body bonds in a low-polarity environment $\approx 30$ times larger than the Bjerrum length.
+Conversely, for \ce{PPyr+I-} and \ce{PPyr+Br-} the size of electrostatic blobs is comparable to $\ell_\mathrm{B}$, confirming an aqueous microenvironment.
 Quantitatively, the bonding energy of an ionic bond is inversely proportional to the relative dielectric constant $\epsilon_r$, therefore we need to suppose a local $\epsilon_r$ around 2 (typical of water-insoluble polymers) instead of 80 for water to recover the large bonding energy of \ce{PIm+Br-}.
@@
 Table~\ref{tab:results} also shows the values of $A$ deduced from Equation~(\ref{eq:A}) and the measured value of $\tau$.
-$A$ varies between $5.5$ for \ce{PPyr+I-}, corresponding to $13$ uncondensed counterions per polymer, to $460$ for \ce{PIm+Br-} indicating a single dissociated counterion every 7 polymers.
+$A$ varies between $10$ for \ce{PPyr+I-}, corresponding to $7$ uncondensed counterions per polymer, to $850$ for \ce{PIm+Br-} indicating a single dissociated counterion every $12$ polymers.
 

### 2016-05-12 PALSE/ion_pairing_paper.tex
Avec corrections de Seb

@@
 \author{Hassan Srour}
-\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}
-
-\author{Mathieu Leocmach}
-\affiliation[Institut Lumière Matière]{Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}
-\email{mathieu.leocmach@univ-lyon1.fr}
-
 \author{Martien Duvall Deffo Ayagou}
@@
 \email{cyrille.monnereau@ens-lyon.fr}
+
+\author{Mathieu Leocmach}
+\affiliation[Institut Lumière Matière]{Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}
+\email{mathieu.leocmach@univ-lyon1.fr}
 
@@
 
-In the present report, we investigate the role of the couterion condensation on the mechanical properties of such hydrogels.
+In the present report, we investigate the role of the counterion condensation on the mechanical properties of such hydrogels.
 In a first part we describe how we take advantage of our postfunctionalization approach to play systematically with the nature (aromatic or not) of the cationic repeating unit and the associated halide counterion (\ce{F-}/\ce{Cl-}/\ce{Br-}/\ce{I-}) varied along the Hofmeister series.
@@
 
-Differential Scanning Calorimetry (DSC) thermograms  of all synthesized polymers also reveals marked differences in their properties, that can only be explained by differences in their couterion features.
+Differential Scanning Calorimetry (DSC) thermograms  of all synthesized polymers also reveals marked differences in their properties, that can only be explained by differences in their counterion features.
 Thus, in the case polyelectrolytes with fluoride and chloride counterion, we observe no exo or endothermic transition below \SI{200}{\celsius} (featureless thermograms, not shown).
@@
 It is generally attributed to desorption of weakly bound water from the polymer network.
-%Its presence in the polymer samples where bromide and iodide are used as couterions suggests that adsorbed water could play a role in solvating the dissociated ion pairs in the latter.
+%Its presence in the polymer samples where bromide and iodide are used as counterions suggests that adsorbed water could play a role in solvating the dissociated ion pairs in the latter.
 Conversely, its absence for chloride and fluoride counterions suggests that dissociation is not effective because of stronger ion pair interaction.
@@
 With fluoride counterions, or in the case of \ce{PIm+Cl-}, we observe neither swelling nor dissolution.
-We recover a 8\%wt suspension as a biphasic mixture of its individual components.
+We recover a 8~\%~wt suspension as a biphasic mixture of its individual components.
 
@@
 Conversely, bulkier and softer ions like \ce{I-} are less adsorbed, resulting in more dissociated charges and higher solubility.
-As a consequence, while dissolution is observed for the four polymers with bromide and iodide counterions, in which ion dissociation takes place to a sufficient extent, the two polymers with fluoride counterions, for which couterion condensation is expected to be strong, cannot be dissolved.
+As a consequence, while dissolution is observed for the four polymers with bromide and iodide counterions, in which ion dissociation takes place to a sufficient extent, the two polymers with fluoride counterions, for which counterion condensation is expected to be strong, cannot be dissolved.
 
@@
 Because of strong intercoulombic repulsion, \ce{\pi+-\pi+} interactions are primarily weaker than their \ce{\pi-\pi} counterparts.
-However it has been established that, in the presence of couterions, the stabilizing effect of \ce{\pi+-\pi+} interaction could reach magnitudes largely exceeding that of the latter\cite{Geronimo2011}.
+However it has been established that, in the presence of counterions, the stabilizing effect of \ce{\pi+-\pi+} interaction could reach magnitudes largely exceeding that of the latter\cite{Geronimo2011}.
 In particular, dimerization of imidazolium-chloride ion pairs have been the object of recent studies, and it has been shown that the involvement of the negatively charged counterion was essential in maximizing the stabilization of the \ce{\pi+-\pi+} complex by minimizing coulombic repulsion between the imidazolium moieties\cite{Matthews2014,Gao2015a}.
@@
 All samples are at 8\%wt.
-The gray area limits of the linear, plastic and shear thinning regimes.}%
+The gray area encompasses the plastic regime that lies between the linear regime at very low amplitudes and the fluidlike shear-thinning regime at high strain amplitude.}%
 \label{fig:strain}%
@@
 We perform rheological studies of the water-swelling materials.
-Briefly, data are recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}\cite{Macosko1994,Larson1999}.
+Briefly, data are recorded at \SI{25}{\celsius} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}\cite{Macosko1994,Larson1999}.
 We place the sample on the plate, and then lower the cone to the measuring position, spreading the sample in the process.
@@
 For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile is that of pure water which confirms the visual observation.
-However as shown in Figure~\ref{fig:strain}, all other samples are solidlike ($G^\prime \gg G^{\prime\prime}$) at small strain and flow at large strains ($G^{\prime\prime} \gg G^\prime$) with a steep decrease of the moduli.
-We confirm solidlike behavior at low strain for all accessible frequencies (see supplementary material S6 Fig.
-1).
+However all other samples behave as shear-thinning yield stress fluids, confirming our previous results with \ce{PIm+Br-}~\cite{Srour2014}.
+As shown in Figure~\ref{fig:strain}, soluble samples are solidlike ($G^\prime \gg G^{\prime\prime}$) at small strain and flow at large strains ($G^{\prime\prime} \gg G^\prime$) with a steep decrease of the moduli.
+We confirm solidlike behavior at low strain for all accessible frequencies (see Supplementary Material S6 and Supplementary Figure 1).
 We checked that the linear mechanical properties are unchanged by the flow history given a few minutes of rest.
@@
 Imidazolium based gels are much weaker than their pyrrolidinium counterpart.
-Comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ value (13 and \SI{8100}{\pascal}, resp.) differ by almost three orders of magnitude.
-On the bottom row of Figure~\ref{fig:strain} we can compare the rheological responses of Poly(pyrrolidinium)-based hydrogels for the three counterions allowing dissolution in water.
+Comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ values (13 and \SI{8100}{\pascal}, resp.) differ by almost three orders of magnitude.
+On the bottom row of Figure~\ref{fig:strain} we can compare the rheological responses of poly(pyrrolidinium)-based hydrogels for the three counterions allowing dissolution in water.
 Gels with bromide and iodide counterions show roughly the same modulus value at vanishing strain for both $G^\prime$ and $G^{\prime\prime}$.
@@
 Each vertical line represents a sample.}
+\label{fig:phasediag}
 \end{figure}
@@
 In the following, we make the link between this theory and the chemical structure of our polymers.
-In absence of strain, the structure of a polyelectrolyte chain is organised at various scales.
+In absence of strain, the structure of a polyelectrolyte chain is organised at various scales, sketched on Figure~\ref{fig:phasediag}.
 
@@
 
-In poor solvent, the thermal length $\xi_T$ is larger than $b$.
-The thermal blob contains $g_T$ monomers and its density is set by the reduced temperature $(g_T/g_\mathrm{K})/\xi_T^3 = \tau/b^3$.
+In poor solvent, the reduced temperature $\tau = 1 -T/\Theta$ is smaller than 1 and the thermal length $\xi_T$ is larger than $b$.
+The thermal blob contains $g_T$ monomers and its density is set by the reduced temperature so that $(g_T/g_\mathrm{K})/\xi_T^3 = \tau/b^3$.
 Between $b$ and $\xi_T$ we have a persistent random walk $\xi_T = b (g_T/g_\mathrm{K})^{1/2}$.
@@
 
-At a scale $D$, the electrostatic energy balances the surface energy of the collapsed polymer.
-We name $g_e$ the number of monomers in the electrostatic blob.
-If there is a charge every $A$ monomers then there are $g_e/A$ charges in the electrostatic blob.
-The surface tension is proportional to the number of thermal blobs exposed on the surface $k_\mathrm{B}T/\xi_T^2$, with $k_\mathrm{B}$ the Boltzmann constant and $T$ the temperature.
-Thus the balance of energies writes:
+On can further define a third characteristic length $D$ at which the electrostatic energy balances the surface energy of the collapsed polymer.
+$D$ defines the size of the electrostatic blob and we name $g_e$ the number of monomers in it.
+Due to counterion condensation, not all monomers are charged.
+If there is a charge every $A$ monomers then there are $g_e/A$ charges in the electrostatic blob and the electrostatic energy write $E_e = (g_e/A)^2 e^2/(4\pi\epsilon)$, with $e$ the elementary charge and $\epsilon$ the dielectric constant of the solvent.
+The surface energy is proportional to the number of thermal blobs exposed on the surface $E_s = k_\mathrm{B}T (D/\xi_T)^2$.
+The balance of energies yields:
 \begin{equation}\left(\frac{D}{\xi_T}\right)^2 = \left(\frac{g_e}{A}\right)^2 \frac{\ell_\mathrm{B}}{D},
@@
 \end{equation}
-where $\ell_\mathrm{B} = e^2/(4\pi\epsilon k_\mathrm{B}T)$ is the Bjerrum length with $e$ the elementary charge, $\epsilon$ the dielectric constant of the solvent.
+where $\ell_\mathrm{B} = e^2/(4\pi\epsilon k_\mathrm{B}T)$ is the Bjerrum length.
 In water $\ell_\mathrm{B} \approx \SI{0.7}{\nano\metre}$.
-Since thermal blobs fill the volume of the electrostatic blob $D = \xi_T \left(g_e/g_T\right)^{1/3}$, we obtain
+Since thermal blobs fill the volume of the electrostatic blob, one has $D = \xi_T \left(g_e/g_T\right)^{1/3}$, we obtain
 \begin{equation}
@@
 
-We define the extension parameter $B$ as the ratio between the length of a chain of $g_e$ monomers fully extended and the actual size of the electrostatic blob: 
+We define the extension parameter $B$ as the ratio between the length of a fully extended chain of $g_e$ monomers ($g_e/g_\mathrm{K}$ Kuhn segments) and the actual size of the electrostatic blob: 
 \begin{equation}
@@
 
-At constant solvent quality and polymer architecture, $B$ is thus monotonically related to the amount of counterion condensation.
+At constant solvent quality and polymer architecture, $B$ is thus monotonically related to $A$ and thus to the amount of counterion condensation.
 In the following we will use $B$ as our main variable and estimate it from rheological measurements.
@@
 
-In dilute solution, the chain is a linear rod of length $L= N b / B g_\mathrm{K}$.
-The overlap concentration is thus $c^* = N/L^3 = B g_\mathrm{K} / b L^2$.
-Following \cite{Dobrynin1995}, for a monomer concentration $c > c^*$ and in absence of salt the screening length is
+The next length scale is the screening length $r_\mathrm{src}$.
+Between $D$ and $r_\mathrm{scr}$ the electrostatic blobs are organised in a linear rod containing $g_\mathrm{scr}$ monomers.
+A rod is $B$ times shorter than the fully extended $g_\mathrm{scr}/g_\mathrm{K}$ Kuhn segments it contains such that
+\begin{equation}
+r_\mathrm{scr}= g_\mathrm{scr} b / (B g_\mathrm{K}).
+\label{eq:rod}
+\end{equation}
+To obtain $r_\mathrm{src}$ we follow \cite{Dobrynin1995} by considering in a first step the case of a dilute solution where the length $L$ of a single chain of $\mathcal{N}$ monomers is shorter than the screening length.
+Equation~(\ref{eq:rod}) becomes $L= \mathcal{N} b / (B g_\mathrm{K})$.
+Therefore the overlap concentration is $c^* = \mathcal{N}/L^3 = B g_\mathrm{K} / (b L^2)$.
+In a second step, when the monomer concentration $c$ is larger than $c^*$ and in absence of salt, the screening length is
 \begin{equation}
@@
 
-In the presence of added salt, electrostatic interactions are screened at the Debye length $r_\mathrm{scr} = \left(\ell_\mathrm{B} c_s\right)^{-1/2}$. For arbitrary salt concentration,
+When charge screening is mainly due to the presence of added salt at a concentration $c_s$ much larger than the dissociated counterions, electrostatic interactions are screened at the Debye length $r_\mathrm{scr} = \left(\ell_\mathrm{B} c_s\right)^{-1/2}$.
+Finally, for arbitrary salt concentration \citet{Dobrynin1995} use the crossover expression,
 \begin{equation}
@@
 
-Between $D$ and $r_\mathrm{scr}$ the electrostatic blobs are organised in a linear rod containing $g_\mathrm{scr}$ monomers.
-The length of this rod is $B$ times shorter than if the electrostatic blobs were extended:
+From Equations~(\ref{eq:rod}) and (\ref{eq:rscr}) we obtain
 \begin{equation}
-g_\mathrm{scr} = B \frac{r_\mathrm{scr}}{b} g_\mathrm{K} = \left(\frac{B^3 g_\mathrm{K}^3}{cb^3}\right)^{1/2} \left(1 + B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
+g_\mathrm{scr} = \left(\frac{B^3 g_\mathrm{K}^3}{cb^3}\right)^{1/2} \left(1 + B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
 \label{eq:gscr}
@@
 
-If between two crosslinks or entanglements the ``procession'' of polymer chains contains more monomers than $g$ or $g_\mathrm{scr}$ it performs a random or self avoiding walk respectively.
-Each chain is thus an entropic spring of constant stiffness $k_\mathrm{B}T$.
-The number density of chains is $c/N$ with $c$ the monomer concentration and $N$ the number of monomers between cross-links or entanglements.
+Let us note $N$ the number of monomers between two crosslinks or entanglements.
+If $N>g$ the ``procession'' of polymer chains performs a random walk of correlation blobs, each being a self-avoiding walk.
+If $g>N>g_\mathrm{scr}$ the procession performs only a self-avoiding walk.
+In any case, each procession is thus an entropic spring of constant stiffness $k_\mathrm{B}T$.
+The number density of procession is $c/N$.
 Therefore, the shear modulus at low strain is given by:
 \begin{equation}
-G = \frac{c}{N}k_\mathrm{B}T,
+G = \frac{c}{N}k_\mathrm{B}T.
 \label{eq:G}
 \end{equation}
-where $c$ is the monomer number density.
 Because of the different weights of the pyrrolidinium and imidazolium moieties and of the different counterions, $c$ is not constant as we chose to conduct our experiments at constant weight fraction $w$ of polymer (generally used to quantify gelation ability of a given gelator).
 This explains why the respective modulus or critical strains of \ce{PPyr+I-} (heavier, lower number density) and \ce{PPyr+Br-} are in reverse order with respect to the Hofmeister series.
-Taking into account the molecular mass $M$ of each polymer and  knowing the density $d$ of the solvent, we obtain
+Taking into account the molecular mass $M$ of each polymer and  knowing the density $d$ of the solvent, we obtain the number $n = N/N_0$ of chains between cross-link
 \begin{equation}
@@
 
-As the material is strained, the random walk is stretched, then the self-avoiding walk.
-When the correlation blobs are fully stretched, the chain is a linear assembly of electrostatic blobs.
-Stretching the chain further means exposing more thermal blobs to the solvent, a process that cannot be reduced to spring of constant stiffness.
+As the material is strained, individual processions are stretched, starting from the larger scales~\cite{Pincus1976}.
+When the the correlation blobs are fully stretched, the procession is a linear assembly of electrostatic blobs.
+Stretching the chain further means exposing more thermal blobs to the solvent, a process that cannot be expressed using the model of a spring of constant stiffness.
 This condition is thus the limit between the linear and non linear regimes.
@@
 For all our samples, the linear regime is narrow $\gamma_0\approx 10^{-2}\ll 1$ if existing.
-It means that procession is just long enough to start performing the self-avoiding walk.
+It means that the procession is just long enough to start performing the self-avoiding walk.
 Incidentally, this implies that the procession is too short to perform the random walk.
@@
 \end{equation}
-with $B_s = c/u c_s g_\mathrm{K}$ and $B_0 = b c^{1/3} N^{2/3} g_\mathrm{K}^{-1} = (bc/g_\mathrm{K}) \left(k_\mathrm{B}T/G\right)^{2/3}$.
-Here, the ``salt'' consists in the phosphonate heads, that is to say two charges per chain.
-Thus $c_s/c = 2/N_0$ and $B_s = N_0/2u g_\mathrm{K} \approx 12$.
+with $B_s = c/(u c_s g_\mathrm{K})$ and $B_0 = b c^{1/3} N^{2/3} g_\mathrm{K}^{-1} = (bc/g_\mathrm{K}) \left(k_\mathrm{B}T/G\right)^{2/3}$.
+Here, we do not have salt per se, however negatively charged phosphonate heads and their counterions play the same role by contributing to screening.
+We thus have two ``salt'' charges per chain.
+Thus $c_s/c = 2/N_0$ and $B_s = N_0/(2u g_\mathrm{K}) \approx 12$.
+$B_0$ is obtained from $G^\prime(\gamma\rightarrow 0)$ for each sample.
 
-Analysing Equation (\ref{eq:solveB}), we identify two physically relevant limit cases: (i) if $B \ll B_s$ (low salt condition) then $B = B_0$; (ii) if $B \gg B_s$ (high salt condition) then $B = \left(B_0^3/B_s\right)^{1/2} > B_0$.
+Analysing Equation (\ref{eq:solveB}), we identify two physically relevant limit cases: (i) if $B \ll B_s$ screening is mainly due to uncondensed counterions and $B = B_0$; (ii) if $B \gg B_s$ screening is mainly due to phosphonate heads and $B = \left(B_0^3/B_s\right)^{1/2} > B_0$.
 
@@
 For the same quality of solvent $B$ is monotonically related to the amount of counterion condensation.
-We thus confirm that our samples are sorted by decreasing counterion condensation (decreasing $B$).
-More quantitatively, the number $A$ of monomers between dissociated charges is
+Since our samples are sorted by increasing $B$ we confirm that they are sorted by decreasing counterion condensation.
+To be more quantitative, we extract the number $A$ of monomers between dissociated charges from Equation~(\ref{eq:B}):
 \begin{equation}
@@
 In water the bonding energy between two ions is typically \SI{5}{\kilo\joule\per\mole}\cite{Schneider1992}, thus $E_c\approx 2 k_\mathrm{B}T$.
-However the medium surrounding the ionic bond is not water.
-One has to keep in mind that the chain is in poor solvent, thus electrostatic blobs have low water content.
+However the medium surrounding the ionic bond cannot in general be described as pure water.
+As the chain is in poor solvent, electrostatic blobs have a low water content.
 Such a low polarity microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}.

### 2016-05-12 PALSE/ion_pairing_paper.tex
Crossover around B=1\tau in the text, including for A Tables not updated

@@
 
+\paragraph{Kuhn length}
 The smallest scale is the Kuhn length $b$ containing $g_\mathrm{K}$ monomers.
@@
 
+\paragraph{Thermal length}
 In poor solvent, the reduced temperature $\tau = 1 -T/\Theta$ is smaller than 1 and the thermal length $\xi_T$ is larger than $b$.
@@
 
-On can further define a third characteristic length $D$ at which the electrostatic energy balances the surface energy of the collapsed polymer.
+\paragraph{Electrostatic length}
+One can further define a third characteristic length $D$ at which the electrostatic energy balances the surface energy of the collapsed polymer.
 $D$ defines the size of the electrostatic blob and we name $g_e$ the number of monomers in it.
@@
 g_e = \left(\frac{B^3}{\tau}\right)^{1/2} g_\mathrm{K},\qquad D = b \left(\frac{B}{\tau}\right)^{1/2}.
+\label{eq:geDfromlargeB}
+\end{equation}
+
+From Equations~(\ref{eq:thermal}) and (\ref{eq:geDfromlargeB}) we observe that for $B<1/\tau$ the size of the electrostatic blob should be smaller than the thermal length and the assumptions behind Equation~(\ref{eq:electrosurface}) break down.
+In the regime $B\ll 1/\tau$, most of the counterions are not condensed and poor solvent attraction competes directly against electrostatic repulsion.
+The size of the electrostatic blob is given by this balance of energies:
+\begin{equation}
+\left(\frac{b}{D}\right)^3\left(\frac{g_e}{g_K}\right)^2 \tau = \left(\frac{ge}{A}\right)^2 \frac{\ell_\mathrm{B}}{D}
+\label{eq:poorelec}
+\end{equation}
+yielding, with $D = b(g_e/g_\mathrm{K})^{1/2}$,
+\begin{equation}
+D = b \left(\frac{A^2\tau}{u g_\mathrm{K}^2}\right)^{1/2}, \qquad g_e = \frac{A^2\tau}{u g_\mathrm{K}},
+\end{equation}
+
+The extension parameter $B$ is expressed in this regime as
+\begin{equation}
+B = \frac{g_e}{g_\mathrm{K}}\frac{b}{D} = \left(\frac{g_e}{g_\mathrm{K}^2}\right)^{1/2}, \qquad D = bB, \qquad g_e = B^2 g_\mathrm{K}
+\label{eq:geDfromsmallB}
+\end{equation}
+
+We describe the crossover between weak and strong counterion condensation regimes by
+\begin{equation}
+D = bB (1+B\tau)^{-1/2}, \qquad
+A = u^{1/2} g_\mathrm{K} \frac{B}{\tau^{1/2}}(1+B\tau)^{-1/4}.
 \label{eq:geDfromB}
@@
 
+\paragraph{Screening length}
 The next length scale is the screening length $r_\mathrm{src}$.
@@
 
+\paragraph{Correlation length}
 The last length scale is the correlation length $\xi$.
@@
 Since our samples are sorted by increasing $B$ we confirm that they are sorted by decreasing counterion condensation.
-To be more quantitative, we extract the number $A$ of monomers between dissociated charges from Equation~(\ref{eq:B}):
-\begin{equation}
-A = u^{1/2} \left(\frac{B}{\tau}\right)^{3/4} g_\mathrm{K}
-\label{eq:A}
-\end{equation}.
-
+To be more quantitative, we extract the number $A$ of monomers between dissociated charges from Equation~(\ref{eq:geDfromB})%:
+%\begin{equation}
+%A = u^{1/2} \left(\frac{B}{\tau}\right)^{3/4} g_\mathrm{K}
+%\label{eq:A}
+%\end{equation}
+.
+%
 We know that $\Theta>\SI{100}{\celsius}$ so $0.2<\tau<1$. In the following we will deduce a more precise measure of $\tau$ and thus values of $A$.
@@
 \begin{equation}
-\frac{E_\mathrm{stretch}}{k_\mathrm{B}T} = \left(\left(\gamma +1\right)^{1/2} -1\right) \frac{N}{g_\mathrm{K}} \left(\frac{\tau^3}{B}\right)^{1/2}.
+\frac{E_\mathrm{stretch}}{k_\mathrm{B}T} = \left(\left(\gamma +1\right)^{1/2} -1\right) \frac{N}{g_\mathrm{K}} (1+B\tau)^{-1/2} \tau^2.
 \label{eq:stretch}
@@
 
-If we consider that the head-to-body bonds break at $\gamma_c$, we can express the reduced temperature as
+If we consider that the head-to-body bonds break at $\gamma_c$, the reduced temperature is solution of
 \begin{equation}
-\tau = B^{1/3} \left(\frac{1}{(\gamma_c +1)^{1/2} -1}\frac{g_\mathrm{K}}{N}\frac{E_c}{k_\mathrm{B}T}\right)^{2/3}
+\left(\left((\gamma_c +1)^{1/2} -1\right)\frac{k_\mathrm{B}T}{E_c}\frac{N}{g_\mathrm{K}}\right)^2 \tau^4 - B\tau -1 = 0
 \label{eq:tau}
@@
 
-Applying Equation~(\ref{eq:tau}) to \ce{PPyr+I-} with $E_c = 2 k_\mathrm{B}T$ yields $\tau \approx 0.31\pm 0.07$, or a $\Theta$-temperature around \SI{160}{\celsius}.
+Applying Equation~(\ref{eq:tau}) to \ce{PPyr+I-} with $E_c = 2 k_\mathrm{B}T$ yields $\tau \approx 0.39\pm 0.07$, or a $\Theta$-temperature around \SI{220}{\celsius}.
 Using this value and Equation (\ref{eq:stretch}) we deduce the bounding energy for each composition, as reported in Table~\ref{tab:results}.
-Indeed, the bonding energy increases with increasing counterion condensation to reach 60\% of the carbon-carbon single bond (\SI{350}{\kilo\joule\per\mole} or $140 k_\mathrm{B}T$) in the case of \ce{PIm+Br-}.
-According to Equation (\ref{eq:geDfromB}), in this extreme case an electrostatic blob contains $\approx 10^3$ polymers which embed the head-to-body bonds in a low-polarity environment $\approx 30$ times larger than the Bjerrum length.
+Indeed, the bonding energy increases with increasing counterion condensation to reach 85\% of the carbon-carbon single bond (\SI{350}{\kilo\joule\per\mole} or $140 k_\mathrm{B}T$) in the case of \ce{PIm+Br-}.
+According to Equation (\ref{eq:geDfromlargeB}), in this extreme case an electrostatic blob contains $\approx 10^3$ polymers which embed the head-to-body bonds in a low-polarity environment $\approx 25$ times larger than the Bjerrum length.
 Conversely, for \ce{PPyr+I-} and \ce{PPyr+Br-} the size of electrostatic blobs is comparable to $\ell_\mathrm{B}$, confirming an aqueous microenvironment.
@@
 
-Table~\ref{tab:results} also shows the values of $A$ deduced from Equation~(\ref{eq:A}) and the measured value of $\tau$.
-$A$ varies between $10$ for \ce{PPyr+I-}, corresponding to $7$ uncondensed counterions per polymer, to $850$ for \ce{PIm+Br-} indicating a single dissociated counterion every $12$ polymers.
+Table~\ref{tab:results} also shows the values of $A$ deduced from Equation~(\ref{eq:geDfromB}) and the measured value of $\tau$.
+$A$ varies between $7$ for \ce{PPyr+I-}, corresponding to $10$ uncondensed counterions per polymer, to $700$ for \ce{PIm+Br-} indicating a single dissociated counterion every $10$ polymers.
 

### 2016-05-18 PALSE/ion_pairing_paper.tex
With crossover
rsc template

--- 2016-05-12 PALSE/ion_pairing_paper.tex
@@
-\documentclass[journal=jacsat,manuscript=article]{achemso}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}       % Use modern font encodings
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%This is the LaTeX ARTICLE template for RSC journals
+%Copyright The Royal Society of Chemistry 2014
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\documentclass[twoside,twocolumn,9pt]{article}
+\usepackage{extsizes}
+\usepackage[super,sort&compress,comma]{natbib} 
+\usepackage[version=3]{mhchem}
+\usepackage[left=1.5cm, right=1.5cm, top=1.785cm, bottom=2.0cm]{geometry}
+\usepackage{balance}
+\usepackage{widetext}
+\usepackage{times,mathptmx}
+\usepackage{sectsty}
+\usepackage{graphicx} 
+\usepackage{lastpage}
+\usepackage[format=plain,justification=raggedright,singlelinecheck=false,font={stretch=1.125,small,sf},labelfont=bf,labelsep=space]{caption}
+\usepackage{float}
+\usepackage{fancyhdr}
+\usepackage{fnpos}
+\usepackage[english]{babel}
+\usepackage{array}
+\usepackage{droidsans}
+\usepackage{charter}
+\usepackage[T1]{fontenc}
+\usepackage[usenames,dvipsnames]{xcolor}
+\usepackage{setspace}
+\usepackage[compact]{titlesec}
+%%%Please don't disable any packages in the preamble, as this may cause the template to display incorrectly.%%%
+
 \usepackage{amssymb}
-\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
 \usepackage{siunitx}
-\usepackage[table]{xcolor}
+\usepackage{hyperref}
 
-\usepackage{xr}
-\externaldocument[S-]{ion_pairing_SI}
+%\usepackage{epstopdf}%This line makes .eps figures into .pdf - please comment out if not required.
 
-\author{Hassan Srour}
-\author{Martien Duvall Deffo Ayagou}
-\author{Thi Thanh-Tam Nguyen}
-\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}
-
-\author{Nicolas Taberlet}
-\author{Sébastien Manneville}
-\affiliation[Laboratoire de Physique de l'ENS de Lyon]{Laboratoire de Physique, Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon, 46 Allée d'Italie, 69007 Lyon}
-
-\author{Chantal Andraud}
-\author{Cyrille Monnereau}
-\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}
-\email{cyrille.monnereau@ens-lyon.fr}
-
-\author{Mathieu Leocmach}
-\affiliation[Institut Lumière Matière]{Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}
-\email{mathieu.leocmach@univ-lyon1.fr}
-
-%\title{Supramolecular crocodile line controls the rheological properties of polyelectrolytes hydrogels}
-\title{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}
-%\title{Ion pairing in crocodile line hydrogels controls rheological properties}
-%\title{In crocodile line hydrogels ion pairing controls rheological properties}
-%\title{Controlling the rheological properties of well-defined polyelectrolytes hydrogels by fine tuning their ion pairing process}
-
-\keywords{ATRP, hydrogel, poly(cations), rheology}
+\definecolor{cream}{RGB}{222,217,201}
 
@@
 
-\begin{tocentry}
+\pagestyle{fancy}
+\thispagestyle{plain}
+\fancypagestyle{plain}{
 
-%9\,cm by 3.5\,cm
-\includegraphics{TOC2.png}
+%%%HEADER%%%
+\fancyhead[C]{\includegraphics[width=18.5cm]{head_foot/header_bar}}
+\fancyhead[L]{\hspace{0cm}\vspace{1.5cm}\includegraphics[height=30pt]{head_foot/journal_name}}
+\fancyhead[R]{\hspace{0cm}\vspace{1.7cm}\includegraphics[height=55pt]{head_foot/RSC_LOGO_CMYK}}
+\renewcommand{\headrulewidth}{0pt}
+}
+%%%END OF HEADER%%%
 
-\end{tocentry}
+%%%PAGE SETUP - Please do not change any commands within this section%%%
+\makeFNbottom
+\makeatletter
+\renewcommand\LARGE{\@setfontsize\LARGE{15pt}{17}}
+\renewcommand\Large{\@setfontsize\Large{12pt}{14}}
+\renewcommand\large{\@setfontsize\large{10pt}{12}}
+\renewcommand\footnotesize{\@setfontsize\footnotesize{7pt}{10}}
+\makeatother
 
-\begin{abstract}
-We demonstrated recently that polyelectrolytes with cationic moieties along the chain and a single anionic head are able to form physical hydrogels due to the reversible nature of the head-to-body ionic bond.
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
+\renewcommand\footnoterule{\vspace*{1pt}% 
+\color{cream}\hrule width 3.5in height 0.4pt \color{black}\vspace*{5pt}} 
+\setcounter{secnumdepth}{5}
+
+\makeatletter 
+\renewcommand\@biblabel[1]{#1}            
+\renewcommand\@makefntext[1]% 
+{\noindent\makebox[0pt][r]{\@thefnmark\,}#1}
+\makeatother 
+\renewcommand{\figurename}{\small{Fig.}~}
+\sectionfont{\sffamily\Large}
+\subsectionfont{\normalsize}
+\subsubsectionfont{\bf}
+\setstretch{1.125} %In particular, please do not alter this line.
+\setlength{\skip\footins}{0.8cm}
+\setlength{\footnotesep}{0.25cm}
+\setlength{\jot}{10pt}
+\titlespacing*{\section}{0pt}{4pt}{4pt}
+\titlespacing*{\subsection}{0pt}{15pt}{1pt}
+%%%END OF PAGE SETUP%%%
+
+%%%FOOTER%%%
+\fancyfoot{}
+\fancyfoot[LO,RE]{\vspace{-7.1pt}\includegraphics[height=9pt]{head_foot/LF}}
+\fancyfoot[CO]{\vspace{-7.1pt}\hspace{13.2cm}\includegraphics{head_foot/RF}}
+\fancyfoot[CE]{\vspace{-7.2pt}\hspace{-14.2cm}\includegraphics{head_foot/RF}}
+\fancyfoot[RO]{\footnotesize{\sffamily{1--\pageref{LastPage} ~\textbar  \hspace{2pt}\thepage}}}
+\fancyfoot[LE]{\footnotesize{\sffamily{\thepage~\textbar\hspace{3.45cm} 1--\pageref{LastPage}}}}
+\fancyhead{}
+\renewcommand{\headrulewidth}{0pt} 
+\renewcommand{\footrulewidth}{0pt}
+\setlength{\arrayrulewidth}{1pt}
+\setlength{\columnsep}{6.5mm}
+\setlength\bibsep{1pt}
+%%%END OF FOOTER%%%
+
+%%%FIGURE SETUP - please do not change any commands within this section%%%
+\makeatletter 
+\newlength{\figrulesep} 
+\setlength{\figrulesep}{0.5\textfloatsep}
+
+\newcommand{\topfigrule}{\vspace*{-1pt}% 
+\noindent{\color{cream}\rule[-\figrulesep]{\columnwidth}{1.5pt}} }
+
+\newcommand{\botfigrule}{\vspace*{-2pt}% 
+\noindent{\color{cream}\rule[\figrulesep]{\columnwidth}{1.5pt}} }
+
+\newcommand{\dblfigrule}{\vspace*{-1pt}% 
+\noindent{\color{cream}\rule[-\figrulesep]{\textwidth}{1.5pt}} }
+
+\makeatother
+%%%END OF FIGURE SETUP%%%
+
+%%%TITLE, AUTHORS AND ABSTRACT%%%
+\twocolumn[
+  \begin{@twocolumnfalse}
+\vspace{3cm}
+\sffamily
+\begin{tabular}{m{4.5cm} p{13.5cm} }
+
+\includegraphics{head_foot/DOI} & \noindent\LARGE{\textbf{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels$^\dag$}} \\%Article title goes here instead of the text "This is the title"
+\vspace{0.3cm} & \vspace{0.3cm} \\
+
+ & \noindent\large{Hassan Srour,\textit{$^{a}$} Martien Duvall Deffo Ayagou,\textit{$^{a}$} Thi Thanh-Tam Nguyen,\textit{$^{a}$}, Nicolas Taberlet,\textit{$^{b\ddag}$} S\'{e}bastien Manneville,\textit{$^{b\ddag}$} Chantal Andraud,\textit{$^{a}$} Cyrille Monnereau,\textit{$^{a}$} and Mathieu Leocmach$^{\ast}$\textit{$^{c}$}} \\%Author names go here instead of "Full name", etc.
+
+\includegraphics{head_foot/dates} & \noindent\normalsize{We demonstrated recently that polyelectrolytes with cationic moieties along the chain and a single anionic head are able to form physical hydrogels due to the reversible nature of the head-to-body ionic bond.
 Here we generate a variety of such polyelectrolytes with various cationic moieties and counterion combinations starting from a common polymeric platform.
 We show that the rheological properties (shear modulus, critical strain) of the final hydrogels can be modulated over three orders of magnitude depending on the cation/anion pair.
-Our data fit remarkably well within a theoretical model involving a supramolecular head-to-tail single file between cross-links, akin to the behaviour of pine-processionary caterpillar, the length of the procession being set by the amount of charge dissociation.
-\end{abstract}
+Our data fit remarkably well within a scaling model involving a supramolecular head-to-tail single file between cross-links, akin to the behaviour of pine-processionary caterpillar.
+This model allows the quantitative measure of the amount of counterion condensation from standard rheology procedure.
+%The abstract should be a single paragraph which summarises the content of the article.
+Any references in the abstract should be written out in full \textit{e.g.} [Surname \textit{et al., Journal Title}, 2000, \textbf{35}, 3523].
+} \\%The abstract goes here instead of the text "The abstract should be..."
 
+\end{tabular}
+
+ \end{@twocolumnfalse} \vspace{0.6cm}
+
+  ]
+%%%END OF TITLE, AUTHORS AND ABSTRACT%%%
+
+%%%FONT SETUP - please do not change any commands within this section
+\renewcommand*\rmdefault{bch}\normalfont\upshape
+\rmfamily
+\section*{}
+\vspace{-1cm}
+
+%%%FOOTNOTES%%%
+
+\footnotetext{\textit{$^{a}$~Univ Lyon, Ens de Lyon, Univ Claude Bernard, CNRS,
+Laboratoire de Chimie, F-69342 Lyon, France. }}%Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
+\footnotetext{\textit{$^{b}$~Univ Lyon, Ens de Lyon, Univ Claude Bernard, CNRS,
+Laboratoire de Physique, F-69342 Lyon, France. }}
+\footnotetext{\textit{$^{c}$~Univ Lyon, Univ Claude Bernard, CNRS, Institut Lumière Matière, F-69622 Villeurbanne, France.
+E-mail: mathieu.leocmach@univ-lyon1.fr}}%Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
+
+%Please use \dag to cite the ESI in the main text of the article.
+%If you article does not have ESI please remove the the \dag symbol from the title and the footnotetext below.
+\footnotetext{\dag~Electronic Supplementary Information (ESI) available: [details of any supplementary information available should be included here]. See DOI: 10.1039/b000000x/}
+%additional addresses can be cited as above using the lower-case letters, c, d, e... If all authors are from the same address, no letter is required
+
+%\footnotetext{\ddag~Additional footnotes to the title and authors can be included \emph{e.g.}\ `Present address:' or `These authors contributed equally to this work' as above using the symbols: \ddag, \textsection, and \P.
+Please place the appropriate symbol next to the author's name and include a \texttt{\textbackslash footnotetext} entry in the the correct place in the list.}
+
+%%%END OF FOOTNOTES%%%
+
+%%%MAIN TEXT%%%%
+\begin{figure*}
+\includegraphics{synthesis_rsc.pdf}
+\caption{Synthesis of \ce{PIm+Br-} and \ce{PPyr+Br-} and their intermediates \ce{POH} and \ce{PBr} with $N_0=70$.}
+\label{sch:synthesis}
+\end{figure*}
 \section{Introduction}
@@
 Contrary to a small electrolyte for which full ion pairs dissociation is generally achieved in water, polyelectrolytes in solution are in general not fully dissociated, with a substantial fraction of the counterions bound to the polymer\cite{Manning1979}.
-The resulting net charge of the chain governs the physical properties of the polyelectrolyte solution\cite{Rubinstein1996}, first of all the ability of the polymer to dissolve.
+The resulting net charge of the chain governs the physical properties of the polyelectrolyte solution~\cite{DeGennes1976,Khokhlov1980,Raphael1990,Dobrynin1995,Rubinstein1996}, first of all the ability of the polymer to dissolve in a poor solvent.
 Counterion condensation is the physical bounding or adsorption of counterions near the polymer chain.
@@
 This so-called Hofmeister series has since proved quite universally valid\cite{Zhang2010a}, including multi-charged polymers such as naturally occurring proteins or synthetic polyelectrolytes as well as charged colloidal particles\cite{Schwierz2010,Oncsik2015} or soft matter interfaces in general\cite{Leontidis2014}.
-However, depending on the nature of the polymer (hydrophilic or hydrophobic, anionic or cationic), the Hofmeister series can be direct (in short, well hydrated ions promote condensation) or reverse\cite{Schwierz2010,Oncsik2015}.
+However, depending on the nature of the polymer (hydrophilic or hydrophobic, anionic or cationic), the Hofmeister series can be direct (in short, well hydrated ions promote condensation) or reverse~\cite{Schwierz2010,Oncsik2015}.
 
@@
 In a fourth part, we rationalize this behaviour by proposing a microscopic model based on the idea that at low dissociation rates, cross-links (defined as a point where three or more polymers meet) are not separated by a single macromolecule but by several (up to hundreds) polymers in a supramolecular chain.
-We call this behaviour ``processionary'' in analogy to pine processionary caterpillar (\textit{Thaumetopoea pityocampa}) behaviour\cite{Fabre1916}.%Finally, we discuss our results and possible applications of such smart hydrogels.
+We call this behaviour ``processionary'' in analogy to pine processionary caterpillar (\textit{Thaumetopoea pityocampa}) behaviour\cite{Fabre1916}.
+In particular, this model enables the quantification of the charge condensation rate from standard rheological measurements.%Finally, we discuss our results and possible applications of such smart hydrogels.
 
@@
 We synthesise the phosphonate terminated polymer hereafter referred to as \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
-We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; see Supplementary Materials and Supplementary Figure~S\ref{S-fig:sec}), with a degree of polymerisation $N_0=70$.
+We obtain a well-controlled linear polymer ($M_w= \SI{8200}{\dalton}$ (NMR),  $M_w/M_n = 1.08$ from GPC; see Supplementary Materials and Supplementary Figure~S\ref{S-fig:sec}), with a degree of polymerisation $N_0=70$(NMR).
 This polymer serves as a common scaffold from which we derive  all studied systems.
@@
 In these cases, as an inherent consequence of the structure of the starting material, charge neutrality is provided by bromide counter ions, and the polymer will be referred to as \ce{PIm+Br-} and \ce{PPyr+Br-}, respectively (Scheme~\ref{sch:synthesis}).
-
-\begin{scheme}
-\includegraphics{synthesis.pdf}
-\caption{Synthesis of \ce{PIm+Br-} and \ce{PPyr+Br-} and their intermediates \ce{POH} and \ce{PBr} with $N_0=70$.}
-\label{sch:synthesis}
-\end{scheme}
 
@@
 
-\begin{scheme}
-\includegraphics{metathesis.pdf}
+\begin{figure}
+\includegraphics{metathesis_rsc.pdf}
 \caption{(a) Anionic metathesis reaction. As above, \ce{Nu+} corresponds either to pyrrolidinium or imidazolium. (b) Summary of the polymers derived from a single batch of \ce{PBr}.}
 \label{sch:metathesis}
-\end{scheme}
+\end{figure}
 
@@
 \begin{figure}
-\includegraphics{dsc.pdf}
+\includegraphics{dsc_rsc.pdf}
 \caption{DSC measurements of polymer samples showing the presence of an endothermic peak.
@@
 No dissolution or swelling was observed, implying that $\Theta>\SI{100}{\celsius}$.
-%Nonetheless, due to charge dissociation, five of our polyelectrolytes are soluble at \SI{25}{\celsius}.
-
-%\begin{figure}
-%\caption{pictures of 8\%wt water solutions of \ce{PIm+X-} and \ce{PPyr+X-} in water.}
-%\label{fig:pictures}
-%\end{figure}
 
 \subsection{Rheological measurements}
-\begin{figure}[t]
+\begin{figure*}
 \includegraphics{strain.pdf}%
@@
 \label{fig:strain}%
-\end{figure}
+\end{figure*}
 
@@
 
-%We define the critical strain value $\gamma_c$ between these two regimes when $G^{\prime\prime}$ reaches a maximum.
-We checked that the linear mechanical properties are unchanged by the flow history given a few minutes of rest.
-%To understand the influence of the ion pairing on the mechanical properties, we compute the cohesive energy (equation 1) that should probe the strength of the physical supramolecular bonds.15,16
-
-%In Figure~\ref{fig:strain}a we compare the rheological responses of Poly(pyrrolidinium)-based hydrogels for the three counterions allowing dissolution in water.
-Gels with bromide and iodide counterions show roughly the same mechanical behavior in the linear regime but \ce{PPyr+I-} yields at slightly higher strains ($\gamma_c=4\%$ instead of $3\%$).
-By contrast, the gel with chloride counterions is weaker by more than one order of magnitude.
-In other words, counterion condensation is correlated to softer gel.
-
-%We also compare the rheology of \ce{PIm+X-} and \ce{PPyr+X-} with the same counterion \ce{X} in Figure~\ref{fig:strain}b,c.
-For both iodide (Figure~\ref{fig:strain}b, X=I) and bromide (Figure~\ref{fig:strain}c, X=Br), imidazolium based gels are much weaker than their pyrrolidinium counterpart with a broader linear regime.
-Comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ value (13 and \SI{8100}{\pascal}, resp.) and $\gamma_c$ value (804\% and 1.6\%, resp.) differ by almost three orders of magnitude.
-Here also, gel toughness is related to charge dissociation efficiency.
-In the following we show that this correlation can be explained on the basis of a microscopic model of interchain interactions.
-
 \subsection{Processionary model}
@@
 
-On the other hand, if every chain has a single foreign head attached to its body, every polymer is linked to two others in a single file\bibnote{We neglect the possibility of the two charges of the phosphonate head to attach to two separate chains}.
+On the other hand, if every chain has a single foreign head attached to its body, every polymer is linked to two others in a single file\footnote[3]{We neglect the possibility of the two charges of the phosphonate head to attach to two separate chains}.
 In addition, if we suppose that in order to minimize the inter-chain repulsion between charged cationic groups, heads are preferentially attached to the tail of their neighbour, we obtain a linear chain of effective polymerisation index $n N_0$ where $n$ is the number of polymers in the supramolecular queue.
@@
 the elasticity of the undamaged gel network, (ii) the strain amplitude $\gamma_0$ corresponding to the end of the linear regime~\cite{Hyun2011}, and (iii) the strain amplitude $\gamma_c$ corresponding to the crossing of the moduli.
-By using these three parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters as the average number of chains between cross-links $n$, the number of monomers between dissociated ion pairs $A$ and the head-to-body bonding energy $E_c$.
+By using these three parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters such as the average number of chains between cross-links $n$, the number of monomers between dissociated ion pairs $A$ and the head-to-body bonding energy $E_c$.
 
-\subsubsection{Conformation at rest}
-\begin{figure}
+\subsubsection{Conformation at rest.}
+\begin{figure*}
 \includegraphics{phase_diag.pdf}
-\caption{Conformation of semi-dilute polyelectrolytes at rest at various scales function of dissociated charge frequency.
-The phase diagram is given for $\tau=0.46$ and for the monomer number density of \ce{PIm+Br-}, although $c$ is slightly different for each sample.
+\caption{Conformation of semi-dilute polyelectrolytes at rest at various scales as a function of dissociated charge frequency $A$.
+The phase diagram is given for a reduced temperature $\tau=0.46$ and for the monomer number density of \ce{PIm+Br-}, although $c$ is slightly different for each sample.
 Each vertical line represents a sample.}
 \label{fig:phasediag}
-\end{figure}
+\end{figure*}
 
-The scaling theory of \citet{Rubinstein1996} describes the structure of a polyelectrolyte chain.
+The scaling theory of \citet{Dobrynin1995} describes the structure of a polyelectrolyte chain.
 In the following, we make the link between this theory and the chemical structure of our polymers.
@@
 
+\paragraph{Kuhn length.}
 The smallest scale is the Kuhn length $b$ containing $g_\mathrm{K}$ monomers.
@@
 
-In poor solvent, the reduced temperature $\tau = 1 -T/\Theta$ is smaller than 1 and the thermal length $\xi_T$ is larger than $b$.
-The thermal blob contains $g_T$ monomers and its density is set by the reduced temperature so that $(g_T/g_\mathrm{K})/\xi_T^3 = \tau/b^3$.
-Between $b$ and $\xi_T$ we have a persistent random walk $\xi_T = b (g_T/g_\mathrm{K})^{1/2}$.
-These conditions yields
+\paragraph{Thermal length.}
+At the thermal length $\xi_T$, the attractive potential between monomers is balanced by thermal energy $k_\mathrm{B}T$, with $k_\mathrm{B}$ the Boltzmann constant.
+This balance of energy can be written function of the reduced temperature $\tau = 1 -T/\Theta$ and the number $g_T$ of monomers in the thermal blob:
+\begin{equation}
+k_\mathrm{B}T = \left(\frac{b}{\xi_T}\right)^3\left(\frac{g_T}{g_K}\right)^2 \tau k_\mathrm{B}T 
+\label{eq:poorkT}
+\end{equation}
+Between $b$ and $\xi_T$ we have a persistent random walk $\xi_T = b (g_T/g_\mathrm{K})^{1/2}$. These conditions yields 
 \begin{equation}
@@
 
-On can further define a third characteristic length $D$ at which the electrostatic energy balances the surface energy of the collapsed polymer.
+\paragraph{Electrostatic length, weak charging case.}
+Following \citet{Khokhlov1980} one can further define a third characteristic length $D$ over which the electrostatic energy become dominant.
+balances the surface energy of the collapsed polymer.
 $D$ defines the size of the electrostatic blob and we name $g_e$ the number of monomers in it.
 Due to counterion condensation, not all monomers are charged.
-If there is a charge every $A$ monomers then there are $g_e/A$ charges in the electrostatic blob and the electrostatic energy write $E_e = (g_e/A)^2 e^2/(4\pi\epsilon)$, with $e$ the elementary charge and $\epsilon$ the dielectric constant of the solvent.
-The surface energy is proportional to the number of thermal blobs exposed on the surface $E_s = k_\mathrm{B}T (D/\xi_T)^2$.
+Counterion condensation is quantified by assuming that there is a charge every $A$ monomers.
+Therefore there are $g_e/A$ charges in the electrostatic blob and the electrostatic energy reads $E_e = (g_e/A)^2 e^2/(4\pi\epsilon)$, with $e$ the elementary charge and $\epsilon$ the dielectric constant of the solvent.
+If $D>\xi_T$ the main opposition to the electrostatic energy can be understood as a surface energy proportional to the number of thermal blobs exposed on the surface $E_s = k_\mathrm{B}T (D/\xi_T)^2$.
 The balance of energies yields:
@@
 In water $\ell_\mathrm{B} \approx \SI{0.7}{\nano\metre}$.
-Since thermal blobs fill the volume of the electrostatic blob, one has $D = \xi_T \left(g_e/g_T\right)^{1/3}$, we obtain
+Since thermal blobs fill the volume of the electrostatic blob, one has $D = \xi_T \left(g_e/g_T\right)^{1/3}$ and using Equation~\eqref{eq:thermal} we obtain
 \begin{equation}
@@
 
-We define the extension parameter $B$ as the ratio between the length of a fully extended chain of $g_e$ monomers ($g_e/g_\mathrm{K}$ Kuhn segments) and the actual size of the electrostatic blob: 
+\citet{Dobrynin1995} introduced the extension parameter $B$ defined as the ratio between the length of a fully extended chain of $g_e$ monomers ($g_e/g_\mathrm{K}$ Kuhn segments) and the actual size of the electrostatic blob:
 \begin{equation}
@@
 
-At constant solvent quality and polymer architecture, $B$ is thus monotonically related to $A$ and thus to the amount of counterion condensation.
+At constant solvent quality and polymer architecture, $B$ is thus monotonically related to the amount of counterion condensation $A$.
 In the following we will use $B$ as our main variable and estimate it from rheological measurements.
@@
 g_e = \left(\frac{B^3}{\tau}\right)^{1/2} g_\mathrm{K},\qquad D = b \left(\frac{B}{\tau}\right)^{1/2}.
+\label{eq:geDfromlargeB}
+\end{equation}
+
+\paragraph{Electrostatic length, strong charging case.}
+From Equations~(\ref{eq:thermal}) and (\ref{eq:geDfromlargeB}) we observe that for $B<1/\tau$ the size of the electrostatic blob should be smaller than the thermal length and the assumptions behind Equation~(\ref{eq:electrosurface}) break down.
+To our knowledge the study of polyelectrolytes in poor solvent by \citet{Khokhlov1980} and subsequent literature~\cite{Dobrynin1995,Rubinstein1996} focused on the weak charging regime and did not treat the small $B$ regime.
+When $B\ll 1/\tau$, most of the counterions are not condensed and attraction potential between monomers competes directly against the electrostatic repulsion.
+The size of the electrostatic blob is given by the following balance of energies:
+\begin{equation}
+\left(\frac{b}{D}\right)^3\left(\frac{g_e}{g_K}\right)^2 \tau = \left(\frac{g_e}{A}\right)^2 \frac{\ell_\mathrm{B}}{D}
+\label{eq:poorelec}
+\end{equation}
+yielding, with $D = b(g_e/g_\mathrm{K})^{1/2}$,
+\begin{equation}
+D = b \left(\frac{A^2\tau}{u g_\mathrm{K}^2}\right)^{1/2}, \qquad g_e = \frac{A^2\tau}{u g_\mathrm{K}},
+\end{equation}
+
+In this weak counterion condensation regime, the definition of the extension parameter $B$ yields a different relation with $A$ and both $D$ and $g_e$:
+\begin{equation}
+B = \frac{g_e}{g_\mathrm{K}}\frac{b}{D} = \left(\frac{A^2\tau}{u g_\mathrm{K}^2}\right)^{1/2}, \qquad D = bB, \qquad g_e = B^2 g_\mathrm{K}
+\label{eq:geDfromsmallB}
+\end{equation}
+
+Here we shall describe the crossover between weak and strong counterion condensation regimes by combining Equations~\eqref{eq:geDfromlargeB} and \eqref{eq:geDfromsmallB} into
+\begin{equation}
+D = bB (1+B\tau)^{-1/2}, \qquad
+g_e = g_\mathrm{K} B^2 (1+B\tau)^{-1/2},
 \label{eq:geDfromB}
 \end{equation}
+and we generalize the relation between $A$ and $B$ as
+\begin{equation}
+A = u^{1/2} g_\mathrm{K} \frac{B}{\tau^{1/2}}(1+B\tau)^{-1/4}.
+\label{eq:A}
+\end{equation}
 
-The next length scale is the screening length $r_\mathrm{src}$.
+\paragraph{Screening length.}
+The next length scale is the screening length $r_\mathrm{scr}$.
 Between $D$ and $r_\mathrm{scr}$ the electrostatic blobs are organised in a linear rod containing $g_\mathrm{scr}$ monomers.
-A rod is $B$ times shorter than the fully extended $g_\mathrm{scr}/g_\mathrm{K}$ Kuhn segments it contains such that
+A rod is $B$ times shorter than the fully extended $g_\mathrm{scr}/g_\mathrm{K}$ Kuhn segments that it contains such that
 \begin{equation}
@@
 \end{equation}
-To obtain $r_\mathrm{src}$ we follow \cite{Dobrynin1995} by considering in a first step the case of a dilute solution where the length $L$ of a single chain of $\mathcal{N}$ monomers is shorter than the screening length.
+To obtain $r_\mathrm{scr}$ we follow \cite{Dobrynin1995} by considering in a first step the case of a dilute solution where the length $L$ of a single chain of $\mathcal{N}$ monomers is shorter than the screening length.
 Equation~(\ref{eq:rod}) becomes $L= \mathcal{N} b / (B g_\mathrm{K})$.
@@
 
-When charge screening is mainly due to the presence of added salt at a concentration $c_s$ much larger than the dissociated counterions, electrostatic interactions are screened at the Debye length $r_\mathrm{scr} = \left(\ell_\mathrm{B} c_s\right)^{-1/2}$.
+When charge screening is mainly due to the presence of added salt at a concentration $c_s$ much larger than the dissociated counterions, electrostatic interactions are screened at the Debye length $r_\mathrm{scr} = \left(\ell_\mathrm{B} c_s\right)^{-1/2} \equiv \kappa^{-1}$.
 Finally, for arbitrary salt concentration \citet{Dobrynin1995} use the crossover expression,
@@
 
+\paragraph{Correlation length.}
 The last length scale is the correlation length $\xi$.
@@
 
-\subsubsection{From modulus to procession length}
+\subsubsection{From modulus to procession length.}
 
@@
 
-We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 880 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
+We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 860 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
 This larger cross-link ratio is consistent with a higher probability of attaching two or more heads on a highly dissociated body.
 
-\begin{table}
+\begin{table*}
 \sisetup{
@@
 %table-format = 1.3(1),
-}
-\begin{tabular}{@{}l|SSS|SSSSSS@{}}
+}\centering
+\begin{tabular}{@{}lSSSSSSSSS@{}}\hline
 & {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_0$} & {$\gamma_c$} & {$n$} & {$B$} & {$E_c$} & {$E_c$} & {$D$} & {$A$}\\
 &	{\si{\pascal}} & {\%} & {\%} & & & {\si{\kilo\joule\per\mole}}& {$k_\mathrm{B}T$} & {$\ell_B$}&\\\hline&&&&&\\[-10pt]
-\ce{PIm+Br-}	& 13	&	1.1	&	810	&	860	&	930	&	210	&	84	&	29	&	850\\
-\ce{PIm+I-}		& 78	&	1.8	&	18	&	122	&	126	&	80	&	32	&	10.6	&	189\\
-\ce{PPyr+Cl-}	& 270 	&	2.8	&	12	&	50.6	&	64	&	46	&	19	&	7.6	&	115\\
-\ce{PPyr+Br-}	& 9800 	& 	0.3	&	2	&	1.15	&	2.8	&	5.05	&	2.02	&	1.6	&	11\\
-\ce{PPyr+I-}	& 8800	&	0.0	&	89	&	1.08	&	2.5	&	5	&	2	&	1.5	&	10\\
+\ce{PIm+Br-}	&	13	&	1.1	&	800	&	800	&	830	&	304	&	123	&	24	&	640\\
+\ce{PIm+I-}	&	78	&	1.8	&	18	&	113	&	113	&	116	&	46.8	&	8.7	&	140\\
+\ce{PPyr+Cl-}	&	272	&	2.8	&	10	&	46.8	&	58.2	&	66	&	26.6	&	6.2	&	90\\
+\ce{PPyr+Br-}	&	9810	&	0.3	&	2	&	1.1	&	2.6	&	5.2	&	2.1	&	1.0	&	7.1\\
+\ce{PPyr+I-}	&	8800	&	0.1	&	90	&	1.0	&	2.4	&	5.0	&	2.0	&	0.9	&	6.5\\
+\hline
 \end{tabular}
-\caption{Summary of rheological measurements and microscopic values deduced from the model. $E_c$, $D$ and $A$ are obtained supposing $\tau=0.31$.}
+\caption{Summary of rheological measurements and microscopic values deduced from the model. $E_c$, $D$ and $A$ are obtained by assuming $\tau=0.40$}
 \label{tab:results}
-\end{table}
+\end{table*}
 
-\subsubsection{Limit of the linear regime}
+\subsubsection{Limit of the linear regime.}
 
@@
 Since our samples are sorted by increasing $B$ we confirm that they are sorted by decreasing counterion condensation.
-To be more quantitative, we extract the number $A$ of monomers between dissociated charges from Equation~(\ref{eq:B}):
-\begin{equation}
-A = u^{1/2} \left(\frac{B}{\tau}\right)^{3/4} g_\mathrm{K}
-\label{eq:A}
-\end{equation}.
-
+To be more quantitative, we extract the number $A$ of monomers between dissociated charges from Equation~(\ref{eq:A})%:
+%\begin{equation}
+%A = u^{1/2} \left(\frac{B}{\tau}\right)^{3/4} g_\mathrm{K}
+%\label{eq:A}
+%\end{equation}
+.
+%
 We know that $\Theta>\SI{100}{\celsius}$ so $0.2<\tau<1$. In the following we will deduce a more precise measure of $\tau$ and thus values of $A$.
@@
 
-\subsubsection{Extension of the electrostatic blobs}
+\subsubsection{Extension of the electrostatic blobs.}
 
@@
 
-From Equation~(\ref{eq:geDfromB}) we know that $D$ increased with $B$.
+From Equation~(\ref{eq:geDfromB}) we know that $D$ increases with $B$.
 We thus expect stronger head-to-body bonds when $B$ is large and the low-polarity environment is extended ($D\gg\ell_\mathrm{B}$).
@@
 
-Applying Equation~(\ref{eq:tau}) to \ce{PPyr+I-} with $E_c = 2 k_\mathrm{B}T$ yields $\tau \approx 0.31\pm 0.07$, or a $\Theta$-temperature around \SI{160}{\celsius}.
-Using this value and Equation (\ref{eq:stretch}) we deduce the bounding energy for each composition, as reported in Table~\ref{tab:results}.
-Indeed, the bonding energy increases with increasing counterion condensation to reach 60\% of the carbon-carbon single bond (\SI{350}{\kilo\joule\per\mole} or $140 k_\mathrm{B}T$) in the case of \ce{PIm+Br-}.
-According to Equation (\ref{eq:geDfromB}), in this extreme case an electrostatic blob contains $\approx 10^3$ polymers which embed the head-to-body bonds in a low-polarity environment $\approx 30$ times larger than the Bjerrum length.
+Applying Equation~(\ref{eq:tau}) to \ce{PPyr+I-} with $E_c = 2 k_\mathrm{B}T$ yields $\tau \approx 0.40\pm 0.07$, or a $\Theta$-temperature around \SI{220}{\celsius}.
+Using this value and Equation (\ref{eq:stretch}) we deduce the bonding energy for each composition, as reported in Table~\ref{tab:results}.
+Indeed, the bonding energy increases with increasing counterion condensation to reach, in the case of \ce{PIm+Br-}, 85\% of the carbon-carbon single bond (\SI{350}{\kilo\joule\per\mole} or $140 k_\mathrm{B}T$).
+According to Equation (\ref{eq:geDfromB}), in this extreme case an electrostatic blob contains $\approx 10^3$ polymers which embed the head-to-body bonds in a low-polarity environment $\approx 24$ times larger than the Bjerrum length.
 Conversely, for \ce{PPyr+I-} and \ce{PPyr+Br-} the size of electrostatic blobs is comparable to $\ell_\mathrm{B}$, confirming an aqueous microenvironment.
-Quantitatively, the bonding energy of an ionic bond is inversely proportional to the relative dielectric constant $\epsilon_r$, therefore we need to suppose a local $\epsilon_r$ around 2 (typical of water-insoluble polymers) instead of 80 for water to recover the large bonding energy of \ce{PIm+Br-}.
+Quantitatively, the bonding energy of an ionic bond being inversely proportional to the relative dielectric constant $\epsilon_r$, we need to suppose a local $\epsilon_r$ around 2 (typical of water-insoluble polymers) instead of 80 for water to recover the large bonding energy of \ce{PIm+Br-}.
+
+\subsubsection{Quantitative phase diagram.}
 
 Table~\ref{tab:results} also shows the values of $A$ deduced from Equation~(\ref{eq:A}) and the measured value of $\tau$.
-$A$ varies between $10$ for \ce{PPyr+I-}, corresponding to $7$ uncondensed counterions per polymer, to $850$ for \ce{PIm+Br-} indicating a single dissociated counterion every $12$ polymers.
+$A$ varies between $7$ for \ce{PPyr+I-}, corresponding to $10$ uncondensed counterions per polymer, to $640$ for \ce{PIm+Br-} indicating a single dissociated counterion every $9$ polymers.
+We stress that even in this case of very strong counterion condensation an electrostatic blob still contains several ($\approx 90$) charges ensuring electrostatic repulsion and solubility,
 
-From purely rheological measurements, we have thus obtained microscopic information on the state of the gel: amount of charge condensation, head-to-body bounding energy and size of a procession.
+Figure~\ref{fig:phasediag} summarizes the different conformations of a procession as a function of scale and counterion condensation for our experimental $c$ and $\tau$.
+The most striking feature of this phase diagram is the narrowing of the regime between $D$ and $r_\mathrm{scr}$ (white region in Figure~\ref{fig:phasediag}) as counterion condensation increases.
+As the size of the electrostatic blob crosses the screening length, electrostatic repulsion is no longer able to sustain solubility.
+This abrupt transition between an uniform gel and a precipitate is what separates our five soluble samples from insoluble \ce{PPyr+F-}, \ce{PIm+Cl-} and \ce{PIm+F-}.
+Our rheological measurements probe the structure of the procession from the Kuhn length to the distance between crosslinks that is approximately $r_\mathrm{scr}$.
+\ce{PPyr+Cl-} and \ce{PIm+I-} display the typical structure of weakly charged polyelectrolytes in poor solvent.
+\ce{PPyr+I-} and \ce{PPyr+Br-} lie right in the crossover to strongly charged regime.
+Their correct description is possible only through Equation~(\ref{eq:geDfromB}).
+Finally \ce{PIm+Br-} is extremely close to the transition to insolubility (within experimental errors) which explains its extreme softness and wide plastic regime.
+
+From purely rheological measurements, we have thus obtained microscopic information on the state of the gel: amount of charge condensation, head-to-body bonding energy and size of a procession.
 
@@
 The hardest gels with the narrowest linear domain are obtained with aliphatic cationic moiety and counterion at the other end of the Hofmeister series that are less bound to the polymer.
-This gives rise to maximal charge dissociation in water, facilitates charge pairing of the peripheral, iterative cations with the terminal anion and thus affords a very high density of cross-links.
-In between, in the case of aliphatic cations with a counterion in the middle of the Hofmeister series (\ce{Cl-}) or in the case of aromatic cations and poorly interacting anion, charge dissociation is limited, polymers associate in long processions with strong head-to-body bonds.
-We thus obtain soft, highly deformable gels.
+This gives rise to minimal counterion condensation, free the peripheral, iterative cations and increase their pairing probability with the terminal anion and thus affords a very high density of cross-links.
+In between, in the case of aliphatic cations with a counterion in the middle of the Hofmeister series (\ce{Cl-}) or in the case of aromatic cations and poorly interacting anion, counterion condensation is important and polymers associate in long processions with strong head-to-body bonds.
+We thus obtain soft gels able to sustain large deformations before flowing.
 Coincidentally, the mesh size of our gels is always close to the procession persistence length, a regime often encountered in networks of biological semiflexible filaments as collagen or actin\cite{Meng2016}.
@@
 To conclude, our procedure yields robust, highly tunable hydrogels from short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials.
-Most importantly, systematic comparisons between the different poly(electrolytes) investigated in this study were used to establish a model linking the microsctructure of the gel to ion pair dissociation efficiency on the individual polymer chains.
+Most importantly, systematic comparisons between the different poly(electrolytes) investigated in this study were used to establish a model linking the microstructure of the gel to ion pair dissociation efficiency on the individual polymer chains.
 We hope that this model will join the procession toward future works in supramolecular assemblies.
 
-\begin{acknowledgement}
-The authors are grateful for the financial support of Université de Lyon through the program ``Investissements d'Avenir'' (ANR-1 1-IDEX-0007).
+\paragraph*{Acknowledgements.}
+The authors are grateful for the financial support of Universit\'{e} de Lyon through the program ``Investissements d'Avenir'' (ANR-1 1-IDEX-0007).
 S.M.
@@
 The Graphical TOC entry is derived from a picture by Arturo Reina, CC BY-SA 3.0, \url{https://commons.wikimedia.org/w/index.php?curid=282898}.
-\end{acknowledgement}
 
-\begin{suppinfo}
+% \subsection{This is the subsection heading style}
+% Section headings can be typeset with and without numbers.\cite{Abernethy2003}
 
-Materials and methods, detailed synthesis and characterizations of gels, GPC, HRMS and DSC data, calculated cohesion energy, and frequency sweep curves.
+% \subsubsection{This is the subsubsection style.~~} These headings should end in a full point.
 
-\end{suppinfo}
+% \paragraph{This is the next level heading.~~} For this level please use \texttt{\textbackslash paragraph}. These headings should also end in a full point.
 
-\bibliography{PALSE.bib}
+% \section{Graphics and tables}
+% \subsection{Graphics}
+% Graphics should be inserted on the page where they are first mentioned (unless they are equations, which appear in the flow of the text).\cite{Cotton1999}
+
+% \begin{figure}[h]
+% \centering
+%   \includegraphics[height=3cm]{example}
+%   \caption{An example figure caption}
+%   \label{fgr:example}
+% \end{figure}
+
+% \begin{figure*}
+%  \centering
+%  \includegraphics[height=3cm]{example2}
+%  \caption{An image from the \textit{Physical Chemistry Chemical Physics} cover gallery, set as a two-column figure.}
+%  \label{fgr:example2col}
+% \end{figure*}
+
+% \subsection{Tables}
+% Tables typeset in RSC house style do not include vertical lines.
+Table footnote symbols are lower-case italic letters and are typeset at the bottom of the table.
+Table captions do not end in a full point.\cite{Arduengo1992,Eisenstein2005}
+
+% \begin{table}[h]
+% \small
+%   \caption{\ An example of a caption to accompany a table}
+%   \label{tbl:example}
+%   \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}lll}
+%     \hline
+%     Header one/units & Header two & Header three \\
+%     \hline
+%     1 & 2 & 3 \\
+%     4 & 5 & 6 \\
+%     7 & 8 & 9 \\
+%     10 & 11 & 12 \\
+%     \hline
+%   \end{tabular*}
+% \end{table}
+
+% Adding notes to tables can be complicated.
+Perhaps the easiest method is to generate these manually.\footnote[4]{Footnotes should appear here.
+These might include comments relevant to but not central to the matter under discussion, limited experimental and spectral data, and crystallographic data.}
+
+% an example of a two-column table
+%\begin{table*}
+%\small
+  %\caption{\ An example of a caption to accompany a table, table captions do not end in a full point}
+  %\label{tbl:example}
+  %\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lllllll}
+    %\hline
+    %Header one & Header two & Header three & Header four & Header five & Header six  & Header seven\\
+    %\hline
+    %1 & 2 & 3 & 4 & 5 & 6  & 7\\
+    %8 & 9 & 10 & 11 & 12 & 13 & 14 \\
+    %15 & 16 & 17 & 18 & 19 & 20 & 21\\
+    %\hline
+  %\end{tabular*}
+%\end{table*}
+
+% \section{Equations}
+
+% Equations can be typeset inline \textit{e.g.} $ y = mx + c$ or displayed with and without numbers:
+
+%  \[ A = \pi r^2 \]
+
+% \begin{equation}
+%   \frac{\mathrm{\gamma}}{\mathrm{\epsilon}x} r^2 = 2r
+% \end{equation}
+% \pagebreak
+% \begin{widetext}
+% \begin{eqnarray} \label{e27}
+% T_{nn''}^{mm''}=
+% \begin{cases}
+% & \dfrac{(q{+}p'')!\,(p{-}p'')!}{(q{-}q'')!\,(p{+}q'')!}\,a^{m-m''}b^{n''}\,P_{p-p''}^{(m-m'',n'')}(1{-}2A^2)P_{p-p''}^{(m-m'',n'')}(2B^2{-}1)~~~ \\ & \text{when}~~~n-n''\geq m-m''\geq0~,\\
+% \\
+% & \dfrac{(p{+}q'')!\,(q{-}q'')!}{(p{-}p'')!\,(q{+}p'')!}\,a^{m''-m}b^{n''}\,P_{q-q''}^{(m''-m,n'')}(1{-}2A^2)\,P_{q-q''}^{(m''-m,n'')}(2B^2{-}1)~~~ \\ & \text{when}~~~ n-n''\geq m''-m\geq0~, \\
+% \\
+% &0 ~~~ \text{otherwise}.
+% \end{cases}
+% \end{eqnarray}
+% \end{widetext}
+
+% You can also put lists into the text. You can have bulleted or numbered lists of almost any kind. 
+% The \texttt{mhchem} package can also be used so that formulae are easy to input: \texttt{\textbackslash ce\{H2SO4\}} gives \ce{H2SO4}.
+
+% For footnotes in the main text of the article please number the footnotes to avoid duplicate symbols.
+e.g.
+\texttt{\textbackslash footnote[num]\{your text\}} the corresponding author $\ast$ counts as footnote 1, ESI as footnote 2, e.g.
+if there is no ESI, please start at [num]=[2], if ESI is cited in the title please start at [num]=[3] etc.
+Please also cite the ESI within the main body of the text using \dag.
+
+% \section{Conclusions}
+% The conclusions section should come at the end of article. For the reference section, the style file rsc.bst can be used to generate the correct reference style.
+%%%END OF MAIN TEXT%%%
+
+%The \balance command can be used to balance the columns on the final page if desired. It should be placed anywhere within the first column of the last page.
+
+%\balance
+
+%If notes are included in your references you can change the title from 'References' to 'Notes and references' using the following command:
+%\renewcommand\refname{Notes and references}
+
+%%%REFERENCES%%%
+\bibliography{PALSE} %You need to replace "rsc" on this line with the name of your .bib file
+\bibliographystyle{rsc} %the RSC's .bst file
 

### 2016-05-20 ion_pairing_SI.tex
proper references to the SI in the main text
Affiliations and detailed GPC

--- 2016-02-01 PALSE/ion_pairing_SI.tex
@@
 
-\usepackage{xr,xc}
-\externaldocument[M-]{ion_pairing_paper}
+\usepackage{zref-xr,xc}
+\usepackage{hyperref}
+\zexternaldocument[M-]{ion_pairing_paper}
 \externalcitedocument{ion_pairing_paper}
@@
 \author{Hassan Srour}
-\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}
+\author{Martien Duvall Deffo Ayagou}
+\author{Thi Thanh-Tam Nguyen}
+\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Univ Lyon, Ens de Lyon, Univ Claude Bernard, CNRS,
+Laboratoire de Chimie, F-69342 Lyon, France.}
+
+\author{Nicolas Taberlet}
+\author{Sébastien Manneville}
+\affiliation[Laboratoire de Physique de l'ENS de Lyon]{Univ Lyon, Ens de Lyon, Univ Claude Bernard, CNRS,
+Laboratoire de Physique, F-69342 Lyon, France.}
+
+\author{Chantal Andraud}
+\author{Cyrille Monnereau}
+\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Univ Lyon, Ens de Lyon, Univ Claude Bernard, CNRS,
+Laboratoire de Chimie, F-69342 Lyon, France.}
 
@@
 \email{mathieu.leocmach@univ-lyon1.fr}
-
-\author{Martien Duvall Deffo Ayagou}
-\author{Thi Thanh-Tam Nguyen}
-\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}
-
-\author{Nicolas Taberlet}
-\author{Sébastien Manneville}
-\affiliation[Laboratoire de Physique de l'ENS de Lyon]{Laboratoire de Physique, Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon, 46 Allée d'Italie, 69007 Lyon}
-
-\author{Chantal Andraud}
-\author{Cyrille Monnereau}
-\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}
-\email{cyrille.monnereau@ens-lyon.fr}
 
@@
 All reactions were performed under an argon atmosphere using schlenk techniques.
-CH2Cl2 and THF were dried and degassed on a solvent station by passage through an activated alumina column followed by argon flush.
+\ce{CH2Cl2} and THF were dried and degassed on a solvent station by passage through an activated alumina column followed by argon flush.
 All other solvents were used without additional purification.
@@
 
-SEC analyses were carried out on ca \SI{5}{\milli\gram\per\milli\litre} solutions of each polymer in DMSO LiBr (\SI{0.01}{M}), using a MALVERN VISCOTEK 430  system equipped with three PSS Gram columns (Polyester Copolymer Network) set in series at \SI{80}{\celsius}, coupled to a Viscotek HT-GPC Module 350 A (refractive and viscosimeter) detector (at \SI{80}{\celsius}) .
+SEC analyses were carried out on ca \SI{5}{\milli\gram\per\milli\litre} solutions of each polymer in DMSO LiBr (\SI{0.01}{M}), using a MALVERN VISCOTEK 430  system equipped with three PSS Gram columns (Polyester Copolymer Network) set in series at \SI{80}{\celsius}, coupled to a Viscotek HT-GPC Module 350 A (refractive and viscosimeter) detector (at \SI{80}{\celsius}).
 The polydispersity indexes (PDI = Mw/Mn) of the samples were derived from the RI signal by a universal calibration curve by pullulan.
 The number-average molar masses (Mn) were calculated from RI signals with the OmniSEC 4.6 software.
+
+The thermal analysis was obtained on Differential Scanning Calorimetry (DSC).
+All samples were investigated using a Setaram DSC 131 apparatus.
+A weighed sample of \SI{5.0\pm 1}{\milli\gram} was placed in a sealed stainless crucible.
+A standard heating rate of \SI{5}{\kelvin\per\minute} was applied from 25 up to \SI{200}{\celsius} under \SI{50}{\milli\litre\per\minute} of nitrogen.
+The start and the onset temperatures were determined in order to assess the thermal stability.
+$T_\mathrm{start}$ is the temperature at which the sample starts to lose some mass and $T_\mathrm{onset}$ was determined from the step tangent method.
 
@@
 
+Frequency sweeps were performed within the linear domain by decreasing the frequency from $f=\SI{10}{\hertz}$ down to $f=\SI{0.01}{\hertz}$ at a strain amplitude $\gamma = 0.1 \%$ and at a temperature (\SI{25}{\celsius}) fixed by a Peltier plate.
+Strain sweeps were performed by ramping up the strain amplitude from $\gamma = 0.1 \%$ up to $\gamma =5000\%$ at $f=\SI{1}{\hertz}$ in the same temperature conditions.
+Most tests had to be stopped before the strain could reach its highest value due to sample ejection from the cone-plate geometry.
+
 \begin{figure}
 \includegraphics[width=\textwidth]{sec_spectra.png}
+\begin{tabular}{llll}
+Peak RV - (ml) & 23.72 & 
+Mn - (Daltons) & 5 155 \\
+Mw - (Daltons) & 5 614 &
+Mz - (Daltons) & 5 985 \\
+Mp - (Daltons) & 7 221 &
+Mw / Mn & 1.089\\
+Percent Above Mw: 0 & 100.0 &
+Percent Below Mw: 0 & 0.0\\
+Mw 10.0\% Low & 3 012 &
+Mw 10.0\% High & 7 087\\
+IV - (dl/g) & 0.6026 &
+Rh(w) - (nm) & 3.899\\
+Wt Fr (Peak) & 1.000 &\\
+Mark-Houwink a & 0.331&
+Mark-Houwink logK & -1.453\\
+Branches & 0.000 &
+Branch Freq. & 0.000\\
+RI Area - (mvml) & 98.76 &
+UV Area - (mvml) & 0.00\\
+RALS Area - (mvml) & 0.00 &
+LALS Area - (mvml) & 0.00\\
+IVDP Area - (mvml) & 105.24\\
+\end{tabular}
 \caption{SEC Spectra of \ce{POH}.}
+\label{fig:sec}
 \end{figure}
@@
 \caption{High resolution mass spectra of \ce{PIm+Br-}. Bromide peak exist at m/z=81.0.}
+\label{fig:massPImBr}
 \end{figure}
@@
 \caption{High resolution mass spectra of \ce{PPy+Br-}. Bromide peak exist at m/z=79-81.}
+\label{fig:massPPyBr}
 \end{figure}
@@
 \caption{High resolution mass spectra of \ce{PIm+I-}. Iodide peak exist at m/z=126.9.}
+\label{fig:massPImI}
 \end{figure}
@@
 \caption{High resolution mass spectra of \ce{PPy+I-}. Iodide peak exist at m/z=127.1.}
+\label{fig:massPPyI}
+\end{figure}
+
+\begin{figure}
+\includegraphics{frequency.pdf}
+\caption{Storage modulus $G^\prime$ ($\bullet$) and loss modulus $G^{\prime\prime}$ ($\square$) measured through oscillatory shear experiments plotted against the frequency $f$.
+The moieties change with rows and the counterions with columns.
+The fixed strain amplitude is $\gamma=0.1\%$.
+All samples are at 7.4~\%~wt }
+\label{fig:frequency}
 \end{figure}

### 2016-07-06 ion_pairing_paper.tex
v2

@@
-\documentclass[journal=jacsat,manuscript=article]{achemso}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}       % Use modern font encodings
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%This is the LaTeX ARTICLE template for RSC journals
+%Copyright The Royal Society of Chemistry 2014
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\documentclass[twoside,twocolumn,9pt]{article}
+\usepackage{extsizes}
+\usepackage[super,sort&compress,comma]{natbib} 
+\usepackage[version=3]{mhchem}
+\usepackage[left=1.5cm, right=1.5cm, top=1.785cm, bottom=2.0cm]{geometry}
+\usepackage{balance}
+\usepackage{widetext}
+\usepackage{times,mathptmx}
+\usepackage{sectsty}
+\usepackage{graphicx} 
+\usepackage{lastpage}
+\usepackage[format=plain,justification=raggedright,singlelinecheck=false,font={stretch=1.125,small,sf},labelfont=bf,labelsep=space]{caption}
+\usepackage{float}
+\usepackage{fancyhdr}
+\usepackage{fnpos}
+\usepackage[english]{babel}
+\usepackage{array}
+\usepackage{droidsans}
+\usepackage{charter}
+\usepackage[T1]{fontenc}
+\usepackage[usenames,dvipsnames]{xcolor}
+\usepackage{setspace}
+\usepackage[compact]{titlesec}
+%%%Please don't disable any packages in the preamble, as this may cause the template to display incorrectly.%%%
+
 \usepackage{amssymb}
-\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
 \usepackage{siunitx}
-\usepackage[table]{xcolor}
+\usepackage{hyperref}
+\usepackage{xspace}
 
-\usepackage{xr}
-\externaldocument[S-]{ion_pairing_SI}
+\newcommand{\figSsec}{S1\xspace}
+\newcommand{\figSmassPImBr}{S2\xspace}
+\newcommand{\figSmassPPyBr}{S3\xspace}
+\newcommand{\figSmassPImI}{S4\xspace}
+\newcommand{\figSmassPPyI}{S5\xspace}
+\newcommand{\figSfrequency}{S6\xspace}
 
-\author{Hassan Srour}
-\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}
+%\usepackage{epstopdf}%This line makes .eps figures into .pdf - please comment out if not required.
 
-\author{Mathieu Leocmach}
-\affiliation[Institut Lumière Matière]{Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}
-\email{mathieu.leocmach@univ-lyon1.fr}
-
-\author{Martien Duvall Deffo Ayagou}
-\author{Thi Thanh-Tam Nguyen}
-\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}
-
-\author{Nicolas Taberlet}
-\author{Sébastien Manneville}
-\affiliation[Laboratoire de Physique de l'ENS de Lyon]{Laboratoire de Physique, Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon, 46 Allée d'Italie, 69007 Lyon}
-
-\author{Chantal Andraud}
-\author{Cyrille Monnereau}
-\affiliation[Laboratoire de Chimie de l'ENS de Lyon]{Laboratoire de Chimie UMR CNRS 5182 Ecole Normale Supérieure de Lyon/ Université Claude Bernard Lyon1/ Université de Lyon 46 Allée d'Italie, 69007 Lyon}
-\email{cyrille.monnereau@ens-lyon.fr}
-
-%\title{Supramolecular crocodile line controls the rheological properties of polyelectrolytes hydrogels}
+\definecolor{cream}{RGB}{222,217,201}
 \title{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}
-%\title{Ion pairing in crocodile line hydrogels controls rheological properties}
-%\title{In crocodile line hydrogels ion pairing controls rheological properties}
-%\title{Controlling the rheological properties of well-defined polyelectrolytes hydrogels by fine tuning their ion pairing process}
-
-\keywords{ATRP, hydrogel, poly(cations), rheology}
 
@@
 
-\begin{tocentry}
+\pagestyle{fancy}
+\thispagestyle{plain}
+\fancypagestyle{plain}{
 
-%9\,cm by 3.5\,cm
-\includegraphics{TOC2.png}
+%%%HEADER%%%
+\fancyhead[C]{\includegraphics[width=18.5cm]{head_foot/header_bar}}
+\fancyhead[L]{\hspace{0cm}\vspace{1.5cm}\includegraphics[height=30pt]{head_foot/journal_name}}
+\fancyhead[R]{\hspace{0cm}\vspace{1.7cm}\includegraphics[height=55pt]{head_foot/RSC_LOGO_CMYK}}
+\renewcommand{\headrulewidth}{0pt}
+}
+%%%END OF HEADER%%%
 
-\end{tocentry}
+%%%PAGE SETUP - Please do not change any commands within this section%%%
+\makeFNbottom
+\makeatletter
+\renewcommand\LARGE{\@setfontsize\LARGE{15pt}{17}}
+\renewcommand\Large{\@setfontsize\Large{12pt}{14}}
+\renewcommand\large{\@setfontsize\large{10pt}{12}}
+\renewcommand\footnotesize{\@setfontsize\footnotesize{7pt}{10}}
+\makeatother
 
-\begin{abstract}
-We demonstrated recently that polyelectrolytes with cationic moieties along the chain and a single anionic head are able to form physical hydrogels due to the reversible nature of the head-to-body ionic bond.
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
+\renewcommand\footnoterule{\vspace*{1pt}% 
+\color{cream}\hrule width 3.5in height 0.4pt \color{black}\vspace*{5pt}} 
+\setcounter{secnumdepth}{5}
+
+\makeatletter 
+\renewcommand\@biblabel[1]{#1}            
+\renewcommand\@makefntext[1]% 
+{\noindent\makebox[0pt][r]{\@thefnmark\,}#1}
+\makeatother 
+\renewcommand{\figurename}{\small{Fig.}~}
+\sectionfont{\sffamily\Large}
+\subsectionfont{\normalsize}
+\subsubsectionfont{\bf}
+\setstretch{1.125} %In particular, please do not alter this line.
+\setlength{\skip\footins}{0.8cm}
+\setlength{\footnotesep}{0.25cm}
+\setlength{\jot}{10pt}
+\titlespacing*{\section}{0pt}{4pt}{4pt}
+\titlespacing*{\subsection}{0pt}{15pt}{1pt}
+%%%END OF PAGE SETUP%%%
+
+%%%FOOTER%%%
+\fancyfoot{}
+\fancyfoot[LO,RE]{\vspace{-7.1pt}\includegraphics[height=9pt]{head_foot/LF}}
+\fancyfoot[CO]{\vspace{-7.1pt}\hspace{13.2cm}\includegraphics{head_foot/RF}}
+\fancyfoot[CE]{\vspace{-7.2pt}\hspace{-14.2cm}\includegraphics{head_foot/RF}}
+\fancyfoot[RO]{\footnotesize{\sffamily{1--\pageref{LastPage} ~\textbar  \hspace{2pt}\thepage}}}
+\fancyfoot[LE]{\footnotesize{\sffamily{\thepage~\textbar\hspace{3.45cm} 1--\pageref{LastPage}}}}
+\fancyhead{}
+\renewcommand{\headrulewidth}{0pt} 
+\renewcommand{\footrulewidth}{0pt}
+\setlength{\arrayrulewidth}{1pt}
+\setlength{\columnsep}{6.5mm}
+\setlength\bibsep{1pt}
+%%%END OF FOOTER%%%
+
+%%%FIGURE SETUP - please do not change any commands within this section%%%
+\makeatletter 
+\newlength{\figrulesep} 
+\setlength{\figrulesep}{0.5\textfloatsep}
+
+\newcommand{\topfigrule}{\vspace*{-1pt}% 
+\noindent{\color{cream}\rule[-\figrulesep]{\columnwidth}{1.5pt}} }
+
+\newcommand{\botfigrule}{\vspace*{-2pt}% 
+\noindent{\color{cream}\rule[\figrulesep]{\columnwidth}{1.5pt}} }
+
+\newcommand{\dblfigrule}{\vspace*{-1pt}% 
+\noindent{\color{cream}\rule[-\figrulesep]{\textwidth}{1.5pt}} }
+
+\makeatother
+%%%END OF FIGURE SETUP%%%
+
+%%%TITLE, AUTHORS AND ABSTRACT%%%
+\twocolumn[
+  \begin{@twocolumnfalse}
+\vspace{3cm}
+\sffamily
+\begin{tabular}{m{4.5cm} p{13.5cm} }
+
+\includegraphics{head_foot/DOI} & \noindent\LARGE{\textbf{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels$^\dag$}} \\%Article title goes here instead of the text "This is the title"
+\vspace{0.3cm} & \vspace{0.3cm} \\
+
+ & \noindent\large{Hassan Srour,\textit{$^{a}$} Martien Duvall Deffo Ayagou,\textit{$^{a}$} Thi Thanh-Tam Nguyen,\textit{$^{a}$}, Nicolas Taberlet,\textit{$^{b}$} S\'{e}bastien Manneville,\textit{$^{b}$} Chantal Andraud,\textit{$^{a}$} Cyrille Monnereau,\textit{$^{a}$} and Mathieu Leocmach$^{\ast}$\textit{$^{c}$}} \\%Author names go here instead of "Full name", etc.
+
+\includegraphics{head_foot/dates} & \noindent\normalsize{We demonstrated recently that polyelectrolytes with cationic moieties along the chain and a single anionic head are able to form physical hydrogels due to the reversible nature of the head-to-body ionic bond.
 Here we generate a variety of such polyelectrolytes with various cationic moieties and counterion combinations starting from a common polymeric platform.
 We show that the rheological properties (shear modulus, critical strain) of the final hydrogels can be modulated over three orders of magnitude depending on the cation/anion pair.
-Our data fit remarkably well within a theoretical model involving a supramolecular head-to-tail single file between cross-links, akin to the behaviour of pine-processionary caterpillar, the length of the procession being set by the amount of charge dissociation.
-\end{abstract}
+Our data fit remarkably well within a scaling model involving a supramolecular head-to-tail single file between cross-links, akin to the behaviour of pine-processionary caterpillar.
+This model allows the quantitative measure of the amount of counterion condensation from standard rheology procedure.
+%The abstract should be a single paragraph which summarises the content of the article.
+Any references in the abstract should be written out in full \textit{e.g.} [Surname \textit{et al., Journal Title}, 2000, \textbf{35}, 3523].
+} \\%The abstract goes here instead of the text "The abstract should be..."
 
+\end{tabular}
+
+ \end{@twocolumnfalse} \vspace{0.6cm}
+
+  ]
+%%%END OF TITLE, AUTHORS AND ABSTRACT%%%
+
+%%%FONT SETUP - please do not change any commands within this section
+\renewcommand*\rmdefault{bch}\normalfont\upshape
+\rmfamily
+\section*{}
+\vspace{-1cm}
+
+%%%FOOTNOTES%%%
+
+\footnotetext{\textit{$^{a}$~Univ Lyon, Ens de Lyon, Univ Claude Bernard, CNRS,
+Laboratoire de Chimie, F-69342 Lyon, France. }}%Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
+\footnotetext{\textit{$^{b}$~Univ Lyon, Ens de Lyon, Univ Claude Bernard, CNRS,
+Laboratoire de Physique, F-69342 Lyon, France. }}
+\footnotetext{\textit{$^{c}$~Univ Lyon, Univ Claude Bernard, CNRS, Institut Lumi\`{e}re Mati\`{e}re, F-69622 Villeurbanne, France.
+E-mail: mathieu.leocmach@univ-lyon1.fr}}%Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
+
+%Please use \dag to cite the ESI in the main text of the article.
+%If you article does not have ESI please remove the the \dag symbol from the title and the footnotetext below.
+\footnotetext{\dag~Electronic Supplementary Information (ESI) available: GPC chromatogram of POH, HRMS spectra before and after metathesis, frequency dependent rheology. See DOI: 10.1039/b000000x/}
+%additional addresses can be cited as above using the lower-case letters, c, d, e... If all authors are from the same address, no letter is required
+
+%\footnotetext{\ddag~Additional footnotes to the title and authors can be included \emph{e.g.}\ `Present address:' or `These authors contributed equally to this work' as above using the symbols: \ddag, \textsection, and \P.
+Please place the appropriate symbol next to the author's name and include a \texttt{\textbackslash footnotetext} entry in the the correct place in the list.}
+
+%%%END OF FOOTNOTES%%%
+
+%%%MAIN TEXT%%%%
+\begin{figure*}
+\includegraphics{synthesis_rsc.pdf}
+\caption{Synthesis of \ce{PIm+Br-} and \ce{PPyr+Br-} and their intermediates \ce{POH} and \ce{PBr} with $N_0=70$.}
+\label{sch:synthesis}
+\end{figure*}
 \section{Introduction}
 Since the pioneering work of \citet{Wichterle1960} which established their relevance in a biomedical context, polymer-based hydrogels have never ceased to be a very active field of research\cite{Calo2015,Hoffman2001}.
-They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargos or scaffolds for wound healing or tissue repair\cite{Vashist2014,Ratner2004}.
+They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargoes or scaffolds for wound healing or tissue repair\cite{Vashist2014,Ratner2004}.
 Polyelectrolytes are being extensively put to use in this particular context\cite{Rosso2003,BinImran2014}.
 Their gel formation mechanism often involves reversible-by-nature electrostatic interactions, which can be used for instance to trigger ``smart'' release of bioactive substances\cite{Khare1993,Lockwood2007}.
-Besides, poly(cationic) gels have been reported to combine scaffold properties for cell’s adherence and growth with antimicrobial activity, and are therefore increasingly used in clinical applications\cite{Hoffman2001,Landers2002}.
+Besides, poly(cationic) gels have been reported to combine antimicrobial activity with scaffold properties for the adherence and growth of cells, and are therefore increasingly used in clinical applications\cite{Hoffman2001,Landers2002}.
 In this framework, injectability is a key feature, as it provides an easier way to gel delivery in vivo\cite{Tibbitt2016}; thus, reversible shear thinning biocompatible hydrogels are a particularly sought after class of materials\cite{Yu2008}.
@@
 Contrary to a small electrolyte for which full ion pairs dissociation is generally achieved in water, polyelectrolytes in solution are in general not fully dissociated, with a substantial fraction of the counterions bound to the polymer\cite{Manning1979}.
-The resulting net charge of the chain governs the physical properties of the polyelectrolyte solution\cite{Rubinstein1996}, first of all the ability of the polymer to dissolve.
+The resulting net charge of the chain governs the physical properties of the polyelectrolytes solution~\cite{DeGennes1976,Khokhlov1980,Raphael1990,Dobrynin1995,Rubinstein1996}, first of all the ability of the polymer to dissolve in a poor solvent.
 Counterion condensation is the physical bounding or adsorption of counterions near the polymer chain.
@@
 This so-called Hofmeister series has since proved quite universally valid\cite{Zhang2010a}, including multi-charged polymers such as naturally occurring proteins or synthetic polyelectrolytes as well as charged colloidal particles\cite{Schwierz2010,Oncsik2015} or soft matter interfaces in general\cite{Leontidis2014}.
-However, depending on the nature of the polymer (hydrophilic or hydrophobic, anionic or cationic), the Hofmeister series can be direct (in short, well hydrated ions promote condensation) or reverse\cite{Schwierz2010,Oncsik2015}.
+However, depending on the nature of the polymer (hydrophilic or hydrophobic, anionic or cationic), the Hofmeister series can be direct (in short, well hydrated ions promote condensation) or reverse~\cite{Schwierz2010,Oncsik2015}.
 
 We have recently reported on a highly reversible poly(cation) based hydrogel\cite{Srour2014}, which formation relies on a new concept of supramolecular  electrostatic interaction.
-Briefly, the atom transfer radical polymerisation (ATRP) is initiated by a phosphonate-terminated ATRP initiator, eventually affording an anion-terminated poly(cationic) polymer.
-The resulting poly(cationic) polymer, when dispersed in an aqueous medium, stabilizes hydrogel formation through the occurrence of a head(anion)-to-body(cations) supramolecular network.
+Briefly, the atom transfer radical polymerisation (ATRP) is initiated by a phosphonate-terminated ATRP initiator, eventually affording an \emph{anion-terminated poly(cationic) polymer}.
+When dispersed in an aqueous medium, this polymer forms an hydrogel through the occurrence of a head(anion)-to-body(cations) supramolecular network.
 This highly dynamic electrostatic network provides the resulting gel with spectacular mechanical and self-healing properties.
@@
 We synthesise the phosphonate terminated polymer hereafter referred to as \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
-We obtain a well-controlled linear polymer ($Mw= \SI{5614}{\dalton}$, $Mw/Mn = 1.08$; see Supplementary Materials and Supplementary Figure~S\ref{S-fig:sec}), with a degree of polymerisation $N_0=70$.
+We obtain a well-controlled linear polymer ($M_w= \SI{8200}{\dalton}$ from NMR, \SI{5614}{\dalton} from GPC with $M_w/M_n = 1.08$; see Supplementary Materials and Supplementary Figure~\figSsec), with a degree of polymerisation $N_0=70$ (NMR).
 This polymer serves as a common scaffold from which we derive  all studied systems.
 Nucleophilic addition of N-methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF affords the corresponding poly(imidazolium) and poly(pyrrolidinium) compounds.
-In these cases, as an inherent consequence of the structure of the starting material, charge neutrality is provided by bromide counter ions, and the polymer will be referred to as \ce{PIm+Br-} and \ce{PPyr+Br-}, respectively (Scheme~\ref{sch:synthesis}).
+In these cases, as an inherent consequence of the structure of the starting material, charge neutrality is provided by bromide counter ions, and the polymer will be referred to as \ce{PIm+Br-} and \ce{PPyr+Br-}, respectively (Figure~\ref{sch:synthesis}).
 
-\begin{scheme}
-\includegraphics{synthesis.pdf}
-\caption{Synthesis of \ce{PIm+Br-} and \ce{PPyr+Br-} and their intermediates \ce{POH} and \ce{PBr} with $N_0=70$.}
-\label{sch:synthesis}
-\end{scheme}
-
-We performed anionic metathesis by pouring an aqueous solution of \ce{PIm+Br-} or \ce{PPyr+Br-} into a saturated aqueous solution of the different sodium halides (\ce{NaF}, \ce{NaCl}, \ce{NaI}) (Scheme~\ref{sch:metathesis}).
+We performed anionic metathesis by pouring an aqueous solution of \ce{PIm+Br-} or \ce{PPyr+Br-} into a saturated aqueous solution of the different sodium halides (\ce{NaF}, \ce{NaCl}, \ce{NaI}) (Figure~\ref{sch:metathesis}).
 In the cases of \ce{PIm+F-}, \ce{PIm+Cl-} , \ce{PPyr+F-} and \ce{PPyr+Cl-}, we obtain a turbid suspension immediately after addition.
@@
 In order to bring evidence for the efficiency of the ionic metathesis, we submit the resulting materials to negative-mode high-resolution mass spectrometry (HRMS, see Supplementary Materials).
-We unambiguously assess complete displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Supplementary Figures~S\ref{S-fig:massPImBr} and S\ref{S-fig:massPPyBr}).
-We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures~S\ref{S-fig:massPImI} and S\ref{S-fig:massPPyI}), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (not shown).
+We unambiguously assess complete displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Supplementary Figures~\figSmassPImBr and \figSmassPPyBr).
+We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures~\figSmassPImI and \figSmassPPyI), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (not shown).
 
-\begin{scheme}
-\includegraphics{metathesis.pdf}
+\begin{figure}
+\includegraphics{metathesis_rsc.pdf}
 \caption{(a) Anionic metathesis reaction. As above, \ce{Nu+} corresponds either to pyrrolidinium or imidazolium. (b) Summary of the polymers derived from a single batch of \ce{PBr}.}
 \label{sch:metathesis}
-\end{scheme}
+\end{figure}
 
-Differential Scanning Calorimetry (DSC) thermograms  of all synthesized polymers also reveals marked differences in their properties, that can only be explained by differences in their couterion features.
+Differential Scanning Calorimetry (DSC) thermograms  of all synthesized polymers also reveals marked differences in their properties, that can only be explained by differences in their counterion features.
 Thus, in the case polyelectrolytes with fluoride and chloride counterion, we observe no exo or endothermic transition below \SI{200}{\celsius} (featureless thermograms, not shown).
@@
 When bromide counterions are present, we observe a significant increase of the peak temperature (114-\SI{115}{\celsius}) but again, with a similar value between \ce{PIm+Br-} and \ce{PPyr+Br-} as shown in Figure~\ref{fig:dsc}.
-The overall shape and position of the DSC peaks is very reminiscent of previously reported data for various naturally occuring or synthetic poly(electrolytes)\cite{Li2005,Sarmento2006,Ostrowska-Czubenko2009a,Moin2015}.
+The overall shape and position of the DSC peaks is very reminiscent of previously reported data for various naturally occurring or synthetic poly(electrolytes)\cite{Li2005,Sarmento2006,Ostrowska-Czubenko2009a,Moin2015}.
 It is generally attributed to desorption of weakly bound water from the polymer network.
-%Its presence in the polymer samples where bromide and iodide are used as couterions suggests that adsorbed water could play a role in solvating the dissociated ion pairs in the latter.
+%Its presence in the polymer samples where bromide and iodide are used as counterions suggests that adsorbed water could play a role in solvating the dissociated ion pairs in the latter.
 Conversely, its absence for chloride and fluoride counterions suggests that dissociation is not effective because of stronger ion pair interaction.
@@
 With fluoride counterions, or in the case of \ce{PIm+Cl-}, we observe neither swelling nor dissolution.
-We recover a 8\%wt suspension as a biphasic mixture of its individual components.
+We recover a 7.4~\%~wt suspension as a biphasic mixture of its individual components.
 
@@
 Locally, the monomers are condensed in collapsed globules to minimize the contacts with water molecules, and solubility on larger scales can only be achieved if the polymer bears enough charges\cite{Khokhlov1980,Raphael1990}.
-To estimate the $\Theta$-temperature, that delimit bad and good solvent behaviour, we held the two fluoride polymers in water at boiling temperature overnight.
+To estimate the $\Theta$-temperature, that delimit poor and good solvent behaviour, we held the two fluoride polymers in water at boiling temperature overnight.
 No dissolution or swelling was observed, implying that $\Theta>\SI{100}{\celsius}$.
-%Nonetheless, due to charge dissociation, five of our polyelectrolytes are soluble at \SI{25}{\celsius}.
-
-%\begin{figure}
-%\caption{pictures of 8\%wt water solutions of \ce{PIm+X-} and \ce{PPyr+X-} in water.}
-%\label{fig:pictures}
-%\end{figure}
 
 \subsection{Rheological measurements}
-\begin{figure}[t]
-\includegraphics{strain.pdf}%
+\begin{figure*}
+\includegraphics{strain_rsc.pdf}%
 \caption{Storage modulus $G^\prime$ ($\bullet$) and loss modulus $G^{\prime\prime}$ ($\square$) measured through oscillatory shear experiments plotted against the strain amplitude $\gamma$.
@@
 The fixed frequency is $f=\SI{1}{\hertz}$.
-All samples are at 8\%wt.
-The gray area limits of the linear, plastic and shear thinning regimes.}%
+All samples are at 7.4~\%~wt.
+The gray area encompasses the plastic regime that lies between the linear regime at very low amplitudes and the fluid-like shear-thinning regime at high strain amplitude.}%
 \label{fig:strain}%
-\end{figure}
+\end{figure*}
 
 We perform rheological studies of the water-swelling materials.
-Briefly, data are recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}\cite{Macosko1994,Larson1999}.
+Briefly, data are recorded at \SI{25}{\celsius} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}\cite{Macosko1994,Larson1999}.
 We place the sample on the plate, and then lower the cone to the measuring position, spreading the sample in the process.
@@
 To minimize water absorption, we cover the geometry with a solvent trap, using light mineral oil as a liquid seal between the rotor and the cap.
-After one minute of equilibration we perform an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$, see Supplementary Figure~S\ref{S-fig:frequency}) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$).
+After one minute of equilibration we perform an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$, see Supplementary Figure~\figSfrequency) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$).
 For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile is that of pure water which confirms the visual observation.
-However as shown in Figure~\ref{fig:strain}, all other samples are solidlike ($G^\prime \gg G^{\prime\prime}$) at small strain and flow at large strains ($G^{\prime\prime} \gg G^\prime$) with a steep decrease of the moduli.
-We confirm solidlike behavior at low strain for all accessible frequencies (see supplementary material S6 Fig.
-1).
+However all other samples behave as shear-thinning yield stress fluids, confirming our previous results with \ce{PIm+Br-}~\cite{Srour2014}.
+As shown in Figure~\ref{fig:strain}, soluble samples are solid-like ($G^\prime \gg G^{\prime\prime}$) at small strain and flow at large strains ($G^{\prime\prime} \gg G^\prime$) with a steep decrease of the moduli.
+We confirm solid-like behavior at low strain for all accessible frequencies (see Supplementary Figure~\figSfrequency).
 We checked that the linear mechanical properties are unchanged by the flow history given a few minutes of rest.
@@
 Imidazolium based gels are much weaker than their pyrrolidinium counterpart.
-Comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ value (13 and \SI{8100}{\pascal}, resp.) differ by almost three orders of magnitude.
-On the bottom row of Figure~\ref{fig:strain} we can compare the rheological responses of Poly(pyrrolidinium)-based hydrogels for the three counterions allowing dissolution in water.
+Comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ values (\SI{13}{\pascal} and \SI{8100}{\pascal}, resp.) differ by almost three orders of magnitude.
+On the bottom row of Figure~\ref{fig:strain} we can compare the rheological responses of poly(pyrrolidinium)-based hydrogels for the three counterions allowing dissolution in water.
 Gels with bromide and iodide counterions show roughly the same modulus value at vanishing strain for both $G^\prime$ and $G^{\prime\prime}$.
@@
 The linear regime is either extremely narrow or non existent in heavily charged \ce{PPyr+Br-} and \ce{PPyr+I-}.
-By contrast the lightly charged \ce{PIm+Br-} displays a broad intermediate regime (tree decades of strain) between the linear regime and the crossing of the moduli.
+By contrast the lightly charged \ce{PIm+Br-} displays a broad plastic regime (tree decades of strain) between the linear regime and the crossing of the moduli.
 In any case, the end of the linear regime corresponds to a strain much smaller than 1.
@@
 \begin{figure}
-\includegraphics{procession.pdf}
+\centering\includegraphics{procession.pdf}
 \caption{Processionary model.
-Sketch of the network in the case of a procession of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{src}\approx 3D$.
+Sketch of the network in the case of a procession of size $n=3$ between cross-links (gray disks) and a persistence length $r_\mathrm{scr}\approx 6D$.
 Empty circles are not individual monomers but electrostatic blobs of size $D$.
 Electrostatic blobs containing an anionic head are shown as black filled circles.}
-\label{fig:dissociation}
+\label{fig:procession}
 \end{figure}
+\begin{figure*}
+\centering\includegraphics{scales.pdf}
+\caption{Conformation of semi-dilute polyelectrolytes at rest at various scales. Length increases from left to right.}
+\label{fig:scales}
+\end{figure*}
 
@@
 Conversely, the latter case should be observed at low charge dissociation.
-In between these two limiting cases, we should observe cross-link points separated by processions of $n N_0$ monomers as sketched in Figure~\ref{fig:dissociation}a.
+In between these two limiting cases, we should observe cross-link points separated by processions of $n N_0$ monomers as sketched in Figure~\ref{fig:procession}.
 
@@
 the elasticity of the undamaged gel network, (ii) the strain amplitude $\gamma_0$ corresponding to the end of the linear regime~\cite{Hyun2011}, and (iii) the strain amplitude $\gamma_c$ corresponding to the crossing of the moduli.
-By using these three parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters as the average number of chains between cross-links $n$, the number of monomers between dissociated ion pairs $A$ and the head-to-body bonding energy $E_c$.
+By using these three parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters such as the average number of chains between cross-links $n$, the number of monomers between dissociated ion pairs $A$ that quantify counterion condensation and the head-to-body bonding energy $E_c$.
 
-\subsubsection{Conformation at rest}
+\subsubsection{Conformation at rest.}
+The scaling theory of \citet{Dobrynin1995} describes the structure of a polyelectrolyte chain.
+In the following, we make the link between this theory and the chemical structure of our polymers.
+In absence of strain, the structure of a polyelectrolyte chain is organised at various scales, sketched on Figure~\ref{fig:scales}.
 
-The scaling theory of \citet{Rubinstein1996} describes the structure of a polyelectrolyte chain.
-In the following, we make the link between this theory and the chemical structure of our polymers.
-In absence of strain, the structure of a polyelectrolyte chain is organised at various scales.
-
+\paragraph{Kuhn length.}
 The smallest scale is the Kuhn length $b$ containing $g_\mathrm{K}$ monomers.
@@
 
-In bad solvent, the thermal length $\xi_T$ is larger than $b$.
-The thermal blob contains $g_T$ monomers and its density is set by the reduced temperature $(g_T/g_\mathrm{K})/\xi_T^3 = \tau/b^3$.
-Between $b$ and $\xi_T$ we have a persistent random walk $\xi_T = b (g_T/g_\mathrm{K})^{1/2}$.
-These conditions yields $\xi_T = b/\tau$ and $g_T = g_\mathrm{K}/\tau^2$.
+\paragraph{Thermal length.}
+At the thermal length $\xi_T$, the attractive potential between monomers is balanced by thermal energy $k_\mathrm{B}T$, with $k_\mathrm{B}$ the Boltzmann constant.
+This balance of energy can be written function of the reduced temperature $\tau = 1 -T/\Theta$ and the number $g_T$ of monomers in the thermal blob:
+\begin{equation}
+k_\mathrm{B}T = \left(\frac{b}{\xi_T}\right)^3\left(\frac{g_T}{g_K}\right)^2 \tau k_\mathrm{B}T 
+\label{eq:poorkT}
+\end{equation}
+Between $b$ and $\xi_T$ we have a persistent random walk $\xi_T = b (g_T/g_\mathrm{K})^{1/2}$. These conditions yields 
+\begin{equation}
+\xi_T = b/\tau,\qquad g_T = g_\mathrm{K}/\tau^2
+\label{eq:thermal}
+\end{equation}
 
-At a scale $D$, the electrostatic energy balances the surface energy of the collapsed polymer.
-We name $g_e$ the number of monomers in the electrostatic blob.
-If there is a charge every $A$ monomers then there are $g_e/A$ charges in the electrostatic blob.
-The surface tension is proportional to the number of thermal blobs exposed on the surface $k_\mathrm{B}T/\xi_T^2$, with $k_\mathrm{B}$ the Boltzmann constant and $T$ the temperature.
-Thus the balance of energies writes:
+\paragraph{Electrostatic length, weak charging case.}
+Following \citet{Khokhlov1980} one can further define a third characteristic length $D$ over which the electrostatic energy become dominant over short range attraction or surface energy of the collapsed polymer.
+$D$ defines the size of the electrostatic blob and we name $g_e$ the number of monomers in it.
+Due to counterion condensation, not all monomers are charged.
+Counterion condensation is quantified by assuming that there is a charge every $A$ monomers.
+Therefore there are $g_e/A$ charges in the electrostatic blob and the electrostatic energy reads $E_e = (g_e/A)^2 e^2/(4\pi\epsilon D)$, with $e$ the elementary charge and $\epsilon$ the dielectric constant of the solvent.
+If $D>\xi_T$ the main opposition to the electrostatic energy can be understood as a surface energy proportional to the number of thermal blobs exposed on the surface $E_s = k_\mathrm{B}T (D/\xi_T)^2$.
+The balance of energies yields:
 \begin{equation}\left(\frac{D}{\xi_T}\right)^2 = \left(\frac{g_e}{A}\right)^2 \frac{\ell_\mathrm{B}}{D},
@@
 \end{equation}
-where $\ell_\mathrm{B} = e^2/(4\pi\epsilon k_\mathrm{B}T)$ is the Bjerrum length with $e$ the elementary charge, $\epsilon$ the dielectric constant of the solvent.
+where $\ell_\mathrm{B} = e^2/(4\pi\epsilon k_\mathrm{B}T)$ is the Bjerrum length.
 In water $\ell_\mathrm{B} \approx \SI{0.7}{\nano\metre}$.
-Since thermal blobs fill the volume of the electrostatic blob $D = \xi_T \left(g_e/g_T\right)^{1/3}$, we obtain
+Since thermal blobs fill the volume of the electrostatic blob, one has $D = \xi_T \left(g_e/g_T\right)^{1/3}$ and using Equation~\eqref{eq:thermal} we obtain
 \begin{equation}
@@
 
-We define the extension parameter $B$ as the ratio between the length of a chain of $g_e$ monomers fully extended and the actual size of the electrostatic blob: 
+\citet{Dobrynin1995} introduced the extension parameter $B$ defined as the ratio between the length of a fully extended chain of $g_e$ monomers ($g_e/g_\mathrm{K}$ Kuhn segments) and the actual size of the electrostatic blob:
 \begin{equation}
-B = \frac{g_e}{g_\mathrm{K}}\frac{b}{D} = \left(\frac{A^2}{u}\right)^{2/3} g_\mathrm{K}^{-1/3}\tau
+B = \frac{g_e}{g_\mathrm{K}}\frac{b}{D} = \left(\frac{A^2}{u g_\mathrm{K}^2}\right)^{2/3} \tau
 \label{eq:B}
@@
 
-At constant solvent quality and polymer architecture, $B$ is thus monotonically related to the amount of counterion condensation. In the following we will estimate $B$ from rheological measurements.
+At constant solvent quality and polymer architecture, $B$ is thus monotonically related to the amount of counterion condensation $A$.
+In the following we will use $B$ as our main variable and estimate it from rheological measurements.
+We conveniently combine Equations~(\ref{eq:geD}) and (\ref{eq:B}) to express $g_e$ and $D$ function of $B$:
+\begin{equation}
+g_e = \left(\frac{B^3}{\tau}\right)^{1/2} g_\mathrm{K},\qquad D = b \left(\frac{B}{\tau}\right)^{1/2}.
+\label{eq:geDfromlargeB}
+\end{equation}
 
-In dilute solution, the chain is a linear rod of length $L= N b / B g_\mathrm{K}$.
-The overlap concentration is thus $c^* = N/L^3 = g_\mathrm{K}B / b L^2$.
-Following \cite{Dobrynin1995}, for a monomer concentration $c > c^*$ and in absence of salt the screening length is
+\paragraph{Electrostatic length, strong charging case.}
+From Equations~(\ref{eq:thermal}) and (\ref{eq:geDfromlargeB}) we observe that for $B<1/\tau$ the size of the electrostatic blob should be smaller than the thermal length and the assumptions behind Equation~(\ref{eq:electrosurface}) break down.
+To our knowledge the study of polyelectrolytes in poor solvent by \citet{Khokhlov1980} and subsequent literature~\cite{Dobrynin1995,Rubinstein1996} focused on the weak charging regime and did not treat the small $B$ regime.
+When $B\ll 1/\tau$, most of the counterions are not condensed and attraction potential between monomers competes directly against the electrostatic repulsion.
+The size of the electrostatic blob is given by the following balance of energies:
 \begin{equation}
-r_\mathrm{src} = L \left(\frac{c^*}{c}\right)^{1/2} =  \left(\frac{B g_\mathrm{K}}{cb}\right)^{1/2}.
+\left(\frac{b}{D}\right)^3\left(\frac{g_e}{g_K}\right)^2 \tau = \left(\frac{g_e}{A}\right)^2 \frac{\ell_\mathrm{B}}{D}
+\label{eq:poorelec}
+\end{equation}
+yielding, with $D = b(g_e/g_\mathrm{K})^{1/2}$,
+\begin{equation}
+D = b \left(\frac{A^2\tau}{u g_\mathrm{K}^2}\right)^{1/2}, \qquad g_e = \frac{A^2\tau}{u g_\mathrm{K}},
+\end{equation}
+
+In this weak counterion condensation regime, the definition of the extension parameter $B$ yields a different relation with $A$ and both $D$ and $g_e$:
+\begin{equation}
+B = \frac{g_e}{g_\mathrm{K}}\frac{b}{D} = \left(\frac{A^2\tau}{u g_\mathrm{K}^2}\right)^{1/2}, \qquad D = bB, \qquad g_e = B^2 g_\mathrm{K}
+\label{eq:geDfromsmallB}
+\end{equation}
+
+Here we shall describe the crossover between weak and strong counterion condensation regimes by combining Equations~\eqref{eq:geDfromlargeB} and \eqref{eq:geDfromsmallB} into
+\begin{equation}
+D = bB (1+B\tau)^{-1/2}, \qquad
+g_e = g_\mathrm{K} B^2 (1+B\tau)^{-1/2},
+\label{eq:geDfromB}
+\end{equation}
+and we generalize the relation between $A$ and $B$ as
+\begin{equation}
+A = u^{1/2} g_\mathrm{K} \frac{B}{\tau^{1/2}}(1+B\tau)^{-1/4}.
+\label{eq:A}
+\end{equation}
+
+\paragraph{Screening length.}
+The next length scale is the screening length $r_\mathrm{scr}$.
+Between $D$ and $r_\mathrm{scr}$ the electrostatic blobs are organised in a linear rod containing $g_\mathrm{scr}$ monomers.
+A rod is $B$ times shorter than the fully extended $g_\mathrm{scr}/g_\mathrm{K}$ Kuhn segments that it contains such that
+\begin{equation}
+r_\mathrm{scr}= g_\mathrm{scr} b / (B g_\mathrm{K}).
+\label{eq:rod}
+\end{equation}
+To obtain $r_\mathrm{scr}$ we follow \cite{Dobrynin1995} by considering in a first step the case of a dilute solution where the length $L$ of a single chain of $\mathcal{N}$ monomers is shorter than the screening length.
+Equation~(\ref{eq:rod}) becomes $L= \mathcal{N} b / (B g_\mathrm{K})$.
+Therefore the overlap concentration is $c^* = \mathcal{N}/L^3 = B g_\mathrm{K} / (b L^2)$.
+In a second step, when the monomer concentration $c$ is larger than $c^*$ and in absence of salt, the screening length is
+\begin{equation}
+r_\mathrm{scr} = L \left(\frac{c^*}{c}\right)^{1/2} =  \left(\frac{B g_\mathrm{K}}{cb}\right)^{1/2}.
 \label{eq:rscrNoSalt}
@@
 
-In the presence of added salt, electrostatic interactions are screened at the Debye length $r_\mathrm{src} = \left(\ell_\mathrm{B} c_s\right)^{-1/2}$. For arbitrary salt concentration,
+When charge screening is mainly due to the presence of added salt at a concentration $c_s$ much larger than the dissociated counterions, electrostatic interactions are screened at the Debye length $r_\mathrm{scr} = \left(\ell_\mathrm{B} c_s\right)^{-1/2} \equiv \kappa^{-1}$.
+Finally, for arbitrary salt concentration \citet{Dobrynin1995} use the crossover expression,
 \begin{equation}
-r_\mathrm{src} = \left(\frac{B g_\mathrm{K}}{cb}\right)^{1/2} \left(1+ B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
+r_\mathrm{scr} = \left(\frac{B g_\mathrm{K}}{cb}\right)^{1/2} \left(1+ B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
 \label{eq:rscr}
@@
 
-Between $D$ and $r_\mathrm{src}$ the electrostatic blobs are organised in a linear rod containing $g_\mathrm{src}$ monomers.
-The length of this rod is $B$ times shorter than if the electrostatic blobs were extended:
+From Equations~(\ref{eq:rod}) and (\ref{eq:rscr}) we obtain
 \begin{equation}
-g_\mathrm{src} = B \frac{r_\mathrm{src}}{b} g_\mathrm{K} = \left(\frac{B^3 g_\mathrm{K}^3}{cb^3}\right)^{1/2} \left(1 + B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
+g_\mathrm{scr} = \left(\frac{B^3 g_\mathrm{K}^3}{cb^3}\right)^{1/2} \left(1 + B u \frac{c_s}{c} g_\mathrm{K}\right)^{-1/2}.
 \label{eq:gscr}
@@
 
+\paragraph{Correlation length.}
 The last length scale is the correlation length $\xi$.
-Between $r_\mathrm{src}$ and $\xi$ the chain forms a self avoiding walk of persistence length $r_\mathrm{src}$.
+Between $r_\mathrm{scr}$ and $\xi$ the chain forms a self avoiding walk of persistence length $r_\mathrm{scr}$.
 Above the correlation length the polyelectrolyte chain forms a random walk of correlation blobs containing $g=c\xi^3$ monomers.
 
-\subsubsection{From modulus to procession length}
+\subsubsection{From modulus to procession length.}
 
-If between two crosslinks or entanglements the ``procession'' of polymer chains contains more monomers than $g$ or $g_\mathrm{src}$ it performs a random or self avoiding walk respectively.
-Each chain is thus an entropic spring of constant stiffness $k_\mathrm{B}T$.
-The number density of chains is $c/N$ with $c$ the monomer concentration and $N$ the number of monomers between cross-links or entanglements.
+Let us note $N$ the number of monomers between two crosslinks or entanglements.
+If $N>g$ the ``procession'' of polymer chains performs a random walk of correlation blobs, each being a self-avoiding walk.
+If $g>N>g_\mathrm{scr}$ the procession performs only a self-avoiding walk.
+In any case, each procession is thus an entropic spring of constant stiffness $k_\mathrm{B}T$.
+The number density of procession is $c/N$.
 Therefore, the shear modulus at low strain is given by:
 \begin{equation}
-G = \frac{c}{N}k_\mathrm{B}T,
+G = \frac{c}{N}k_\mathrm{B}T.
 \label{eq:G}
 \end{equation}
-where $c$ is the monomer number density.
 Because of the different weights of the pyrrolidinium and imidazolium moieties and of the different counterions, $c$ is not constant as we chose to conduct our experiments at constant weight fraction $w$ of polymer (generally used to quantify gelation ability of a given gelator).
 This explains why the respective modulus or critical strains of \ce{PPyr+I-} (heavier, lower number density) and \ce{PPyr+Br-} are in reverse order with respect to the Hofmeister series.
-Taking into account the molecular mass $M$ of each polymer and  knowing the density $d$ of the solvent, we obtain
+Taking into account the molecular mass $M$ of each polymer, the Avogadro number $\mathcal{N_A}$ and  knowing the density $d$ of the solvent, we obtain the number $n = N/N_0$ of chains between cross-link
 \begin{equation}
@@
 
-We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 880 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
+We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 800 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
 This larger cross-link ratio is consistent with a higher probability of attaching two or more heads on a highly dissociated body.
 
-\begin{table}
+\begin{table*}
 \sisetup{
@@
 %table-format = 1.3(1),
-}
-\begin{tabular}{@{}l|SSS|SSSSSS@{}}
+}\centering
+\begin{tabular}{@{}lSSSSSSSSS@{}}\hline
 & {$G^\prime(\gamma\rightarrow 0)$} & {$\gamma_0$} & {$\gamma_c$} & {$n$} & {$B$} & {$E_c$} & {$E_c$} & {$D$} & {$A$}\\
 &	{\si{\pascal}} & {\%} & {\%} & & & {\si{\kilo\joule\per\mole}}& {$k_\mathrm{B}T$} & {$\ell_B$}&\\\hline&&&&&\\[-10pt]
-\ce{PIm+Br-}	& 13	&	1.1	&	810	&	860	&	930	&	210	&	84	&	13	&	460\\
-\ce{PIm+I-}		& 78	&	1.8	&	18	&	122	&	126	&	80	&	32	&	5	&	103\\
-\ce{PPyr+Cl-}	& 270 	&	2.8	&	12	&	50.6	&	64	&	46	&	19	&	3.4	&	62\\
-\ce{PPyr+Br-}	& 9800 	& 	0.3	&	2	&	1.15	&	2.8	&	5.05	&	2.02	&	0.7	&	6\\
-\ce{PPyr+I-}	& 8800	&	0.0	&	89	&	1.08	&	2.5	&	5	&	2	&	0.7	&	5.5\\
+\ce{PIm+Br-}	&	13	&	1.1	&	800	&	800	&	830	&	304	&	123	&	24	&	640\\
+\ce{PIm+I-}	&	78	&	1.8	&	18	&	113	&	113	&	116	&	46.8	&	8.7	&	140\\
+\ce{PPyr+Cl-}	&	272	&	2.8	&	10	&	46.8	&	58.2	&	66	&	26.6	&	6.2	&	90\\
+\ce{PPyr+Br-}	&	9810	&	0.3	&	2	&	1.1	&	2.6	&	5.2	&	2.1	&	1.0	&	7.1\\
+\ce{PPyr+I-}	&	8800	&	0.1	&	90	&	1.0	&	2.4	&	5.0	&	2.0	&	0.9	&	6.5\\
+\hline
 \end{tabular}
-\caption{Summary of rheological measurements and microscopic values deduced from the model. $E_c$, $D$ and $A$ are obtained supposing $\tau=0.46$.}
+\caption{Summary of rheological measurements and microscopic values deduced from the model. $E_c$, $D$ and $A$ are obtained by assuming $\tau=0.40$}
 \label{tab:results}
-\end{table}
+\end{table*}
 
-\subsubsection{Limit of the linear regime}
+\subsubsection{Limit of the linear regime.}
 
-As the material is strained, the random walk is stretched, then the self-avoiding walk.
-When the correlation blobs are fully stretched, the chain is a linear assembly of electrostatic blobs.
-Stretching the chain further means exposing more thermal blobs to the solvent, a process that cannot be reduced to spring of constant stiffness.
+As the material is strained, individual processions are stretched, starting from the larger scales~\cite{Pincus1976}.
+As sketched in Figure~\ref{fig:stretch}, when the the correlation blobs are fully stretched, the procession is a linear assembly of electrostatic blobs.
+Stretching the chain further means exposing more thermal blobs to the solvent, a process that cannot be expressed using the model of a spring of constant stiffness.
 This condition is thus the limit between the linear and non linear regimes.
@@
 For the same quality of solvent $B$ is monotonically related to the amount of counterion condensation.
-We thus confirm that our samples are sorted by decreasing counterion condensation (decreasing $B$).
-More quantitatively, the number $A$ of monomers between dissociated charges is
-\begin{equation}
-A = u^{1/2} \left(\frac{B}{\tau}\right)^{3/4} g_\mathrm{K}^{1/4}
-\label{eq:A}
-\end{equation}.
-
+Since our samples are sorted by increasing $B$ we confirm that they are sorted by decreasing counterion condensation.
+To be more quantitative, we extract the number $A$ of monomers between dissociated charges from Equation~(\ref{eq:A})%:
+%\begin{equation}
+%A = u^{1/2} \left(\frac{B}{\tau}\right)^{3/4} g_\mathrm{K}
+%\label{eq:A}
+%\end{equation}
+.
+%
 We know that $\Theta>\SI{100}{\celsius}$ so $0.2<\tau<1$. In the following we will deduce a more precise measure of $\tau$ and thus values of $A$.
@@
 
-\subsubsection{Extension of the electrostatic blobs}
+\subsubsection{Extension of the electrostatic blobs.}
+\begin{figure}
+\centering\includegraphics{stretch.pdf}
+\caption{Shearing the gel translates into stretching individual processions.
+From top to bottom: configuration at rest; full extension of the self avoiding walk; stretching of the electrostatic blobs.}
+\label{fig:stretch}
+\end{figure}
 
-The fully extended self-avoiding walk is a cylinder of length $R_0 = Nb/Bg_\mathrm{K}$, of diameter $D$, of volume $V = R_0 D^2$ and of area $\Sigma_0 = R_0 D = (R_0 V)^{1/2}$.
+The fully extended self-avoiding walk is a cylinder of length $R_0 = Nb/(Bg_\mathrm{K})$, of diameter $D$, of volume $V = R_0 D^2$ and of area $\Sigma_0 = R_0 D = (R_0 V)^{1/2}$.
 Extending the cylinder to $R = (\gamma+1) R_0$ while keeping the volume $V$ constant thus creates an extra area
@@
 $
-that translates into a stretching energy due to the surface tension $k_\mathrm{B}T/\xi_T^2$:
+that translates into a stretching energy due to the surface tension $k_\mathrm{B}T/\xi_T^2$, via Equations~(\ref{eq:thermal}) and (\ref{eq:geDfromB}):
 \begin{equation}
-\frac{E_\mathrm{stretch}}{k_\mathrm{B}T} = \left(\left(\gamma +1\right)^{1/2} -1\right) N \left(\frac{\tau^3}{B g_\mathrm{K}^5}\right)^{1/2}.
+\frac{E_\mathrm{stretch}}{k_\mathrm{B}T} = \left(\left(\gamma +1\right)^{1/2} -1\right) \frac{N}{g_\mathrm{K}} (1+B\tau)^{-1/2} \tau^2.
 \label{eq:stretch}
@@
 
-If we consider that the head-to-body bonds break at $\gamma_c$, we can express the reduced temperature as
+If we consider that the head-to-body bonds break at $\gamma_c$, the reduced temperature is solution of
 \begin{equation}
-\tau = B^{1/3} g_\mathrm{K}^{5/3} \left(\frac{1}{(\gamma_c +1)^{1/2} -1}\frac{E_c}{N k_\mathrm{B}T}\right)^{2/3}
+\left(\left((\gamma_c +1)^{1/2} -1\right)\frac{k_\mathrm{B}T}{E_c}\frac{N}{g_\mathrm{K}}\right)^2 \tau^4 - B\tau -1 = 0
 \label{eq:tau}
@@
 In water the bonding energy between two ions is typically \SI{5}{\kilo\joule\per\mole}\cite{Schneider1992}, thus $E_c\approx 2 k_\mathrm{B}T$.
-However the medium surrounding the ionic bond is not water.
-One has to keep in mind that the chain is in bad solvent, thus electrostatic blobs have low water content.
+However the medium surrounding the ionic bond cannot in general be described as pure water.
+As the chain is in poor solvent, electrostatic blobs have a low water content.
 Such a low polarity microenvironement is well known to enhance otherwise weak electrostatic interactions in protein folding or engineered self-assembly\cite{Rehm2010}.
 
-Using Equations~(\ref{eq:geD}) and (\ref{eq:B}) we express the size $D$ of electrostatic blobs function of $B$ and $\tau$:
-\begin{equation}
-g_e = \left(\frac{B^3}{g_\mathrm{K}\tau}\right)^{1/2},\qquad 
-D = b\left(\frac{B}{g_\mathrm{K}^3\tau}\right)^{1/2}.
-\label{eq:DB}
-\end{equation}
+From Equation~(\ref{eq:geDfromB}) we know that $D$ increases with $B$.
+We thus expect stronger head-to-body bonds when $B$ is large and the low-polarity environment is extended ($D\gg\ell_\mathrm{B}$).
+A contrario, the weakest ionic links, closest to their strength in water, should be found in \ce{PPyr+I-} where $B$ is minimum ($D\approx\ell_\mathrm{B}$).
 
-We thus expect stronger head-to-body bonds when $B$ is large and the low-polarity environment is extended ($D\gg\ell_\mathrm{B}$).
-A contrario, the weakest ionic links, closest to their strength in water, should be found in \ce{PPyr+I-} where $B$ is minimum (hopefully $D<\ell_\mathrm{B}$).
+Applying Equation~(\ref{eq:tau}) to \ce{PPyr+I-} with $E_c = 2 k_\mathrm{B}T$ yields $\tau \approx 0.40\pm 0.07$, or a $\Theta$-temperature around \SI{220}{\celsius}.
+Using this value and Equation (\ref{eq:stretch}) we deduce the bonding energy for each composition, as reported in Table~\ref{tab:results}.
+Indeed, the bonding energy increases with increasing counterion condensation to reach, in the case of \ce{PIm+Br-}, 85\% of the carbon-carbon single bond (\SI{350}{\kilo\joule\per\mole} or $140 k_\mathrm{B}T$).
+According to Equation (\ref{eq:geDfromB}), in this extreme case an electrostatic blob contains $\approx 10^3$ polymers which embed the head-to-body bonds in a low-polarity environment $\approx 24$ times larger than the Bjerrum length.
+Conversely, for \ce{PPyr+I-} and \ce{PPyr+Br-} the size of electrostatic blobs is comparable to $\ell_\mathrm{B}$, confirming an aqueous microenvironment.
+Quantitatively, the bonding energy of an ionic bond being inversely proportional to the relative dielectric constant $\epsilon_r$, we need to suppose a local $\epsilon_r$ around 2 (typical of water-insoluble polymers) instead of 80 for water to recover the large bonding energy of \ce{PIm+Br-}.
 
-Applying Equation~(\ref{eq:tau}) to \ce{PPyr+I-} with $E_c = 2 k_\mathrm{B}T$ yields $\tau \approx 0.46\pm 0.1$.
-Using this value and Equation (\ref{eq:stretch}) we deduce the bounding energy for each composition, as reported in Table~\ref{tab:results}.
-Indeed, the bonding energy increases with increasing counterion condensation to reach 60\% of the carbon-carbon single bond (\SI{350}{\kilo\joule\per\mole} or $140 k_\mathrm{B}T$) in the case of \ce{PIm+Br-}.
-According to Equation (\ref{eq:DB}), in this extreme case an electrostatic blob contains $\approx 500$ polymers which embed the head-to-body bonds in a low-polarity environment 13 times larger than the Bjerrum length.
-Conversely, for \ce{PPyr+I-} and \ce{PPyr+Br-} electrostatic blobs are smaller than $\ell_\mathrm{B}$, confirming an aqueous microenvironment.
-Quantitatively, the bonding energy of an ionic bond is inversely proportional to the relative dielectric constant $\epsilon_r$, therefore we need to suppose a local $\epsilon_r$ around 2 (typical of water-insoluble polymers) instead of 80 for water to recover the large bonding energy of \ce{PIm+Br-}.
+\subsubsection{Quantitative phase diagram.}
+
+\begin{figure}
+\includegraphics{phase_diag.pdf}
+\caption{Conformation of semi-dilute polyelectrolytes at rest at various scales as a function of counterion condensation parameter $A$.
+The phase diagram is given for a reduced temperature $\tau=0.46$ and for the monomer number density of \ce{PIm+Br-}, although $c$ is slightly different for each sample.
+Each vertical line represents a sample.}
+\label{fig:phasediag}
+\end{figure}
 
 Table~\ref{tab:results} also shows the values of $A$ deduced from Equation~(\ref{eq:A}) and the measured value of $\tau$.
-$A$ varies between $5.5$ for \ce{PPyr+I-}, corresponding to $13$ uncondensed counterions per polymer, to $460$ for \ce{PIm+Br-} indicating a single dissociated counterion every 7 polymers.
+$A$ varies between $7$ for \ce{PPyr+I-}, corresponding to $10$ uncondensed counterions per polymer, to $640$ for \ce{PIm+Br-} indicating a single dissociated counterion every $9$ polymers.
+We stress that even in this case of very strong counterion condensation an electrostatic blob still contains several ($\approx 90$) charges ensuring electrostatic repulsion and solubility,
 
-From purely rheological measurements, we have thus obtained microscopic information on the state of the gel: amount of charge condensation, head-to-body bounding energy and size of a procession.
+Figure~\ref{fig:phasediag} summarizes the different conformations of a procession as a function of scale and counterion condensation for our experimental $c$ and $\tau$.
+The most striking feature of this phase diagram is the narrowing of the regime between $D$ and $r_\mathrm{scr}$ (white region in Figure~\ref{fig:phasediag}) as counterion condensation increases.
+As the size of the electrostatic blob crosses the screening length, electrostatic repulsion is no longer able to sustain solubility.
+This abrupt transition between an uniform gel and a precipitate is what separates our five soluble samples from insoluble \ce{PPyr+F-}, \ce{PIm+Cl-} and \ce{PIm+F-}.
+Our rheological measurements probe the structure of the procession from the Kuhn length to the distance between crosslinks that is approximately $r_\mathrm{scr}$ (upper limit of the white region in Figure~\ref{fig:phasediag}).
+\ce{PPyr+Cl-} and \ce{PIm+I-} display the typical structure of weakly charged polyelectrolytes in poor solvent.
+\ce{PPyr+I-} and \ce{PPyr+Br-} lie right in the crossover to strongly charged regime.
+Their correct description is possible only through Equation~(\ref{eq:geDfromB}).
+Finally \ce{PIm+Br-} is extremely close to the transition to insolubility (within experimental errors) which explains its extreme softness and wide plastic regime.
+
+From purely rheological measurements, we have thus obtained microscopic information on the state of the gel: amount of charge condensation, head-to-body bonding energy and size of a procession.
 
@@
 To conclude, our procedure yields robust, highly tunable hydrogels from short, linear polymer chains and in the absence of any additive, which could find interesting applications especially in the context of biomaterials.
-Most importantly, systematic comparisons between the different poly(electrolytes) investigated in this study were used to establish a model linking the microsctructure of the gel to ion pair dissociation efficiency on the individual polymer chains.
+Most importantly, systematic comparisons between the different poly(electrolytes) investigated in this study were used to establish a model linking the microstructure of the gel to ion pair dissociation efficiency on the individual polymer chains.
 We hope that this model will join the procession toward future works in supramolecular assemblies.
 
-\begin{acknowledgement}
-The authors are grateful for the financial support of Université de Lyon through the program ``Investissements d'Avenir'' (ANR-1 1-IDEX-0007).
+\paragraph*{Acknowledgements.}
+The authors are grateful for the financial support of Universit\'{e} de Lyon through the program ``Investissements d'Avenir'' (ANR-1 1-IDEX-0007).
 S.M.
 and N.T.
-acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007–2013)/ERC grant agreement No.258803.
+acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013)/ERC grant agreement No.258803.
 The Graphical TOC entry is derived from a picture by Arturo Reina, CC BY-SA 3.0, \url{https://commons.wikimedia.org/w/index.php?curid=282898}.
-\end{acknowledgement}
 
-\begin{suppinfo}
+% \subsection{This is the subsection heading style}
+% Section headings can be typeset with and without numbers.\cite{Abernethy2003}
 
-Materials and methods, detailed synthesis and characterizations of gels, GPC, HRMS and DSC data, calculated cohesion energy, and frequency sweep curves.
+% \subsubsection{This is the subsubsection style.~~} These headings should end in a full point.
 
-\end{suppinfo}
+% \paragraph{This is the next level heading.~~} For this level please use \texttt{\textbackslash paragraph}. These headings should also end in a full point.
 
-\bibliography{PALSE.bib}
+% \section{Graphics and tables}
+% \subsection{Graphics}
+% Graphics should be inserted on the page where they are first mentioned (unless they are equations, which appear in the flow of the text).\cite{Cotton1999}
+
+% \begin{figure}[h]
+% \centering
+%   \includegraphics[height=3cm]{example}
+%   \caption{An example figure caption}
+%   \label{fgr:example}
+% \end{figure}
+
+% \begin{figure*}
+%  \centering
+%  \includegraphics[height=3cm]{example2}
+%  \caption{An image from the \textit{Physical Chemistry Chemical Physics} cover gallery, set as a two-column figure.}
+%  \label{fgr:example2col}
+% \end{figure*}
+
+% \subsection{Tables}
+% Tables typeset in RSC house style do not include vertical lines.
+Table footnote symbols are lower-case italic letters and are typeset at the bottom of the table.
+Table captions do not end in a full point.\cite{Arduengo1992,Eisenstein2005}
+
+% \begin{table}[h]
+% \small
+%   \caption{\ An example of a caption to accompany a table}
+%   \label{tbl:example}
+%   \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}}lll}
+%     \hline
+%     Header one/units & Header two & Header three \\
+%     \hline
+%     1 & 2 & 3 \\
+%     4 & 5 & 6 \\
+%     7 & 8 & 9 \\
+%     10 & 11 & 12 \\
+%     \hline
+%   \end{tabular*}
+% \end{table}
+
+% Adding notes to tables can be complicated.
+Perhaps the easiest method is to generate these manually.\footnote[4]{Footnotes should appear here.
+These might include comments relevant to but not central to the matter under discussion, limited experimental and spectral data, and crystallographic data.}
+
+% an example of a two-column table
+%\begin{table*}
+%\small
+  %\caption{\ An example of a caption to accompany a table, table captions do not end in a full point}
+  %\label{tbl:example}
+  %\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lllllll}
+    %\hline
+    %Header one & Header two & Header three & Header four & Header five & Header six  & Header seven\\
+    %\hline
+    %1 & 2 & 3 & 4 & 5 & 6  & 7\\
+    %8 & 9 & 10 & 11 & 12 & 13 & 14 \\
+    %15 & 16 & 17 & 18 & 19 & 20 & 21\\
+    %\hline
+  %\end{tabular*}
+%\end{table*}
+
+% \section{Equations}
+
+% Equations can be typeset inline \textit{e.g.} $ y = mx + c$ or displayed with and without numbers:
+
+%  \[ A = \pi r^2 \]
+
+% \begin{equation}
+%   \frac{\mathrm{\gamma}}{\mathrm{\epsilon}x} r^2 = 2r
+% \end{equation}
+% \pagebreak
+% \begin{widetext}
+% \begin{eqnarray} \label{e27}
+% T_{nn''}^{mm''}=
+% \begin{cases}
+% & \dfrac{(q{+}p'')!\,(p{-}p'')!}{(q{-}q'')!\,(p{+}q'')!}\,a^{m-m''}b^{n''}\,P_{p-p''}^{(m-m'',n'')}(1{-}2A^2)P_{p-p''}^{(m-m'',n'')}(2B^2{-}1)~~~ \\ & \text{when}~~~n-n''\geq m-m''\geq0~,\\
+% \\
+% & \dfrac{(p{+}q'')!\,(q{-}q'')!}{(p{-}p'')!\,(q{+}p'')!}\,a^{m''-m}b^{n''}\,P_{q-q''}^{(m''-m,n'')}(1{-}2A^2)\,P_{q-q''}^{(m''-m,n'')}(2B^2{-}1)~~~ \\ & \text{when}~~~ n-n''\geq m''-m\geq0~, \\
+% \\
+% &0 ~~~ \text{otherwise}.
+% \end{cases}
+% \end{eqnarray}
+% \end{widetext}
+
+% You can also put lists into the text. You can have bulleted or numbered lists of almost any kind. 
+% The \texttt{mhchem} package can also be used so that formulae are easy to input: \texttt{\textbackslash ce\{H2SO4\}} gives \ce{H2SO4}.
+
+% For footnotes in the main text of the article please number the footnotes to avoid duplicate symbols.
+e.g.
+\texttt{\textbackslash footnote[num]\{your text\}} the corresponding author $\ast$ counts as footnote 1, ESI as footnote 2, e.g.
+if there is no ESI, please start at [num]=[2], if ESI is cited in the title please start at [num]=[3] etc.
+Please also cite the ESI within the main body of the text using \dag.
+
+% \section{Conclusions}
+% The conclusions section should come at the end of article. For the reference section, the style file rsc.bst can be used to generate the correct reference style.
+%%%END OF MAIN TEXT%%%
+
+%The \balance command can be used to balance the columns on the final page if desired. It should be placed anywhere within the first column of the last page.
+
+%\balance
+
+%If notes are included in your references you can change the title from 'References' to 'Notes and references' using the following command:
+%\renewcommand\refname{Notes and references}
+
+%%%REFERENCES%%%
+\bibliography{PALSE} %You need to replace "rsc" on this line with the name of your .bib file
+\bibliographystyle{rsc} %the RSC's .bst file
 

### 2016-07-07 PALSE/talk_Kyoto.tex
Polymerisation and post-functionalisation, as caterpillar and as chemistry

\documentclass[xcolor=table]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage[super]{nth}
%\usetheme{Boadilla}
%\usecolortheme{rose}
%\usecolortheme{crane}
%\usefonttheme{structuresmallcapsserif}
%\setbeamertemplate{navigation symbols}{}

\definecolor{Main}{rgb}{0.74, 0.13, 0.19}
\definecolor{Accent1}{rgb}{0.76,0.36,0.13}
\definecolor{Accent2}{rgb}{0.54,0.1,0.4}

\mode<presentation>{\usetheme{ilm}}
%\usecolortheme{rose}
%\useinnertheme[shadow]{circles}
%\usecolortheme{whale}
%\useoutertheme{infolines}

%\usecolortheme[named=Accent1]{structure}

%\setbeamerfont{page number in head/foot}{size=\large}
%\setbeamercolor{page number in head/foot}{fg=Main}
%% page/total
%%\setbeamertemplate{footline}[frame number]
%% pas de total
%\setbeamertemplate{footline}{%
%    	\hfill%
%	\usebeamercolor[fg]{page number in head/foot}%
%	\usebeamerfont{page number in head/foot}%
%	\insertframenumber\kern1em\vskip2pt%
%}
%
%\setbeamersize{text margin left=1em}
%\setbeamersize{text margin right=1em}

\usepackage[overlay,absolute]{textpos}
\setlength{\TPHorizModule}{10mm}
\setlength{\TPVertModule}{\TPHorizModule}
\textblockorigin{10mm}{10mm} % start everything near the top-left corner
\setlength{\parindent}{0pt}

%font
\usepackage[T1]{fontenc}
\usepackage{times}
%\usepackage[oldstylenums]{kpfonts}

%\include{macros}
\usepackage{ifthen}

%proper math and math symbols
\usepackage{amsmath,mathtools}
\usepackage{amssymb}

\usepackage{datenumber,fp}

\usepackage[separate-uncertainty = true]{siunitx}

\usepackage{tabu}
\usepackage{multirow}
\usepackage{booktabs}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, decorations.text, shapes.geometric, matrix,patterns,mindmap}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.9}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{external}
\makeatletter
\newcommand*{\overlaynumber}{\number\beamer@slideinframe}
\tikzset{
  beamer externalizing/.style={%
    execute at end picture={%
      \tikzifexternalizing{%
        \ifbeamer@anotherslide
        \pgfexternalstorecommand{\string\global\string\beamer@anotherslidetrue}%
        \fi
      }{}%
    }%
  },
  external/optimize=false
}
\let\orig@tikzsetnextfilename=\tikzsetnextfilename
\renewcommand\tikzsetnextfilename[1]{\orig@tikzsetnextfilename{#1-\overlaynumber}}
\makeatother

\tikzset{every picture/.style={beamer externalizing}}
\tikzexternalize
\tikzsetexternalprefix{fig_presentation/}
%\tikzset{external/optimize=false}
%\tikzset{external/force remake}

%link or play movies
\usepackage{multimedia}

%chemistry
\usepackage[version=3]{mhchem}
\usepackage{chemfig}
\usepackage{setspace}

%beamer related package

\usepackage{todonotes}
\presetkeys{todonotes}{inline}{}

\tikzset{onslide/.code args={<#1>#2}{%
  \only<#1>{\pgfkeysalso{#2}} %
}}%

%bibliography
\usepackage[style=authoryear-comp, language=british,eprint=false, url=false, doi=false, sortcites=true, sorting=none, isbn=false, firstinits=true,maxcitenames=6]{biblatex}
%minimal citations
\AtEveryCitekey{%
	\clearfield{title}
	\clearfield{pages}
	\clearfield{volume}
	\clearfield{number}
	\clearfield{month}}
\newcommand{\myfullcite}[1]{{\scriptsize\fullcite{#1}}}
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}
%\bibliography{biblio}

\newcolumntype{P}[1]{>{\raggedright}p{#1}}

\institute[iLM]{Univ Lyon, Université Claude Bernard Lyon 1, CNRS, Institut Lumière Matière}
\title[processionary gel]{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}
\author[M. Leocmach]{Mathieu Leocmach {\usebeamerfont{normal text}\texttt{\usebeamercolor[fg]{normal text}\footnotesize @LamSonLeoc}}\vspace{-\baselineskip}}
\date{%10 August 2016
}
\titlegraphic{\vspace{-\baselineskip}
	\begin{tabu}{X[c]X[c]X[c]X[c]}
		\multicolumn{2}{c}{\usebeamerfont{institute}\usebeamercolor[fg]{institute}ENS de Lyon, Laboratoire de Chimie}&
		\multicolumn{2}{c}{\usebeamerfont{institute}\usebeamercolor[fg]{institute}ENS de Lyon, Laboratoire de Physique}\\\cmidrule(r){1-2}\cmidrule(l){3-4}
		\includegraphics[height=3\baselineskip]{presentation/Hassan}&
		\includegraphics[height=3\baselineskip]{presentation/Cyrille}&
		\includegraphics[height=3\baselineskip]{presentation/Nicolas_Taberlet}&
		\includegraphics[height=3\baselineskip]{../Yaourt/Seb}\\
		Hassan & Cyrille & Nicolas & S\'{e}bastien\\
		Srour & Monnereau & Taberlet & Manneville\\\cmidrule(r){1-2}
		\multicolumn{2}{c}{Martien Duvall Deffo Ayagou}&\\
		\multicolumn{2}{c}{Thi Thanh-Tam Nguyen}&\raisebox{-0.1em}[0pt][0pt]{\includegraphics[height=2\baselineskip]{presentation/invest-avenir.jpg}}&\raisebox{-0.1em}[0pt][0pt]{\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{../Yaourt/NEW-Logo-ERC-OUTLINE}}\\
	\end{tabu}

\bigskip
Martien Duvall Deffo Ayagou, Thi Thanh-Tam Nguyen%, Chantal Andraud

	%\vfill
	%\includegraphics[height=2\baselineskip]{logo_ens-lyon}\quad
	%\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
	%\includegraphics[height=2\baselineskip]{../../Yaourt/CNRSfilaire-Q}\quad
	%\includegraphics[height=2\baselineskip]{CRPP}\quad
	%\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}
	}

\newlength{\mylength}

%\includeonly{creep_beamer}

\begin{document}
\tikzset{every mark/.append style={scale=0.8}}
\pgfplotsset{every axis/.append style={footnotesize}}

\pgfplotscreateplotcyclelist{earthy}{%
{red!40!black,mark=o},
{red!60!black,mark=triangle, every mark/.append style={rotate=180}},
{red!80!black,mark=square},
{red,mark=triangle},
{red!80!yellow, mark=diamond},
red!60!yellow,
red!40!yellow,
}

\AtBeginSection[]{
	\addtocounter{framenumber}{-1}
	\begin{frame}[plain]
		\tableofcontents[currentsection, hideothersubsections]
	\end{frame}
}

\begin{frame}{\pgfuseimage{cnrs-logo}\hspace*{0.3cm}\pgfuseimage{ucbl-logo}\pgfuseimage{univlyon-logo}}%[plain]
	\titlepage
\end{frame}

\setcounter{framenumber}{0}

\begin{frame}{Fine chemistry}
	\vspace{\baselineskip}
	\tikzsetnextfilename{polymerisation}%
	\let\mrad\relax%
	\newlength\mrad%
	\setlength{\mrad}{1ex}%
	% The face style, can be changed
	\tikzset{face/.style={
		shape=circle,minimum size=2\mrad,shading=radial,outer sep=0pt,
	    inner color=white!50!yellow,outer color= yellow!70!orange
		}}%
	\begin{tikzpicture}
		\foreach \i [evaluate=\i as \dist using \i*0.45] in {0,1,2}{
			\begin{scope}[xshift=\dist\textwidth]
			%face
			\node[face,inner color=white!50!ilmcolor,outer color= ilmcolor!90!black] (emoticon) {};
			%% The eyes are fixed.
			\draw[fill=white] 
				(-0.5\mrad,0ex) ..controls (-0.25\mrad,0.1\mrad)and(0.25\mrad,0.1\mrad)..
		        (0.5\mrad,0.0pt) ..controls (0.75\mrad,0.75\mrad)and(0.1\mrad,0.85\mrad)..
		        (0pt,0.2\mrad) ..controls (-0.1\mrad,0.85\mrad)and(-0.75\mrad,0.75\mrad)..
		        (-0.5\mrad,0pt)--cycle;
			%% standard pupils
			\node[fill, ellipse,inner xsep=0.075\mrad, inner ysep=0.0375\mrad, rotate=80] at (0.25\mrad,0.25\mrad) {};
			\node[fill, ellipse,inner xsep=0.075\mrad, inner ysep=0.0375\mrad, rotate=100] at (-0.25\mrad,0.25\mrad) {};
			%% mouth
			\draw[thick,line cap=round] (-0.5\mrad,-0.5\mrad)
			     ..controls (-0.25\mrad,-0.75\mrad)and(0.25\mrad,-0.75\mrad)..(0.5\mrad,-0.5\mrad);
			%% horns
			\draw ( emoticon.80)..controls ( 0.3\mrad,1.2\mrad)..(0.5\mrad,1.25\mrad)
			      ..controls ( 0.4\mrad,1.15\mrad)..(emoticon.70);
			\draw (emoticon.100)..controls (-0.3\mrad,1.2\mrad)..(-0.5\mrad,1.25\mrad)
			      ..controls (-0.4\mrad,1.15\mrad)..(emoticon.110);
			%body segments
			\ifthenelse{\i=0}{}{
			\foreach \x in {0,2,...,10}{
				\node[face,inner color=white!50!ilmorange,outer color= ilmorange!90!black, left=\x\mrad of emoticon] (monomer){};
				%legs
				\ifthenelse{\i=1}{}{
					\draw[line width=0.25\mrad, gray] (monomer.south) ++(0,0.5\mrad) to[bend left] +(0,-1\mrad);}
			};}
			\ifthenelse{\i=1}{
				\draw[|-|] ($(monomer.west)+(0,-1.5\mrad)$) -- ($(emoticon.west)+(0,-1.5\mrad)$) node[midway,below, font=\scriptsize] {$N_0$};
			}{}
			\end{scope}
			};
			%arrows
			\draw[-stealth, thick] ++(3\mrad,0) -- +(6\mrad,0) node[midway, above=0.2\mrad, face,inner color=white!50!ilmorange,outer color= ilmorange!90!black]{} node[midway, below, font=\scriptsize]{ATRP};
			\draw[-stealth, thick] ++(0.45\textwidth,0) ++(3\mrad,0) -- +(6\mrad,0) coordinate[midway, above=0.2\mrad] (leg) node[midway, below, font=\scriptsize,text width=10\mrad,align=center]{post-functionalization};
			\draw[line width=0.25\mrad, gray] (leg) to[bend right] +(0,1\mrad);
		\end{tikzpicture}

\bigskip
\definesubmol{head}{(-[::-105,,,,ilmcolor])(-[::-15,,,,ilmcolor])-[::60,,,,ilmcolor](=[::60,,,,ilmcolor]\textcolor{ilmcolor}{O})-[::-60,,,,ilmcolor]\textcolor{ilmcolor}{O}-[::-60,,,,ilmcolor]-[::60,,,,ilmcolor]-[::60,,,,ilmcolor]}
\definesubmol{mono}{-[::-60,,,,ilmorange](=[,,,,ilmorange]\textcolor{ilmorange}{O})-[::-60,,,,ilmorange]\textcolor{ilmorange}{O}-[::60,,,,ilmorange]-[,,,,ilmorange]-[::-60,,,,ilmorange]}
\tikzsetnextfilename{polymerisation_PImBr}
\begin{tikzpicture}[every node/.style={font=\tiny\setatomsep{1.5em}}]
	\node (init) {\chemfig{Br-[:-30]!{head}\textcolor{ilmcolor}{P}|\textcolor{ilmcolor}{O_3}Et_2}};

	\node[below left=5\baselineskip and 0 of init.east] (POH) {%
	\chemfig{[:-30]-[@{left,0.25}](%
		!{mono}OH%
		)-[::60,,,,ilmorange]-[@{right,0.25}]!{head}\textcolor{ilmcolor}{P}|\textcolor{ilmcolor}{O_3}Et_2}%
	};
	\node[below right=-0.6em and -0.25em of POH.base west] {$\prescript{}{N_0}{\left[\vrule height 1em depth 0.25em width 0pt\hspace{1.75em}\right]}$};
	\node[below right=-0.6em and -0.25em of POH.base west] {};

	\node[right=0.12\textwidth of init.base east, anchor=base west] (PBr) {%
	\chemfig{[:-30]-[@{left,0.25}](%
		!{mono}Br%
		)-[::60,,,,ilmorange]-[@{right,0.25}]!{head}\textcolor{ilmcolor}{P}|\textcolor{ilmcolor}{O_3^{2-}}}%
	};
	\node[below right=-0.6em and -0.25em of PBr.base west] {$\prescript{}{N_0}{\left[\vrule height 1em depth 0.25em width 0pt\hspace{1.75em}\right]}$};

	\node[right=\textwidth of init.base west, anchor=base east] (PNuBr) {%
	\chemfig{[:-30]-[@{left,0.25}](%
		!{mono}\textcolor{gray}{N}|^{+}*5(=[,,,,thick, gray]-[,,,,thick, gray]\textcolor{gray}{N}(-[,,,,thick, gray])-[,,,,thick, gray]=[,,,,thick, gray]-[,,,,thick, gray])(-[0,2,,,draw=none]Br^{-})
		)-[::60,,,,ilmorange]-[@{right,0.25}]!{head}\textcolor{ilmcolor}{P}|\textcolor{ilmcolor}{O_3^{2-}}}%
	};
	\node[below right=-0.6em and -0.25em of PNuBr.base west] {$\prescript{}{N_0}{\left[\vrule height 1em depth 0.25em width 0pt\hspace{1.75em}\right]}$};

	\draw[-stealth, thick] (init.south-|POH.north) -- (POH) node[midway, left=0, -]{\chemfig{[:-30]=[,,,,ilmorange]!{mono}OH}} node[midway, right=0, text width=0.13\textwidth, align=center]{\ce{CuBr}\linebreak 4-4'-bipyridine\linebreak \SI{85}{\celsius}};
	\draw[-stealth, thick] (POH) -| (PBr) node[pos=0.25, above]{\ce{TMS-Br}} node[pos=0.25, below, sloped, text width=0.2\textwidth, align=center]{\ce{CH2Cl2}\linebreak \SI{85}{\celsius}};
	\draw[-stealth, thick] (PBr) -- (PBr-|PNuBr.west) node[midway, above,align=center,-]{\chemfig[thick, gray]{[2]*5(-N(-)-=N-=)}} node[midway, below, text width=0.2\textwidth, align=center]{THF\linebreak \SI{85}{\celsius}};;
	\node[anchor=south] at (POH.south) {\ce{POH}};
	\node[anchor=south] at (PBr.south) {\ce{PBr}};
\end{tikzpicture}
\end{frame}

\end{document}

### 2016-07-08 PALSE/talk_Kyoto.tex
switch color scheme: body segment in gray and leg (Nucleophile) in orange

@@
 	\tikzsetnextfilename{polymerisation}%
-	\let\mrad\relax%
-	\newlength\mrad%
-	\setlength{\mrad}{1ex}%
-	% The face style, can be changed
-	\tikzset{face/.style={
-		shape=circle,minimum size=2\mrad,shading=radial,outer sep=0pt,
-	    inner color=white!50!yellow,outer color= yellow!70!orange
-		}}%
+	\input{fig_presentation/polymerisation.pgf}
+
+	\bigskip
+	\tikzsetnextfilename{polymerisation_PImBr}%
+	\input{fig_presentation/polymerisation_PImBr.pgf}
+\end{frame}
+
+\begin{frame}{Physical polyelectrolytes hydrogel}
+\let\mrad\relax
+\newlength\mrad
+\setlength{\mrad}{0.45em}
+\tikzsetnextfilename{polyelectrolytes}%
 \begin{tikzpicture}
-		\foreach \i [evaluate=\i as \dist using \i*0.45] in {0,1,2}{
-			\begin{scope}[xshift=\dist\textwidth]
-			%face
-			\node[face,inner color=white!50!ilmcolor,outer color= ilmcolor!90!black] (emoticon) {};
-			%% The eyes are fixed.
-			\draw[fill=white] 
-				(-0.5\mrad,0ex) ..controls (-0.25\mrad,0.1\mrad)and(0.25\mrad,0.1\mrad)..
-		        (0.5\mrad,0.0pt) ..controls (0.75\mrad,0.75\mrad)and(0.1\mrad,0.85\mrad)..
-		        (0pt,0.2\mrad) ..controls (-0.1\mrad,0.85\mrad)and(-0.75\mrad,0.75\mrad)..
-		        (-0.5\mrad,0pt)--cycle;
-			%% standard pupils
-			\node[fill, ellipse,inner xsep=0.075\mrad, inner ysep=0.0375\mrad, rotate=80] at (0.25\mrad,0.25\mrad) {};
-			\node[fill, ellipse,inner xsep=0.075\mrad, inner ysep=0.0375\mrad, rotate=100] at (-0.25\mrad,0.25\mrad) {};
-			%% mouth
-			\draw[thick,line cap=round] (-0.5\mrad,-0.5\mrad)
-			     ..controls (-0.25\mrad,-0.75\mrad)and(0.25\mrad,-0.75\mrad)..(0.5\mrad,-0.5\mrad);
-			%% horns
-			\draw ( emoticon.80)..controls ( 0.3\mrad,1.2\mrad)..(0.5\mrad,1.25\mrad)
-			      ..controls ( 0.4\mrad,1.15\mrad)..(emoticon.70);
-			\draw (emoticon.100)..controls (-0.3\mrad,1.2\mrad)..(-0.5\mrad,1.25\mrad)
-			      ..controls (-0.4\mrad,1.15\mrad)..(emoticon.110);
-			%body segments
-			\ifthenelse{\i=0}{}{
-			\foreach \x in {0,2,...,10}{
-				\node[face,inner color=white!50!ilmorange,outer color= ilmorange!90!black, left=\x\mrad of emoticon] (monomer){};
-				%legs
-				\ifthenelse{\i=1}{}{
-					\draw[line width=0.25\mrad, gray] (monomer.south) ++(0,0.5\mrad) to[bend left] +(0,-1\mrad);}
-			};}
-			\ifthenelse{\i=1}{
-				\draw[|-|] ($(monomer.west)+(0,-1.5\mrad)$) -- ($(emoticon.west)+(0,-1.5\mrad)$) node[midway,below, font=\scriptsize] {$N_0$};
-			}{}
-			\end{scope}
-			};
-			%arrows
-			\draw[-stealth, thick] ++(3\mrad,0) -- +(6\mrad,0) node[midway, above=0.2\mrad, face,inner color=white!50!ilmorange,outer color= ilmorange!90!black]{} node[midway, below, font=\scriptsize]{ATRP};
-			\draw[-stealth, thick] ++(0.45\textwidth,0) ++(3\mrad,0) -- +(6\mrad,0) coordinate[midway, above=0.2\mrad] (leg) node[midway, below, font=\scriptsize,text width=10\mrad,align=center]{post-functionalization};
-			\draw[line width=0.25\mrad, gray] (leg) to[bend right] +(0,1\mrad);
+\draw[every node/.style={circle,draw,font=\tiny, inner sep=0,minimum height=0.5em}] (0,0) --(30:\mrad) node[above right,ilmcolor]{-} (0,0)--(-30:\mrad) node[below right,ilmcolor]{-} (0,0)--++(-180:\mrad)\foreach \x [evaluate=\x as \r using 1/(rnd+0.1), evaluate=\x as \rr using 1/(rnd+0.1)] in {1,2,...,35}{--++(150:\mrad)node[above,ilmorange]{+} node[above=\r\mrad]{-}--++(-150:\mrad)node[below=0,ilmorange]{+} node[below=\rr\mrad]{-}};
 \end{tikzpicture}
 
-\bigskip
-\definesubmol{head}{(-[::-105,,,,ilmcolor])(-[::-15,,,,ilmcolor])-[::60,,,,ilmcolor](=[::60,,,,ilmcolor]\textcolor{ilmcolor}{O})-[::-60,,,,ilmcolor]\textcolor{ilmcolor}{O}-[::-60,,,,ilmcolor]-[::60,,,,ilmcolor]-[::60,,,,ilmcolor]}
-\definesubmol{mono}{-[::-60,,,,ilmorange](=[,,,,ilmorange]\textcolor{ilmorange}{O})-[::-60,,,,ilmorange]\textcolor{ilmorange}{O}-[::60,,,,ilmorange]-[,,,,ilmorange]-[::-60,,,,ilmorange]}
-\tikzsetnextfilename{polymerisation_PImBr}
-\begin{tikzpicture}[every node/.style={font=\tiny\setatomsep{1.5em}}]
-	\node (init) {\chemfig{Br-[:-30]!{head}\textcolor{ilmcolor}{P}|\textcolor{ilmcolor}{O_3}Et_2}};
-
-	\node[below left=5\baselineskip and 0 of init.east] (POH) {%
-	\chemfig{[:-30]-[@{left,0.25}](%
-		!{mono}OH%
-		)-[::60,,,,ilmorange]-[@{right,0.25}]!{head}\textcolor{ilmcolor}{P}|\textcolor{ilmcolor}{O_3}Et_2}%
-	};
-	\node[below right=-0.6em and -0.25em of POH.base west] {$\prescript{}{N_0}{\left[\vrule height 1em depth 0.25em width 0pt\hspace{1.75em}\right]}$};
-	\node[below right=-0.6em and -0.25em of POH.base west] {};
-
-	\node[right=0.12\textwidth of init.base east, anchor=base west] (PBr) {%
-	\chemfig{[:-30]-[@{left,0.25}](%
-		!{mono}Br%
-		)-[::60,,,,ilmorange]-[@{right,0.25}]!{head}\textcolor{ilmcolor}{P}|\textcolor{ilmcolor}{O_3^{2-}}}%
-	};
-	\node[below right=-0.6em and -0.25em of PBr.base west] {$\prescript{}{N_0}{\left[\vrule height 1em depth 0.25em width 0pt\hspace{1.75em}\right]}$};
-
-	\node[right=\textwidth of init.base west, anchor=base east] (PNuBr) {%
-	\chemfig{[:-30]-[@{left,0.25}](%
-		!{mono}\textcolor{gray}{N}|^{+}*5(=[,,,,thick, gray]-[,,,,thick, gray]\textcolor{gray}{N}(-[,,,,thick, gray])-[,,,,thick, gray]=[,,,,thick, gray]-[,,,,thick, gray])(-[0,2,,,draw=none]Br^{-})
-		)-[::60,,,,ilmorange]-[@{right,0.25}]!{head}\textcolor{ilmcolor}{P}|\textcolor{ilmcolor}{O_3^{2-}}}%
-	};
-	\node[below right=-0.6em and -0.25em of PNuBr.base west] {$\prescript{}{N_0}{\left[\vrule height 1em depth 0.25em width 0pt\hspace{1.75em}\right]}$};
-
-	\draw[-stealth, thick] (init.south-|POH.north) -- (POH) node[midway, left=0, -]{\chemfig{[:-30]=[,,,,ilmorange]!{mono}OH}} node[midway, right=0, text width=0.13\textwidth, align=center]{\ce{CuBr}\linebreak 4-4'-bipyridine\linebreak \SI{85}{\celsius}};
-	\draw[-stealth, thick] (POH) -| (PBr) node[pos=0.25, above]{\ce{TMS-Br}} node[pos=0.25, below, sloped, text width=0.2\textwidth, align=center]{\ce{CH2Cl2}\linebreak \SI{85}{\celsius}};
-	\draw[-stealth, thick] (PBr) -- (PBr-|PNuBr.west) node[midway, above,align=center,-]{\chemfig[thick, gray]{[2]*5(-N(-)-=N-=)}} node[midway, below, text width=0.2\textwidth, align=center]{THF\linebreak \SI{85}{\celsius}};;
-	\node[anchor=south] at (POH.south) {\ce{POH}};
-	\node[anchor=south] at (PBr.south) {\ce{PBr}};
-\end{tikzpicture}
+\begin{columns}
+\column{0.5\textwidth}
+\structure{Reversible gel}
+\begin{itemize}
+	\item in deionised, neutral water
+	\item yield stress fuild
+\end{itemize}
+\structure{Solution} if
+\begin{itemize}
+	\item ionic strength, pH
+	\item non ionic head
+\end{itemize}
+$\Rightarrow$ Head-to-body ionic bonds
+\column{0.5\textwidth}
+\end{columns}
 \end{frame}

### 2016-07-08 PALSE/talk_Kyoto.tex
Flow curves + caterpillar, cite Srour MRC 2015

@@
 \begin{frame}{Physical polyelectrolytes hydrogel}
-\let\mrad\relax
-\newlength\mrad
-\setlength{\mrad}{0.45em}
 	\tikzsetnextfilename{polyelectrolytes}%
-\begin{tikzpicture}
-\draw[every node/.style={circle,draw,font=\tiny, inner sep=0,minimum height=0.5em}] (0,0) --(30:\mrad) node[above right,ilmcolor]{-} (0,0)--(-30:\mrad) node[below right,ilmcolor]{-} (0,0)--++(-180:\mrad)\foreach \x [evaluate=\x as \r using 1/(rnd+0.1), evaluate=\x as \rr using 1/(rnd+0.1)] in {1,2,...,35}{--++(150:\mrad)node[above,ilmorange]{+} node[above=\r\mrad]{-}--++(-150:\mrad)node[below=0,ilmorange]{+} node[below=\rr\mrad]{-}};
-\end{tikzpicture}
-
+\input{fig_presentation/polyelectrolytes.pgf}
 \begin{columns}
@@
 $\Rightarrow$ Head-to-body ionic bonds
+
 \column{0.5\textwidth}
+\hfill Herschel-Buckley law\\
+\hfill $\sigma =\sigma_c + A \dot{\gamma}^n$
+\tikzsetnextfilename{flowcurves}%
+\input{fig_presentation/flowcurves.pgf}
 \end{columns}
+\vspace{-\baselineskip}\textit{\scriptsize Srour et al. Macromol. Rapid Comm., 36, 55 (2015)}
 \end{frame}

### 2016-07-11 PALSE/talk_Kyoto.tex
start Counterion condensation

@@
 
+\begin{frame}{Linear rheology}
+	\begin{columns}
+	\column{0.5\textwidth}
+	\tikzsetnextfilename{linear_rheology_PImBr}%
+	\input{fig_presentation/linear_rheology_PImBr.pgf}
+	\[ G = \frac{c}{N}k_\mathrm{B}T \]
+	\begin{description}[$N$]
+	\item[$c$] monomer concentration
+	\item[$N$] \# monomers between CL
+	\end{description}
+	\column{0.45\textwidth}
+	\pause
+	\begin{itemize}
+	\item $N \approx 800 N_0$ !
+	\item 800 chains between CL
+	\end{itemize}
+	\begin{flushright}\includegraphics[height=14\baselineskip]{presentation/Thaumetopea_pityocampa_01.jpg}\\
+	\vspace{-1.5em}\colorbox{lightgray}{\scriptsize Wikimedia commons}
+	\end{flushright}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Tweak repeated ions + counterions}
+	\vspace{1.5\baselineskip}
+	\tikzsetnextfilename{postfunc}%
+	\input{fig_presentation/postfunc.pgf}
+
+	\bigskip
+	\tikzsetnextfilename{postfunc_chem}%
+	\input{fig_presentation/postfunc_chem.pgf}
+\end{frame}
+
+\begin{frame}\frametitle{Counterion condensation}
+\begin{itemize}
+\item Direct Hoffmeister series $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$
+\item \ce{\pi+-\pi+} interactions {\tiny\setatomsep{1.5em}$\raisebox{-0.5em}{\chemfig{[2]*5(--N(-)---)}}>\raisebox{-0.5em}{\chemfig{[2]*5(-N(-)-=N-=)}}$}
+\end{itemize}
+\begin{tikzpicture}
+\pgfdeclarelayer{background}
+\pgfdeclarelayer{foreground}
+\pgfsetlayers{background,main,foreground}
+\node at (0.33\columnwidth, -8\baselineskip) {PBr} [->, level distance=5em] 
+	child [grow=east] {node(PImBr) {\ce{PIm+Br-}} [level distance=5em]
+		child[grow=-45] {node(PImF) {\ce{PIm+F-}}} 
+		child {node(PImCl) {\ce{PIm+Cl-}}} 
+		child[grow=45] {node(PImI) {\ce{PIm+I-}}}} 
+	child [grow=west] {node(PPyrBr) {\ce{PPyr+Br-}} [level distance=5em]
+		child[grow=-135] {node(PPyrF) {\ce{PPyr+F-}}} 
+		child {node(PPyrCl) {\ce{PPyr+Cl-}}} 
+		child[grow=135] {node(PPyrI) {\ce{PPyr+I-}}}};
+\begin{pgfonlayer}{background}
+	\draw[ilmorange, line width=0.3em, ->] (PImI) to[bend right] (PImBr.north) ..controls (PImBr.center) and (PImCl.center) .. (PImCl.base) to[bend left] (PImF);
+	\draw[ilmorange, line width=0.3em, ->] (PPyrI) to[bend left] (PPyrBr.north) ..controls (PPyrBr.center) and (PPyrCl.center) .. (PPyrCl.base) to[bend right] (PPyrF);
+\end{pgfonlayer}
+	\foreach \x in {PImF,PImCl,PPyrF}{
+	\draw[ilmcolor] (\x.north west) -- (\x.base east) (\x.north east) -- (\x.base west) ;
+	}
+\end{tikzpicture}
+
 \end{document}

### 2016-07-11 PALSE/talk_Kyoto.tex
solubility

--- 2016-07-08 PALSE/talk_Kyoto.tex
@@
 \title[processionary gel]{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}
-\author[M. Leocmach]{Mathieu Leocmach {\usebeamerfont{normal text}\texttt{\usebeamercolor[fg]{normal text}\footnotesize @LamSonLeoc}}\vspace{-\baselineskip}}
+\author[M.
+Leocmach]{Mathieu Leocmach \hfill{\usebeamerfont{normal text}\texttt{\usebeamercolor[fg]{normal text}\footnotesize @LamSonLeoc\includegraphics[height=1em]{Twitter_Bird}}}\vspace{-\baselineskip}}
 \date{%10 August 2016
@@
 
+\begin{frame}{Linear rheology}
+	\begin{columns}
+	\column{0.5\textwidth}
+	\tikzsetnextfilename{linear_rheology_PImBr}%
+	\input{fig_presentation/linear_rheology_PImBr.pgf}
+	\[ G = \frac{c}{N}k_\mathrm{B}T \]
+	\begin{description}[$N$]
+	\item[$c$] monomer concentration
+	\item[$N$] \# monomers between CL
+	\end{description}
+	\column{0.45\textwidth}
+	\pause
+	\begin{itemize}
+	\item $N \approx 800 N_0$ !
+	\item 800 chains between CL
+	\end{itemize}
+	\begin{flushright}\includegraphics[height=14\baselineskip]{presentation/Thaumetopea_pityocampa_01.jpg}\\
+	\vspace{-1.5em}\colorbox{lightgray}{\scriptsize Wikimedia commons}
+	\end{flushright}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Tweak repeated ions + counterions}
+	\vspace{1.5\baselineskip}
+	\tikzsetnextfilename{postfunc}%
+	\input{fig_presentation/postfunc.pgf}
+
+	\bigskip
+	\tikzsetnextfilename{postfunc_chem}%
+	\input{fig_presentation/postfunc_chem.pgf}
+\end{frame}
+
+\begin{frame}{Counterion condensation}
+	\tikzsetnextfilename{solubility}%
+	\input{fig_presentation/solubility.pgf}
+	\begin{itemize}
+		\item Too much condensation prevents dissolution
+		\item<2-> Direct Hoffmeister series $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$
+		\item<3> \ce{\pi+-\pi+} interactions {\tiny\setatomsep{1.5em}$\raisebox{-0.5em}{\chemfig{[2]*5(--N(-)---)}}<\raisebox{-0.5em}{\chemfig{[2]*5(-N(-)-=N-=)}}$}
+		\item<3> \ce{PIm+Br-} < \ce{PIm+I-} < \ce{PPyr+Cl-} < \ce{PPyr+Br-} < \ce{PPyr+I-}
+	\end{itemize}
+\end{frame}
+
 \end{document}

### 2016-07-12 PALSE/talk_Kyoto.tex
solubility in a row embed chemfig in a named tikzpicture
polyelectrolytes physics -> can we measure B from rheology?

@@
 \usepackage{tikz}
-\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, decorations.text, shapes.geometric, matrix,patterns,mindmap}
+\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, decorations.text, shapes.geometric, matrix,patterns,mindmap,fit}
 %\usetikzlibrary{positioning, patterns,topaths,chains,matrix}
@@
 
-\begin{frame}{Counterion condensation}
+\begin{frame}<1>[label=condensationranking]{Counterion condensation}
 	\tikzsetnextfilename{solubility}%
@@
 		\item Too much condensation prevents dissolution
-		\item<2-> Direct Hoffmeister series $\ce{I-}<\ce{Br-}<\ce{Cl-}<\ce{F-}$
-		\item<3> \ce{\pi+-\pi+} interactions {\tiny\setatomsep{1.5em}$\raisebox{-0.5em}{\chemfig{[2]*5(--N(-)---)}}<\raisebox{-0.5em}{\chemfig{[2]*5(-N(-)-=N-=)}}$}
-		\item<3> \ce{PIm+Br-} < \ce{PIm+I-} < \ce{PPyr+Cl-} < \ce{PPyr+Br-} < \ce{PPyr+I-}
+		\begin{itemize}
+			\item water poor solvent for polymer backbone
+			\item polyelectrolytes solubilized by electrostatic repulsion
+		\end{itemize}
+		\item<2-> Direct Hoffmeister series $\ce{I-}\alert{<}\ce{Br-}\alert{<}\ce{Cl-}\alert{<}\ce{F-}$\\
+	\hfill\textit{\scriptsize Hofmeister 1888; Zhang Annu Rev Phys Chem 2010; Schwierz Langmuir 2010}
+		\item<3-> \ce{\pi+-\pi+} interactions trap counterions\raisebox{-0.5em}{%
+			\begin{tikzpicture}[every node/.style={inner ysep=0,font=\tiny\setatomsep{1.5em}}]
+			\node[left=0.5em] (Pyr){\chemfig{[0]*5(--N(-)---)}};
+			\node[right=0.5em] (Im){\chemfig{[2]*5(-N(-)-=N-=)}};
+			\node[font=\normalsize]{$\alert{<}$};
+			\end{tikzpicture}}\\
+			\hfill\textit{\scriptsize
+			Geronimo PCCP 2011;
+			Matthews PCCP 2014;
+			Gao J Mol Graph Model 2015}
 	\end{itemize}
@@
 
+\begin{frame}{Condensation and solubility}
+	\begin{itemize}
+		\item All counterions condensed $\Rightarrow$ neutral $\Rightarrow$ not soluble
+	\end{itemize}
+	\tikzsetnextfilename{polyelectrolytes_condensed}%
+	\input{fig_presentation/polyelectrolytes_condensed.pgf}
+	\begin{itemize}
+		\item ``Enough'' free counterions $\Rightarrow$ repulsion $\Rightarrow$ soluble
+	\end{itemize}
+	\tikzsetnextfilename{polyelectrolytes}%
+	\input{fig_presentation/polyelectrolytes.pgf}
+\end{frame}
+
+\againframe<2-3>{condensationranking}
+
+\begin{frame}{\# chains between cross links}%
+\tikzsetnextfilename{frequency}%
+\input{fig_presentation/frequency.pgf}
+\begin{itemize}
+\item Procession length increases with counterion condensation
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Polyelectrolytes physics}
+	\structure{Can we measure counterion condensation from rheology?}
+
+	\begin{itemize}
+		\item 1 dissociated counterion every \alert{$A$} monomers
+		\begin{itemize}\item large $A$, more condensed\end{itemize}
+		\item poor solvent: reduced temperature $\alert{\tau} = 1 -T/\Theta$
+		\begin{itemize}\item \ce{PIm+F-} non soluble at \SI{100}{\celsius} $\Rightarrow\tau>0.2$ \end{itemize}
+		\smallskip
+		\item extension parameter $\alert{B} = \frac{\text{extended length}}{\text{diameter}}$ of electrostatic blob
+		\smallskip
+		\item from scaling theory: $B \sim A^{4/3} \tau$\hfill \textit{\scriptsize Rubinstein Macromol. 1996}
+	\end{itemize}
+
+	\tikzsetnextfilename{scales}%
+	\input{fig_presentation/scales.pgf}
+	\begin{center}
+	\structure{Can we measure $B$ from rheology?}
+	\end{center}
+\end{frame}
+
 \end{document}

### 2016-07-13 PALSE/talk_Kyoto.tex
scaling theory in the middle of the slide
forgotten strain.pgf extention of the electrostatic blobs => tau => E_c => A

@@
 \begin{frame}{Polyelectrolytes physics}
+	\vspace{\baselineskip}
 	\structure{Can we measure counterion condensation from rheology?}
@@
 		\begin{itemize}\item \ce{PIm+F-} non soluble at \SI{100}{\celsius} $\Rightarrow\tau>0.2$ \end{itemize}
-		\smallskip
+	\end{itemize}
+	\alert{Scaling theory}\hfill \textit{\scriptsize Rubinstein Macromol. 1996}
+	\tikzsetnextfilename{scales}%
+	\input{fig_presentation/scales.pgf}
+
+	\begin{itemize}
 		\item extension parameter $\alert{B} = \frac{\text{extended length}}{\text{diameter}}$ of electrostatic blob
 		\smallskip
-		\item from scaling theory: $B \sim A^{4/3} \tau$\hfill \textit{\scriptsize Rubinstein Macromol. 1996}
+		\item scales $\geq D$ depend only on $B \sim A^{4/3} \tau$
 	\end{itemize}
 
-	\tikzsetnextfilename{scales}%
-	\input{fig_presentation/scales.pgf}
 	\begin{center}
@@
 
+\begin{frame}<1>[label=endlinear]{Stretching a polyelectrolyte chain}
+	\vspace{\baselineskip}
+	\begin{itemize}
+		\item Unfold largest scales first\hfill \textit{\scriptsize Pincus Macromol. 1976}
+	\end{itemize}
+
+	\tikzsetnextfilename{scales_stretch}%
+	\begin{tikzpicture}
+	\node[inner sep=0](scales){\input{fig_presentation/scales.pgf}};
+	\draw[->, ultra thick, ilmcolor] (scales.south east) ++(0,-0.3em) -- +(-0.33\textwidth,0) node[midway, below] {linear};
+	\draw[->, ultra thick,ilmorange] (scales.south) ++(0,-0.3em) -- +(-0.35\textwidth,0) node[midway, below] {non linear};
+	\end{tikzpicture}
+
+	\pause
+	\begin{itemize}
+		\item narrow linear regime $\Rightarrow R_0\approx r_\mathrm{scr} \Rightarrow \left(\frac{B}{B_0}\right)^3 = 1 + \frac{B}{B_s}$
+		\begin{itemize}
+			\item \alert{$B_s$} screening from \ce{-PO_3^{2-}} heads\tikzsetnextfilename{head}
+				\begin{tikzpicture}
+					\let\mrad\relax%
+					\newlength\mrad%
+					\setlength{\mrad}{1ex}%
+					%face
+					\input{fig_presentation/head.pgf}
+					%% horns
+					\draw[every node/.style={shape=circle,inner sep=0.2\mrad,shading=radial,inner color=white!50!ilmcolor,outer color= ilmcolor!90!black}] (emoticon.70) -- ++(70:0.4\mrad) node{} (emoticon.110) -- ++(110:0.4\mrad) node{};
+				\end{tikzpicture}
+			\item if $B \ll B_s$ screening from uncondensed counterions \[B = B_0 \sim bc \left(\frac{k_\mathrm{B}T}{G}\right)^{2/3}\]
+			\item if $B \gg B_s$ screening from heads: $B = \left(B_0^3/B_s\right)^{1/2} > B_0$
+		\end{itemize}
+	\end{itemize}
+\end{frame}
+
+\begin{frame}<1>[label=strainsweep]{Strain sweep}
+	\vspace{0.5\baselineskip}
+	\tikzsetnextfilename{strain}%
+	\input{fig_presentation/strain.pgf}
+
+	\vspace{-0.5\baselineskip}
+	\begin{overprint}%
+	\onslide<-2>%
+	\begin{itemize}
+		\item narrow linear regime $\gamma_0\ll 1$
+		\item<2> extension parameter increases with condensation
+	\end{itemize}
+	\onslide<3->%
+	\begin{itemize}
+		\item bonding energy increases with $B$
+		\item<4> measure of counterion condensation $A$
+%		\begin{description}
+%			\item[\ce{PPyr+I-}] $10$ uncondensed counterions per polymer
+%			\item[\ce{PIm+Br-}] a single dissociated counterion every $9$ polymers
+%		\end{description}
+	\end{itemize}
+	\end{overprint}
+\end{frame}
+
+\againframe<2>{endlinear}
+\againframe<2>{strainsweep}
+
+\begin{frame}{Extension of the electrostatic blobs}
+\vspace{\baselineskip}
+
+\tikzsetnextfilename{stretch}%
+	\tikzsetnextfilename{stretch}%
+	\input{fig_presentation/stretch.pgf}
+	\[E_\mathrm{stretch} = \text{area increment}(\gamma,B) \times \text{surface tension}(\tau,B)\]
+
+	\begin{itemize}
+		\item fluidisation = rupture $\Rightarrow E_\mathrm{stretch}(\gamma_c) = E_\text{head-body}$
+		\begin{itemize}
+			\item \structure{$\tau(E_\text{head-body}, B, \gamma_c)$}
+		\end{itemize}
+		\item ionic bond energy depends on local $\epsilon_r$%\\\hfill\textit{\scriptsize Rehm Chem Soc Rev. 2010}
+		\begin{itemize}
+			\item in water $E_\text{head-body}\approx 2k_\mathrm{B}T$
+			\item large $B\Rightarrow$ low $\epsilon_r$ $\Rightarrow$ high $E_\text{head-body}$
+		\end{itemize}
+		\item smallest $B$ is \ce{PPyr+I-} $\Rightarrow \textcolor{ilmcolor}{\tau = 0.4\pm0.07}, \Theta\approx\SI{220}{\celsius}$
+	\end{itemize}
+\end{frame}
+
+\againframe<3-4>{strainsweep}
+
 \end{document}

### 2016-08-08 PALSE/talk_Kyoto.tex
fixes from Sebastien comments: coherent colors between strain sweeps and stretch
funny pics for intro on interdisciplinary
Help the flow Counterion condensation -> Behaviour in water extension parameter -> compaction parameter Extension of the electrostatic blobs -> Measure reduced temperature Reorganise conclusion order

@@
 \pgfplotsset{compat=1.9}
+\usepgfplotslibrary{fillbetween}
 \usepgfplotslibrary{groupplots}
@@
 
+\begin{frame}{What chemists/physicists think I do}
+	\begin{columns}[t]
+	\column{0.5\textwidth}
+	\begin{flushleft}\structure{chemistry}
+
+	``This counterion is more chaotropic''
+
+	\vspace{\baselineskip}
+	\includegraphics[width=\textwidth]{presentation/sorciere}
+	\end{flushleft}
+	\column{0.5\textwidth}
+	\begin{flushright}\structure{rheology}
+
+	``We impose a constant shear stress''
+	\end{flushright}
+	\hspace{0.25\textwidth}\includegraphics[height=10\baselineskip]{presentation/rascar}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Physical chemistry positive feedback}
+\tikzsetnextfilename{chem_rheol}%
+	\input{fig_presentation/chem_rheol.pgf}
+\end{frame}
+
 \begin{frame}{Fine chemistry}
@@
 	\item in deionised, neutral water
-	\item yield stress fuild
+	\item yield stress fluid
 \end{itemize}
@@
 \column{0.5\textwidth}
-\hfill Herschel-Buckley law\\
+\hfill Herschel-Bulkley law\\
 \hfill $\sigma =\sigma_c + A \dot{\gamma}^n$
@@
 
-\begin{frame}{Linear rheology}
+\begin{frame}{Linear rheology: A physicist problem}
 	\begin{columns}
@@
 
-\begin{frame}{Tweak repeated ions + counterions}
+\begin{frame}{Tweaking repeated ions + counterions}
 	\vspace{1.5\baselineskip}
@@
 	\input{fig_presentation/postfunc_chem.pgf}
+
+	\vspace{-3\baselineskip}\hfill \begin{minipage}{0.35\textwidth}``Can you put that in the rheometer?''\end{minipage}
 \end{frame}
 
-\begin{frame}<1>[label=condensationranking]{Counterion condensation}
+\begin{frame}<1>[label=condensationranking]{Behaviour in water}
 	\tikzsetnextfilename{solubility}%
@@
 			Gao J Mol Graph Model 2015}
+		\item<3-> 5 gels ranked by counterion condensation
 	\end{itemize}
@@
 	\begin{itemize}
-		\item extension parameter $\alert{B} = \frac{\text{extended length}}{\text{diameter}}$ of electrostatic blob
+		\item compaction parameter $\alert{B} = \frac{\text{extended length}}{\text{diameter}}$ of electrostatic blob
 		\smallskip
@@
 		\item narrow linear regime $\gamma_0\ll 1$
-		\item<2> extension parameter increases with condensation
+		\item<2> compaction parameter increases with condensation
 	\end{itemize}
-	\onslide<3->%
+	\onslide<3>%
 	\begin{itemize}
-		\item bonding energy increases with $B$
-		\item<4> measure of counterion condensation $A$
-%		\begin{description}
-%			\item[\ce{PPyr+I-}] $10$ uncondensed counterions per polymer
-%			\item[\ce{PIm+Br-}] a single dissociated counterion every $9$ polymers
-%		\end{description}
+		\item bonding energy increases with compaction (condensation)
+		\item allows larger strains
+	\end{itemize}
+	\onslide<4>%
+	\begin{itemize}
+		\item $A \sim (B/\tau)^{3/4}$
+		\item measure of counterion condensation $A$
 	\end{itemize}
@@
 
-\begin{frame}{Extension of the electrostatic blobs}
+\begin{frame}{Measure reduced temperature}
 \vspace{\baselineskip}
-
+\begin{itemize}
+		\item fluidisation = rupture: stretching breaks head-to-body
+\end{itemize}
 \tikzsetnextfilename{stretch}%
@@
 	\input{fig_presentation/stretch.pgf}
-	\[E_\mathrm{stretch} = \text{area increment}(\gamma,B) \times \text{surface tension}(\tau,B)\]
+	\[E_\mathrm{stretch} = \text{area increment}(\gamma,B) \times \text{surface tension}(\textcolor{ilmcolor}{\tau},B)\]
 
 	\begin{itemize}
-		\item fluidisation = rupture $\Rightarrow E_\mathrm{stretch}(\gamma_c) = E_\text{head-body}$
+		%\item \structure{$\tau(E_\text{head-body}, B, \gamma_c)$}
+		\item for \ce{PPyr+I-}
 		\begin{itemize}
-			\item \structure{$\tau(E_\text{head-body}, B, \gamma_c)$}
+			\item little compaction $B\approx 2.4$
+			\item ionic bond in water: $E_\text{head-body}\approx 2k_\mathrm{B}T$
+			\item $\textcolor{ilmcolor}{\tau = 0.4\pm0.07}\qquad \Theta\approx\SI{220}{\celsius}$
 		\end{itemize}
@@
 		\begin{itemize}
-			\item in water $E_\text{head-body}\approx 2k_\mathrm{B}T$
-			\item large $B\Rightarrow$ low $\epsilon_r$ $\Rightarrow$ high $E_\text{head-body}$
+			\item large $B\Rightarrow$ large electrostatic blobs with low $\epsilon_r$
+			\item stronger head-to-body bonds
 		\end{itemize}
-		\item smallest $B$ is \ce{PPyr+I-} $\Rightarrow \textcolor{ilmcolor}{\tau = 0.4\pm0.07}, \Theta\approx\SI{220}{\celsius}$
 	\end{itemize}
@@
 
+\begin{frame}{Conclusion}
+	\begin{itemize}
+		\item post-functionalisation allows to change $A\times 100$
+		\begin{itemize}
+			\item \ce{PPyr+I-}: 10 free counterions per polymer
+			\item \ce{PIm+Br-}: 1 free counterion every 9 polymers
+		\end{itemize}
+		\item mechanical properties $\times 1000$
+		\begin{itemize}
+			\item \ce{PPyr+I-}: $G\approx\SI{10}{\kilo\pascal}$, fluidizes at $\gamma_c\approx 2\%$
+			\item \ce{PIm+Br-}: $G\approx\SI{10}{\pascal}$, fluidizes at $\gamma_c\approx 800\%$
+		\end{itemize}
+		\item measure of microscopic parameters from rheology
+		\begin{itemize}
+			\item counterion condensation
+			\item head-to-body bonding energy
+		\end{itemize}
+		\item highly tunable hydrogels from short, linear polymer chains
+		\begin{itemize}
+			\item no additive
+			\item yield stress fluids
+		\end{itemize}
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Processionary phase diagram}
+	\begin{columns}
+	\column{0.65\textwidth}
+	\vspace{\baselineskip}
+	\tikzsetnextfilename{phase_diag}%
+	\input{fig_presentation/phase_diag.pgf}
+	\column{0.35\textwidth}
+	\begin{flushright}\includegraphics[height=14\baselineskip]{presentation/Thaumetopea_pityocampa_01.jpg}\\
+	\vspace{-1.5em}\colorbox{lightgray}{\scriptsize Wikimedia commons}
+	\end{flushright}
+	\end{columns}
+\end{frame}
+
 \end{document}

### 2016-09-01 PALSE/cover_letter.tex
cover letter of processionnary paper in Soft Matter

\documentclass[a4paper, parskip=true, firsthead=false, fromemail=true, foldmarks=false]{scrlttr2}
\usepackage[utf8]{inputenc}

\setkomavar{fromemail}{mathieu.leocmach@univ-lyon1.fr}
\setkomavar{signature}{Hassan Srour, Martien Duvall Deffo Ayagou, Thi Thanh-Tam Nguyen, Nicolas Taberlet, S\'{e}bastien Manneville, Chantal Andraud, Cyrille Monnereau, and Mathieu Leocmach}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[british]{babel}

\usepackage{kpfonts}
\usepackage{tabu}
\usepackage{wrapfig}
\setlength{\columnsep}{0.05\textwidth}
\usepackage{ifthen}
\usepackage{hyperref}
\usepackage{enumitem}
\setlist{nosep}

%\usepackage{tikz}
%\definecolor{ilmcolor}{RGB}{177,37,159}
%\definecolor{ilmorange}{RGB}{255,116,40}
%\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, decorations.text, shapes.geometric, matrix,patterns,mindmap,fit}

\newcommand{\journal}{Soft Matter}

\begin{document}
\begin{letter}{From:\\
Mathieu Leocmach,\\
Univ Lyon,\\ 
Universit\'e Claude Bernard Lyon~1, CNRS,\\
Institut Lumi\`ere Mati\`ere,\\
F-69622, VILLEURBANNE,\\
France\\
\texttt{mathieu.leocmach@univ-lyon1.fr}
}
\opening{\bf Dear Editor,}

\begin{wrapfigure}[16]{r}[0pt]{0.25\textwidth}
\hfill
\includegraphics[width=\linewidth]{presentation/Thaumetopea_pityocampa_01.jpg}

\raggedright
Pine processionary caterpillars
\end{wrapfigure}

Please find enclosed our manuscript, entitled \textsc{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}, which we are submitting for publication as an article to \journal.

Polymer-based hydrogels have promising biomedical applications, eg. as a soft scaffold for cellular growth. To be used in vivo, solid pieces of gel have to be implanted surgically. By contrast a physical gel with a yield stress rheology would allow a a much less traumatic delivery via injection: solid in the syringe, the gel could flow through the needle and reform as a solid once in the body. In the present work we design and study such a polymeric physical gel.

The system we investigate is made of short (70 units) linear polymer chains. In aqueous solution, a single extremity (the `head') is charged negatively whereas the repeated unit of the `body' carries a positive charge. We demonstrate that these chains self assemble into long processions (up to 800 chains) due to electrostatic interactions. %These processions are entangled or cross-linked on the scale of the screening length. 
The resulting physical hydrogel behaves as a yield stress fluid. In this article we build a microscopic understanding of the chemical physics involved in this gel that allows us engineer its rheological properties.

We have demonstrated previously [\href{ https://doi.org/10.1002/marc.201400478}{Srour et. al, Macromol Rapid Commun. 2015.}] how to obtain well-controlled anion-terminated poly(cationic) polymer via atom transfer radical polymerisation and post-functionalisation. Here we harness the power of this post-functionalisation approach to systematically alter the repeated unit. From a common scaffold we generate a family of such polyelectrolytes where the cationic moieties can be either all aromatic (imidazolium) or all aliphatic (pyrrolidinium), with counterions ranging from fluoride to iodide along the Hofmeister series. We show that the resulting difference in counterion condensation alters the rheological properties of the resulting hydrogel over three orders of magnitudes.

From these experimental results, we develop the ``processionary'' model briefly exposed above. This model builds on the scaling theory of polyelectrolytes gels and allows us to quantify microscopic observables (e.g. the amount of counterion condensation, head-to-body bonding energy) from standard macroscopic rheology procedures (oscillatory strain amplitude sweep). Conversely this model exposes all the relevant parameters we can tune to design physical hydrogels with desired rheological properties.

Our findings are at the interface between polymer chemistry and rheology, bridged by an understanding of the underlying physics, with possible biomedical applications.

We hope that you will find that the paper is interesting enough to be published in \journal.

We suggest the following scientists as possible reviewers, by alphabetical order:
\begin{enumerate} \setlength\itemsep{0em}
\item Andrey V. Dobrynin, University of Connecticut, 
avd@ims.uconn.edu
\item Elie Raphaël, UMR CNRS Gulliver 7083, ESPCI, elie.raphael@espci.fr
\item Michael Rubinstein, University of North Carolina, mr@unc.edu
\item Annette M. Schmidt, University of Cologne, annette.schmidt@uni-koeln.de
\item Nadine Schwierz, University of California, Berkeley, nschwierz@berkeley.edu
\end{enumerate}

Warm thanks to you for your kind attention and cooperation.

\closing{\bf Sincerely yours,}

\end{letter} 
\end{document}

### 2016-10-23 response.tex
Réponse aux referees: mise en forme et questions de rhéologie

\documentclass[a4paper, parskip=true, firsthead=false, fromemail=true, foldmarks=false]{scrlttr2}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[british]{babel}
\usepackage{kpfonts}
\usepackage{siunitx}
\usepackage[version=3]{mhchem}
\usepackage{hyperref}

\newenvironment{quotationi}
{\begin{quotation}\itshape}
{\end{quotation}}

\setkomavar{fromemail}{mathieu.leocmach@univ-lyon1.fr}
\setkomavar{signature}{Hassan Srour, Martien Duvall Deffo Ayagou, Thi Thanh-Tam Nguyen, Nicolas Taberlet, S\'{e}bastien Manneville, Chantal Andraud, Cyrille Monnereau, and Mathieu Leocmach}

\newcommand{\journal}{Soft Matter}

\begin{document}
\begin{letter}{From:\\
Mathieu Leocmach,\\
Univ Lyon,\\ 
Universit\'e Claude Bernard Lyon~1, CNRS,\\
Institut Lumi\`ere Mati\`ere,\\
F-69622, VILLEURBANNE,\\
France\\
\texttt{mathieu.leocmach@univ-lyon1.fr}
}
\opening{\bf Prof. Paul Janmey,}

Thank you very much for your e-mail concerning our manuscript (SM\nobreakdash-ART\nobreakdash-09\nobreakdash-2016-002022) together with the comments of the Reviewers.

Following the constructive comments and suggestions of the Reviewers, we have revised our manuscript. 
We believe that we have been able to answer the comments of both reviewers on a satisfactory level and thus the revised manuscript has been much improved, thanks to the valuable comments of both Reviewers. We note that the parts revised to answer the Reviewers have been highlighted with blue characters in the revised manuscript.

We hope that you and your reviewers would find that the revised manuscript is now suitable for publication in \journal.

\closing{\bf Sincerely yours,} 
\clearpage

\textsf{\textbf{Replies to the comments of Reviewer \#1}}

\begin{quotationi}
The submitted work combines synthesis with theoretical calculation of polymer based hydrogels and is a very interesting manuscript to be published in the journal Soft Matter. The authors explain in a well-structured way their experiments and the conclusions are supported by a variety of analytical techniques. Besides synthetic work, the reader also gets an insight into the processionary model. Nevertheless, I have some suggestions that might help improving this publication:
\end{quotationi}

We are very glad to see the positive comments of the Reviewer. Below we reply to the comments of the Reviewer one by one.

\begin{quotationi}
Page 5, Fig. 1: Br end group of the polymer POH/PBr and Nu end group in PNu+Br- is missing in the structure.
\end{quotationi}

\begin{quotationi}
Page 5, chapter 2.1, beginning of 1st section (polymer analysis): Why is Mw mentioned there? I suggest that Mn is presented. Was Mn calculated via monomer conversion or end group analysis via 1H NMR? Mw/Mn = 1.089 should be rounded up to 1.09.
\end{quotationi}

\begin{quotationi}
Page 5, chapter 2.1, end of 2nd section (peaks of Br-/I-, detection limits for F-/Cl-): Where do 89/91 and 121 come from? The peaks of Br- and I- should be 79/81 and 126. The content of the text does not fit to what is shown in S2 and S3. It also should be found an alternative analysis method to proof the successful anion exchange for F- and Cl-.
\end{quotationi}

\begin{quotationi}
Page 6, chapter 2.2, beginning of 1st section (listing of compounds): Abbreviations can be used here for bromide, iodide 
\end{quotationi}

This is done.

\begin{quotationi}
Page 7, Fig. 4: Why are all the amplitude sweeps measured until different deformation values and not all from 10-3 till e.g. 102? What is the unit there [\%]? Further, weight percent is abbreviated as wt\%.
\end{quotationi}

As the Reviewer kindly suggested, the amplitude sweeps were programmed to run between \num{1e-3} and \num{1e2} for all samples. However the rheometer can only apply a torque larger than a minimum value to perform reliable measurements. Beyond the plastic regime, the samples are shear thinning and the rheometer always reaches its minimum torque at strain amplitudes smaller than \num{1e2}. Since further measurement points are not reliable we do not display them.

As it is common in the rheology literature, strains are given without units since they are non dimensional. A strain of 1 is 100\%.

\begin{quotationi}
Page 7, chapter 2.3: In the text you mention the materials are shear thinning. The amplitude sweeps shown indicate that the structure of the material is changed when leaving the linear viscoelastic region. Because both moduli are decreasing this is indicating shear thinning properties. A clear prove for shear thinning would be given using rotational experiments (viscosity over shear rate plots).
\end{quotationi}

As mensioned in the text, our samples display a strong Weissenberg effect. In our cone-plane geometry, it results in outward ejection of the sample at moderately high shear rates. To free us from Weissenberg effect we could have used a Couette geometry, but the amount of material to synthesise to fill this geometry would have been prohibitive (about 20 times the cone-plate volume).

\ce{PIm+Br-} was the most well-behaved gel by this standard and we were able to perform flow curves on a reasonable range of shear rates (see Fig 2 of our previous paper \href{http://doi.org/10.1002/marc.201400478}{Srour et al. Macromolecular Rapid Communications 2015}). Indeed viscosity drops as shear rate increases confirming our claim of shear thinning.

Qualitatively the other gels follow this trend, but the range of accessible shear rates is much more limited before Weissenberg-induced sample ejection kicks in (less than a decade for \ce{PPyr+I-}). Therefore we decided not to present flow curve data and instead show only the more reliable amplitude sweeps.

\begin{quotationi}
Page 7, chapter 2.3, beginning of 2nd section (imidazolium and pyrrolidinium based gels): Please insert here a short explanation why immidazolium based gels are much weaker than their pyrrolidinium counterpart.
\end{quotationi}

\begin{quotationi}
Page 10, chapter 2.4.3: typing error (“the the correlation”)
\end{quotationi}

\begin{quotationi}
Supporting Information
Please insert 1H NMR spectrum of POH in order to understand the calculation of Mn.
\end{quotationi}

\begin{quotationi}
Page 2, 1st chapter (SEC analysis): Please change polydispersity index PDI to dispersity Ð.
\end{quotationi}

\begin{quotationi}
Page 8: Weight percent is abbreviated as wt\%.
\end{quotationi}

We hope that the Reviewer would think that the revised manuscript has been much improved and is now suitable for publication in \journal.

\clearpage

\textsf{\textbf{Replies to the comments of Reviewer \#3}}

First we would like to thank the reviewer for having carefully read our manuscript and provided valuable comments and suggestions.

\begin{quotationi}
The authors report on rheological characterization and theoretical interpretation of properties of polyelectrolyte assemblies prepared from molecules with a single anionic headgroup and cationic groups along the chain. The presented work including preparation of the polyelectrolytes, their rheological characterization in solution as well as theoretical interpretation, is considered to be conducted in a nice way, it indicate the role of counterions in affecting the assembly modes of the structures and novel assembly modes are suggested. There are, however, some concerns that should be addressed before publication can be recommended:

Although offering novel handles to generate molecular assemblies, and for this reason, is considered to be general interest, the suggestion for the particular structures being assembled could be further bolstered by scattering tools.
\end{quotationi}

\begin{quotationi}
In applying network theory to interpret the data, it appears that e.g. loose end corrections are not employed. What is the rational for that?
\end{quotationi}

\begin{quotationi}
In the introduction, the authors refer to interest in hydrogels in general is related to their application potential in biomedicine, but do not explicitly relate the particular conditions used to modulate properties also to such an application domain later in the manuscript. Are the various solution conditions studied compatible with such applications? Are the results sensitive to addition of a background electrolyte (e.g., similar to physiological ionic strength)?

\end{quotationi}

We hope that the Reviewer would think that the revised manuscript is now suitable for publication in \journal.

%\cc{Cclist} 
%\ps{adding a postscript} 
%\encl{list of enclosed material} 
\end{letter} 
\end{document}

### 2016-10-23 response.tex
add supplementary figure on 1H NMR of POH

@@
 
+Although the methodology used for polymer chain length determination has been reported in full in one of our previous papers (Appukuttan et al.,2012), we decided, following referee's recommendation, and for reader's comfort, to include a 1H NMR spectrum of the precursor POH polymer used in this study.
+In the caption, the calculation used to determine polymer chain length is explained.
+
 \begin{quotationi}
@@
 The peaks of Br- and I- should be 79/81 and 126.
-The content of the text does not fit to what is shown in S2 and S3.
+The content of the text does not fit to what is shown in S2 and S3.\end{quotationi}
+
+We thank the referee for spotting this typo.
+
+\begin{quotationi}
  It also should be found an alternative analysis method to proof the successful anion exchange for F- and Cl-.
@@
 
+This paragraph was meant to be descriptive and not explicative.
+At this point in the text we do not have an explanation on why immidazolium based gels are much weaker than their pyrrolidinium counterpart or why \ce{Cl-} gives a weaker gel than \ce{Br-}.
+We only note that gel shear modulus follows counter-ion condensation.
+Explanations will come with the model.
+We rephrased the first sentence of the paragraph to make its purpose clearer.
+
 \begin{quotationi}
@@
 \end{quotationi}
+
+Fixed.
 
@@
 \end{quotationi}
+
+Done.
 

### 2016-10-23 PALSE/response.tex
remove old parenthesis position labels Add Br on tails of POH and PBr Add Nu+Br- on tail of PNuBr (not in presentation to keep people focused)
dispersity \DJ

--- 2016-10-23 response.tex
@@
 
+We thank the Reviewer for spotting this omission.
+Actually we are not completely sure that nucleophile substitution actually takes place on a tertiary carbon, but indeed a \ce{Nu+Br-} tail group is our best guess.
+
 \begin{quotationi}
@@
 
+%we changed the text to give Mn instead of Mw
+
 \begin{quotationi}
@@
 
-We thank the referee for spotting this typo.
+We thank the Reviewer for spotting this typo.
 
@@
 
-As the Reviewer kindly suggested, the amplitude sweeps were programmed to run between \num{1e-3} and \num{1e2} for all samples.
+As the Reviewer kindly suggested, the amplitude sweeps were programmed to run between $10^{-3}$ and $10^2$ for all samples.
 However the rheometer can only apply a torque larger than a minimum value to perform reliable measurements.
-Beyond the plastic regime, the samples are shear thinning and the rheometer always reaches its minimum torque at strain amplitudes smaller than \num{1e2}.
+Beyond the plastic regime, the samples are shear thinning and the rheometer always reaches its minimum torque at strain amplitudes smaller than $10^2$.
 Since further measurement points are not reliable we do not display them.
@@
 
-As mensioned in the text, our samples display a strong Weissenberg effect.
+As mentioned in the text, our samples display a strong Weissenberg effect.
 In our cone-plane geometry, it results in outward ejection of the sample at moderately high shear rates.
@@
 \begin{quotationi}
-Page 2, 1st chapter (SEC analysis): Please change polydispersity index PDI to dispersity Ð.
+Page 2, 1st chapter (SEC analysis): Please change polydispersity index PDI to dispersity \DJ.
 \end{quotationi}

### 2016-10-23 PALSE/ion_pairing_SI.tex
dispersity \DJ

--- 2016-02-01 PALSE/ion_pairing_SI.tex
@@
 
-SEC analyses were carried out on ca \SI{5}{\milli\gram\per\milli\litre} solutions of each polymer in DMSO LiBr (\SI{0.01}{M}), using a MALVERN VISCOTEK 430  system equipped with three PSS Gram columns (Polyester Copolymer Network) set in series at \SI{80}{\celsius}, coupled to a Viscotek HT-GPC Module 350 A (refractive and viscosimeter) detector (at \SI{80}{\celsius}) .
-The polydispersity indexes (PDI = Mw/Mn) of the samples were derived from the RI signal by a universal calibration curve by pullulan.
+SEC analyses were carried out on ca \SI{5}{\milli\gram\per\milli\litre} solutions of each polymer in DMSO LiBr (\SI{0.01}{M}), using a MALVERN VISCOTEK 430  system equipped with three PSS Gram columns (Polyester Copolymer Network) set in series at \SI{80}{\celsius}, coupled to a Viscotek HT-GPC Module 350 A (refractive and viscosimeter) detector (at \SI{80}{\celsius}).
+The dispersity (\DJ = Mw/Mn) of the samples were derived from the RI signal by a universal calibration curve by pullulan.
 The number-average molar masses (Mn) were calculated from RI signals with the OmniSEC 4.6 software.
+
+The thermal analysis was obtained on Differential Scanning Calorimetry (DSC).
+All samples were investigated using a Setaram DSC 131 apparatus.
+A weighed sample of \SI{5.0\pm 1}{\milli\gram} was placed in a sealed stainless crucible.
+A standard heating rate of \SI{5}{\kelvin\per\minute} was applied from 25 up to \SI{200}{\celsius} under \SI{50}{\milli\litre\per\minute} of nitrogen.
+The start and the onset temperatures were determined in order to assess the thermal stability.
+$T_\mathrm{start}$ is the temperature at which the sample starts to lose some mass and $T_\mathrm{onset}$ was determined from the step tangent method.
 
@@
 
+Frequency sweeps were performed within the linear domain by decreasing the frequency from $f=\SI{10}{\hertz}$ down to $f=\SI{0.01}{\hertz}$ at a strain amplitude $\gamma = 0.1 \%$ and at a temperature (\SI{25}{\celsius}) fixed by a Peltier plate.
+Strain sweeps were performed by ramping up the strain amplitude from $\gamma = 0.1 \%$ up to $\gamma =5000\%$ at $f=\SI{1}{\hertz}$ in the same temperature conditions.
+Most tests had to be stopped before the strain could reach its highest value due to sample ejection from the cone-plate geometry.
+
 \begin{figure}
 \includegraphics[width=\textwidth]{sec_spectra.png}
+\begin{tabular}{llll}
+Peak RV - (ml) & 23.72 & 
+Mn - (Daltons) & 5 155 \\
+Mw - (Daltons) & 5 614 &
+Mz - (Daltons) & 5 985 \\
+Mp - (Daltons) & 7 221 &
+Mw / Mn & 1.089\\
+Percent Above Mw: 0 & 100.0 &
+Percent Below Mw: 0 & 0.0\\
+Mw 10.0\% Low & 3 012 &
+Mw 10.0\% High & 7 087\\
+IV - (dl/g) & 0.6026 &
+Rh(w) - (nm) & 3.899\\
+Wt Fr (Peak) & 1.000 &\\
+Mark-Houwink a & 0.331&
+Mark-Houwink logK & -1.453\\
+Branches & 0.000 &
+Branch Freq. & 0.000\\
+RI Area - (mvml) & 98.76 &
+UV Area - (mvml) & 0.00\\
+RALS Area - (mvml) & 0.00 &
+LALS Area - (mvml) & 0.00\\
+IVDP Area - (mvml) & 105.24\\
+\end{tabular}
 \caption{SEC Spectra of \ce{POH}.}
+\label{fig:sec}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{POH_1H_NMR.png}
+\caption{1H NMR of POH in deuterated methanol, with signal attribution.
+Calibration of the integrals intensity, relative to the distinct chain-end signals (a and b, 6 protons each), allows us to determine polymer chain length $n$ (here, considering integration of signal d, $n=63$)}
 \end{figure}

### 2016-10-23 response.tex
wt\%

@@
 \begin{quotationi}
-Page 2, 1st chapter (SEC analysis): Please change polydispersity index PDI to dispersity Ð.
+Page 2, 1st chapter (SEC analysis): Please change polydispersity index PDI to dispersity \DJ.
 \end{quotationi}
+
+Done.
 

### 2016-10-23 response.tex
DLS loose ends

@@
 \usepackage{hyperref}
+\usepackage{graphicx}
 
@@
 
+In many instances we tried to perform DLS measurements on our samples and, indeed, signals were observed.
+However the obtained results were largely unreproducible (even for a given sample upon slight change of the concentration, not to mention when two different samples were used).
+An example is shown below.
+
+\includegraphics[width=\textwidth]{DLS.png}
+\textit{Figure: \textbf{a} DLS data and best fit \textbf{b} size distribution by diffusion intensity and \textbf{c} size distribution by number for PImBr (top row) and PPyBr (bottom row) in water at 10\% weight concentrations}
+
+As the Reviewer can see, dispersion was very high, and signal quality so low that cumulant fit analyses were absolutely irrelevant.
+In our opinion, there is nothing relevant in these data, this is why we finally chose not to include it in the SI.
+If the referee thinks that it might be worth adding some of these data to the article, we are ready to do it.
+However, we do know whether a reasonable interpretation could be provided from the data.
+We leave the referee judge; top is PImBr bottom PPyBr.
+Although a pattern seems to exist, it seems difficult to us to get its exact meaning, and to separate the artefact from the fact\ldots
+
 \begin{quotationi}
@@
 \end{quotationi}
+
+To our understanding, loose end corrections describe a situation where pre-existing linear chains of size $M$ are randomly cross-linked with an average of $M_c$ monomers between cross links with $M_c\leq M/2$.
+The modulus then scales as
+\begin{equation}
+\frac{G}{k_\mathrm{B}T} = \frac{c}{M_c}  \left(1 - \frac{2M_c}{M}\right).
+\end{equation}
+For example if $M_c$ is only one third of $M$, then two third of the original linear chain are loose ends and do not contribute to the elasticity.
+
+This is rather different from our situation where the number of monomers between cross links is larger or equal to the number of monomers in the original chain.
+This situation stems from the reactivity of the heads: a cross link necessarily occurs at the head of a polymer, preventing one of the possible loose end, but leaving the possibility of many loose ends at the tails of each polymer.
+
+We can better compare our situation to a gel formed by branched polymers, for example bottle brush polymers.
+In this situation, Papagiannopoulos et al.
+(\href{http://doi.org/10.1021/bm060287d}{Biomacromolecules 2006} or \href{http://doi.org/10.1039/b714864j}{Faraday Discussions 2008}) have shown that the shear modulus scales as
+\begin{equation}
+\frac{G}{k_\mathrm{B}T} = \frac{c}{N_\mathrm{tot}}.
+\end{equation}
+where $N_\mathrm{tot}$ is the total number of monomers between cross-links, including side chains that are topological `loose ends'.
+
+In other words, one can consider two possible definitions of ``cross link''.
+In Flory's definition, every branching point is a cross link, therefore a brush has many loose ends.
+In the more coarse-grained definition used by Papagiannopoulos et al., branching points leading to a loose end are not considered meaningful cross links, but every monomer of the side chains actually contribute to entropy and thus to elasticity.
 

### 2016-10-24 PALSE/cover_letter.tex
Final version of the cover letter (forgot to commit)

@@
 
-Please find enclosed our manuscript, entitled \textsc{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}, which we are submitting for publication as an article to \journal.
+Please find enclosed our manuscript, entitled \textsc{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}, which we are submitting for publication as a full paper to \journal.
 
@@
 By contrast a physical gel with a yield stress rheology would allow a a much less traumatic delivery via injection: solid in the syringe, the gel could flow through the needle and reform as a solid once in the body.
-In the present work we design and study such a polymeric physical gel.
+In the present work we design and study a family of such polymeric physical gels.
 
-The system we investigate is made of short (70 units) linear polymer chains.
+The systems we investigate are made of short (70 units) linear polymer chains.
 In aqueous solution, a single extremity (the `head') is charged negatively whereas the repeated unit of the `body' carries a positive charge.
@@
 Our findings are at the interface between polymer chemistry and rheology, bridged by an understanding of the underlying physics, with possible biomedical applications.
+In our opinion they are relevant and of importance for a broad community of scientists that encompass polymer and material chemists, physicochemists and soft matter physicists.
 
@@
 avd@ims.uconn.edu
+\item Kwang S. Kim, Ulsan National Institute of Science and Technology, kimks@unist.ac.kr
 \item Elie Raphaël, UMR CNRS Gulliver 7083, ESPCI, elie.raphael@espci.fr
-\item Michael Rubinstein, University of North Carolina, mr@unc.edu
+%\item Michael Rubinstein, University of North Carolina, mr@unc.edu
 \item Annette M. Schmidt, University of Cologne, annette.schmidt@uni-koeln.de
 \item Nadine Schwierz, University of California, Berkeley, nschwierz@berkeley.edu
+%\item István Szilágyi, University of Geneva, Istvan.Szilagyi@unige.ch
 \end{enumerate}

### 2016-10-24 ion_pairing_SI.tex
add supplementary figure on 1H NMR of POH
Title: SI of...

@@
 
-%\title{Supramolecular crocodile line controls the rheological properties of polyelectrolytes hydrogels}
-\title{Supplementary Informations to Ion pairing controls rheological properties of ``crocodile-line'' polyelectrolyte hydrogels}
-%\title{Ion pairing in crocodile line hydrogels controls rheological properties}
-%\title{In crocodile line hydrogels ion pairing controls rheological properties}
-%\title{Controlling the rheological properties of well-defined polyelectrolytes hydrogels by fine tuning their ion pairing process}
+\title{Supplementary Informations to\\ Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}
 
@@
 SEC analyses were carried out on ca \SI{5}{\milli\gram\per\milli\litre} solutions of each polymer in DMSO LiBr (\SI{0.01}{M}), using a MALVERN VISCOTEK 430  system equipped with three PSS Gram columns (Polyester Copolymer Network) set in series at \SI{80}{\celsius}, coupled to a Viscotek HT-GPC Module 350 A (refractive and viscosimeter) detector (at \SI{80}{\celsius}).
-The polydispersity indexes (PDI = Mw/Mn) of the samples were derived from the RI signal by a universal calibration curve by pullulan.
+The dispersity (\DJ = Mw/Mn) of the samples were derived from the RI signal by a universal calibration curve by pullulan.
 The number-average molar masses (Mn) were calculated from RI signals with the OmniSEC 4.6 software.
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{POH_1H_NMR.png}
+\caption{1H NMR of POH in deuterated methanol, with signal attribution.
+Calibration of the integrals intensity, relative to the distinct chain-end signals (a and b, 6 protons each), allows us to determine polymer chain length $n$ (here, considering integration of signal d, $n=63$)}
+\end{figure}
+
+\begin{figure}
 \includegraphics[height=\textheight-2\baselineskip]{mass_PImBr.png}
@@
 The fixed strain amplitude is $\gamma=0.1\%$.
-All samples are at 7.4~\%~wt }
+All samples are at 7.4~wt\%.
+}
 \label{fig:frequency}

### 2016-10-24 PALSE/ion_pairing_paper.tex
Cyrille correspondant

@@
 \usepackage{hyperref}
+\usepackage{xspace}
+
+\newcommand{\figSsec}{S1\xspace}
+\newcommand{\figSmassPImBr}{S2\xspace}
+\newcommand{\figSmassPPyBr}{S3\xspace}
+\newcommand{\figSmassPImI}{S4\xspace}
+\newcommand{\figSmassPPyI}{S5\xspace}
+\newcommand{\figSfrequency}{S6\xspace}
 
@@
 \definecolor{cream}{RGB}{222,217,201}
+\title{Ion pairing controls rheological properties of ``processionary'' polyelectrolyte hydrogels}
 
@@
 
- & \noindent\large{Hassan Srour,\textit{$^{a}$} Martien Duvall Deffo Ayagou,\textit{$^{a}$} Thi Thanh-Tam Nguyen,\textit{$^{a}$}, Nicolas Taberlet,\textit{$^{b\ddag}$} S\'{e}bastien Manneville,\textit{$^{b\ddag}$} Chantal Andraud,\textit{$^{a}$} Cyrille Monnereau,\textit{$^{a}$} and Mathieu Leocmach$^{\ast}$\textit{$^{c}$}} \\%Author names go here instead of "Full name", etc.
+ & \noindent\large{Hassan Srour,\textit{$^{a}$} Martien Duvall Deffo Ayagou,\textit{$^{a}$} Thi Thanh-Tam Nguyen,\textit{$^{a}$}, Nicolas Taberlet,\textit{$^{b}$} S\'{e}bastien Manneville,\textit{$^{b}$} Chantal Andraud,\textit{$^{a}$} Cyrille Monnereau,$^{\ast}$\textit{$^{a}$} and Mathieu Leocmach$^{\ast}$\textit{$^{c}$}} \\%Author names go here instead of "Full name", etc.
 
@@
 
-\footnotetext{\textit{$^{a}$~Univ Lyon, Ens de Lyon, Univ Claude Bernard, CNRS,
-Laboratoire de Chimie, F-69342 Lyon, France. }}%Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
-\footnotetext{\textit{$^{b}$~Univ Lyon, Ens de Lyon, Univ Claude Bernard, CNRS,
+\footnotetext{\textit{$^{a}$~Univ Lyon, Ens de Lyon, Universit\'e Claude Bernard Lyon 1, CNRS,
+Laboratoire de Chimie, F-69342 Lyon, France. E-mail: cyrille.monnereau@ens-lyon.fr}}%Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
+\footnotetext{\textit{$^{b}$~Univ Lyon, Ens de Lyon, Universit\'e Claude Bernard Lyon 1, CNRS,
 Laboratoire de Physique, F-69342 Lyon, France. }}
-\footnotetext{\textit{$^{c}$~Univ Lyon, Univ Claude Bernard, CNRS, Institut Lumière Matière, F-69622 Villeurbanne, France.
-E-mail: mathieu.leocmach@univ-lyon1.fr}}%Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
+\footnotetext{\textit{$^{c}$~Univ Lyon, Universit\'e Claude Bernard Lyon 1, CNRS, Institut Lumi\`ere Mati\`ere, F-69622, VILLEURBANNE, France.
+E-mail: mathieu.leocmach@univ-lyon1.fr.
+Twitter: @LamSonLeoc}}%Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
 
@@
 %If you article does not have ESI please remove the the \dag symbol from the title and the footnotetext below.
-\footnotetext{\dag~Electronic Supplementary Information (ESI) available: [details of any supplementary information available should be included here]. See DOI: 10.1039/b000000x/}
+\footnotetext{\dag~Electronic Supplementary Information (ESI) available: GPC chromatogram of POH, HRMS spectra before and after metathesis, frequency dependent rheology. See DOI: 10.1039/b000000x/}
 %additional addresses can be cited as above using the lower-case letters, c, d, e... If all authors are from the same address, no letter is required
@@
 Since the pioneering work of \citet{Wichterle1960} which established their relevance in a biomedical context, polymer-based hydrogels have never ceased to be a very active field of research\cite{Calo2015,Hoffman2001}.
-They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargos or scaffolds for wound healing or tissue repair\cite{Vashist2014,Ratner2004}.
+They have recently experienced a burst of interest among the biomedical community as controlled drug delivery cargoes or scaffolds for wound healing or tissue repair\cite{Vashist2014,Ratner2004}.
 Polyelectrolytes are being extensively put to use in this particular context\cite{Rosso2003,BinImran2014}.
 Their gel formation mechanism often involves reversible-by-nature electrostatic interactions, which can be used for instance to trigger ``smart'' release of bioactive substances\cite{Khare1993,Lockwood2007}.
-Besides, poly(cationic) gels have been reported to combine scaffold properties for cell’s adherence and growth with antimicrobial activity, and are therefore increasingly used in clinical applications\cite{Hoffman2001,Landers2002}.
+Besides, poly(cationic) gels have been reported to combine antimicrobial activity with scaffold properties for the adherence and growth of cells, and are therefore increasingly used in clinical applications\cite{Hoffman2001,Landers2002}.
 In this framework, injectability is a key feature, as it provides an easier way to gel delivery in vivo\cite{Tibbitt2016}; thus, reversible shear thinning biocompatible hydrogels are a particularly sought after class of materials\cite{Yu2008}.
@@
 Contrary to a small electrolyte for which full ion pairs dissociation is generally achieved in water, polyelectrolytes in solution are in general not fully dissociated, with a substantial fraction of the counterions bound to the polymer\cite{Manning1979}.
-The resulting net charge of the chain governs the physical properties of the polyelectrolyte solution~\cite{DeGennes1976,Khokhlov1980,Raphael1990,Dobrynin1995,Rubinstein1996}, first of all the ability of the polymer to dissolve in a poor solvent.
+The resulting net charge of the chain governs the physical properties of the polyelectrolytes solution~\cite{DeGennes1976,Khokhlov1980,Raphael1990,Dobrynin1995,Rubinstein1996}, first of all the ability of the polymer to dissolve in a poor solvent.
 Counterion condensation is the physical bounding or adsorption of counterions near the polymer chain.
@@
 We have recently reported on a highly reversible poly(cation) based hydrogel\cite{Srour2014}, which formation relies on a new concept of supramolecular  electrostatic interaction.
-Briefly, the atom transfer radical polymerisation (ATRP) is initiated by a phosphonate-terminated ATRP initiator, eventually affording an anion-terminated poly(cationic) polymer.
-The resulting poly(cationic) polymer, when dispersed in an aqueous medium, stabilizes hydrogel formation through the occurrence of a head(anion)-to-body(cations) supramolecular network.
+Briefly, the atom transfer radical polymerisation (ATRP) is initiated by a phosphonate-terminated ATRP initiator, eventually affording an \emph{anion-terminated poly(cationic) polymer}.
+When dispersed in an aqueous medium, this polymer forms an hydrogel through the occurrence of a head(anion)-to-body(cations) supramolecular network.
 This highly dynamic electrostatic network provides the resulting gel with spectacular mechanical and self-healing properties.
@@
 We synthesise the phosphonate terminated polymer hereafter referred to as \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
-We obtain a well-controlled linear polymer ($M_w= \SI{8200}{\dalton}$ (NMR),  $M_w/M_n = 1.08$ from GPC; see Supplementary Materials and Supplementary Figure~S\ref{S-fig:sec}), with a degree of polymerisation $N_0=70$(NMR).
+We obtain a well-controlled linear polymer ($M_w= \SI{8200}{\dalton}$ from NMR, \SI{5614}{\dalton} from GPC with $M_w/M_n = 1.08$; see Supplementary Materials and Supplementary Figure~\figSsec), with a degree of polymerisation $N_0=70$ (NMR).
 This polymer serves as a common scaffold from which we derive  all studied systems.
 Nucleophilic addition of N-methylimidazole or N-methylpyrrolidine to a heated solution of \ce{PBr} in THF affords the corresponding poly(imidazolium) and poly(pyrrolidinium) compounds.
-In these cases, as an inherent consequence of the structure of the starting material, charge neutrality is provided by bromide counter ions, and the polymer will be referred to as \ce{PIm+Br-} and \ce{PPyr+Br-}, respectively (Scheme~\ref{sch:synthesis}).
+In these cases, as an inherent consequence of the structure of the starting material, charge neutrality is provided by bromide counter ions, and the polymer will be referred to as \ce{PIm+Br-} and \ce{PPyr+Br-}, respectively (Figure~\ref{sch:synthesis}).
 
-We performed anionic metathesis by pouring an aqueous solution of \ce{PIm+Br-} or \ce{PPyr+Br-} into a saturated aqueous solution of the different sodium halides (\ce{NaF}, \ce{NaCl}, \ce{NaI}) (Scheme~\ref{sch:metathesis}).
+We performed anionic metathesis by pouring an aqueous solution of \ce{PIm+Br-} or \ce{PPyr+Br-} into a saturated aqueous solution of the different sodium halides (\ce{NaF}, \ce{NaCl}, \ce{NaI}) (Figure~\ref{sch:metathesis}).
 In the cases of \ce{PIm+F-}, \ce{PIm+Cl-} , \ce{PPyr+F-} and \ce{PPyr+Cl-}, we obtain a turbid suspension immediately after addition.
@@
 In order to bring evidence for the efficiency of the ionic metathesis, we submit the resulting materials to negative-mode high-resolution mass spectrometry (HRMS, see Supplementary Materials).
-We unambiguously assess complete displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Supplementary Figures~S\ref{S-fig:massPImBr} and S\ref{S-fig:massPPyBr}).
-We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures~S\ref{S-fig:massPImI} and S\ref{S-fig:massPPyI}), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (not shown).
+We unambiguously assess complete displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Supplementary Figures~\figSmassPImBr and \figSmassPPyBr).
+We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures~\figSmassPImI and \figSmassPPyI), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (not shown).
 
@@
 When bromide counterions are present, we observe a significant increase of the peak temperature (114-\SI{115}{\celsius}) but again, with a similar value between \ce{PIm+Br-} and \ce{PPyr+Br-} as shown in Figure~\ref{fig:dsc}.
-The overall shape and position of the DSC peaks is very reminiscent of previously reported data for various naturally occuring or synthetic poly(electrolytes)\cite{Li2005,Sarmento2006,Ostrowska-Czubenko2009a,Moin2015}.
+The overall shape and position of the DSC peaks is very reminiscent of previously reported data for various naturally occurring or synthetic poly(electrolytes)\cite{Li2005,Sarmento2006,Ostrowska-Czubenko2009a,Moin2015}.
 It is generally attributed to desorption of weakly bound water from the polymer network.
@@
 With fluoride counterions, or in the case of \ce{PIm+Cl-}, we observe neither swelling nor dissolution.
-We recover a 8~\%~wt suspension as a biphasic mixture of its individual components.
+We recover a 7.4~\%~wt suspension as a biphasic mixture of its individual components.
 
@@
 \begin{figure*}
-\includegraphics{strain.pdf}%
+\includegraphics{strain_rsc.pdf}%
 \caption{Storage modulus $G^\prime$ ($\bullet$) and loss modulus $G^{\prime\prime}$ ($\square$) measured through oscillatory shear experiments plotted against the strain amplitude $\gamma$.
@@
 The fixed frequency is $f=\SI{1}{\hertz}$.
-All samples are at 8\%wt.
-The gray area encompasses the plastic regime that lies between the linear regime at very low amplitudes and the fluidlike shear-thinning regime at high strain amplitude.}%
+All samples are at 7.4~\%~wt.
+The gray area encompasses the plastic regime that lies between the linear regime at very low amplitudes and the fluid-like shear-thinning regime at high strain amplitude.}%
 \label{fig:strain}%
@@
 To minimize water absorption, we cover the geometry with a solvent trap, using light mineral oil as a liquid seal between the rotor and the cap.
-After one minute of equilibration we perform an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$, see Supplementary Figure~S\ref{S-fig:frequency}) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$).
+After one minute of equilibration we perform an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$, see Supplementary Figure~\figSfrequency) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$).
 For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile is that of pure water which confirms the visual observation.
 However all other samples behave as shear-thinning yield stress fluids, confirming our previous results with \ce{PIm+Br-}~\cite{Srour2014}.
-As shown in Figure~\ref{fig:strain}, soluble samples are solidlike ($G^\prime \gg G^{\prime\prime}$) at small strain and flow at large strains ($G^{\prime\prime} \gg G^\prime$) with a steep decrease of the moduli.
-We confirm solidlike behavior at low strain for all accessible frequencies (see Supplementary Material S6 and Supplementary Figure 1).
+As shown in Figure~\ref{fig:strain}, soluble samples are solid-like ($G^\prime \gg G^{\prime\prime}$) at small strain and flow at large strains ($G^{\prime\prime} \gg G^\prime$) with a steep decrease of the moduli.
+We confirm solid-like behavior at low strain for all accessible frequencies (see Supplementary Figure~\figSfrequency).
 We checked that the linear mechanical properties are unchanged by the flow history given a few minutes of rest.
@@
 Imidazolium based gels are much weaker than their pyrrolidinium counterpart.
-Comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ values (13 and \SI{8100}{\pascal}, resp.) differ by almost three orders of magnitude.
+Comparison between \ce{PIm+Br-} and \ce{PPyr+Br-} is particularly illustrative of this trend, as their $G^\prime$ values (\SI{13}{\pascal} and \SI{8100}{\pascal}, resp.) differ by almost three orders of magnitude.
 On the bottom row of Figure~\ref{fig:strain} we can compare the rheological responses of poly(pyrrolidinium)-based hydrogels for the three counterions allowing dissolution in water.
@@
 \begin{figure}
-\includegraphics{procession.pdf}
+\centering\includegraphics{procession.pdf}
 \caption{Processionary model.
@@
 Electrostatic blobs containing an anionic head are shown as black filled circles.}
-\label{fig:dissociation}
+\label{fig:procession}
 \end{figure}
+\begin{figure*}
+\centering\includegraphics{scales.pdf}
+\caption{Conformation of semi-dilute polyelectrolytes at rest at various scales. Length increases from left to right.}
+\label{fig:scales}
+\end{figure*}
 
@@
 the elasticity of the undamaged gel network, (ii) the strain amplitude $\gamma_0$ corresponding to the end of the linear regime~\cite{Hyun2011}, and (iii) the strain amplitude $\gamma_c$ corresponding to the crossing of the moduli.
-By using these three parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters such as the average number of chains between cross-links $n$, the number of monomers between dissociated ion pairs $A$ and the head-to-body bonding energy $E_c$.
+By using these three parameters, we show that it is possible, with reasonable assumptions, to estimate microscopic parameters such as the average number of chains between cross-links $n$, the number of monomers between dissociated ion pairs $A$ that quantify counterion condensation and the head-to-body bonding energy $E_c$.
 
 \subsubsection{Conformation at rest.}
-\begin{figure*}
-\includegraphics{phase_diag.pdf}
-\caption{Conformation of semi-dilute polyelectrolytes at rest at various scales as a function of dissociated charge frequency $A$.
-The phase diagram is given for a reduced temperature $\tau=0.46$ and for the monomer number density of \ce{PIm+Br-}, although $c$ is slightly different for each sample.
-Each vertical line represents a sample.}
-\label{fig:phasediag}
-\end{figure*}
-
 The scaling theory of \citet{Dobrynin1995} describes the structure of a polyelectrolyte chain.
 In the following, we make the link between this theory and the chemical structure of our polymers.
-In absence of strain, the structure of a polyelectrolyte chain is organised at various scales, sketched on Figure~\ref{fig:phasediag}.
+In absence of strain, the structure of a polyelectrolyte chain is organised at various scales, sketched on Figure~\ref{fig:scales}.
 
@@
 \paragraph{Electrostatic length, weak charging case.}
-Following \citet{Khokhlov1980} one can further define a third characteristic length $D$ over which the electrostatic energy become dominant.
-balances the surface energy of the collapsed polymer.
+Following \citet{Khokhlov1980} one can further define a third characteristic length $D$ over which the electrostatic energy become dominant over short range attraction or surface energy of the collapsed polymer.
 $D$ defines the size of the electrostatic blob and we name $g_e$ the number of monomers in it.
@@
 Counterion condensation is quantified by assuming that there is a charge every $A$ monomers.
-Therefore there are $g_e/A$ charges in the electrostatic blob and the electrostatic energy reads $E_e = (g_e/A)^2 e^2/(4\pi\epsilon)$, with $e$ the elementary charge and $\epsilon$ the dielectric constant of the solvent.
+Therefore there are $g_e/A$ charges in the electrostatic blob and the electrostatic energy reads $E_e = (g_e/A)^2 e^2/(4\pi\epsilon D)$, with $e$ the elementary charge and $\epsilon$ the dielectric constant of the solvent.
 If $D>\xi_T$ the main opposition to the electrostatic energy can be understood as a surface energy proportional to the number of thermal blobs exposed on the surface $E_s = k_\mathrm{B}T (D/\xi_T)^2$.
@@
 This explains why the respective modulus or critical strains of \ce{PPyr+I-} (heavier, lower number density) and \ce{PPyr+Br-} are in reverse order with respect to the Hofmeister series.
-Taking into account the molecular mass $M$ of each polymer and  knowing the density $d$ of the solvent, we obtain the number $n = N/N_0$ of chains between cross-link
+Taking into account the molecular mass $M$ of each polymer, the Avogadro number $\mathcal{N_A}$ and  knowing the density $d$ of the solvent, we obtain the number $n = N/N_0$ of chains between cross-link
 \begin{equation}
@@
 
-We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 860 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
+We find that the number $n$ of chains between cross-link point goes from 1 in \ce{PPyr+I-} and \ce{PPy+Br-} to 800 in \ce{PIm+Br-}, following the \textit{a priori} ranking of charge dissociation, see Table~\ref{tab:results}.
 This larger cross-link ratio is consistent with a higher probability of attaching two or more heads on a highly dissociated body.
@@
 As the material is strained, individual processions are stretched, starting from the larger scales~\cite{Pincus1976}.
-When the the correlation blobs are fully stretched, the procession is a linear assembly of electrostatic blobs.
+As sketched in Figure~\ref{fig:stretch}, when the the correlation blobs are fully stretched, the procession is a linear assembly of electrostatic blobs.
 Stretching the chain further means exposing more thermal blobs to the solvent, a process that cannot be expressed using the model of a spring of constant stiffness.
@@
 \subsubsection{Extension of the electrostatic blobs.}
+\begin{figure}
+\centering\includegraphics{stretch.pdf}
+\caption{Shearing the gel translates into stretching individual processions.
+From top to bottom: configuration at rest; full extension of the self avoiding walk; stretching of the electrostatic blobs.}
+\label{fig:stretch}
+\end{figure}
 
@@
 
+\begin{figure}
+\includegraphics{phase_diag.pdf}
+\caption{Conformation of semi-dilute polyelectrolytes at rest at various scales as a function of counterion condensation parameter $A$.
+The phase diagram is given for a reduced temperature $\tau=0.46$ and for the monomer number density of \ce{PIm+Br-}, although $c$ is slightly different for each sample.
+Each vertical line represents a sample.}
+\label{fig:phasediag}
+\end{figure}
+
 Table~\ref{tab:results} also shows the values of $A$ deduced from Equation~(\ref{eq:A}) and the measured value of $\tau$.
@@
 This abrupt transition between an uniform gel and a precipitate is what separates our five soluble samples from insoluble \ce{PPyr+F-}, \ce{PIm+Cl-} and \ce{PIm+F-}.
-Our rheological measurements probe the structure of the procession from the Kuhn length to the distance between crosslinks that is approximately $r_\mathrm{scr}$.
+Our rheological measurements probe the structure of the procession from the Kuhn length to the distance between crosslinks that is approximately $r_\mathrm{scr}$ (upper limit of the white region in Figure~\ref{fig:phasediag}).
 \ce{PPyr+Cl-} and \ce{PIm+I-} display the typical structure of weakly charged polyelectrolytes in poor solvent.
@@
 and N.T.
-acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007–2013)/ERC grant agreement No.258803.
+acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013)/ERC grant agreement No.258803.
 The Graphical TOC entry is derived from a picture by Arturo Reina, CC BY-SA 3.0, \url{https://commons.wikimedia.org/w/index.php?curid=282898}.

### 2016-10-24 ion_pairing_SI.tex
NMR integration at n=70

@@
 \caption{1H NMR of POH in deuterated methanol, with signal attribution.
-Calibration of the integrals intensity, relative to the distinct chain-end signals (a and b, 6 protons each), allows us to determine polymer chain length $n$ (here, considering integration of signal d, $n=63$)}
+Calibration of the integrals intensity, relative to the distinct chain-end signals (a and b, 6 protons each), allows us to determine polymer chain length $n$ (here, considering integration of signal d,  $n=69-70$, $Mn=\SI{8335}{\dalton}$)}
 \end{figure}

### 2016-10-25 PALSE/response.tex
Réponse de Cyrille remaniées

@@
 }
-\opening{\bf Prof. Paul Janmey,}
+\opening{\bf Dear Prof. Paul Janmey,}
 
@@
 
-Although the methodology used for polymer chain length determination has been reported in full in one of our previous papers (Appukuttan et al.,2012), we decided, following referee's recommendation, and for reader's comfort, to include a 1H NMR spectrum of the precursor POH polymer used in this study.
-In the caption, the calculation used to determine polymer chain length is explained.
+This is indeed a difficult question ; although most papers dealing with polymer chemistry present the molecular weight calculated by NMR as a Mn, we personally think that Mw better fits: the methodology is indeed based on the integration of repeating unit protons, and heavy chains are overexpressed in this integration.
+However, on the basis that (1) The NMR calculated value can anyway, strictly speaking, neither be considered as a ``pure''  Mn or Mw (2) Mn agrees better with existing literature (3) Polydispersity is anyway low, we changed the sentence to ``Mn =\SI{8300}{\dalton} from NMR, \SI{5125}{\dalton} from GPC''.
 
@@
 The peaks of Br- and I- should be 79/81 and 126.
-The content of the text does not fit to what is shown in S2 and S3.\end{quotationi}
-
-We thank the Reviewer for spotting this typo.
+The content of the text does not fit to what is shown in S2 and S3.
+\end{quotationi}
+We thank the Reviewer for spotting this typo. We quite fortunately agree with the periodic table.
 
@@
 \end{quotationi}
+We think that the complete disappearance of any trace of Br- from the MS spectrum after addition of either NaCl or NaF is a clear evidence that the process occurred successfully.
+We initially thought of other alternative methods, namely atomic absorption  and elemental analysis: the former is no more available in our laboratories; and we know that the latter is practically really tricky on those highly hygroscopic polymers, as water tends to accumulate within the sample during either storage or preparation (as seen in the DSC spectrum) making it very difficult to get reliable data.
 
@@
 
-Done.
+Thank you for the suggestion, the figure has been added as Figure S2; the caption depicts the methodology used for Dp/Mn calculation.
 
@@
 
+Done.
+
 \begin{quotationi}
@@
 
+In many instances we tried to perform DLS measurements on our samples and, indeed, signals were observed.
+However the obtained results were largely unreproducible (even for a given sample upon slight change of the concentration, not to mention when two different samples were used).
+An example is shown below.
+
+\includegraphics[width=\textwidth]{DLS.png}
+\textit{Figure: \textbf{a} DLS data and best fit \textbf{b} size distribution by diffusion intensity and \textbf{c} size distribution by number for PImBr (top row) and PPyrBr (bottom row) in water at 10\% weight concentrations}
+
+As the Reviewer can see, dispersion was very high, and signal quality so low that cumulant fit analyses were absolutely irrelevant.
+In our opinion, there is nothing relevant in these data, this is why we finally chose not to include it in the SI.
+If the referee thinks that it might be worth adding some of these data to the article, we are ready to do it.
+However, we do know whether a reasonable interpretation could be provided from the data.
+We leave the referee judge; top is PImBr bottom PPyrBr.
+Although a pattern seems to exist, it seems difficult to us to get its exact meaning, and to separate the artefact from the fact\ldots
+
 \begin{quotationi}
@@
 \end{quotationi}
+
+To our understanding, loose end corrections describe a situation where pre-existing linear chains of size $M$ are randomly cross-linked with an average of $M_c$ monomers between cross links with $M_c\leq M/2$.
+The modulus then scales as
+\begin{equation}
+\frac{G}{k_\mathrm{B}T} = \frac{c}{M_c}  \left(1 - \frac{2M_c}{M}\right).
+\end{equation}
+For example if $M_c$ is only one third of $M$, then two third of the original linear chain are loose ends and do not contribute to the elasticity.
+
+This is rather different from our situation where the number of monomers between cross links is larger or equal to the number of monomers in the original chain.
+This situation stems from the reactivity of the heads: a cross link necessarily occurs at the head of a polymer, preventing one of the possible loose end, but leaving the possibility of many loose ends at the tails of each polymer.
+
+We can better compare our situation to a gel formed by branched polymers, for example bottle brush polymers.
+In this situation, Papagiannopoulos et al.
+(\href{http://doi.org/10.1021/bm060287d}{Biomacromolecules 2006} or \href{http://doi.org/10.1039/b714864j}{Faraday Discussions 2008}) have shown that the shear modulus scales as
+\begin{equation}
+\frac{G}{k_\mathrm{B}T} = \frac{c}{N_\mathrm{tot}}.
+\end{equation}
+where $N_\mathrm{tot}$ is the total number of monomers between cross-links, including side chains that are topological `loose ends'.
+
+In other words, one can consider two possible definitions of ``cross link''.
+In Flory's definition, every branching point is a cross link, therefore a brush has many loose ends.
+In the more coarse-grained definition used by Papagiannopoulos et al., branching points leading to a loose end are not considered meaningful cross links, but every monomer of the side chains actually contribute to entropy and thus to elasticity.
+
+We added a paragraph at the end of 2.4.2 to stress this difference.
 
@@
 Are the results sensitive to addition of a background electrolyte (e.g., similar to physiological ionic strength)?
+\end{quotationi}
 
-\end{quotationi}
+We thank the Reviewer for this excellent suggestion.
+We added a paragraph and Figure 8b to describe the predictions of our model concerning the consequences of a background electrolyte concentration similar to physiological ionic strength.
+We think it demonstrates quite well the power of our model as well as the robustness of our claims concerning potential biomedical applications.
+However making our gels biocompatible would demand a careful selection of the ion pairs to ensure lack of toxicity, a feature that is out of the scope of the present paper.
 

### 2016-10-27 response.tex
resubmitted version

@@
 
-\textsf{\textbf{Replies to the comments of Reviewer \#3}}
+\textsf{\textbf{Replies to the comments of Reviewer \#2}}
 
@@
 \includegraphics[width=\textwidth]{DLS.png}
-\textit{Figure: \textbf{a} DLS data and best fit \textbf{b} size distribution by diffusion intensity and \textbf{c} size distribution by number for PImBr (top row) and PPyBr (bottom row) in water at 10\% weight concentrations}
+\textit{Figure: \textbf{a} DLS data and best fit \textbf{b} size distribution by diffusion intensity and \textbf{c} size distribution by number for PImBr (top row) and PPyrBr (bottom row) in water at 10\% weight concentrations}
 
@@
 However, we do know whether a reasonable interpretation could be provided from the data.
-We leave the referee judge; top is PImBr bottom PPyBr.
+We leave the referee judge; top is PImBr bottom PPyrBr.
 Although a pattern seems to exist, it seems difficult to us to get its exact meaning, and to separate the artefact from the fact\ldots
@@
 
+We added a paragraph at the end of 2.4.2 to stress this difference.
+
 \begin{quotationi}

### 2016-10-27 ion_pairing_paper.tex
resubmitted version

@@
 \newcommand{\figSsec}{S1\xspace}
-\newcommand{\figSmassPImBr}{S2\xspace}
-\newcommand{\figSmassPPyBr}{S3\xspace}
-\newcommand{\figSmassPImI}{S4\xspace}
-\newcommand{\figSmassPPyI}{S5\xspace}
-\newcommand{\figSfrequency}{S6\xspace}
+\newcommand{\figSNMR}{S2\xspace}
+\newcommand{\figSmassPImBr}{S3\xspace}
+\newcommand{\figSmassPPyBr}{S4\xspace}
+\newcommand{\figSmassPImI}{S5\xspace}
+\newcommand{\figSmassPPyI}{S6\xspace}
+\newcommand{\figSfrequency}{S7\xspace}
 
@@
 
- & \noindent\large{Hassan Srour,\textit{$^{a}$} Martien Duvall Deffo Ayagou,\textit{$^{a}$} Thi Thanh-Tam Nguyen,\textit{$^{a}$}, Nicolas Taberlet,\textit{$^{b}$} S\'{e}bastien Manneville,\textit{$^{b}$} Chantal Andraud,\textit{$^{a}$} Cyrille Monnereau,\textit{$^{a}$} and Mathieu Leocmach$^{\ast}$\textit{$^{c}$}} \\%Author names go here instead of "Full name", etc.
+ & \noindent\large{Hassan Srour,\textit{$^{a}$} Martien Duvall Deffo Ayagou,\textit{$^{a}$} Thi Thanh-Tam Nguyen,\textit{$^{a}$}, Nicolas Taberlet,\textit{$^{b}$} S\'{e}bastien Manneville,\textit{$^{b}$} Chantal Andraud,\textit{$^{a}$} Cyrille Monnereau,$^{\ast}$\textit{$^{a}$} and Mathieu Leocmach$^{\ast}$\textit{$^{c}$}} \\%Author names go here instead of "Full name", etc.
 
@@
 
-\footnotetext{\textit{$^{a}$~Univ Lyon, Ens de Lyon, Univ Claude Bernard, CNRS,
-Laboratoire de Chimie, F-69342 Lyon, France. }}%Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
-\footnotetext{\textit{$^{b}$~Univ Lyon, Ens de Lyon, Univ Claude Bernard, CNRS,
+\footnotetext{\textit{$^{a}$~Univ Lyon, Ens de Lyon, Universit\'e Claude Bernard Lyon 1, CNRS,
+Laboratoire de Chimie, F-69342 Lyon, France. E-mail: cyrille.monnereau@ens-lyon.fr}}%Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
+\footnotetext{\textit{$^{b}$~Univ Lyon, Ens de Lyon, Universit\'e Claude Bernard Lyon 1, CNRS,
 Laboratoire de Physique, F-69342 Lyon, France. }}
-\footnotetext{\textit{$^{c}$~Univ Lyon, Univ Claude Bernard, CNRS, Institut Lumi\`{e}re Mati\`{e}re, F-69622 Villeurbanne, France.
-E-mail: mathieu.leocmach@univ-lyon1.fr}}%Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
+\footnotetext{\textit{$^{c}$~Univ Lyon, Universit\'e Claude Bernard Lyon 1, CNRS, Institut Lumi\`ere Mati\`ere, F-69622, VILLEURBANNE, France.
+E-mail: mathieu.leocmach@univ-lyon1.fr.
+Twitter: @LamSonLeoc}}%Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
 
@@
 %If you article does not have ESI please remove the the \dag symbol from the title and the footnotetext below.
-\footnotetext{\dag~Electronic Supplementary Information (ESI) available: GPC chromatogram of POH, HRMS spectra before and after metathesis, frequency dependent rheology. See DOI: 10.1039/b000000x/}
+\footnotetext{\dag~Electronic Supplementary Information (ESI) available: GPC chromatogram and 1H NMR of POH, HRMS spectra before and after metathesis, frequency-dependent rheology.
+See DOI: 10.1039/b000000x/}
 %additional addresses can be cited as above using the lower-case letters, c, d, e... If all authors are from the same address, no letter is required
@@
 We call this behaviour ``processionary'' in analogy to pine processionary caterpillar (\textit{Thaumetopoea pityocampa}) behaviour\cite{Fabre1916}.
-In particular, this model enables the quantification of the charge condensation rate from standard rheological measurements.%Finally, we discuss our results and possible applications of such smart hydrogels.
+In particular, this model enables the quantification of the charge condensation rate from standard rheological measurements.
+\textcolor{blue}{In a fifth part we apply this model to discuss the consequences of a background electrolyte at physiological ionic strength.} %Finally, we discuss our results and possible applications of such smart hydrogels.
 
@@
 We synthesise the phosphonate terminated polymer hereafter referred to as \ce{PBr} according to our previously reported methodology\cite{Srour2014,Appukuttan2012}.
-We obtain a well-controlled linear polymer ($M_w= \SI{8200}{\dalton}$ from NMR, \SI{5614}{\dalton} from GPC with $M_w/M_n = 1.08$; see Supplementary Materials and Supplementary Figure~\figSsec), with a degree of polymerisation $N_0=70$ (NMR).
+We obtain a well-controlled linear polymer (\textcolor{blue}{$M_n= \SI{8300}{\dalton}$ from NMR, \SI{5125}{\dalton} from GPC with $M_w/M_n = 1.09$}; see Supplementary Materials and Supplementary Figures~\figSsec and~\figSNMR), with a degree of polymerisation $N_0=70$ (NMR).
 This polymer serves as a common scaffold from which we derive  all studied systems.
@@
 In order to bring evidence for the efficiency of the ionic metathesis, we submit the resulting materials to negative-mode high-resolution mass spectrometry (HRMS, see Supplementary Materials).
-We unambiguously assess complete displacement of bromide counterions by the full disappearance of the diisotopic mass peak (89/\SI{91}{\dalton}, Supplementary Figures~\figSmassPImBr and \figSmassPPyBr).
-We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{121}{\dalton}, Supplementary Figures~\figSmassPImI and \figSmassPPyI), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (not shown).
+We unambiguously assess complete displacement of bromide counterions by the full disappearance of the diisotopic mass peak (79/\SI{81}{\dalton}, Supplementary Figures~\figSmassPImBr and \figSmassPPyBr).
+We further confirm substitution by iodide by the concomitant apparition of a characteristic monoisotopic peak (\SI{126}{\dalton}, Supplementary Figures~\figSmassPImI and \figSmassPPyI), while fluoride and chloride anions signals, which are below the detection limits of the spectrometer, could not be observed in the corresponding polymers (not shown).
 
@@
 Water-swelling properties turn out to be extremely dependent on the polymer composition.
-With bromide or iodide conterions, both poly(pyrrolidinium) and poly(immidazolium) polymers afford homogeneous and optically transparent gels upon swelling with deionised water.
+With \ce{Br-} or \ce{I-} conterions, both poly(pyrrolidinium) and poly(immidazolium) polymers afford homogeneous and optically transparent gels upon swelling with deionised water.
 We observe a limited swelling for \ce{PPyr+Cl-}, resulting in a granular, inhomogeneous gel.
-With fluoride counterions, or in the case of \ce{PIm+Cl-}, we observe neither swelling nor dissolution.
+With \ce{F-} counterions, or in the case of \ce{PIm+Cl-}, we observe neither swelling nor dissolution.
 We recover a 7.4~\%~wt suspension as a biphasic mixture of its individual components.
@@
 The fixed frequency is $f=\SI{1}{\hertz}$.
-All samples are at 7.4~\%~wt.
+All samples are at 7.4~wt\%.
 The gray area encompasses the plastic regime that lies between the linear regime at very low amplitudes and the fluid-like shear-thinning regime at high strain amplitude.}%
@@
 
-The mechanical properties at low strain evolve with the factors influencing counter-ion condensation.
+\textcolor{blue}{We observe that} the mechanical properties at low strain evolve with the factors influencing counter-ion condensation.
 Imidazolium based gels are much weaker than their pyrrolidinium counterpart.
@@
 
-Let us note $N$ the number of monomers between two crosslinks or entanglements.
+Let us note $N$ the number of monomers between two cross-links or entanglements.
 If $N>g$ the ``procession'' of polymer chains performs a random walk of correlation blobs, each being a self-avoiding walk.
@@
 This larger cross-link ratio is consistent with a higher probability of attaching two or more heads on a highly dissociated body.
+
+\textcolor{blue}{We stress that $N$ in Equation~(\ref{eq:G}) is the total number of monomers between effective cross-links, including possible side chains and loose ends.
+In that we follow results from the literature of branched polymer gels~\cite{Papagiannopoulos2008} that are more tractable than the original theory from Flory~\cite{Flory1944} and Case~\cite{Case1960} for randomly cross-linked linear chains with loose end corrections, where every junction between two polymers is a cross-link.}
 
@@
 As the material is strained, individual processions are stretched, starting from the larger scales~\cite{Pincus1976}.
-As sketched in Figure~\ref{fig:stretch}, when the the correlation blobs are fully stretched, the procession is a linear assembly of electrostatic blobs.
+As sketched in Figure~\ref{fig:stretch}, when the correlation blobs are fully stretched, the procession is a linear assembly of electrostatic blobs.
 Stretching the chain further means exposing more thermal blobs to the solvent, a process that cannot be expressed using the model of a spring of constant stiffness.
@@
 Incidentally, this implies that the procession is too short to perform the random walk.
-In other words the number $N$ of monomers between crosslinks is comparable to $g_\mathrm{scr}$.
-Using equation (\ref{eq:gscr}), this condition translates into
+In other words the number $N$ of monomers between cross-links is comparable to $g_\mathrm{scr}$.
+Using Equation (\ref{eq:gscr}), this condition translates into
 \begin{equation}
@@
 \caption{Conformation of semi-dilute polyelectrolytes at rest at various scales as a function of counterion condensation parameter $A$.
-The phase diagram is given for a reduced temperature $\tau=0.46$ and for the monomer number density of \ce{PIm+Br-}, although $c$ is slightly different for each sample.
-Each vertical line represents a sample.}
+Each vertical line represents a sample.
+The phase diagrams are given for a reduced temperature $\tau=0.46$.
+\textcolor{blue}{(a) In pure water.
+Background screening length $\kappa^{-1}$ comes from the \ce{PO3^2-} heads.
+$c$ is set by \ce{PIm+Br-}.
+(b) With a background electrolyte at $c_s=\SI{100}{\milli M}$.
+$c$ is set by \ce{PPyr+Cl-}.
+Variations in $c$ values from sample to sample have little impact on the phase boundaries, except near the crossing of $r_\mathrm{scr}$ and $D$.}}
 \label{fig:phasediag}
@@
 
-Figure~\ref{fig:phasediag} summarizes the different conformations of a procession as a function of scale and counterion condensation for our experimental $c$ and $\tau$.
-The most striking feature of this phase diagram is the narrowing of the regime between $D$ and $r_\mathrm{scr}$ (white region in Figure~\ref{fig:phasediag}) as counterion condensation increases.
+Figure~\ref{fig:phasediag}a summarizes the different conformations of a procession as a function of scale and counterion condensation for our experimental $c$ and $\tau$.
+The most striking feature of this phase diagram is the narrowing of the regime between $D$ and $r_\mathrm{scr}$ (white region in Figure~\ref{fig:phasediag}a) as counterion condensation increases.
 As the size of the electrostatic blob crosses the screening length, electrostatic repulsion is no longer able to sustain solubility.
 This abrupt transition between an uniform gel and a precipitate is what separates our five soluble samples from insoluble \ce{PPyr+F-}, \ce{PIm+Cl-} and \ce{PIm+F-}.
-Our rheological measurements probe the structure of the procession from the Kuhn length to the distance between crosslinks that is approximately $r_\mathrm{scr}$ (upper limit of the white region in Figure~\ref{fig:phasediag}).
+Our rheological measurements probe the structure of the procession from the Kuhn length to the distance between cross-links that is approximately $r_\mathrm{scr}$ (upper limit of the white region in Figure~\ref{fig:phasediag}a).
 \ce{PPyr+Cl-} and \ce{PIm+I-} display the typical structure of weakly charged polyelectrolytes in poor solvent.
@@
 From purely rheological measurements, we have thus obtained microscopic information on the state of the gel: amount of charge condensation, head-to-body bonding energy and size of a procession.
+
+{\color{blue}
+\subsection{Application to physiological ionic strength}
+
+To demonstrate the versatility of our model we expose briefly the consequences of having a background electrolyte with ionic strength similar to the physiological one and with the same counterion (e.g.
+\ce{Na+X-}).
+We set the salt concentration $c_s$ to \SI{100}{\milli M}, implying a background Debye length $\kappa^{-1}\approx 13 b$ and $B_s$ close to unity.
+We show the resulting phase diagram in Figure~\ref{fig:phasediag}b.
+The background electrolyte lowers the horizontal asymptote of $r_\mathrm{scr}$, which strongly reduces the white region of the phase diagram.
+\ce{PIm+I-} and \ce{PIm+Br-} have now lost the possibility to form linear chains, such that their electrostatic blobs are now space-filling.
+Therefore these two gels should precipitate at biological ionic strength.
+By contrast \ce{PPyr+Br-} and \ce{PPyr+I-} would still have linear chains between $D$ and $r_\mathrm{scr}$ and would stay soluble in a robust gel state.
+
+Let us consider the case of \ce{PPyr+Cl-} whose chloride counterions are interesting from a physiological point of view.
+If cross-links still occur at the scale of the now shorter screening length, we expect less monomers between cross-links and thus larger shear modulus.
+Combining Equation (\ref{eq:gscr}) and (\ref{eq:G}) we obtain a modulus of \SI{780}{\pascal}, almost four times larger than \ce{PPyr+Cl-} in pure water.
+However \ce{PPyr+Cl-} would be a limit case, stable for an ionic strength of \SI{100}{\milli M} but precipitating at \SI{150}{\milli M} which is still in the range of physiological ionic strength.
+To ensure gel stability, it would thus be necessary to use cationic moeties with even less counterion condensation than pyrrolidinium.
+}
 

### 2017-02-17 PALSE/talk_Kyoto.tex
Replace white-backgroung png of Investissement d'avernir logo with vectorial pdf

@@
 		\multicolumn{2}{c}{Martien Duvall Deffo Ayagou}&\\
-		\multicolumn{2}{c}{Thi Thanh-Tam Nguyen}&\raisebox{-0.1em}[0pt][0pt]{\includegraphics[height=2\baselineskip]{presentation/invest-avenir.jpg}}&\raisebox{-0.1em}[0pt][0pt]{\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{../Yaourt/NEW-Logo-ERC-OUTLINE}}\\
+		\multicolumn{2}{c}{Thi Thanh-Tam Nguyen}&\raisebox{-0.1em}[0pt][0pt]{\includegraphics[height=2\baselineskip]{presentation/invest-avenir}}&\raisebox{-0.1em}[0pt][0pt]{\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{../Yaourt/NEW-Logo-ERC-OUTLINE}}\\
 	\end{tabu}
